### 2017-01-26 bare_conf.tex
layout

%% bare_conf.tex
%% V1.4b
%% 2015/08/26
%% by Michael Shell
%% See:
%% http://www.michaelshell.org/
%% for current contact information.
%%
%% This is a skeleton file demonstrating the use of IEEEtran.cls
%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
%% conference paper.
%%
%% Support sites:
%% http://www.michaelshell.org/tex/ieeetran/
%% http://www.ctan.org/pkg/ieeetran
%% and
%% http://www.ieee.org/

%%*************************************************************************
%% Legal Notice:
%% This code is offered as-is without any warranty either expressed or
%% implied; without even the implied warranty of MERCHANTABILITY or
%% FITNESS FOR A PARTICULAR PURPOSE!
%% User assumes all risk.
%% In no event shall the IEEE or any contributor to this code be liable for
%% any damages or losses, including, but not limited to, incidental,
%% consequential, or any other damages, resulting from the use or misuse
%% of any information contained here.
%%
%% All comments are the opinions of their respective authors and are not
%% necessarily endorsed by the IEEE.
%%
%% This work is distributed under the LaTeX Project Public License (LPPL)
%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
%% distributed and modified. A copy of the LPPL, version 1.3, is included
%% in the base LaTeX documentation of all distributions of LaTeX released
%% 2003/12/01 or later.
%% Retain all contribution notices and credits.
%% ** Modified files should be clearly indicated as such, including  **
%% ** renaming them and changing author support contact information. **
%%*************************************************************************

% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
% *** with production work. The IEEE's font choices and paper sizes can   ***
% *** trigger bugs that do not appear when using other class files.       ***                          ***
% The testflow support page is at:
% http://www.michaelshell.org/tex/testflow/

\documentclass[conference]{IEEEtran}
% Some Computer Society conferences also require the compsoc mode option,
% but others use the standard conference format.
%
% If IEEEtran.cls has not been installed into the LaTeX system files,
% manually specify the path to it like:
% \documentclass[conference]{../sty/IEEEtran}

% Some very useful LaTeX packages include:
% (uncomment the ones you want to load)

% *** MISC UTILITY PACKAGES ***
%
%\usepackage{ifpdf}
% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
% compilation based on whether the output is pdf or dvi.
% usage:
% \ifpdf
%   % pdf code
% \else
%   % dvi code
% \fi
% The latest version of ifpdf.sty can be obtained from:
% http://www.ctan.org/pkg/ifpdf
% Also, note that IEEEtran.cls V1.7 and later provides a builtin
% \ifCLASSINFOpdf conditional that works the same way.
% When switching from latex to pdflatex and vice-versa, the compiler may
% have to be run twice to clear warning/error messages.

% *** CITATION PACKAGES ***
%
%\usepackage{cite}
% cite.sty was written by Donald Arseneau
% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
% \cite{} output to follow that of the IEEE. Loading the cite package will
% result in citation numbers being automatically sorted and properly
% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
% \cite will automatically add leading space, if needed. Use cite.sty's
% noadjust option (cite.sty V3.8 and later) if you want to turn this off
% such as if a citation ever needs to be enclosed in parenthesis.
% cite.sty is already installed on most LaTeX systems. Be sure and use
% version 5.0 (2009-03-20) and later if using hyperref.sty.
% The latest version can be obtained at:
% http://www.ctan.org/pkg/cite
% The documentation is contained in the cite.sty file itself.

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
  % \usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../pdf/}{../jpeg/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else
  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
  % will default to the driver specified in the system graphics.cfg if no
  % driver is specified.
  % \usepackage[dvips]{graphicx}
  % declare the path(s) where your graphic files are
  % \graphicspath{{../eps/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
  % \DeclareGraphicsExtensions{.eps}
\fi
% graphicx was written by David Carlisle and Sebastian Rahtz. It is
% required if you want graphics, photos, etc. graphicx.sty is already
% installed on most LaTeX systems. The latest version and documentation
% can be obtained at:
% http://www.ctan.org/pkg/graphicx
% Another good source of documentation is "Using Imported Graphics in
% LaTeX2e" by Keith Reckdahl which can be found at:
% http://www.ctan.org/pkg/epslatex
%
% latex, and pdflatex in dvi mode, support graphics in encapsulated
% postscript (.eps) format. pdflatex in pdf mode supports graphics
% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
% which can result in "jaggedy"/blurry rendering of lines and letters as
% well as large increases in file sizes.
%
% You can find documentation about the pdfTeX application at:
% http://www.tug.org/applications/pdftex

% *** MATH PACKAGES ***
%
%\usepackage{amsmath}
% A popular package from the American Mathematical Society that provides
% many useful and powerful commands for dealing with mathematics.
%
% Note that the amsmath package sets \interdisplaylinepenalty to 10000
% thus preventing page breaks from occurring within multiline equations. Use:
%\interdisplaylinepenalty=2500
% after loading amsmath to restore such page breaks as IEEEtran.cls normally
% does. amsmath.sty is already installed on most LaTeX systems. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/amsmath

% *** SPECIALIZED LIST PACKAGES ***
%
%\usepackage{algorithmic}
% algorithmic.sty was written by Peter Williams and Rogerio Brito.
% This package provides an algorithmic environment fo describing algorithms.
% You can use the algorithmic environment in-text or within a figure
% environment to provide for a floating algorithm. Do NOT use the algorithm
% floating environment provided by algorithm.sty (by the same authors) or
% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
% algorithm float types and packages that provide these will not provide
% correct IEEE style captions. The latest version and documentation of
% algorithmic.sty can be obtained at:
% http://www.ctan.org/pkg/algorithms
% Also of interest may be the (relatively newer and more customizable)
% algorithmicx.sty package by Szasz Janos:
% http://www.ctan.org/pkg/algorithmicx

% *** ALIGNMENT PACKAGES ***
%
%\usepackage{array}
% Frank Mittelbach's and David Carlisle's array.sty patches and improves
% the standard LaTeX2e array and tabular environments to provide better
% appearance and additional user controls. As the default LaTeX2e table
% generation code is lacking to the point of almost being broken with
% respect to the quality of the end results, all users are strongly
% advised to use an enhanced (at the very least that provided by array.sty)
% set of table tools. array.sty is already installed on most systems. The
% latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/array

% IEEEtran contains the IEEEeqnarray family of commands that can be used to
% generate multiline equations as well as matrices, tables, etc., of high
% quality.

% *** SUBFIGURE PACKAGES ***
%\ifCLASSOPTIONcompsoc
%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
%\else
%  \usepackage[caption=false,font=footnotesize]{subfig}
%\fi
% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
% for subfigure.sty, the latter of which is no longer maintained and is
% incompatible with some LaTeX packages including fixltx2e. However,
% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
% which will override IEEEtran.cls' handling of captions and this will result
% in non-IEEE style figure/table captions. To prevent this problem, be sure
% and invoke subfig.sty's "caption=false" package option (available since
% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
% handling of captions.
% Note that the Computer Society format requires a larger sans serif font
% than the serif footnote size font used in traditional IEEE formatting
% and thus the need to invoke different subfig.sty package options depending
% on whether compsoc mode has been enabled.
%
% The latest version and documentation of subfig.sty can be obtained at:
% http://www.ctan.org/pkg/subfig

% *** FLOAT PACKAGES ***
%
%\usepackage{fixltx2e}
% fixltx2e, the successor to the earlier fix2col.sty, was written by
% Frank Mittelbach and David Carlisle. This package corrects a few problems
% in the LaTeX2e kernel, the most notable of which is that in current
% LaTeX2e releases, the ordering of single and double column floats is not
% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
% single column figure to be placed prior to an earlier double column
% figure.
% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
% corrections already built into the system in which case a warning will
% be issued if an attempt is made to load fixltx2e.sty as it is no longer
% needed.
% The latest version and documentation can be found at:
% http://www.ctan.org/pkg/fixltx2e

%\usepackage{stfloats}
% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
% the ability to do double column floats at the bottom of the page as well
% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
% LaTeX2e). It also provides a command:
%\fnbelowfloat
% to enable the placement of footnotes below bottom floats (the standard
% LaTeX2e kernel puts them above bottom floats). This is an invasive package
% which rewrites many portions of the LaTeX2e float routines. It may not work
% with other packages that modify the LaTeX2e float routines. The latest
% version and documentation can be obtained at:
% http://www.ctan.org/pkg/stfloats
% Do not use the stfloats baselinefloat ability as the IEEE does not allow
% \baselineskip to stretch. Authors submitting work to the IEEE should note
% that the IEEE rarely uses double column equations and that authors should try
% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
% such ways.
% Do not attempt to use stfloats with fixltx2e as they are incompatible.
% Instead, use Morten Hogholm'a dblfloatfix which combines the features
% of both fixltx2e and stfloats:
%
% \usepackage{dblfloatfix}
% The latest version can be found at:
% http://www.ctan.org/pkg/dblfloatfix

% *** PDF, URL AND HYPERLINK PACKAGES ***
%
%\usepackage{url}
% url.sty was written by Donald Arseneau. It provides better support for
% handling and breaking URLs. url.sty is already installed on most LaTeX
% systems. The latest version and documentation can be obtained at:
% http://www.ctan.org/pkg/url
% Basically, \url{my_url_here}.

% *** Do not adjust lengths that control margins, column widths, etc. ***
% *** Do not use packages that alter fonts (such as pslatex).         ***
% There should be no need to do such things with IEEEtran.cls V1.6 and later.
% (Unless specifically asked to do so by the journal or conference you plan
% to submit to, of course. )

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}
%
% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{How Important are Usage Examples in REST API Documentation?}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Michael Shell}
\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
Georgia Institute of Technology\\
Atlanta, Georgia 30332--0250\\
Email: http://www.michaelshell.org/contact.html}
\and
\IEEEauthorblockN{Homer Simpson}
\IEEEauthorblockA{Twentieth Century Fox\\
Springfield, USA\\
Email: homer@thesimpsons.com}
\and
\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
\IEEEauthorblockA{Starfleet Academy\\
San Francisco, California 96678--2391\\
Telephone: (800) 555--1212\\
Fax: (888) 555--1212}}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

% for over three affiliations, or if they all won't fit within the width
% of the page, use this alternative format:
%
%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
%Homer Simpson\IEEEauthorrefmark{2},
%James Kirk\IEEEauthorrefmark{3},
%Montgomery Scott\IEEEauthorrefmark{3} and
%Eldon Tyrell\IEEEauthorrefmark{4}}
%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
%Georgia Institute of Technology,
%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
%Email: homer@thesimpsons.com}
%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
%Telephone: (800) 555--1212, Fax: (888) 555--1212}
%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area
\maketitle

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
\end{abstract}

% no keywords

% For peer review papers, you can put extra information on the cover
% page as needed:
% \ifCLASSOPTIONpeerreview
% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
% \fi
%
% For peerreview papers, this IEEEtran command inserts a page break and
% creates the second title. It will be ignored for other modes.
\IEEEpeerreviewmaketitle

\section{Introduction}
\section{Methodolody}
\subsection{Research Goals} % (fold)
\subsection{Requirements} % (fold)
\subsection{Study Object} % (fold)
\subsection{Study Design} % (fold)
\subsubsection{Tasks} % (fold)
  For each task, define the task. Information that provided to the participants and what is required to perform it successfully.

\subsubsection{Subject Selection} % (fold)
\subsubsection{Process} % (fold)
\subsubsection{Data Collection} % (fold)
\subsubsection{Data Analysis: Success, Failure, Partial} % (fold)

\section{Results}

  Observation and Implications
    For each task, describe the results. Provide examples of failed attempts with counts, and relate to the documentation sources to share an observation. Imply what needs to be done.

\section{Threats to Validity}
\section{Related Work}

\section{Conclusion}
\section*{Acknowledgment}

% that's all folks
\end{document}

### 2017-01-29 bare_conf.tex
author names

@@
 % affiliations
-\author{\IEEEauthorblockN{Michael Shell}
-\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
-Georgia Institute of Technology\\
-Atlanta, Georgia 30332--0250\\
-Email: http://www.michaelshell.org/contact.html}
+\author{\IEEEauthorblockN{S M Sohan, Frank Maurer\IEEEauthorrefmark{1}}
+\IEEEauthorblockA{\IEEEauthorrefmark{1}
+Dept. of Computer Science\\
+University of Calgary\\
+Canada\\
+\{smsohan, frank.maurer\}@ucalgary.ca}
 \and
-\IEEEauthorblockN{Homer Simpson}
-\IEEEauthorblockA{Twentieth Century Fox\\
-Springfield, USA\\
-Email: homer@thesimpsons.com}
+\IEEEauthorblockN{Craig Anslow}
+\IEEEauthorblockA{School of Eng. and Computer Science\\
+Victoria University of Wellington\\
+New Zealand\\
+craig@ecs.vuw.ac.nz}
 \and
-\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
-\IEEEauthorblockA{Starfleet Academy\\
-San Francisco, California 96678--2391\\
-Telephone: (800) 555--1212\\
-Fax: (888) 555--1212}}
+\IEEEauthorblockN{Martin Robillard}
+\IEEEauthorblockA{
+School of Computer Science\\
+McGill University\\
+Canada\\
+martin@cs.mcgill.ca}
+}
 
@@
 \section{Methodolody}
-\subsection{Research Goals} % (fold)
+\subsection{Research Goals}
+
+This research is aimed at answering the following research questions:
+
+\begin{itemize}
+  \item What impact does usage examples have on REST API client developer productivity?
+  \item What what problems API client developers face when using a REST API where documentation lacks usage examples?
+\end{itemize}
+
 \subsection{Requirements} % (fold)

### 2017-01-29 bare_conf.tex
requirements

@@
 \begin{itemize}
-  \item What impact does usage examples have on REST API client developer productivity?
-  \item What what problems API client developers face when using a REST API where documentation lacks usage examples?
+  \item \textbf{RQ1}. What impact does usage examples have on REST API client developer productivity?
+  \item \textbf{RQ2}. What problems API client developers face when using a REST API documentation that lacks usage examples?
+\end{itemize}
+.
+\subsection{Requirements} % (fold)
+
+To answer the aforementioned research questions, the study has the following requirements:
+
+\begin{itemize}
+  \item \textbf{R1.
+Realistic.} We had to choose an existing REST API that is currently used by API client developers.
+Selecting a mature REST API for this study reduces the possibility of an underdeveloped solution that may be found if a new or a seldom used API is selected.
+A familiar domain needed to be selected so that participants are able to relate to the API features without requiring upfront training.
+In addition to selecting the API, we had to select tasks that are related to the core features provided by the API to represent reality.
+  \item \textbf{R2.
+Time bound.} The total time spent by each participant is required to be time bound to measure productivity.
+As a result, the study needs to be setup such that participants are able to focus on performing the tasks minimizing any overhead.
+  \item \textbf{R3.
+Experienced developers.} Developers with prior experience on REST APIs need to be recruited as study participants to perform the study within a limited amount of time and in a realistic setup
 \end{itemize}
 
-\subsection{Requirements} % (fold)
 \subsection{Study Object} % (fold)

### 2017-01-29 bare_conf.tex
add doc process

@@
 \ifCLASSINFOpdf
-  % \usepackage[pdftex]{graphicx}
+  \usepackage{listings}
+  \usepackage[pdftex]{graphicx}
+  \usepackage{url}
   % declare the path(s) where your graphic files are
@@
   \item \textbf{R3.
-Experienced developers.} Developers with prior experience on REST APIs need to be recruited as study participants to perform the study within a limited amount of time and in a realistic setup
+Experienced developers.} Developers with prior experience on REST APIs need to be recruited as study participants to perform the study within a limited amount of time and in a realistic setup.
+Furthermore, to reduce a learning bias, only participants with no prior experince of using the WordPress REST API V2 are accepted for this study.
 \end{itemize}
@@
 \subsection{Study Object} % (fold)
+
+We selected the WordPress REST API V2 for this study.
+WordPress is a blog-like open-source framework that is used by over 409 million people to visit 23.6 billion pages each month.
+The API allows programmatic access to list, create, update, and delete WordPress data such as blog posts, comments, users, images, tags, etc.
+
+The WordPress REST API V2 has been published and maintained since May 2015.
+Before January 2017, the WordPress REST API V2 was distributed as a plug-in where WordPress users could optionally install the API component.
+The following statistics are for the plug-in install numbers between May 2015 and October 2016:
+
+\begin{itemize}
+  \item Total installs: aprox. 248K installs of the plug-in.
+  \item Average daily installs: approx. 500.
+\end{itemize}
+
+Starting January 2017, the WordPress REST API V2 is no longer required to be installed as a separate plug-in since it's pre-bundled with every WordPress install.
+As per the code repository on GitHub, there are a total of 99 and 46 contributors that had at least one commit to the code repository behind the API and and it's documentation, respectively.
+These properties satisfy R1, our requirements for being realistic.
+
+By selecting an open-source project we are able to access the source-code to inspect the implementation and documentation technique of the WordPress REST API.
+The API implements a self-documenting feature where API developers expose API endpoints over HTTP OPTIONS verb to explain the API elements.
+To implement this feature, the API developers describe the API elements in the code.
+For example:
+
+\lstset{basicstyle=\footnotesize}
+\begin{lstlisting}[language=php,breaklines=true,showspaces=false,showstringspaces=false,numbers=left,xleftmargin=2em,caption={Example of self-documenting API Code},label=wordpress_code]
+public function get_item_schema() {
+  $schema = array(
+    '$schema'    => 'http://json-schema.org/draft-04/schema#',
+    'title'      => $this->post_type,
+    'type'       => 'object',
+    /*
+     * Base properties for every Post.
+     */
+    'properties' => array(
+        'date'            => array(
+            'description' => __( "The date the object was published, in the site's timezone." ),
+            'type'        => 'string',
+            'format'      => 'date-time',
+            'context'     => array( 'view', 'edit', 'embed' ),
+        ),...
+\end{lstlisting}
+
+On Listing \ref{wordpress_code}, line 4 specifies that this is a schema definition for the API element $Post$.
+Then, on Line 10, it defines $date$, one of the properties of $Post$, followed by a human readable description and and type information.
+Then, on line 14, the context of this property is mentioned as $view, edit, embed$, meaning that this property will be returned when the $Post$ object is returned, embedded, or can be used as an input for editing.
+
+This self-documenting feature is leveraged to generate and publish the official API documentation as an HTML website.
+Figure \ref{fig:schema} shows a screenshot of the published documentation for the $date$ attribute of the $Post$ API element\footnote{\url{http://v2.wp-api.org/reference/posts/}}.
+
+\begin{figure}[h!]
+\begin{center}
+\includegraphics[width=\linewidth,keepaspectratio]{schema.png}
+\end{center}
+\caption{Screenshot showing auto-generated documentation for Post/date}
+\label{fig:schema}
+\end{figure}
+
+In addition to the auto-generated documentation of the the schema and API actions, custom content is added by the contributors to the API documentation project to provide prosaic overviews and usage examples.
+
 \subsection{Study Design} % (fold)

### 2017-01-29 bare_conf.tex
first pass

@@
 \end{itemize}
-.
+
 \subsection{Requirements} % (fold)
@@
 In addition to selecting the API, we had to select tasks that are related to the core features provided by the API to represent reality.
+  \item \textbf{Open source.} To be able to understand the impact of usage examples, we needed to select an open-source API where we can add new examples to the documentation for performing this study.
+For proprietary APIs, it won't be possible to retrofit usage examples on existing API documentation without separating the examples from the documentation.
   \item \textbf{R2.
@@
 We selected the WordPress REST API V2 for this study.
-WordPress is a blog-like open-source framework that is used by over 409 million people to visit 23.6 billion pages each month.
+WordPress is a blog-like open-source (R2) framework used by over 409 million people to visit 23.6 billion pages each month.
 The API allows programmatic access to list, create, update, and delete WordPress data such as blog posts, comments, users, images, tags, etc.

### 2017-01-30 bare_conf.tex
tasks

@@
 \subsubsection{Tasks} % (fold)
-  For each task, define the task. Information that provided to the participants and what is required to perform it successfully.
+
+The participants are requested to perform a total of six tasks using the API including one practice task.
+All of the tasks are related to a single API element, $Post$.
+The tasks get progressively difficult, and all but the last task can be performed independently of each other.
+Participants are requested to limit the total time on the study to a maximum of one hour.
+Participants are encouraged to proceed to the next task when they are either satisfied about or stuck on the right answer.
+
+To use a REST API, the API client developers need to work with the following four inputs over HTTP:
+
+\begin{itemize}
+  \item I1. Request method
+  \item I2. Request URL
+  \item I3. Request headers
+  \item I4. Request body
+\end{itemize}
+
+To verify the response of an API call, the API client developers can use HTTP response headers and/or HTTP response body.
+To perform the tasks, the participants are required to use one or more of the inputs I1-4.
+In the following paragraphs, we describe each task with it's description and the intended test scenario against the aforementioned API input and output information.
+For each of the tasks, the participants are required to use the same WordPress REST API and its documentation.
+
+\textbf{T1: ListAllPostsTask}.
+We ask the participants to use the WordPress REST API following its documentation to get a list of the blog posts from a live WordPress site.
+This is the practice task, and the inputs to answer this task are pre-filled for the participants.
+It allows the participants to understand the tools used for this study as well as get familiarity with the Post API.
+The answer for this task makes use of I1 and I2.
+
+\textbf{T2: FilterPostsByAuthorTask.} The participants are asked to use the API to filter the list of posts by an author given the author's user name.
+To answer this correctly, the participants are required to first make an API call to get the numeric ID of the author given the string based user name.
+Then, the ID needs to be used on the Post API to filter posts by the author.
+This task allows us to understand the impact of usage examples on API client developers when multiple API calls need to be made to perform a task using the API.
+Inputs I1-2 are required to complete this task successfully.
+
+\textbf{T3: ExcludePostsByIdsTask.} We ask the participants to use the API to get a list of all posts excluding posts with IDs 1 and 4.
+Participants need to use the inputs I1-2, and use a desired format on I2 to pass an array of IDs as a parameter.
+This task allows us to understand how API client developers identify the format for using an array within the URL with respect to the usage examples in the API documentation.
+
+\textbf{T4: FindTotalPostsTasks.} This task requires the participants to use the API to find a total number of posts.
+Participants need to use the inputs I1-2 and inspect the HTTP response headers to successfully complete this task.
+This task allows us to understand how API client developer productivity is affected with respect to missing examples about HTTP response headers in the API documentation.
+
+\textbf{T5: PublishPostTask.} We ask the participants to use the API to publish a blog post with a specific title, content, and a published date.
+To successfully complete this task, the participants are required to use all four input types and inspect both the HTTP response header and the response body.
+Additionally, the participants are required to use a specific date format that the API accepts as a valid format for date specification.
+Answers to this task allows us to study API client developer productivity with respect to the usage examples lacking details about the inputs I3-4.
+
+\textbf{T6: UpdatePostTask.} We ask the participants to use the API to update a blog post that they published in T5 with a new excerpt.
+Similar to T5, this task requires the use of inputs I1-4, but with different values for the inputs.
+This task allows us to understand API client developer productivity on inter-dependent tasks with respect to usage examples.
+
+To summarize, the tasks allow us to understand how REST API client developers approach API tasks of different complexity levels involving various input types and available output information with respect to the usage examples in the API documentation.
 

### 2017-01-30 bare_conf.tex
pilot

@@
   \usepackage{url}
+
   % declare the path(s) where your graphic files are
@@
 
-\subsection{Study Object} % (fold)
+\subsection{Study API} % (fold)
 
@@
 
-\subsubsection{Subject Selection} % (fold)
-\subsubsection{Process} % (fold)
+\subsubsection{Participant Selection} % (fold)
+
+To satisfy the requirement of developers with REST API experience (R3), we have used the following criteria for recruiting the study participants:
+
+\begin{itemize}
+  \item Currently working as a software engineer.
+  \item At least 1 year of industry experience as a software engineer.
+  \item At least 1 year of industry experience with REST APIs.
+  \item No prior experience with WordPress REST API.
+\end{itemize}
+
+Participants were recruited through online announcements posted on Twitter, Facebook, and software developer focused mailing lists.
+
+\subsubsection{Process}
+
+Two pilot studies were performed to evaluate and understand an effective process for performing this study.
+The first pilot study involved four participants that were invited to join the first author on this paper in-person or using a video conferencing software.
+The study involved tasks using two APIs, the WordPress REST API V2 and the GMail REST API.
+Each participant was given one of the two APIs and a set of tasks to complete using the API within an hour.
+Participants were given an online answer form to record the answers to the tasks.
+The primary findings from this pilot are as follows: 1) asking participants to use an API to perform the tasks required significant overhead time for them to setup a development environment with the proper API credentials, 2) the intermediate trial attempts of using the API are potentially more valuable than the final answer as it allows us to understand API client developer information needs that may are not answered by the documentation, 3) the number of tasks for the study had to be reduced so the participants could complete the tasks within the one hour limit, and 4) for GMail API, participants used up a large portion of their time on setting up their API credentials that requires understanding of OAuth.
+
+\begin{figure*}[h!]
+\begin{center}
+\includegraphics[width=\linewidth,keepaspectratio]{api_explorer.png}
+\end{center}
+\caption{Screenshot of the web-based REST API explorer}
+\label{fig:api_explorer}
+\end{figure*}
+
+To overcome the shortcomings found from the first pilot study, we decided to develop a web-based REST API explorer as shown on Figure \ref{fig:api_explorer} that allows participants to use their browser to make the API calls without setting up any development environment.
+The web-based API explorer only requires the inputs I1-4, and displays the HTTP response headers and body on the click of a button.
+Thus the participants could focus on using the right input and verifying the output without having to write any code.
+The web-base REST API explorer also allowed us to automatically collect all the trial API calls that the participants make for each API task.
+A second pilot study involving seven new participants was performed to understand the features of the web-based REST API explorer in practice and to improve the user interface based on feedback from the participants.
+Participants completed the study on their own without having to meet in-person or over video conferencing.
+Only the WordPress REST API was used to focus on REST APIs without the required learning curve associated with OAuth.
+We found encouraging results from this pilot study as the collected data showed patterns of mistakes that API client developers make that can be reduced by adding usage examples in the API documentation.
+
 \subsubsection{Data Collection} % (fold)

### 2017-01-31 bare_conf.tex
data collection

@@
 In addition to selecting the API, we had to select tasks that are related to the core features provided by the API to represent reality.
-  \item \textbf{Open source.} To be able to understand the impact of usage examples, we needed to select an open-source API where we can add new examples to the documentation for performing this study.
+  \item \textbf{R2.
+Open source.} To be able to understand the impact of usage examples, we needed to select an open-source API where we can add new examples to the documentation for performing this study.
 For proprietary APIs, it won't be possible to retrofit usage examples on existing API documentation without separating the examples from the documentation.
-  \item \textbf{R2.
+  \item \textbf{R3.
 Time bound.} The total time spent by each participant is required to be time bound to measure productivity.
 As a result, the study needs to be setup such that participants are able to focus on performing the tasks minimizing any overhead.
-  \item \textbf{R3.
+  \item \textbf{R4.
 Experienced developers.} Developers with prior experience on REST APIs need to be recruited as study participants to perform the study within a limited amount of time and in a realistic setup.
@@
 
-\begin{figure*}[h!]
+\begin{figure*}[t]
 \begin{center}
@@
 
+\begin{figure*}[t]
+\begin{center}
+\includegraphics[width=\linewidth,keepaspectratio]{original.png}
+\end{center}
+\caption{Screenshot of the original WordPress REST API documentation}
+\label{fig:original}
+\end{figure*}
+
+\begin{figure*}[tb]
+\begin{center}
+\includegraphics[width=\linewidth,keepaspectratio]{forked.png}
+\end{center}
+\caption{Screenshot of the forked WordPress REST API documentation with a usage example}
+\label{fig:forked}
+\end{figure*}
+
+Based on the lessons learned from the pilot studies, we designed the actual study process as follows: the original WordPress REST API documentation was forked and a total of 3 API usage examples were added to show listing of blog posts with query parameters for filtering, a request to create a blog post and a request to update a blog post.
+Figure \ref{fig:original} shows a screenshot of the original API documentation related to T4 where the API client developers are provided with a reference table describing the different properties that can be used to create a $Post$ object\footnote{\url{http://v2.wp-api.org/reference/posts/}}.
+Figure \ref{fig:forked} shows a screenshot of the forked API documentation with a cURL based usage example.
+cURL is used because it's used by elsewhere in the original API documentation.
+In the forked API documentation, the example shows one possible API call with realistic values for the data that is described in the reference table and associated API response headers and body.
+
+Participants were divided into two groups, G1 and G2.
+G1 participants were provided with a link to the official API documentation on the web-based API explorer, and G2 participants were provided with a link to the forked API documentation with usage examples.
+The web-based API explorer allocated more participants to G1 compared to G2 because we wanted to better understand the impact of the lack of usage examples on API client developer productivity.
+However, each individual participant was randomly assigned to a group by the web-based API explorer.
+All participants were given the same set of API tasks and were requested to limit their participation time to a maximum of one hour.
+No task specific time limit was imposed because we wanted participants to spend sufficient time on each task without forcing them to move the next one.
+Participants were allowed to access the internet and external resources alongside the provided API documentation to perform the tasks as they'd normally use on a typical work day.
+
 \subsubsection{Data Collection} % (fold)
+
+The collected data for this study is both quantitative and qualitative.
+The quantitative part comprises of demographic information, statistics about the API tasks, and a numeric rating of the given API documentation to answer R1.
+We collected the demographic information to record the business and industry experience information.
+For each participant, we recorded all trial API calls with timestamps, and compared against the required API call to annotate the tasks answers as successful, partially successful, or unsuccessful.
+The quantitative data is used to measure REST API client developer productivity as a tuple of success rate, average number of trial API calls for each successful API call, and the average time taken for each successful API task completion by participants in G1 and G2.
+
+The data from qualitative analysis comprises of a text document that is exported from the database behind the web-based REST API explorer.
+For each API call made by the participants, inputs I1-4 along with the HTTP response headers and body are automatically saved on a database by the web-based API explorer.
+Additionally, the participants provide a free-form feedback about the experience of using the given REST API documentation to perform the tasks.
+The qualitative data was analyzed to answer R2.
+
 \subsubsection{Data Analysis: Success, Failure, Partial} % (fold)

### 2017-02-05 bare_conf.tex
minor edits
protocol complete

@@
 \begin{itemize}
-  \item \textbf{RQ1}. What impact does usage examples have on REST API client developer productivity?
+  \item \textbf{RQ1}.
+What impact does usage examples have on REST API client developer productivity?
+In this study, we define API client developer productivity by the following metrics: a) number of API tasks performed successfully within a limited amount of time, b) number of trial API calls before a successful API call is found, and c) time spent to successfully complete an API task.
   \item \textbf{RQ2}. What problems API client developers face when using a REST API documentation that lacks usage examples?
@@
   \item \textbf{R1.
-Realistic.} We had to choose an existing REST API that is currently used by API client developers.
+Representative API.} We had to choose an existing REST API that is currently used by API client developers.
 Selecting a mature REST API for this study reduces the possibility of an underdeveloped solution that may be found if a new or a seldom used API is selected.
@@
 As per the code repository on GitHub, there are a total of 99 and 46 contributors that had at least one commit to the code repository behind the API and and it's documentation, respectively.
-These properties satisfy R1, our requirements for being realistic.
+These properties satisfy R1, our requirements for using a representative API.
 
@@
 \subsubsection{Data Collection} % (fold)
+The data collected by the web-based REST API explorer for each participant is exported into a text file as the raw data artifact.
+For each API task and each participant that attempted the task, the exported text file contains the request inputs I1-4 and associated API response headers and body for each API call made by the participant.
+For each participant, we recorded all trial API calls with timestamps.
+Additionally, the demographic information, experience rating and the free-form feedback for each participant is also included in the artifact.
 
-The collected data for this study is both quantitative and qualitative.
-The quantitative part comprises of demographic information, statistics about the API tasks, and a numeric rating of the given API documentation to answer R1.
-We collected the demographic information to record the business and industry experience information.
-For each participant, we recorded all trial API calls with timestamps, and compared against the required API call to annotate the tasks answers as successful, partially successful, or unsuccessful.
-The quantitative data is used to measure REST API client developer productivity as a tuple of success rate, average number of trial API calls for each successful API call, and the average time taken for each successful API task completion by participants in G1 and G2.
+\subsubsection{Data Analysis}
+The same data artifact is used to analyze both quantitatively and qualitatively to answer the aforementioned research questions.
+To answer RQ1, for each API task attempted by each participant, we annotated the artifact with one of the following labels: successful, partially successful, and unsuccessful.
+Task participations are marked successful when I1-4 matches the required values for performing the given task.
+If a participant is able to use the correct I1-4 for one of the two tasks required to complete a single task (T2), we marked it as a partially successful.
+Otherwise, it's marked as unsuccessful.
+Based on these annotations, the answers for RQ1 was computed using the following:
 
-The data from qualitative analysis comprises of a text document that is exported from the database behind the web-based REST API explorer.
-For each API call made by the participants, inputs I1-4 along with the HTTP response headers and body are automatically saved on a database by the web-based API explorer.
-Additionally, the participants provide a free-form feedback about the experience of using the given REST API documentation to perform the tasks.
-The qualitative data was analyzed to answer R2.
+success rate of group = (total number of successful tasks by group) / (number of participants in group * number of API tasks)
 
-\subsubsection{Data Analysis: Success, Failure, Partial} % (fold)
+API calls per success of group = (number of API calls annotated as successful) / (total number of trial API calls made by group)
+
+average time taken for success = (number of successful API tasks by group) / (total time spent by all participants in the group)
+
+To answer RQ2, we analyzed the data artifact qualitatively.
+The data artifact exported from the web-based REST API explorer included a table with raw data.
+For each row, the table contains the following columns: API task, participant identifier, timestamp, API trial number, I1-4, response headers and body.
+It allowed us to observe and categorize the different input values that the participants used to perform the given API tasks.
+The values for I1-4 used by the participants were manually coded to group the participant responses into categories.
+The categories help us determine the types of information that API client developers need in the API documentation to perform API tasks successfully.
+The analysis started with an empty set of codes and new codes were introduced to describe scenarios that didn't fall under  codes that were already applied.
+The first author on this paper applied the codes on the raw data artifact and the co-authors verified the codes randomly.
+To resolve disagreements, the authors discussed and updated the codes accordingly.
+The codes were then further categorized to higher level grouping to represent common problems faced by the study participants for each group.
 

### 2017-02-05 bare_conf.tex
abstract raw

@@
 \begin{abstract}
+
+[Brain-dump mode]
+
+What is the the problem? Effort and cost associated to REST API documentation with usage examples.
+Why is this a problem? Cost vs. benefit of usage examples is not known. What to include in the examples?
+What have we done?
+Performed an empirical study with 20 professional software engineers with REST API experience to understand how their productivity and types of problems faced changes with respect to usage examples.
+What have we learned? (Hoping) examples help developer perform task with fewer problems with greater success, with fewer errors, and faster in terms of time.
+What does it mean? (expected) API developers should include usage examples in REST API documentation that include specific examples based on the problems faced.
+
 \end{abstract}
@@
 
+\subsection{RQ1: REST API Client Developer Productivity} % (fold)
+
+A graph showing the difference in productivity metrics between G1 and G2.
+Implications
+
+\subsection{RQ2: Common Problems Faced by REST API Client Developers} % (fold)
+
+% subsection rq1 (end)
+
   Observation and Implications

### 2017-02-19 bare_conf.tex
RQ 1&2 merged into RQ

@@
 \begin{itemize}
-  \item \textbf{RQ1}.
-What impact does usage examples have on REST API client developer productivity?
-In this study, we define API client developer productivity by the following metrics: a) number of API tasks performed successfully within a limited amount of time, b) number of trial API calls before a successful API call is found, and c) time spent to successfully complete an API task.
-  \item \textbf{RQ2}. What problems API client developers face when using a REST API documentation that lacks usage examples?
+  \item \textbf{RQ}. What obstacles API client developers face when using a REST API documentation that lacks usage examples?
 \end{itemize}
@@
 \subsubsection{Data Analysis}
-The same data artifact is used to analyze both quantitatively and qualitatively to answer the aforementioned research questions.
-To answer RQ1, for each API task attempted by each participant, we annotated the artifact with one of the following labels: successful, partially successful, and unsuccessful.
+The same data artifact is used to analyze both quantitatively and qualitatively to answer the aforementioned research question.We analyzed the data artifact qualitatively to understand the obstacles that REST API client developers face when using an API documentation that lacks usage examples.
+The data artifact exported from the web-based REST API explorer included a table with raw data.
+For each row, the table contains the following columns: API task, participant identifier, timestamp, API trial number, I1-4, response headers and body.
+It allowed us to observe and categorize the different input values that the participants used to perform the given API tasks.
+The values for I1-4 used by the participants were manually coded to group the participant responses into categories.
+The categories help us determine the types of information that API client developers need in the API documentation to perform API tasks successfully.
+The analysis started with an empty set of codes and new codes were introduced to describe scenarios that didn't fall under  codes that were already applied.
+The first author on this paper applied the codes on the raw data artifact and the co-authors verified the codes randomly.
+To resolve disagreements, the authors discussed and updated the codes accordingly.
+The codes were then further categorized to higher level grouping to represent common problems faced by the study participants for each group.
+
+For each API task attempted by each participant, we annotated the artifact with one of the following labels: successful, partially successful, and unsuccessful.
 Task participations are marked successful when I1-4 matches the required values for performing the given task.
@@
 
-To answer RQ2, we analyzed the data artifact qualitatively.
-The data artifact exported from the web-based REST API explorer included a table with raw data.
-For each row, the table contains the following columns: API task, participant identifier, timestamp, API trial number, I1-4, response headers and body.
-It allowed us to observe and categorize the different input values that the participants used to perform the given API tasks.
-The values for I1-4 used by the participants were manually coded to group the participant responses into categories.
-The categories help us determine the types of information that API client developers need in the API documentation to perform API tasks successfully.
-The analysis started with an empty set of codes and new codes were introduced to describe scenarios that didn't fall under  codes that were already applied.
-The first author on this paper applied the codes on the raw data artifact and the co-authors verified the codes randomly.
-To resolve disagreements, the authors discussed and updated the codes accordingly.
-The codes were then further categorized to higher level grouping to represent common problems faced by the study participants for each group.
-
 \section{Results}
 
-\subsection{RQ1: REST API Client Developer Productivity} % (fold)
+\subsection{Qualitative Analysis Results} % (fold)
+
+  Observation and Implications
+    For each task, describe the results. Provide examples of failed attempts with counts, and relate to the documentation sources to share an observation. Imply what needs to be done.
+
+\subsection{Quantitative Analysis Results} % (fold)
 
@@
 
-\subsection{RQ2: Common Problems Faced by REST API Client Developers} % (fold)
-
 % subsection rq1 (end)
-
-  Observation and Implications
-    For each task, describe the results. Provide examples of failed attempts with counts, and relate to the documentation sources to share an observation. Imply what needs to be done.
 

### 2017-02-26 controlled_study.tex
refs + related work pass 1

--- 2017-02-19 bare_conf.tex
@@
-%% bare_conf.tex
-%% V1.4b
-%% 2015/08/26
-%% by Michael Shell
-%% See:
-%% http://www.michaelshell.org/
-%% for current contact information.
-%%
-%% This is a skeleton file demonstrating the use of IEEEtran.cls
-%% (requires IEEEtran.cls version 1.8b or later) with an IEEE
-%% conference paper.
-%%
-%% Support sites:
-%% http://www.michaelshell.org/tex/ieeetran/
-%% http://www.ctan.org/pkg/ieeetran
-%% and
-%% http://www.ieee.org/
+\documentclass[conference]{IEEEtran}
 
-%%*************************************************************************
-%% Legal Notice:
-%% This code is offered as-is without any warranty either expressed or
-%% implied; without even the implied warranty of MERCHANTABILITY or
-%% FITNESS FOR A PARTICULAR PURPOSE!
-%% User assumes all risk.
-%% In no event shall the IEEE or any contributor to this code be liable for
-%% any damages or losses, including, but not limited to, incidental,
-%% consequential, or any other damages, resulting from the use or misuse
-%% of any information contained here.
-%%
-%% All comments are the opinions of their respective authors and are not
-%% necessarily endorsed by the IEEE.
-%%
-%% This work is distributed under the LaTeX Project Public License (LPPL)
-%% ( http://www.latex-project.org/ ) version 1.3, and may be freely used,
-%% distributed and modified. A copy of the LPPL, version 1.3, is included
-%% in the base LaTeX documentation of all distributions of LaTeX released
-%% 2003/12/01 or later.
-%% Retain all contribution notices and credits.
-%% ** Modified files should be clearly indicated as such, including  **
-%% ** renaming them and changing author support contact information. **
-%%*************************************************************************
-
-% *** Authors should verify (and, if needed, correct) their LaTeX system  ***
-% *** with the testflow diagnostic prior to trusting their LaTeX platform ***
-% *** with production work. The IEEE's font choices and paper sizes can   ***
-% *** trigger bugs that do not appear when using other class files.       ***                          ***
-% The testflow support page is at:
-% http://www.michaelshell.org/tex/testflow/
-
-\documentclass[conference]{IEEEtran}
-% Some Computer Society conferences also require the compsoc mode option,
-% but others use the standard conference format.
-%
-% If IEEEtran.cls has not been installed into the LaTeX system files,
-% manually specify the path to it like:
-% \documentclass[conference]{../sty/IEEEtran}
-
-% Some very useful LaTeX packages include:
-% (uncomment the ones you want to load)
-
-% *** MISC UTILITY PACKAGES ***
-%
-%\usepackage{ifpdf}
-% Heiko Oberdiek's ifpdf.sty is very useful if you need conditional
-% compilation based on whether the output is pdf or dvi.
-% usage:
-% \ifpdf
-%   % pdf code
-% \else
-%   % dvi code
-% \fi
-% The latest version of ifpdf.sty can be obtained from:
-% http://www.ctan.org/pkg/ifpdf
-% Also, note that IEEEtran.cls V1.7 and later provides a builtin
-% \ifCLASSINFOpdf conditional that works the same way.
-% When switching from latex to pdflatex and vice-versa, the compiler may
-% have to be run twice to clear warning/error messages.
-
-% *** CITATION PACKAGES ***
-%
-%\usepackage{cite}
-% cite.sty was written by Donald Arseneau
-% V1.6 and later of IEEEtran pre-defines the format of the cite.sty package
-% \cite{} output to follow that of the IEEE. Loading the cite package will
-% result in citation numbers being automatically sorted and properly
-% "compressed/ranged". e.g., [1], [9], [2], [7], [5], [6] without using
-% cite.sty will become [1], [2], [5]--[7], [9] using cite.sty. cite.sty's
-% \cite will automatically add leading space, if needed. Use cite.sty's
-% noadjust option (cite.sty V3.8 and later) if you want to turn this off
-% such as if a citation ever needs to be enclosed in parenthesis.
-% cite.sty is already installed on most LaTeX systems. Be sure and use
-% version 5.0 (2009-03-20) and later if using hyperref.sty.
-% The latest version can be obtained at:
-% http://www.ctan.org/pkg/cite
-% The documentation is contained in the cite.sty file itself.
-
-% *** GRAPHICS RELATED PACKAGES ***
-%
 \ifCLASSINFOpdf
@@
 
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../pdf/}{../jpeg/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.pdf,.jpeg,.png}
 \else
-  % or other class option (dvipsone, dvipdf, if not using dvips). graphicx
-  % will default to the driver specified in the system graphics.cfg if no
-  % driver is specified.
-  % \usepackage[dvips]{graphicx}
-  % declare the path(s) where your graphic files are
-  % \graphicspath{{../eps/}}
-  % and their extensions so you won't have to specify these with
-  % every instance of \includegraphics
-  % \DeclareGraphicsExtensions{.eps}
 \fi
-% graphicx was written by David Carlisle and Sebastian Rahtz. It is
-% required if you want graphics, photos, etc. graphicx.sty is already
-% installed on most LaTeX systems. The latest version and documentation
-% can be obtained at:
-% http://www.ctan.org/pkg/graphicx
-% Another good source of documentation is "Using Imported Graphics in
-% LaTeX2e" by Keith Reckdahl which can be found at:
-% http://www.ctan.org/pkg/epslatex
-%
-% latex, and pdflatex in dvi mode, support graphics in encapsulated
-% postscript (.eps) format. pdflatex in pdf mode supports graphics
-% in .pdf, .jpeg, .png and .mps (metapost) formats. Users should ensure
-% that all non-photo figures use a vector format (.eps, .pdf, .mps) and
-% not a bitmapped formats (.jpeg, .png). The IEEE frowns on bitmapped formats
-% which can result in "jaggedy"/blurry rendering of lines and letters as
-% well as large increases in file sizes.
-%
-% You can find documentation about the pdfTeX application at:
-% http://www.tug.org/applications/pdftex
-
-% *** MATH PACKAGES ***
-%
-%\usepackage{amsmath}
-% A popular package from the American Mathematical Society that provides
-% many useful and powerful commands for dealing with mathematics.
-%
-% Note that the amsmath package sets \interdisplaylinepenalty to 10000
-% thus preventing page breaks from occurring within multiline equations. Use:
-%\interdisplaylinepenalty=2500
-% after loading amsmath to restore such page breaks as IEEEtran.cls normally
-% does. amsmath.sty is already installed on most LaTeX systems. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/amsmath
-
-% *** SPECIALIZED LIST PACKAGES ***
-%
-%\usepackage{algorithmic}
-% algorithmic.sty was written by Peter Williams and Rogerio Brito.
-% This package provides an algorithmic environment fo describing algorithms.
-% You can use the algorithmic environment in-text or within a figure
-% environment to provide for a floating algorithm. Do NOT use the algorithm
-% floating environment provided by algorithm.sty (by the same authors) or
-% algorithm2e.sty (by Christophe Fiorio) as the IEEE does not use dedicated
-% algorithm float types and packages that provide these will not provide
-% correct IEEE style captions. The latest version and documentation of
-% algorithmic.sty can be obtained at:
-% http://www.ctan.org/pkg/algorithms
-% Also of interest may be the (relatively newer and more customizable)
-% algorithmicx.sty package by Szasz Janos:
-% http://www.ctan.org/pkg/algorithmicx
-
-% *** ALIGNMENT PACKAGES ***
-%
-%\usepackage{array}
-% Frank Mittelbach's and David Carlisle's array.sty patches and improves
-% the standard LaTeX2e array and tabular environments to provide better
-% appearance and additional user controls. As the default LaTeX2e table
-% generation code is lacking to the point of almost being broken with
-% respect to the quality of the end results, all users are strongly
-% advised to use an enhanced (at the very least that provided by array.sty)
-% set of table tools. array.sty is already installed on most systems. The
-% latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/array
-
-% IEEEtran contains the IEEEeqnarray family of commands that can be used to
-% generate multiline equations as well as matrices, tables, etc., of high
-% quality.
-
-% *** SUBFIGURE PACKAGES ***
-%\ifCLASSOPTIONcompsoc
-%  \usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
-%\else
-%  \usepackage[caption=false,font=footnotesize]{subfig}
-%\fi
-% subfig.sty, written by Steven Douglas Cochran, is the modern replacement
-% for subfigure.sty, the latter of which is no longer maintained and is
-% incompatible with some LaTeX packages including fixltx2e. However,
-% subfig.sty requires and automatically loads Axel Sommerfeldt's caption.sty
-% which will override IEEEtran.cls' handling of captions and this will result
-% in non-IEEE style figure/table captions. To prevent this problem, be sure
-% and invoke subfig.sty's "caption=false" package option (available since
-% subfig.sty version 1.3, 2005/06/28) as this is will preserve IEEEtran.cls
-% handling of captions.
-% Note that the Computer Society format requires a larger sans serif font
-% than the serif footnote size font used in traditional IEEE formatting
-% and thus the need to invoke different subfig.sty package options depending
-% on whether compsoc mode has been enabled.
-%
-% The latest version and documentation of subfig.sty can be obtained at:
-% http://www.ctan.org/pkg/subfig
-
-% *** FLOAT PACKAGES ***
-%
-%\usepackage{fixltx2e}
-% fixltx2e, the successor to the earlier fix2col.sty, was written by
-% Frank Mittelbach and David Carlisle. This package corrects a few problems
-% in the LaTeX2e kernel, the most notable of which is that in current
-% LaTeX2e releases, the ordering of single and double column floats is not
-% guaranteed to be preserved. Thus, an unpatched LaTeX2e can allow a
-% single column figure to be placed prior to an earlier double column
-% figure.
-% Be aware that LaTeX2e kernels dated 2015 and later have fixltx2e.sty's
-% corrections already built into the system in which case a warning will
-% be issued if an attempt is made to load fixltx2e.sty as it is no longer
-% needed.
-% The latest version and documentation can be found at:
-% http://www.ctan.org/pkg/fixltx2e
-
-%\usepackage{stfloats}
-% stfloats.sty was written by Sigitas Tolusis. This package gives LaTeX2e
-% the ability to do double column floats at the bottom of the page as well
-% as the top. (e.g., "\begin{figure*}[!b]" is not normally possible in
-% LaTeX2e). It also provides a command:
-%\fnbelowfloat
-% to enable the placement of footnotes below bottom floats (the standard
-% LaTeX2e kernel puts them above bottom floats). This is an invasive package
-% which rewrites many portions of the LaTeX2e float routines. It may not work
-% with other packages that modify the LaTeX2e float routines. The latest
-% version and documentation can be obtained at:
-% http://www.ctan.org/pkg/stfloats
-% Do not use the stfloats baselinefloat ability as the IEEE does not allow
-% \baselineskip to stretch. Authors submitting work to the IEEE should note
-% that the IEEE rarely uses double column equations and that authors should try
-% to avoid such use. Do not be tempted to use the cuted.sty or midfloat.sty
-% packages (also by Sigitas Tolusis) as the IEEE does not format its papers in
-% such ways.
-% Do not attempt to use stfloats with fixltx2e as they are incompatible.
-% Instead, use Morten Hogholm'a dblfloatfix which combines the features
-% of both fixltx2e and stfloats:
-%
-% \usepackage{dblfloatfix}
-% The latest version can be found at:
-% http://www.ctan.org/pkg/dblfloatfix
-
-% *** PDF, URL AND HYPERLINK PACKAGES ***
-%
-%\usepackage{url}
-% url.sty was written by Donald Arseneau. It provides better support for
-% handling and breaking URLs. url.sty is already installed on most LaTeX
-% systems. The latest version and documentation can be obtained at:
-% http://www.ctan.org/pkg/url
-% Basically, \url{my_url_here}.
-
-% *** Do not adjust lengths that control margins, column widths, etc. ***
-% *** Do not use packages that alter fonts (such as pslatex).         ***
-% There should be no need to do such things with IEEEtran.cls V1.6 and later.
-% (Unless specifically asked to do so by the journal or conference you plan
-% to submit to, of course. )
-
-% correct bad hyphenation here
 \hyphenation{op-tical net-works semi-conduc-tor}
@@
 \begin{document}
-%
-% paper title
-% Titles are generally capitalized except for words such as a, an, and, as,
-% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
-% not capitalized unless they are the first or last word of the title.
-% Linebreaks \\ can be used within to get better formatting as desired.
-% Do not put math or special symbols in the title.
 \title{How Important are Usage Examples in REST API Documentation?}
 
-% author names and affiliations
-% use a multiple column layout for up to three different
-% affiliations
 \author{\IEEEauthorblockN{S M Sohan, Frank Maurer\IEEEauthorrefmark{1}}
@@
 
-% conference papers do not typically use \thanks and this command
-% is locked out in conference mode. If really needed, such as for
-% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
-% after \documentclass
-
-% for over three affiliations, or if they all won't fit within the width
-% of the page, use this alternative format:
-%
-%\author{\IEEEauthorblockN{Michael Shell\IEEEauthorrefmark{1},
-%Homer Simpson\IEEEauthorrefmark{2},
-%James Kirk\IEEEauthorrefmark{3},
-%Montgomery Scott\IEEEauthorrefmark{3} and
-%Eldon Tyrell\IEEEauthorrefmark{4}}
-%\IEEEauthorblockA{\IEEEauthorrefmark{1}School of Electrical and Computer Engineering\\
-%Georgia Institute of Technology,
-%Atlanta, Georgia 30332--0250\\ Email: see http://www.michaelshell.org/contact.html}
-%\IEEEauthorblockA{\IEEEauthorrefmark{2}Twentieth Century Fox, Springfield, USA\\
-%Email: homer@thesimpsons.com}
-%\IEEEauthorblockA{\IEEEauthorrefmark{3}Starfleet Academy, San Francisco, California 96678-2391\\
-%Telephone: (800) 555--1212, Fax: (888) 555--1212}
-%\IEEEauthorblockA{\IEEEauthorrefmark{4}Tyrell Inc., 123 Replicant Street, Los Angeles, California 90210--4321}}
-
-% use for special paper notices
-%\IEEEspecialpapernotice{(Invited Paper)}
-
-% make the title area
 \maketitle
 
-% As a general rule, do not put math, special symbols or citations
-% in the abstract
 \begin{abstract}
@@
 
+\cite{Robillard_a_field_study}
+
 \end{abstract}
 
-% no keywords
-
-% For peer review papers, you can put extra information on the cover
-% page as needed:
-% \ifCLASSOPTIONpeerreview
-% \begin{center} \bfseries EDICS Category: 3-BBND \end{center}
-% \fi
-%
-% For peerreview papers, this IEEEtran command inserts a page break and
-% creates the second title. It will be ignored for other modes.
 \IEEEpeerreviewmaketitle
@@
 \section{Threats to Validity}
+  API Selection
+  Task Selection
+  Participant Selection
+  Experiment Design
+
 \section{Related Work}
+
+\subsubsection{API Learning Obstacles}
+  Researchers have published several papers on the topic of API learning obstacles.
+Robillard et al.
+used a mixed approach of surveys and in-person interviews with 440 professional software engineers to understand the API learning obstacles \cite{Robillard_a_field_study}.
+They found that API documentation was related to one of the most severe obstacles for API learnabilty and recommended considering five factors for API documentation as follows: documentation of intent, code examples, matching APIs with scenarios, penetrability of the API, and format and presentation.
+Gias et al.
+surveyed 323 professional software engineers to understand the factors that fail a API documentation \cite{g_uddin}.
+They identified ambiguity, incompleteness, and incorrectness as the three severest problems that lead API documentations to fail to answer the information needs of developers.
+Duala-Ekoko et al.
+performed a study with twenty programmers to understand the types of questions that developers have while using an unfamiliar API \cite{Duala-Ekoko:2012:AAQ:2337223.2337255}.
+They categorized the API related information needs into twenty generic questions that can be used to analyze the quality of API documentation.
+
+  In this paper, we focused on the usage examples part of API documentation to further understand the obstacles that API client developers face when the documentation contains relevant information about the API, it's intent, description of the API actions and elements, but lacks code examples.
+
+\subsubsection{Crowd-Sourced API Usage Examples}
+Several existing research has focused on sourcing API usage examples from the Internet.
+Wang et al.
+performed an exploratory study to understand the current state of API related knowledge available on the Internet that can be leveraged by API client developers \cite{6462686}.
+They searched the web for usage examples related to the API of five popular Java libraries and found that on an average API examples could be found for 77\% of the 4,637 APIs included in the selected libraries.
+Moreover, they found that the crowd-sourced API documented accounted for 93.7\% of the usage examples compared to 6.13\% that were published on the official documentation.
+Incorporating a corpus of crowd-sourced API examples from the web and using placeholders for commonly related API elements, an evaluation of Jadite found that developers were three times faster to complete API tasks \cite{5295283}.
+Nasehi et al.
+analyzed StackOverflow threads to understand the characteristics of good and bad code examples based on user provided votes \cite{Nasehi_what_makes}.
+They found that commonly down-voted API related answers lacked code examples, explanations, and shortcoming of solutions.
+They recommended API developers to evolve the documentation with usage examples to answer API client developer questions that were not anticipated in the existing documentation.
+Treude et al.
+proposed a machine learning based approach to find relevant API documentation from StackOverflow using both the text and metadata found from the questions and answers \cite{Treude:2016:AAD:2884781.2884800}.
+Kim et al.
+presented a technique to automatically augment code examples from the web using code search tools to Java APIs \cite{Kim:2009:AEJ:1747491.1747552}.
+Jiau et al.
+observed a severe inequality within the context of crowd-sourced API usage examples where most of the content were related to the popular APIs \cite{jiau}.
+
+Existing research on crowd-sourced API documentation has mostly focused finding crowd-sourced usage examples of local APIs.
+In this paper, we focused on the impact of usage examples for REST APIs, where the API client development programming language is agnostic to the language used to implement the API.
+This important distinction with local APIs makes it hard to leverage the techniques proposed in the aforementioned papers to find relevant REST API usage examples.
+
+\subsubsection{Controlled Studies}
+Several authors published the results of controlled studies on the topic of API usability.
+Nasehi et al.
+performed a controlled study to understand if API unit tests can be used to provide as usage examples to facilitate API client developers \cite{5609553}.
+They grouped participants into two groups, and one of the groups was provided with API unit tests in addition to the documentation.
+The two groups performed the same set of API tasks.
+The researchers found that the examples from the unit tests helped understanding the API concepts better but it was challenging for the participants to locate relevant examples from the corpus of unit tests.
+They recommended automated extraction of relevant high-level API usage examples from API unit tests.
+Endrikat performed a controlled study with four groups of participants to understand the impact of API documentation on APIs that are implemented using programming languages with static and dynamic type systems \cite{Endrikat:2014:ADS:2568225.2568299}.
+Participants were given a set of failing unit tests and were asked to make them pass by writing code using the studied API.
+They defined the API client developer productivity in terms of the time to get the tests passing.
+They found that the participant group using a statically typed API with explicit documentation were more productive than the other groups.
+Ellis et al.
+performed a controlled study with two groups of Java developers to understand the impact of using a constructor vs.
+a factory method design pattern on API client development time \cite{Ellis:2007:FPA:1248820.1248863}.
+They found that API client developers needed more time to complete API tasks when the API requires the use a factory method to instantiate objects compared to using a constructor.
+
+Our work in this paper is based on a controlled study and shares part of the setup that were used by the aforementioned API related controlled studies with the following differences: our goal is to understand the obstacles faced by API client developers without usage examples;  we focus on REST APIs instead of a local API; and, the participants are professional software engineers.
 
@@
 
-% that's all folks
+\bibliographystyle{IEEEtran}
+\bibliography{IEEEabrv,refs}
+
 \end{document}

### 2017-02-27 controlled_study.tex
related work v1

@@
 What does it mean? (expected) API developers should include usage examples in REST API documentation that include specific examples based on the problems faced.
-
-\cite{Robillard_a_field_study}
 
@@
 Robillard et al.
-used a mixed approach of surveys and in-person interviews with 440 professional software engineers to understand the API learning obstacles \cite{Robillard_a_field_study}.
+used a mixed approach of surveys and in-person interviews with professional software engineers to understand the API learning obstacles \cite{Robillard_what_makes} \cite{Robillard_a_field_study}.
 They found that API documentation was related to one of the most severe obstacles for API learnabilty and recommended considering five factors for API documentation as follows: documentation of intent, code examples, matching APIs with scenarios, penetrability of the API, and format and presentation.
@@
 
+\subsubsection{Measuring API Usability}
+Several papers have been published on the topic of measuring API usability.
+Rama et al.
+presented a set of formulas for computing a measure of API usability based on the API's structural components such as classes, methods, parameters, return values, thread-safety, etc \cite{rama}.
+Scheller et al.
+presented a framework for automatically measuring the complexity of an API \cite{scheller2015automated}.
+They identified a list of measurable API properties and provided formulas to compute complexities related to the interfaces, implementation and setup of an API.
+Grill et al.
+proposed an HCI based approach that can be used by API developers to understand and improve on the problem areas related to an API \cite{Grill2012}.
+They suggested using a combination of expert opinion and developer workshops on APIs to identify and collect feedback about API and its documentation related problems.
+In our work, we've focused on the usability related properties of an API and proved relationship between API client developer success and the presence of usage examples on API documentation.
+
 \subsubsection{Controlled Studies}

### 2017-02-27 controlled_study.tex
better ending sentence

@@
 They suggested using a combination of expert opinion and developer workshops on APIs to identify and collect feedback about API and its documentation related problems.
-In our work, we've focused on the usability related properties of an API and proved relationship between API client developer success and the presence of usage examples on API documentation.
+In our work, instead of measuring overall API usability, we have focused specifically on the relationship of REST API usability with respect to usage examples.
 

### 2017-02-27 controlled_study.tex
probably all refs added

@@
 They categorized the API related information needs into twenty generic questions that can be used to analyze the quality of API documentation.
+Myers et al.
+performed a study of enterprise SOA API usability and found that the study participants had limited success finding the relevant API elements to perform an API task \cite{Myers_study}.
+Among other factors, they recommended providing code examples in the API documentation to help API client developers with sample solutions to different patterns of API tasks.
 
@@
 They suggested using a combination of expert opinion and developer workshops on APIs to identify and collect feedback about API and its documentation related problems.
+To improve API usability, Farooq et al.
+proposed using peer reviews of API code in addition to API usability studies to uncover API related bugs and incorporate feedback regarding the API elements \cite{Farooq:2010:AUP:1753326.1753677}.
 In our work, instead of measuring overall API usability, we have focused specifically on the relationship of REST API usability with respect to usage examples.

### 2017-03-05 rename.rb
adds the rename file

@@ -1,0 +1,17 @@
+map = [["P1", "P1.1"], ["P2", "P1.2"], ["P3", "P1.3"], ["P4", "P1.4"], ["P5", "P1.5"], ["P6", "P1.6"], ["P7", "P1.7"], ["P8", "P1.8"], ["P9", "P1.9"], ["P16", "P1.10"], ["P17", "P1.11"], ["P21", "P1.12"], ["P23", "P1.13"], ["P24", "P1.14"], ["P25", "P1.15"], ["P26", "P1.16"]]
+
+
+files = Dir.glob('*/*original*').each do |file|
+  map.each do |old, new_code|
+    `sed -i '' 's/#{old}\ using/#{new_code}\ using/' #{file}`
+  end
+end
+
+map = [["P10", "P2.1"], ["P11", "P2.2"], ["P13", "P2.4"], ["P14", "P2.5"], ["P15", "P2.6"], ["P18", "P2.7"], ["P19", "P2.8"], ["P20", "P2.9"], ["P22", "P2.10"], ["P12", "P2.3"]]
+
+files = Dir.glob('*/*forked*').each do |file|
+  map.each do |old, new_code|
+    `sed -i '' 's/#{old}\ using/#{new_code}\ using/' #{file}`
+  end
+end
+

### 2017-03-05 rename.rb
updates the code for the remaining file

@@ -1,17 +1,17 @@
 map = [["P1", "P1.1"], ["P2", "P1.2"], ["P3", "P1.3"], ["P4", "P1.4"], ["P5", "P1.5"], ["P6", "P1.6"], ["P7", "P1.7"], ["P8", "P1.8"], ["P9", "P1.9"], ["P16", "P1.10"], ["P17", "P1.11"], ["P21", "P1.12"], ["P23", "P1.13"], ["P24", "P1.14"], ["P25", "P1.15"], ["P26", "P1.16"]]
 
 
-files = Dir.glob('*/*original*').each do |file|
+files = Dir.glob('*/*task-2-data*').each do |file|
   map.each do |old, new_code|
-    `sed -i '' 's/#{old}\ using/#{new_code}\ using/' #{file}`
+    `sed -i '' 's/#{old}\ using\ original/#{new_code}\ using\ original/' #{file}`
   end
 end
 
 map = [["P10", "P2.1"], ["P11", "P2.2"], ["P13", "P2.4"], ["P14", "P2.5"], ["P15", "P2.6"], ["P18", "P2.7"], ["P19", "P2.8"], ["P20", "P2.9"], ["P22", "P2.10"], ["P12", "P2.3"]]
 
-files = Dir.glob('*/*forked*').each do |file|
+files = Dir.glob('*/*task-2-data*').each do |file|
   map.each do |old, new_code|
-    `sed -i '' 's/#{old}\ using/#{new_code}\ using/' #{file}`
+    `sed -i '' 's/#{old}\ using\ forked/#{new_code}\ using\ forked/' #{file}`
   end
 end
 

### 2017-03-05 controlled_study.tex
participants table addded

@@
 
-Participants were recruited through online announcements posted on Twitter, Facebook, and software developer focused mailing lists.
+A total 26 participants from sixteen different companies and six different countries (Canada, USA, Germany, Ireland, Brazil, and,
+Bangladesh) were recruited through online announcements posted on Twitter, Facebook, and software developer focused mailing lists.
+Table \ref{table:participants} shows a summary of the experience level of the participants on each group.
+
+\begin{table}[!tbh]
+  \caption{Participant Profile}
+  \label{table:participants}
+
+\begin{tabular}{|p{0.8in} p{.6in} p{0.7in} p{0.7in}|}
+
+\hline
+& & \textbf{Group 1} & \textbf{Group 2}\\
+\hline
+\multicolumn{2}{|l}{\textbf{Number of Participants}} & 16 (P1.1-P1.16) & 10 (P2.1-P2.10) \\
+\hline
+\multicolumn{4}{|l|}{\textbf{Industry Experience}} \\
+& 1-5 years & 5 & 1 \\
+& 5-10 years & 6 & 5 \\
+& 10+ years & 5 & 4 \\
+& Average & 9.1 & 10.6 \\
+\hline
+
+\multicolumn{4}{|l|}{\textbf{REST API Experience}} \\
+&   1-3 years & 5 & 3 \\
+&   3-5 years & 7 & 4 \\
+&   5+ years & 4 & 3 \\
+&   Average & 4.5 & 4 \\
+
+\hline
+\multicolumn{2}{|l}{\textbf{Number of Companies}} & 10 & 8 \\
+\hline
+\end{tabular}
+
+\end{table}
 

### 2017-03-06 controlled_study.tex
add the observations + implications

@@
 
-  Observation and Implications
-    For each task, describe the results. Provide examples of failed attempts with counts, and relate to the documentation sources to share an observation. Imply what needs to be done.
+  Key Observations
+
+  API client developers face problem with using the right data type
+    Proof: Number of codes in task 2 for G1 vs. G2
+    Implication: Example to show a happy path scenario
+
+  API client developers face problem with using the right data format
+    Proof: Number of codes in task 3 for G1 vs. G2
+    Implication: Example showing expected data format
+
+  API client developers face problem with using request headers
+    Proof: Number of codes in task 4 for G1 vs. G2
+    Implication: Example showing all 4 inputs
+
+  API client developers face problem with using inter-dependent APIs
+    Proof: Number of successful cases on G1. vs. G2
+    Implication: Example showing dependency
 

### 2017-03-06 extract_code.rb
extracts the codes into separate files for counting

@@ -1,0 +1,11 @@
+Dir.glob("*/*/*task*-data.md").each do |file|
+  dir = File.dirname(file)
+  file_name = File.basename(file)
+
+  extrated_code_file = File.join(dir, "codes-from-#{file_name}")
+
+  command =  "grep -i \"Participant\\|\\`CODE\" #{file} > #{extrated_code_file}"
+
+  `#{command}`
+
+end

### 2017-03-06 controlled_study.tex
skeleton graph added

@@
   \usepackage{url}
+  \usepackage{tikz,pgfplots,pgfplotstable}
+  \usepgfplotslibrary{statistics}
 
@@
 
-A graph showing the difference in productivity metrics between G1 and G2.
-Implications
+\pgfplotsset{compat=1.14}
+\usetikzlibrary{patterns}
+\pgfplotstableread[col sep=comma]{groupone.csv}\groupone
+
+\begin{tikzpicture}
+    \begin{axis}[
+        ybar,
+        xtick=data,
+        xticklabels from table={\groupone}{Task},
+        legend style={at={(0.5,-0.15)},
+        anchor=north,legend columns=-1},
+        ylabel=Average Trial API Calls,
+        title=Comparison of Trial vs. Successful API Calls,
+        width=2.8in
+    ]
+    \addplot[pattern=crosshatch dots, pattern color=gray] table [col sep=comma, x expr=\coordindex, y=AverageAPICalls] {groupone.csv};
+    \addplot[draw=black] table [col sep=comma, x expr=\coordindex, y=AverageAPICalls] {grouptwo.csv};
+    \legend{G1: Using Original, G2: Using Forked}
+    \end{axis}
+
+    \begin{axis}[
+      axis y line*=right,
+      axis x line=none,
+      ymin=0, ymax=100,
+      ylabel= Success Rate,
+      yticklabel={\pgfmathparse{\tick}\pgfmathprintnumber{\pgfmathresult}\%},
+      width=2.8in
+    ]
+    \addplot[dashed,mark=*,gray]
+      coordinates{
+        (0,0)
+        (0.0148,48)
+        (0.0295,66)
+        (0.0441,66)
+        (0.059,45.0)
+    }; \addlegendentry{G1 Success Rate}
+
+    \addplot[smooth,mark=*,black]
+      coordinates{
+        (0,0)
+        (0.0148,20)
+        (0.0295,10)
+        (0.0441,6)
+        (0.059,5.0)
+    }; \addlegendentry{G2 Success Rate}
+    \end{axis}
+
+\end{tikzpicture}
 

### 2017-03-06 controlled_study.tex
two plots complete

@@
     ]
-    \addplot[dashed,mark=*,gray]
-      coordinates{
-        (0,0)
-        (0.0148,48)
-        (0.0295,66)
-        (0.0441,66)
-        (0.059,45.0)
-    }; \addlegendentry{G1 Success Rate}
+    \addplot[dashed,mark=*,gray] table [col sep=comma, x expr=\coordindex, y=SuccessRate] {groupone_success.csv};
+    \addlegendentry{G1 Success Rate}
 
-    \addplot[smooth,mark=*,black]
-      coordinates{
-        (0,0)
-        (0.0148,20)
-        (0.0295,10)
-        (0.0441,6)
-        (0.059,5.0)
-    }; \addlegendentry{G2 Success Rate}
+    \addplot[smooth,mark=*,black] table [col sep=comma, x expr=\coordindex, y=SuccessRate] {grouptwo_success.csv};
+      \addlegendentry{G2 Success Rate}
     \end{axis}

### 2017-03-06 controlled_study.tex
legends are good now

@@
 
+\begin{figure}
+
 \begin{tikzpicture}
@@
         ylabel=Average Trial API Calls,
-        title=Comparison of Trial vs. Successful API Calls,
         width=2.8in
@@
       yticklabel={\pgfmathparse{\tick}\pgfmathprintnumber{\pgfmathresult}\%},
-      width=2.8in
+      width=2.8in,
+      legend style={at={(1.0,1.25)},anchor=north east}
     ]
@@
 \end{tikzpicture}
+
+\caption{DUMMY Comparison of Trial vs. Successful API Calls}
+\end{figure}
 

### 2017-03-06 controlled_study.tex
labels are uniform
adds real data for quantitative section of the results

@@
       width=2.8in,
-      legend style={at={(1.0,1.25)},anchor=north east}
+      legend style={at={(1.0,1.25)},anchor=north east},
     ]
     \addplot[dashed,mark=*,gray] table [col sep=comma, x expr=\coordindex, y=SuccessRate] {groupone_success.csv};
-    \addlegendentry{G1 Success Rate}
+    \addlegendentry{G1: Using Original}
 
     \addplot[smooth,mark=*,black] table [col sep=comma, x expr=\coordindex, y=SuccessRate] {grouptwo_success.csv};
-      \addlegendentry{G2 Success Rate}
+      \addlegendentry{G2: Using Forked}
     \end{axis}
@@
 
-\caption{DUMMY Comparison of Trial vs. Successful API Calls}
+\caption{Comparison of Trial vs. Successful API Calls Between the Groups}
+\label{graph:productivity}
 \end{figure}

### 2017-03-06 controlled_study.tex
quantitative first pass done

@@
 
-% subsection rq1 (end)
+We present a summary of the quantitative results as found by analyzing a total of 539 API calls (385 from G1, 152 from G2) made by the participants from this study in Figure \ref{graph:productivity}.
+The graph juxtaposes the average number of trial API calls made by each group against their rate of success in performing the given tasks T2-T6.
+The average number of trial API calls and success rate for each group is computed using the following formulas:
+
+AverageTrialAPICall(task, group) = Total number of API calls made by group on task / Number of participants in group
+
+SuccessRate(task, group) = Number of participants that correctly completed task / Number of participants in group
+
+In Figure \ref{graph:productivity} we see that, for the simpler tasks T2 and T3, G1 had used more trial API calls on average and reached successful answers with a lower percentage compared to G2.
+For T4, where the participants are required to inspect a HTTP response header, G2 needed more API calls on average, but also yielded 24\% higher success rate.
+For the more complicated tasks T5 and T6, we see that participants in G1 had very low success rate and very high number of average API trials compared to G2.
+The data here clearly depicts that without usage examples REST API client developers require more trial and error attempts yet resulting in lower success rates, and the situation gets worse for relatively complicated tasks.
+Based on these results, we recommend REST API developers to study the usability of their API documentation with prospective API client developers and provide required usage examples in the documentation.
 

### 2017-03-06 controlled_study.tex
qualitative results

@@
 What have we done?
-Performed an empirical study with 20 professional software engineers with REST API experience to understand how their productivity and types of problems faced changes with respect to usage examples.
+Performed an empirical study with 26 professional software engineers with REST API experience to understand how their productivity and types of problems faced changes with respect to usage examples.
 What have we learned? (Hoping) examples help developer perform task with fewer problems with greater success, with fewer errors, and faster in terms of time.
@@
 
-  Key Observations
+\textbf{Observation 1.} API client developers face problem with using right data type without examples.
+To successfully complete T1 participants had to use an integer ID representing an $Author$ object given the author name.
+In the original API documentation, the following is mentioned about the author parameter: ``author: Limit result set to posts assigned to specific authors.'' The documentation doesn't mention that the data type required is the numeric author ID, not the author name.
+We found 10 of the 16 participants from G1 made API calls with the author name instead of the ID.
+P1.12 mentioned the following: \textbf{``...to find posts for the author, I had to inspect the response to see that indeed the $author\_id$ was in there''}.
+The forked documentation showed one example of using numeric ID.
+3 of the 10 participants from G2 ran into the same error.
+When faced with this error, participants used trial and error to eventually complete the task.
 
-  API client developers face problem with using the right data type
-    Proof: Number of codes in task 2 for G1 vs. G2
-    Implication: Example to show a happy path scenario
+\textbf{Implication 1.} REST API documentation needs to include examples of data types for each API field to satisfy API client developer information need.
 
-  API client developers face problem with using the right data format
-    Proof: Number of codes in task 3 for G1 vs. G2
-    Implication: Example showing expected data format
+\textbf{Observation 2.} API client developers face problem with using the right data format without examples.
+Tasks T3 required the use of a correct format to represent an array of numeric post IDs to be excluded using the API.
+The original documentation mentions the following about this API query parameter: ``exclude: Ensure result set excludes specific IDs.'' Participants attempted to solve this task using multiple different formats for specifying the array.
+For example: they tried with $exclude=[1,4]$, $exclude=1\&4$, $exclude=4\&\&exclude=1$, $id!=1\&id!=4$ and other alternatives before eventually finding the right format as follows $exclude=1,4$.
+Participant P1.6 mentioned the following: \textbf{``It was difficult to figure out whether some of the inputs needed to be arrays or just a comma-separated list.''} 7 out of the 16 participants in G1 faced this error.
+The forked documentation showed one example of using multiple IDs and all G2 participants could use the right format.
+
+Task T5 also required the participants to use an ISO8601 formatted date for publishing a post with a specific date.
+Even though the documentation mentions ISO8601 format, it doesn't provide an example.
+So, participants had to use online search.
+P1.8 mentioned the following: \textbf{``Got hung up trying to figure out the formatting of the date (which doesn't appear to be ISO8601, despite what it says)''}.
+After analyzing the response, we found that P1.8 used the ISO8601 formatted date in the API calls, but didn't provide the $second$ portion of the time as required by the API.
+8 of the 16 participants in G1 had an error in the date formats.
+With an example, 1 of the 10 participants had the same error from G2.
+
+\textbf{Implication 2.} REST API developers need to include examples showing the valid data format for the API elements.
+
+\textbf{Observation 3.} REST API client developers face problem with using request headers without usage examples. Tasks T5 and T6 required the use of a HTTP request header named $Content-Type$
 

### 2017-03-07 controlled_study.tex
complete the observations

@@
 
-\textbf{Observation 3.} REST API client developers face problem with using request headers without usage examples. Tasks T5 and T6 required the use of a HTTP request header named $Content-Type$
+\textbf{Observation 3.} REST API client developers face problem with using request headers without usage examples.
+Tasks T5 and T6 required the use of a HTTP request header named $Content-Type$.
+In the original API documentation, the following was mentioned: ``The API uses JSON exclusively as the request and response format, including error responses.''.
+Experienced software engineers still had problem using this HTTP header as we found from the participants in G1.
+P1.12 mentioned the following: \textbf{``I didn't know how to specify application/json so that I could send POST data via the body instead in the URL parameters...
+Examples would really help - with a sample request and a sample response.''} We were surprised to see that 14 of the 15 participants that attempted T5 and T6 from G1 had the error of not specifying the $Content-Type$ header even though they have years of experience with REST APIs.
+In contrast, 2 out of the 10 participants from G2 failed used this header correctly as indicated in the example in the forked documentation.
+Moreover, 1 of the 2 G2 participants that didn't use the header first time, later used it correctly on a subsequent attempt.
 
-  API client developers face problem with using request headers
-    Proof: Number of codes in task 4 for G1 vs. G2
-    Implication: Example showing all 4 inputs
+\textbf{Implication 3.} If API requests need to use HTTP Request headers, in addition to the request method and body, REST API develpers need to include examples of the HTTP headers.
 
-  API client developers face problem with using inter-dependent APIs
-    Proof: Number of successful cases on G1. vs. G2
-    Implication: Example showing dependency
+\textbf{Observation 4.} API client developers face problem with performing API tasks that require multiple API calls.
+Tasks T2 and T6 required the use of multiple API calls to complete successfully.
+Both tasks require an initial API call to find data using the API that is needed make a second API call.
+For example, P1.5 mentioned the following about T6, \textbf{``The post I created is not visible and not editable, though I got 201.
+Not sure why''}.
+After analyzing the response, we observed that P1.5 was able to create a blog post, but failed to publish it because it didn't have the $publish$ status.
+As a result, the post was not returned via the API for T6.
+P2.2 mentioned the following about T2: \textbf{``In case of search posts by author, it's not clear how to find the author\_id.
+Although, I realized that users and authors are same in this case.''}
+
+\textbf{Implication 4.} If there are prerequisites for making an API call, REST API developers need to provide examples showing how to get the prerequisites in the API documentation.
+
+\textbf{Observation 5.} REST API client developers want usage examples along with the definition of API elements.
+From analyzing the participant feedback, we've observed usage examples mentioned as the most frequent topic of interest.
+11 of the 16 G2 participants mentioned the lack of examples with HTTP headers, request and response as the primary problem with the API documentation.
+On the other hand, 7 out of the 10 G2 participants mentioned the examples being helpful from the forked documentation.
+
+\textbf{Implication 5.} REST API developers need to provide usage examples showing sample requests and responses with HTTP headers and body for API actions to help API usability.
 

### 2017-03-07 controlled_study.tex
complete first pass of results

@@
   \usepackage{url}
+  % \usepackage{caption}
+  \usepackage{subcaption}
   \usepackage{tikz,pgfplots,pgfplotstable}
@@
 
-\textbf{Observation 1.} API client developers face problem with using right data type without examples.
+\textbf{Observation 1.} REST API client developers want usage examples along with the definition of API elements.
+From analyzing the participant feedback, we've observed usage examples mentioned as the most frequent and central topic of interest.
+11 of the 16 G2 participants mentioned the lack of examples with HTTP headers, request and response as the primary problem with the API documentation.
+On the other hand, 7 out of the 10 G2 participants mentioned the examples being helpful from the forked documentation.
+This is also supported by our quantitative analysis as discussed later in this section.
+
+\textbf{Implication 1.} REST API developers need to provide usage examples showing sample requests and responses with HTTP headers and body for API actions to help API usability.
+
+\textbf{Observation 2.} API client developers face problem with using right data type without examples.
 To successfully complete T1 participants had to use an integer ID representing an $Author$ object given the author name.
@@
 
-\textbf{Implication 1.} REST API documentation needs to include examples of data types for each API field to satisfy API client developer information need.
+\textbf{Implication 2.} REST API documentation needs to include examples of data types for each API field to satisfy API client developer information need.
 
-\textbf{Observation 2.} API client developers face problem with using the right data format without examples.
+\textbf{Observation 3.} API client developers face problem with using the right data format without examples.
 Tasks T3 required the use of a correct format to represent an array of numeric post IDs to be excluded using the API.
@@
 
-\textbf{Implication 2.} REST API developers need to include examples showing the valid data format for the API elements.
+\textbf{Implication 3.} REST API developers need to include examples showing the valid data format for the API elements.
 
-\textbf{Observation 3.} REST API client developers face problem with using request headers without usage examples.
+\textbf{Observation 4.} REST API client developers face problem with using request headers without usage examples.
 Tasks T5 and T6 required the use of a HTTP request header named $Content-Type$.
@@
 
-\textbf{Implication 3.} If API requests need to use HTTP Request headers, in addition to the request method and body, REST API develpers need to include examples of the HTTP headers.
+\textbf{Implication 4.} If API requests need to use HTTP Request headers, in addition to the request method and body, REST API develpers need to include examples of the HTTP headers.
 
-\textbf{Observation 4.} API client developers face problem with performing API tasks that require multiple API calls.
+\textbf{Observation 5.} API client developers face problem with performing API tasks that require multiple API calls.
 Tasks T2 and T6 required the use of multiple API calls to complete successfully.
@@
 
-\textbf{Implication 4.} If there are prerequisites for making an API call, REST API developers need to provide examples showing how to get the prerequisites in the API documentation.
+\textbf{Implication 5.} If there are prerequisites for making an API call, REST API developers need to provide examples showing how to get the prerequisites in the API documentation.
 
-\textbf{Observation 5.} REST API client developers want usage examples along with the definition of API elements.
-From analyzing the participant feedback, we've observed usage examples mentioned as the most frequent topic of interest.
-11 of the 16 G2 participants mentioned the lack of examples with HTTP headers, request and response as the primary problem with the API documentation.
-On the other hand, 7 out of the 10 G2 participants mentioned the examples being helpful from the forked documentation.
-
-\textbf{Implication 5.} REST API developers need to provide usage examples showing sample requests and responses with HTTP headers and body for API actions to help API usability.
+To summarize, we recommend API developers to consider usage examples as an essential requirement for REST API documentation.
+As such, sufficient resources and a high priority should be devoted to generate and maintain the usage examples.
+Based on our observations, we recommend REST API documentation tools to provide first-class support for including usage examples with realistic data.
 
@@
 
+\centering
+
+\begin{subfigure}{0.5\textwidth}
+\centering
 \begin{tikzpicture}
@@
 \end{tikzpicture}
+\caption{Trial vs. Successful API Calls}
+\label{graph:productivity}
 
-\caption{Comparison of Trial vs. Successful API Calls Between the Groups}
-\label{graph:productivity}
+\end{subfigure}
+
+\begin{subfigure}{0.5\textwidth}
+\centering
+
+\begin{tikzpicture}
+  \begin{axis}
+    [
+    ytick={1,2,3},
+    yticklabels={G1, G2},
+    width=2.8in,
+    height=1in
+    ]
+    \addplot+[
+    boxplot prepared={
+      median=6,
+      upper quartile=7,
+      lower quartile=3,
+      upper whisker=9,
+      lower whisker=2
+    }, color=black
+    ] coordinates {};
+    \addplot+[
+    boxplot prepared={
+      median=7,
+      upper quartile=8,
+      lower quartile=7,
+      upper whisker=9,
+      lower whisker=7
+    }, color = black
+    ] coordinates {};
+  \end{axis}
+\end{tikzpicture}
+\caption{Participation Satisfaction Ratings}
+\label{fig:rating}
+
+\end{subfigure}
+
+\caption{Comparison Between G1 and G2}
+\label{fig:comparison}
 \end{figure}
 
-We present a summary of the quantitative results as found by analyzing a total of 539 API calls (385 from G1, 152 from G2) made by the participants from this study in Figure \ref{graph:productivity}.
-The graph juxtaposes the average number of trial API calls made by each group against their rate of success in performing the given tasks T2-T6.
+We present a summary of the quantitative results as found by analyzing a total of 539 API calls (385 from G1, 152 from G2) made by the participants from this study in Figure \ref{fig:comparison}.
+Figure \ref{graph:productivity} juxtaposes the average number of trial API calls made by each group against their rate of success in performing the given tasks T2-T6.
 The average number of trial API calls and success rate for each group is computed using the following formulas:
@@
 The data here clearly depicts that without usage examples REST API client developers require more trial and error attempts yet resulting in lower success rates, and the situation gets worse for relatively complicated tasks.
-Based on these results, we recommend REST API developers to study the usability of their API documentation with prospective API client developers and provide required usage examples in the documentation.
+
+In Figure \ref{fig:rating}, we see that the G2 participants provided higher satisfaction rating compared to G1 participants.
+P1.1 provided a satisfaction rating of 9, and analyzing the participant feedback against the success, we see a false positive case, where the participant perceived that the tasks were successfully completed even though only 2 of the 5 tasks were correctly answered.
+Other than this case, the satisfaction ratings provided by G1 participants consistently fall behind the G2 participants' ratings.
+
+Based on these results, we recommend REST API developers to study the usability of their API documentation with prospective API client developers and provide required usage examples in the documentation to improve API client developer productivity and satisfaction.
 

### 2017-03-07 controlled_study.tex
box around the images
better images
abstract first pass

@@
 \begin{abstract}
-
-[Brain-dump mode]
-
-What is the the problem? Effort and cost associated to REST API documentation with usage examples.
-Why is this a problem? Cost vs. benefit of usage examples is not known. What to include in the examples?
-What have we done?
-Performed an empirical study with 26 professional software engineers with REST API experience to understand how their productivity and types of problems faced changes with respect to usage examples.
-What have we learned? (Hoping) examples help developer perform task with fewer problems with greater success, with fewer errors, and faster in terms of time.
-What does it mean? (expected) API developers should include usage examples in REST API documentation that include specific examples based on the problems faced.
+Generating and maintaining REST API documentation with usage examples can be a time consuming and expensive process for evolving APIs.
+Most REST API documentation tools have focused on automating the documentation of the API objects, but require manual effort for capturing realistic usage examples.
+So, REST API developers need to know the cost vs.
+benefit of providing usage examples in the documentation as well as what to include in the usage examples to prioritize the documentation tasks.
+In this paper, we present our observations about the obstacles that REST API client developers face without usage examples in the documentation.
+To this regard, we have performed a controlled study with 26 experienced software engineers and found that REST API client developers face productivity problems using correct data types, data formats, required HTTP headers and request body when documentation lacks usage examples.
+The REST API documentation guidelines shared on this paper will help practitioners to improve API client developer productivity and usability.
 
@@
 \begin{center}
-\includegraphics[width=\linewidth,keepaspectratio]{schema.png}
+\fbox{\includegraphics[width=\linewidth,keepaspectratio]{schema.png}}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\linewidth,keepaspectratio]{api_explorer.png}
+\fbox{\includegraphics[width=\linewidth,keepaspectratio]{api_explorer.png}}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\linewidth,keepaspectratio]{original.png}
+\fbox{\includegraphics[width=\linewidth,height=2.5in,keepaspectratio]{original.png}}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\linewidth,keepaspectratio]{forked.png}
+\fbox{\includegraphics[width=\linewidth,height=2.5in,keepaspectratio]{forked.png}}
 \end{center}
@@
 \centering
+
 \begin{tikzpicture}
@@
 
-\caption{Comparison Between G1 and G2}
+\caption{Quantitative Comparison between G1 and G2}
 \label{fig:comparison}

### 2017-03-07 controlled_study.tex
stronger implications

@@
 To this regard, we have performed a controlled study with 26 experienced software engineers and found that REST API client developers face productivity problems using correct data types, data formats, required HTTP headers and request body when documentation lacks usage examples.
-The REST API documentation guidelines shared on this paper will help practitioners to improve API client developer productivity and usability.
+By following our REST API documentation guidelines shared on this paper, practitioners can reduce the errors, improve success rate and satisfaction of REST API client developers.
 

### 2017-03-08 controlled_study.tex
better abstract

--- 2017-03-07 controlled_study.tex
@@
 Generating and maintaining REST API documentation with usage examples can be a time consuming and expensive process for evolving APIs.
-Most REST API documentation tools have focused on automating the documentation of the API objects, but require manual effort for capturing realistic usage examples.
+Most REST API documentation tools focus on automating the documentation of the API objects, but require manual effort for capturing realistic usage examples.
 So, REST API developers need to know the cost vs.
-benefit of providing usage examples in the documentation as well as what to include in the usage examples to prioritize the documentation tasks.
+benefit of providing usage examples in the documentation as well as what to include in the usage examples to prioritize the documentation efforts.
 In this paper, we present our observations about the obstacles that REST API client developers face without usage examples in the documentation.
-To this regard, we have performed a controlled study with 26 experienced software engineers and found that REST API client developers face productivity problems using correct data types, data formats, required HTTP headers and request body when documentation lacks usage examples.
-The REST API documentation guidelines shared on this paper will help practitioners to improve API client developer productivity and usability.
+To this regard, we have performed a controlled study with 26 experienced software engineers and found that REST API client developers face productivity problems with using correct data types, data formats, required HTTP headers and request body when documentation lacks usage examples.
+By following the REST API documentation guidelines from this paper, practitioners can reduce the errors, improve success rate and satisfaction of REST API client developers.
 
@@
 \subsection{Research Goals}
-
+user
 This research is aimed at answering the following research questions:

### 2017-03-08 controlled_study.tex
time graph addedxt

@@
 
-\begin{figure}
+\begin{figure*}
 
@@
 \begin{tikzpicture}
+    \begin{axis}[
+        ybar,
+        xtick=data,
+        xticklabels from table={\groupone}{Task},
+        legend style={at={(0.5,-0.15)},
+        anchor=north,legend columns=-1},
+        ylabel=Average Time Spent (minutes),
+        width=2.8in
+    ]
+    \addplot[pattern=crosshatch dots, pattern color=gray] table [col sep=comma, x expr=\coordindex, y=Time] {groupone_time.csv};
+    \addplot[draw=black] table [col sep=comma, x expr=\coordindex, y=Time] {grouptwo_time.csv};
+    \legend{G1: Using Original, G2: Using Forked}
+    \end{axis}
+
+\end{tikzpicture}
+\caption{Average Time Spent}
+\label{graph:time}
+
+\end{subfigure}
+
+\begin{subfigure}{0.5\textwidth}
+\centering
+
+\begin{tikzpicture}
   \begin{axis}
@@
 \label{fig:comparison}
-\end{figure}
+\end{figure*}
 

### 2017-03-08 controlled_study.tex
add graph about time

@@
 \centering
-
 \begin{tikzpicture}
@@
     \end{axis}
-
 \end{tikzpicture}
@@
 \label{graph:productivity}
-
 \end{subfigure}
-
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}{0.4\textwidth}
 \centering
-
 \begin{tikzpicture}
@@
 \label{graph:time}
-
 \end{subfigure}
 
-\begin{subfigure}{0.5\textwidth}
+\caption{Quantitative Comparison between G1 and G2}
+\label{fig:comparison}
+\end{figure*}
+
+We present a summary of the quantitative results as found by analyzing a total of 539 API calls (385 from G1, 152 from G2) made by the participants from this study in Figure \ref{fig:comparison}.
+Figure \ref{graph:productivity} juxtaposes the average number of trial API calls made by each group against their rate of success in performing the given tasks T2-T6.
+The average number of trial API calls and success rate for each group is computed using the following formulas:
+
+AverageTrialAPICall(task, group) = Total number of API calls made by group on task / Number of participants in group
+
+SuccessRate(task, group) = Number of participants that correctly completed task / Number of participants in group
+
+In Figure \ref{graph:productivity} and  \ref{graph:time} we see that, for the simpler tasks T2 and T3, G1 had used more trial API calls and more time on average and reached successful answers with a lower percentage compared to G2.
+For T4, where the participants are required to inspect a HTTP response header, G2 needed more API calls on average, but also yielded 24\% higher success rate.
+For the more complicated tasks T5 and T6, we see that participants in G1 had very low success rate and very high number of average API trials compared to G2.
+The data here clearly depicts that without usage examples REST API client developers spend more time and require more trial and error attempts yet resulting in lower success rates, and the situation gets worse for relatively complicated tasks.
+
+\begin{figure}[h]
 \centering
@@
 
-\end{subfigure}
-
-\caption{Quantitative Comparison between G1 and G2}
-\label{fig:comparison}
-\end{figure*}
-
-We present a summary of the quantitative results as found by analyzing a total of 539 API calls (385 from G1, 152 from G2) made by the participants from this study in Figure \ref{fig:comparison}.
-Figure \ref{graph:productivity} juxtaposes the average number of trial API calls made by each group against their rate of success in performing the given tasks T2-T6.
-The average number of trial API calls and success rate for each group is computed using the following formulas:
-
-AverageTrialAPICall(task, group) = Total number of API calls made by group on task / Number of participants in group
-
-SuccessRate(task, group) = Number of participants that correctly completed task / Number of participants in group
-
-In Figure \ref{graph:productivity} we see that, for the simpler tasks T2 and T3, G1 had used more trial API calls on average and reached successful answers with a lower percentage compared to G2.
-For T4, where the participants are required to inspect a HTTP response header, G2 needed more API calls on average, but also yielded 24\% higher success rate.
-For the more complicated tasks T5 and T6, we see that participants in G1 had very low success rate and very high number of average API trials compared to G2.
-The data here clearly depicts that without usage examples REST API client developers require more trial and error attempts yet resulting in lower success rates, and the situation gets worse for relatively complicated tasks.
+\end{figure}
 

### 2017-03-08 controlled_study.tex
better caption for rating graph
first pass complete

@@
 \section{Introduction}
-\section{Methodolody}
+Researchers have emphasized API documentation as key factor that impacts API usability both positively and negatively.
+To improve API documentation, researchers have recommended incorporating usage examples in the API documentation.
+REST APIs are used as the predominant application integration mechanism over the Internet.
+Generating and maintaining REST API documentation with usage examples can be an expensive process because most API documentation tools have a lack of support for capturing realistic usage examples.
+
+In a resource constrained environment, it is important to understand the value of usage examples on REST API usability to allocate sufficient attention and efforts towards incorporating examples in the API documentation.
+While it is no surprise that examples help API client developers, REST API developers need to know what to include in the examples and why.
+To our knowledge, there's a lack of published papers that investigated the impact of different types of content that can be included as usage examples in REST API documentation.
+
+The documentation of REST APIs has distinctive features compared to the documentation of local APIs such as Java libraries.
+For example, local API documentation commonly comprises of describing the modules, classes and interfaces with their methods.
+On the other hand, REST API documentation needs to include information about HTTP headers, request and response body and the data representation formats such as JSON, XML, etc.
+In our work we have focused on understanding the impact of usage examples within the realm of the distinctive REST API features.
+
+We have designed and performed a controlled study with experienced software engineers.
+Participants were divided into two groups and given the same set of API tasks to complete.
+One group was given the official WordPress REST API documentation and another group was given a forked version of the documentation where three usage examples were added.
+Using a novel technique, we have collected 539 API calls made by the participants.
+We have analyzed the data and observed recurring obstacles faced by the participants while performing the API tasks using the official documentation that lacks usage examples.
+Our contributions on this paper is as follows:
+
+\begin{itemize}
+  \item We provide a list of obstacles that REST API client developers face while performing API tasks using documentation that lacks usage examples.
+  \item We provide a list of recommendations for REST developers to be used as a guideline to incorporate usage examples in the API documentation.
+  \item We provide an empirical evidence that usage examples in REST API documentation helps API client developers perform API tasks with higher developer satisfaction, less time, and better success rate.
+\end{itemize}
+
+The remainder of this paper is organized as follows: In Section \ref{sec:method} we discuss our research goals, requirements, the object of the study and the participant selection, the study setup, data collection and analysis methods.
+In Section \ref{sec:results} we discuss the results of our analysis.
+We discuss the threats to validity and provide a summary of the related work in Sections \ref{sec:threat} and  \ref{sec:related_work} respectively.
+
+\section{Methodology}
+\label{sec:method}
 \subsection{Research Goals}
-user
 This research is aimed at answering the following research questions:
@@
 \section{Results}
-
+\label{sec:results}
 \subsection{Qualitative Analysis Results} % (fold)
@@
 \end{tikzpicture}
-\caption{Participation Satisfaction Ratings}
+\caption{Participation Satisfaction Ratings on a Scale of 0-10}
 \label{fig:rating}
@@
 \section{Threats to Validity}
-  API Selection
-  Task Selection
-  Participant Selection
-  Experiment Design
+\label{sec:threat}
+  The selection of the API, the tasks, and the participants may introduce selection bias.
+We have developed and followed a set of requirements as discussed earlier on this paper to reduce this bias.
+This study needs to be replicated with different sets of APIs, tasks, and participants to generalize the results.
+The analysis of the raw data may introduce a bias.
+To reduce this, we have involved multiple researchers in the analysis process.
 
 \section{Related Work}
-
+\label{sec:related_work}
 \subsubsection{API Learning Obstacles}
@@
 \section{Conclusion}
+In this paper, we presented a set of problems that experienced REST API client developers face while performing API tasks using API documentation that lacks usage examples.
+We identified that, without examples, REST API client developers have trouble using the correct data types, correct data formats, and required HTTP headers and request body.
+We've also presented empirical evidence that adding usage examples it is possible to reduce mistakes, improve success rate and developer satisfaction of using the API.
+Practitioners can follow our guidelines while documenting REST APIs to improve API usability.
+
 \section*{Acknowledgment}

### 2017-03-08 controlled_study.tex
better image position
smaller abstract
add example of the wrong date formats

@@
 benefit of providing usage examples in the documentation as well as what to include in the usage examples to prioritize the documentation efforts.
-In this paper, we present our observations about the obstacles that REST API client developers face without usage examples in the documentation.
 To this regard, we have performed a controlled study with 26 experienced software engineers and found that REST API client developers face productivity problems with using correct data types, data formats, required HTTP headers and request body when documentation lacks usage examples.
@@
 \label{sec:results}
-\subsection{Qualitative Analysis Results} % (fold)
-
-\textbf{Observation 1.} REST API client developers want usage examples along with the definition of API elements.
-From analyzing the participant feedback, we've observed usage examples mentioned as the most frequent and central topic of interest.
-11 of the 16 G2 participants mentioned the lack of examples with HTTP headers, request and response as the primary problem with the API documentation.
-On the other hand, 7 out of the 10 G2 participants mentioned the examples being helpful from the forked documentation.
-This is also supported by our quantitative analysis as discussed later in this section.
-
-\textbf{Implication 1.} REST API developers need to provide usage examples showing sample requests and responses with HTTP headers and body for API actions to help API usability.
-
-\textbf{Observation 2.} API client developers face problem with using right data type without examples.
-To successfully complete T1 participants had to use an integer ID representing an $Author$ object given the author name.
-In the original API documentation, the following is mentioned about the author parameter: ``author: Limit result set to posts assigned to specific authors.'' The documentation doesn't mention that the data type required is the numeric author ID, not the author name.
-We found 10 of the 16 participants from G1 made API calls with the author name instead of the ID.
-P1.12 mentioned the following: \textbf{``...to find posts for the author, I had to inspect the response to see that indeed the $author\_id$ was in there''}.
-The forked documentation showed one example of using numeric ID.
-3 of the 10 participants from G2 ran into the same error.
-When faced with this error, participants used trial and error to eventually complete the task.
-
-\textbf{Implication 2.} REST API documentation needs to include examples of data types for each API field to satisfy API client developer information need.
-
-\textbf{Observation 3.} API client developers face problem with using the right data format without examples.
-Tasks T3 required the use of a correct format to represent an array of numeric post IDs to be excluded using the API.
-The original documentation mentions the following about this API query parameter: ``exclude: Ensure result set excludes specific IDs.'' Participants attempted to solve this task using multiple different formats for specifying the array.
-For example: they tried with $exclude=[1,4]$, $exclude=1\&4$, $exclude=4\&\&exclude=1$, $id!=1\&id!=4$ and other alternatives before eventually finding the right format as follows $exclude=1,4$.
-Participant P1.6 mentioned the following: \textbf{``It was difficult to figure out whether some of the inputs needed to be arrays or just a comma-separated list.''} 7 out of the 16 participants in G1 faced this error.
-The forked documentation showed one example of using multiple IDs and all G2 participants could use the right format.
-
-Task T5 also required the participants to use an ISO8601 formatted date for publishing a post with a specific date.
-Even though the documentation mentions ISO8601 format, it doesn't provide an example.
-So, participants had to use online search.
-P1.8 mentioned the following: \textbf{``Got hung up trying to figure out the formatting of the date (which doesn't appear to be ISO8601, despite what it says)''}.
-After analyzing the response, we found that P1.8 used the ISO8601 formatted date in the API calls, but didn't provide the $second$ portion of the time as required by the API.
-8 of the 16 participants in G1 had an error in the date formats.
-With an example, 1 of the 10 participants had the same error from G2.
-
-\textbf{Implication 3.} REST API developers need to include examples showing the valid data format for the API elements.
-
-\textbf{Observation 4.} REST API client developers face problem with using request headers without usage examples.
-Tasks T5 and T6 required the use of a HTTP request header named $Content-Type$.
-In the original API documentation, the following was mentioned: ``The API uses JSON exclusively as the request and response format, including error responses.''.
-Experienced software engineers still had problem using this HTTP header as we found from the participants in G1.
-P1.12 mentioned the following: \textbf{``I didn't know how to specify application/json so that I could send POST data via the body instead in the URL parameters...
-Examples would really help - with a sample request and a sample response.''} We were surprised to see that 14 of the 15 participants that attempted T5 and T6 from G1 had the error of not specifying the $Content-Type$ header even though they have years of experience with REST APIs.
-In contrast, 2 out of the 10 participants from G2 failed used this header correctly as indicated in the example in the forked documentation.
-Moreover, 1 of the 2 G2 participants that didn't use the header first time, later used it correctly on a subsequent attempt.
-
-\textbf{Implication 4.} If API requests need to use HTTP Request headers, in addition to the request method and body, REST API develpers need to include examples of the HTTP headers.
-
-\textbf{Observation 5.} API client developers face problem with performing API tasks that require multiple API calls.
-Tasks T2 and T6 required the use of multiple API calls to complete successfully.
-Both tasks require an initial API call to find data using the API that is needed make a second API call.
-For example, P1.5 mentioned the following about T6, \textbf{``The post I created is not visible and not editable, though I got 201.
-Not sure why''}.
-After analyzing the response, we observed that P1.5 was able to create a blog post, but failed to publish it because it didn't have the $publish$ status.
-As a result, the post was not returned via the API for T6.
-P2.2 mentioned the following about T2: \textbf{``In case of search posts by author, it's not clear how to find the author\_id.
-Although, I realized that users and authors are same in this case.''}
-
-\textbf{Implication 5.} If there are prerequisites for making an API call, REST API developers need to provide examples showing how to get the prerequisites in the API documentation.
-
-To summarize, we recommend API developers to consider usage examples as an essential requirement for REST API documentation.
-As such, sufficient resources and a high priority should be devoted to generate and maintain the usage examples.
-Based on our observations, we recommend REST API documentation tools to provide first-class support for including usage examples with realistic data.
-
-\subsection{Quantitative Analysis Results} % (fold)
 
@@
 
-\begin{figure*}
+\begin{figure*}[htb]
 
@@
 \end{figure*}
+\subsection{Qualitative Analysis Results} % (fold)
+
+\textbf{Observation 1.} REST API client developers want usage examples along with the definition of API elements.
+From analyzing the participant feedback, we've observed usage examples mentioned as the most frequent and central topic of interest.
+11 of the 16 G2 participants mentioned the lack of examples with HTTP headers, request and response as the primary problem with the API documentation.
+On the other hand, 7 out of the 10 G2 participants mentioned the examples being helpful from the forked documentation.
+This is also supported by our quantitative analysis as discussed later in this section.
+
+\textbf{Implication 1.} REST API developers need to provide usage examples showing sample requests and responses with HTTP headers and body for API actions to help API usability.
+
+\textbf{Observation 2.} API client developers face problem with using right data type without examples.
+To successfully complete T1 participants had to use an integer ID representing an $Author$ object given the author name.
+In the original API documentation, the following is mentioned about the author parameter: ``author: Limit result set to posts assigned to specific authors.'' The documentation doesn't mention that the data type required is the numeric author ID, not the author name.
+We found 10 of the 16 participants from G1 made API calls with the author name instead of the ID.
+P1.12 mentioned the following: \textbf{``...to find posts for the author, I had to inspect the response to see that indeed the $author\_id$ was in there''}.
+The forked documentation showed one example of using numeric ID.
+3 of the 10 participants from G2 ran into the same error.
+When faced with this error, participants used trial and error to eventually complete the task.
+
+\textbf{Implication 2.} REST API documentation needs to include examples of data types for each API field to satisfy API client developer information need.
+
+\textbf{Observation 3.} API client developers face problem with using the right data format without examples.
+Tasks T3 required the use of a correct format to represent an array of numeric post IDs to be excluded using the API.
+The original documentation mentions the following about this API query parameter: ``exclude: Ensure result set excludes specific IDs.'' Participants attempted to solve this task using multiple different formats for specifying the array.
+For example: they tried with $exclude=[1,4]$, $exclude=1\&4$, $exclude=4\&\&exclude=1$, $id!=1\&id!=4$ and other alternatives before eventually finding the right format as follows $exclude=1,4$.
+Participant P1.6 mentioned the following: \textbf{``It was difficult to figure out whether some of the inputs needed to be arrays or just a comma-separated list.''} 7 out of the 16 participants in G1 faced this error.
+The forked documentation showed one example of using multiple IDs and all G2 participants could use the right format.
+
+Task T5 also required the participants to use an ISO8601 formatted date for publishing a post with a specific date.
+Even though the original documentation mentions ISO8601 format, it doesn't provide an example.
+Instead of using ``2016-01-01T00:00:00-00:00'', participants from G1 used date formats as follows: ``January 1, 2016'', ``2016-01-01T12:00:00'', ``20160101'', etc.
+P1.8 mentioned the following: \textbf{``Got hung up trying to figure out the formatting of the date (which doesn't appear to be ISO8601, despite what it says)''}.
+After analyzing the response, we found that P1.8 used the ISO8601 formatted date in the API calls, but didn't provide the $second$ portion of the time as required by the API.
+8 of the 16 participants in G1 had an error in the date formats.
+With an example, 1 of the 10 participants had the same error from G2.
+
+\textbf{Implication 3.} REST API developers need to include examples showing the valid data format for the API elements.
+
+\textbf{Observation 4.} REST API client developers face problem with using request headers without usage examples.
+Tasks T5 and T6 required the use of a HTTP request header named $Content-Type$.
+In the original API documentation, the following was mentioned: ``The API uses JSON exclusively as the request and response format, including error responses.''.
+Experienced software engineers still had problem using this HTTP header as we found from the participants in G1.
+P1.12 mentioned the following: \textbf{``I didn't know how to specify application/json so that I could send POST data via the body instead in the URL parameters...
+Examples would really help - with a sample request and a sample response.''} We were surprised to see that 14 of the 15 participants that attempted T5 and T6 from G1 had the error of not specifying the $Content-Type$ header even though they have years of experience with REST APIs.
+In contrast, 2 out of the 10 participants from G2 failed used this header correctly as indicated in the example in the forked documentation.
+Moreover, 1 of the 2 G2 participants that didn't use the header first time, later used it correctly on a subsequent attempt.
+
+\textbf{Implication 4.} If API requests need to use HTTP Request headers, in addition to the request method and body, REST API develpers need to include examples of the HTTP headers.
+
+\textbf{Observation 5.} API client developers face problem with performing API tasks that require multiple API calls.
+Tasks T2 and T6 required the use of multiple API calls to complete successfully.
+Both tasks require an initial API call to find data using the API that is needed make a second API call.
+For example, P1.5 mentioned the following about T6, \textbf{``The post I created is not visible and not editable, though I got 201.
+Not sure why''}.
+After analyzing the response, we observed that P1.5 was able to create a blog post, but failed to publish it because it didn't have the $publish$ status.
+As a result, the post was not returned via the API for T6.
+P2.2 mentioned the following about T2: \textbf{``In case of search posts by author, it's not clear how to find the author\_id.
+Although, I realized that users and authors are same in this case.''}
+
+\textbf{Implication 5.} If there are prerequisites for making an API call, REST API developers need to provide examples showing how to get the prerequisites in the API documentation.
+
+To summarize, we recommend API developers to consider usage examples as an essential requirement for REST API documentation.
+As such, sufficient resources and a high priority should be devoted to generate and maintain the usage examples.
+Based on our observations, we recommend REST API documentation tools to provide first-class support for including usage examples with realistic data.
+
+\subsection{Quantitative Analysis Results} % (fold)
 

### 2017-03-19 controlled_study.tex
add acknowledgement
+keywords
better positioning for screenshots
made all suggested editing changes

@@
   \usepgfplotslibrary{statistics}
-
 \else
@@
 \begin{document}
-\title{How Important are Usage Examples in REST API Documentation?}
+\title{A Study of the Effectiveness of Usage Examples in REST API Documentation}
 
@@
 Most REST API documentation tools focus on automating the documentation of the API objects, but require manual effort for capturing realistic usage examples.
-So, REST API developers need to know the cost vs.
+Consequently, REST API developers need to know the cost vs.
 benefit of providing usage examples in the documentation as well as what to include in the usage examples to prioritize the documentation efforts.
-To this regard, we have performed a controlled study with 26 experienced software engineers and found that REST API client developers face productivity problems with using correct data types, data formats, required HTTP headers and request body when documentation lacks usage examples.
+To this end, we have performed a controlled study with 26 experienced software engineers and found that REST API client developers face productivity problems with using correct data types, data formats, required HTTP headers and request body when documentation lacks usage examples.
 By following the REST API documentation guidelines from this paper, practitioners can reduce the errors, improve success rate and satisfaction of REST API client developers.
@@
 
+\begin{IEEEkeywords}
+API; REST; Documentation; Usage Examples; Empirical Study; Controlled Study; Productivity;
+
+\end{IEEEkeywords}
+
 \IEEEpeerreviewmaketitle
@@
 \section{Introduction}
-Researchers have emphasized API documentation as key factor that impacts API usability both positively and negatively.
+Researchers have emphasized API documentation as a key factor that impacts API usability both positively and negatively.
 To improve API documentation, researchers have recommended incorporating usage examples in the API documentation.
 REST APIs are used as the predominant application integration mechanism over the Internet.
-Generating and maintaining REST API documentation with usage examples can be an expensive process because most API documentation tools have a lack of support for capturing realistic usage examples.
+Generating and maintaining REST API documentation with usage examples can be an expensive process because most API documentation tools do not support effective usage examples.
 
@@
 While it is no surprise that examples help API client developers, REST API developers need to know what to include in the examples and why.
-To our knowledge, there's a lack of published papers that investigated the impact of different types of content that can be included as usage examples in REST API documentation.
 
 The documentation of REST APIs has distinctive features compared to the documentation of local APIs such as Java libraries.
-For example, local API documentation commonly comprises of describing the modules, classes and interfaces with their methods.
-On the other hand, REST API documentation needs to include information about HTTP headers, request and response body and the data representation formats such as JSON, XML, etc.
+For example, local API documentation commonly comprises the description of classes and interfaces with their methods.
+In contrast, REST API documentation needs to include information about HTTP headers, request and response body and the data representation formats such as JSON, XML, etc.
 In our work we have focused on understanding the impact of usage examples within the realm of the distinctive REST API features.
 
-We have designed and performed a controlled study with experienced software engineers.
+We designed and performed a controlled study with experienced software engineers.
 Participants were divided into two groups and given the same set of API tasks to complete.
-One group was given the official WordPress REST API documentation and another group was given a forked version of the documentation where three usage examples were added.
-Using a novel technique, we have collected 539 API calls made by the participants.
-We have analyzed the data and observed recurring obstacles faced by the participants while performing the API tasks using the official documentation that lacks usage examples.
+One group was given the official WordPress REST API documentation and another group was given a enhanced version of the documentation where three usage examples were added.
+Using a novel technique, we collected 539 API calls made by the participants.
+We analyzed the data and observed recurring obstacles faced by the participants while performing the API tasks using the official documentation that lacks usage examples.
 Our contributions on this paper is as follows:
@@
   \item We provide a list of recommendations for REST developers to be used as a guideline to incorporate usage examples in the API documentation.
-  \item We provide an empirical evidence that usage examples in REST API documentation helps API client developers perform API tasks with higher developer satisfaction, less time, and better success rate.
+  \item We provide empirical evidence that usage examples in REST API documentation help API client developers perform API tasks with higher developer satisfaction, less time, and better success rate.
 \end{itemize}
@@
 
-To answer the aforementioned research questions, the study has the following requirements:
+To answer the aforementioned research question, the study has the following requirements:
 
@@
 Representative API.} We had to choose an existing REST API that is currently used by API client developers.
-Selecting a mature REST API for this study reduces the possibility of an underdeveloped solution that may be found if a new or a seldom used API is selected.
+Selecting a mature REST API for this study reduces the possibility that the obstacles we observe in the study are the results of insignificant accidental problems symptomatic of an immature technology.
 A familiar domain needed to be selected so that participants are able to relate to the API features without requiring upfront training.
@@
 Open source.} To be able to understand the impact of usage examples, we needed to select an open-source API where we can add new examples to the documentation for performing this study.
-For proprietary APIs, it won't be possible to retrofit usage examples on existing API documentation without separating the examples from the documentation.
   \item \textbf{R3.
@@
 Before January 2017, the WordPress REST API V2 was distributed as a plug-in where WordPress users could optionally install the API component.
-The following statistics are for the plug-in install numbers between May 2015 and October 2016:
+The following statistics are for the plug-in installation numbers between May 2015 and October 2016:
 
 \begin{itemize}
-  \item Total installs: aprox. 248K installs of the plug-in.
-  \item Average daily installs: approx. 500.
+  \item Total installations: ~248K installs of the plug-in.
+  \item Average daily installations: ~500.
 \end{itemize}
 
-Starting January 2017, the WordPress REST API V2 is no longer required to be installed as a separate plug-in since it's pre-bundled with every WordPress install.
-As per the code repository on GitHub, there are a total of 99 and 46 contributors that had at least one commit to the code repository behind the API and and it's documentation, respectively.
+Starting January 2017, the WordPress REST API V2 is no longer required to be installed as a separate plug-in since it's bundled with every WordPress installation.
+According to the code repository on GitHub, there are a total of 99 and 46 contributors that had at least one commit to the code repository of the API and its documentation, respectively.
 These properties satisfy R1, our requirements for using a representative API.
@@
 \lstset{basicstyle=\footnotesize}
-\begin{lstlisting}[language=php,breaklines=true,showspaces=false,showstringspaces=false,numbers=left,xleftmargin=2em,caption={Example of self-documenting API Code},label=wordpress_code]
+\begin{lstlisting}[language=php,breaklines=true,showspaces=false,showstringspaces=false,numbers=left,xleftmargin=2em,caption={Example of self-documenting API Code in PHP},label=wordpress_code]
 public function get_item_schema() {
@@
 On Listing \ref{wordpress_code}, line 4 specifies that this is a schema definition for the API element $Post$.
-Then, on Line 10, it defines $date$, one of the properties of $Post$, followed by a human readable description and and type information.
-Then, on line 14, the context of this property is mentioned as $view, edit, embed$, meaning that this property will be returned when the $Post$ object is returned, embedded, or can be used as an input for editing.
+On Line 10, it defines $date$, one of the properties of $Post$, followed by a human readable description and and type information.
+On line 14, the context of this property is mentioned as $view, edit, embed$, meaning that this property will be returned when the $Post$ object is returned, embedded, or can be used as an input for editing.
 
-This self-documenting feature is leveraged to generate and publish the official API documentation as an HTML website.
-Figure \ref{fig:schema} shows a screenshot of the published documentation for the $date$ attribute of the $Post$ API element\footnote{\url{http://v2.wp-api.org/reference/posts/}}.
+The WordPress team leverages the this self-documenting feature to generate and publish the official API documentation as an HTML website.
+Figure \ref{fig:schema} shows a screenshot of the published documentation for the $date$ attribute of the $Post$ API element.\footnote{\url{http://v2.wp-api.org/reference/posts/}}
 
@@
 The participants are requested to perform a total of six tasks using the API including one practice task.
-All of the tasks are related to a single API element, $Post$.
-The tasks get progressively difficult, and all but the last task can be performed independently of each other.
+All of the tasks are related to a single API element, $Post$ \footnote{A Post identifies a blog post within WordPress}.
+The tasks get progressively more difficult, and all but the last task can be performed independently of each other.
 Participants are requested to limit the total time on the study to a maximum of one hour.
-Participants are encouraged to proceed to the next task when they are either satisfied about or stuck on the right answer.
+Participants are encouraged to proceed to the next task when they are either satisfied with their answer or feel stuck and unable to make progress.
 
@@
 To perform the tasks, the participants are required to use one or more of the inputs I1-4.
-In the following paragraphs, we describe each task with it's description and the intended test scenario against the aforementioned API input and output information.
+In the following paragraphs, we describe each task with its description and the study observation goals against the aforementioned API input and output information.
 For each of the tasks, the participants are required to use the same WordPress REST API and its documentation.
 
-\textbf{T1: ListAllPostsTask}.
-We ask the participants to use the WordPress REST API following its documentation to get a list of the blog posts from a live WordPress site.
+\textbf{T1: List All Posts Task}.
+We ask the participants to use the WordPress REST API to get a list of the blog posts from a given WordPress site.
 This is the practice task, and the inputs to answer this task are pre-filled for the participants.
@@
 
-\textbf{T2: FilterPostsByAuthorTask.} The participants are asked to use the API to filter the list of posts by an author given the author's user name.
+\textbf{T2: FilterPostsByAuthorTask.} The participants are asked to use the API to filter the list of posts obtained in T1 by an author given the author's user name.
 To answer this correctly, the participants are required to first make an API call to get the numeric ID of the author given the string based user name.
@@
 
-To satisfy the requirement of developers with REST API experience (R3), we have used the following criteria for recruiting the study participants:
+To satisfy the requirement of developers with REST API experience (R3), we used the following criteria for recruiting the study participants:
 
@@
 Bangladesh) were recruited through online announcements posted on Twitter, Facebook, and software developer focused mailing lists.
-Table \ref{table:participants} shows a summary of the experience level of the participants on each group.
+Table \ref{table:participants} shows a summary of the experience level of the participants in each group.
 
@@
 
-\subsubsection{Process}
+\begin{figure*}[t]
+\begin{center}
+\fbox{\includegraphics[width=\linewidth,height=2.5in,keepaspectratio]{original.png}}
+\end{center}
+\caption{Screenshot of the original WordPress REST API documentation}
+\label{fig:original}
+\end{figure*}
 
-Two pilot studies were performed to evaluate and understand an effective process for performing this study.
-The first pilot study involved four participants that were invited to join the first author on this paper in-person or using a video conferencing software.
+\begin{figure*}[tb]
+\begin{center}
+\fbox{\includegraphics[width=\linewidth,height=2.5in,keepaspectratio]{forked.png}}
+\end{center}
+\caption{Screenshot of the enhanced WordPress REST API documentation with a usage example}
+\label{fig:enhanced}
+\end{figure*}
+
+\subsubsection{Development Environment}
+
+We conducted two pilot studies to evaluate and understand an effective process for performing this study.
+The first pilot study involved four participants that were invited to join the first author in-person or using a video conferencing software.
 The study involved tasks using two APIs, the WordPress REST API V2 and the GMail REST API.
 Each participant was given one of the two APIs and a set of tasks to complete using the API within an hour.
-Participants were given an online answer form to record the answers to the tasks.
+We gave the participants an online answer form to record the answers to the tasks.
 The primary findings from this pilot are as follows: 1) asking participants to use an API to perform the tasks required significant overhead time for them to setup a development environment with the proper API credentials, 2) the intermediate trial attempts of using the API are potentially more valuable than the final answer as it allows us to understand API client developer information needs that may are not answered by the documentation, 3) the number of tasks for the study had to be reduced so the participants could complete the tasks within the one hour limit, and 4) for GMail API, participants used up a large portion of their time on setting up their API credentials that requires understanding of OAuth.
@@
 
-\begin{figure*}[t]
-\begin{center}
-\fbox{\includegraphics[width=\linewidth,height=2.5in,keepaspectratio]{original.png}}
-\end{center}
-\caption{Screenshot of the original WordPress REST API documentation}
-\label{fig:original}
-\end{figure*}
+\subsubsection{Protocol}
 
-\begin{figure*}[tb]
-\begin{center}
-\fbox{\includegraphics[width=\linewidth,height=2.5in,keepaspectratio]{forked.png}}
-\end{center}
-\caption{Screenshot of the forked WordPress REST API documentation with a usage example}
-\label{fig:forked}
-\end{figure*}
+Based on the lessons learned from the pilot studies, we designed the actual study process as follows: the original WordPress REST API documentation was enhanced and a total of 3 API usage examples were added to show listing of blog posts with query parameters for filtering, a request to create a blog post and a request to update a blog post.
+Figure \ref{fig:original} shows a screenshot of the original API documentation related to T4 where the API client developers are provided with a reference table describing the different properties that can be used to create a $Post$ object\footnote{\url{http://v2.wp-api.org/reference/posts/}}.
+Figure \ref{fig:enhanced} shows a screenshot of the enhanced API documentation with a cURL based usage example.
+cURL is used because it's used by elsewhere in the original API documentation.
+In the enhanced API documentation, the example shows one possible API call with realistic values for the data that is described in the reference table and associated API response headers and body.
 
-Based on the lessons learned from the pilot studies, we designed the actual study process as follows: the original WordPress REST API documentation was forked and a total of 3 API usage examples were added to show listing of blog posts with query parameters for filtering, a request to create a blog post and a request to update a blog post.
-Figure \ref{fig:original} shows a screenshot of the original API documentation related to T4 where the API client developers are provided with a reference table describing the different properties that can be used to create a $Post$ object\footnote{\url{http://v2.wp-api.org/reference/posts/}}.
-Figure \ref{fig:forked} shows a screenshot of the forked API documentation with a cURL based usage example.
-cURL is used because it's used by elsewhere in the original API documentation.
-In the forked API documentation, the example shows one possible API call with realistic values for the data that is described in the reference table and associated API response headers and body.
-
-Participants were divided into two groups, G1 and G2.
-G1 participants were provided with a link to the official API documentation on the web-based API explorer, and G2 participants were provided with a link to the forked API documentation with usage examples.
+We divided the participants into two groups, G1 and G2.
+G1 participants were provided with a link to the official API documentation on the web-based API explorer, and G2 participants were provided with a link to the enhanced API documentation with usage examples.
 The web-based API explorer allocated more participants to G1 compared to G2 because we wanted to better understand the impact of the lack of usage examples on API client developer productivity.
@@
 All participants were given the same set of API tasks and were requested to limit their participation time to a maximum of one hour.
-No task specific time limit was imposed because we wanted participants to spend sufficient time on each task without forcing them to move the next one.
-Participants were allowed to access the internet and external resources alongside the provided API documentation to perform the tasks as they'd normally use on a typical work day.
+No task specific time limit was imposed except an overall limit one hour for the entire study because we wanted participants to spend sufficient time on each task without forcing them to move the next one.
+Participants were allowed to access the internet and external resources alongside the provided API documentation to perform the tasks as they would normally use on a typical work day.
+We used the web-based API explorer to also collect an experience rating on a scale of 0-10, 10 being the best possible, of using the given REST API documentation and a free-form feedback from each participant as an exit survey.
 
@@
 Additionally, the demographic information, experience rating and the free-form feedback for each participant is also included in the artifact.
-
-\subsubsection{Data Analysis}
-The same data artifact is used to analyze both quantitatively and qualitatively to answer the aforementioned research question.We analyzed the data artifact qualitatively to understand the obstacles that REST API client developers face when using an API documentation that lacks usage examples.
-The data artifact exported from the web-based REST API explorer included a table with raw data.
-For each row, the table contains the following columns: API task, participant identifier, timestamp, API trial number, I1-4, response headers and body.
-It allowed us to observe and categorize the different input values that the participants used to perform the given API tasks.
-The values for I1-4 used by the participants were manually coded to group the participant responses into categories.
-The categories help us determine the types of information that API client developers need in the API documentation to perform API tasks successfully.
-The analysis started with an empty set of codes and new codes were introduced to describe scenarios that didn't fall under  codes that were already applied.
-The first author on this paper applied the codes on the raw data artifact and the co-authors verified the codes randomly.
-To resolve disagreements, the authors discussed and updated the codes accordingly.
-The codes were then further categorized to higher level grouping to represent common problems faced by the study participants for each group.
-
-For each API task attempted by each participant, we annotated the artifact with one of the following labels: successful, partially successful, and unsuccessful.
-Task participations are marked successful when I1-4 matches the required values for performing the given task.
-If a participant is able to use the correct I1-4 for one of the two tasks required to complete a single task (T2), we marked it as a partially successful.
-Otherwise, it's marked as unsuccessful.
-Based on these annotations, the answers for RQ1 was computed using the following:
-
-success rate of group = (total number of successful tasks by group) / (number of participants in group * number of API tasks)
-
-API calls per success of group = (number of API calls annotated as successful) / (total number of trial API calls made by group)
-
-average time taken for success = (number of successful API tasks by group) / (total time spent by all participants in the group)
-
-\section{Results}
-\label{sec:results}
 
@@
     \addplot[draw=black] table [col sep=comma, x expr=\coordindex, y=AverageAPICalls] {grouptwo.csv};
-    \legend{G1: Using Original, G2: Using Forked}
+    \legend{G1: Using Original, G2: Using Enhanced}
     \end{axis}
@@
     \addplot[smooth,mark=*,black] table [col sep=comma, x expr=\coordindex, y=SuccessRate] {grouptwo_success.csv};
-      \addlegendentry{G2: Using Forked}
+      \addlegendentry{G2: Using Enhanced}
     \end{axis}
@@
     \addplot[draw=black] table [col sep=comma, x expr=\coordindex, y=Time] {grouptwo_time.csv};
-    \legend{G1: Using Original, G2: Using Forked}
+    \legend{G1: Using Original, G2: Using Enhanced}
     \end{axis}
@@
 \end{figure*}
-\subsection{Qualitative Analysis Results} % (fold)
 
-\textbf{Observation 1.} REST API client developers want usage examples along with the definition of API elements.
-From analyzing the participant feedback, we've observed usage examples mentioned as the most frequent and central topic of interest.
-11 of the 16 G2 participants mentioned the lack of examples with HTTP headers, request and response as the primary problem with the API documentation.
-On the other hand, 7 out of the 10 G2 participants mentioned the examples being helpful from the forked documentation.
-This is also supported by our quantitative analysis as discussed later in this section.
+\subsubsection{Data Analysis}
+We analyzed the data artifact qualitatively to understand the obstacles that REST API client developers face when using an API documentation that lacks usage examples.
+The data artifact exported from the web-based REST API explorer included a table with raw data.
+For each row, the table contains the following columns: API task, participant identifier, timestamp, API trial number, I1-4, response headers and body.
+It allowed us to observe and categorize the different input values that the participants used to perform the given API tasks.
+The values for I1-4 used by the participants were manually coded to group the participant responses into categories.
+The categories help us determine the types of information that API client developers need in the API documentation to perform API tasks successfully.
+The analysis started with an empty set of codes and new codes were introduced to describe scenarios that did not fall under  codes that were already applied.
+The first author applied the codes on the raw data artifact and the co-authors verified the codes randomly.
+To resolve disagreements, the authors discussed and updated the codes accordingly.
+The codes were then further categorized to higher level grouping to represent common problems faced by the study participants for each group.
 
-\textbf{Implication 1.} REST API developers need to provide usage examples showing sample requests and responses with HTTP headers and body for API actions to help API usability.
+For each API task attempted by each participant, we annotated the artifact with one of the following labels: successful, partially successful, and unsuccessful.
+Task participations are marked successful when I1-4 matches the required values for performing the given task.
+If a participant is able to use the correct I1-4 for one of the two tasks required to complete a single task (T2), we marked it as a partially successful.
+Otherwise, it's marked as unsuccessful.
+Based on these annotations, the answers for RQ was computed using the following:
 
-\textbf{Observation 2.} API client developers face problem with using right data type without examples.
-To successfully complete T1 participants had to use an integer ID representing an $Author$ object given the author name.
-In the original API documentation, the following is mentioned about the author parameter: ``author: Limit result set to posts assigned to specific authors.'' The documentation doesn't mention that the data type required is the numeric author ID, not the author name.
-We found 10 of the 16 participants from G1 made API calls with the author name instead of the ID.
-P1.12 mentioned the following: \textbf{``...to find posts for the author, I had to inspect the response to see that indeed the $author\_id$ was in there''}.
-The forked documentation showed one example of using numeric ID.
-3 of the 10 participants from G2 ran into the same error.
-When faced with this error, participants used trial and error to eventually complete the task.
+success rate of group = (total number of successful tasks by group) / (number of participants in group * number of API tasks)
 
-\textbf{Implication 2.} REST API documentation needs to include examples of data types for each API field to satisfy API client developer information need.
+API calls per success of group = (number of API calls annotated as successful) / (total number of trial API calls made by group)
 
-\textbf{Observation 3.} API client developers face problem with using the right data format without examples.
-Tasks T3 required the use of a correct format to represent an array of numeric post IDs to be excluded using the API.
-The original documentation mentions the following about this API query parameter: ``exclude: Ensure result set excludes specific IDs.'' Participants attempted to solve this task using multiple different formats for specifying the array.
-For example: they tried with $exclude=[1,4]$, $exclude=1\&4$, $exclude=4\&\&exclude=1$, $id!=1\&id!=4$ and other alternatives before eventually finding the right format as follows $exclude=1,4$.
-Participant P1.6 mentioned the following: \textbf{``It was difficult to figure out whether some of the inputs needed to be arrays or just a comma-separated list.''} 7 out of the 16 participants in G1 faced this error.
-The forked documentation showed one example of using multiple IDs and all G2 participants could use the right format.
+average time taken for success = (number of successful API tasks by group) / (total time spent by all participants in the group)
 
-Task T5 also required the participants to use an ISO8601 formatted date for publishing a post with a specific date.
-Even though the original documentation mentions ISO8601 format, it doesn't provide an example.
-Instead of using ``2016-01-01T00:00:00-00:00'', participants from G1 used date formats as follows: ``January 1, 2016'', ``2016-01-01T12:00:00'', ``20160101'', etc.
-P1.8 mentioned the following: \textbf{``Got hung up trying to figure out the formatting of the date (which doesn't appear to be ISO8601, despite what it says)''}.
-After analyzing the response, we found that P1.8 used the ISO8601 formatted date in the API calls, but didn't provide the $second$ portion of the time as required by the API.
-8 of the 16 participants in G1 had an error in the date formats.
-With an example, 1 of the 10 participants had the same error from G2.
-
-\textbf{Implication 3.} REST API developers need to include examples showing the valid data format for the API elements.
-
-\textbf{Observation 4.} REST API client developers face problem with using request headers without usage examples.
-Tasks T5 and T6 required the use of a HTTP request header named $Content-Type$.
-In the original API documentation, the following was mentioned: ``The API uses JSON exclusively as the request and response format, including error responses.''.
-Experienced software engineers still had problem using this HTTP header as we found from the participants in G1.
-P1.12 mentioned the following: \textbf{``I didn't know how to specify application/json so that I could send POST data via the body instead in the URL parameters...
-Examples would really help - with a sample request and a sample response.''} We were surprised to see that 14 of the 15 participants that attempted T5 and T6 from G1 had the error of not specifying the $Content-Type$ header even though they have years of experience with REST APIs.
-In contrast, 2 out of the 10 participants from G2 failed used this header correctly as indicated in the example in the forked documentation.
-Moreover, 1 of the 2 G2 participants that didn't use the header first time, later used it correctly on a subsequent attempt.
-
-\textbf{Implication 4.} If API requests need to use HTTP Request headers, in addition to the request method and body, REST API develpers need to include examples of the HTTP headers.
-
-\textbf{Observation 5.} API client developers face problem with performing API tasks that require multiple API calls.
-Tasks T2 and T6 required the use of multiple API calls to complete successfully.
-Both tasks require an initial API call to find data using the API that is needed make a second API call.
-For example, P1.5 mentioned the following about T6, \textbf{``The post I created is not visible and not editable, though I got 201.
-Not sure why''}.
-After analyzing the response, we observed that P1.5 was able to create a blog post, but failed to publish it because it didn't have the $publish$ status.
-As a result, the post was not returned via the API for T6.
-P2.2 mentioned the following about T2: \textbf{``In case of search posts by author, it's not clear how to find the author\_id.
-Although, I realized that users and authors are same in this case.''}
-
-\textbf{Implication 5.} If there are prerequisites for making an API call, REST API developers need to provide examples showing how to get the prerequisites in the API documentation.
-
-To summarize, we recommend API developers to consider usage examples as an essential requirement for REST API documentation.
-As such, sufficient resources and a high priority should be devoted to generate and maintain the usage examples.
-Based on our observations, we recommend REST API documentation tools to provide first-class support for including usage examples with realistic data.
+\section{Results}
+\label{sec:results}
 
@@
 
+\subsection{Qualitative Analysis Results} % (fold)
+
+\textbf{Observation 1.} REST API client developers want usage examples along with the definition of API elements.
+From analyzing the participant feedback, we've observed usage examples mentioned as the most frequent and central topic of interest.
+11 of the 16 G2 participants mentioned the lack of examples with HTTP headers, request and response as the primary problem with the API documentation.
+On the other hand, 7 out of the 10 G2 participants mentioned the examples being helpful from the enhanced documentation.
+This is also supported by our quantitative analysis as discussed later in this section.
+
+\textbf{Implication 1.} REST API developers need to provide usage examples showing sample requests and responses with HTTP headers and body for API actions to help API usability.
+
+\textbf{Observation 2.} API client developers face problem with using right data type without examples.
+To successfully complete T1 participants had to use an integer ID representing an $Author$ object given the author name.
+In the original API documentation, the following is mentioned about the author parameter: ``author: Limit result set to posts assigned to specific authors.'' The documentation doesn't mention that the data type required is the numeric author ID, not the author name.
+We found 10 of the 16 participants from G1 made API calls with the author name instead of the ID.
+P1.12 mentioned the following: \textbf{``...to find posts for the author, I had to inspect the response to see that indeed the $author\_id$ was in there''}.
+The enhanced documentation showed one example of using numeric ID.
+3 of the 10 participants from G2 ran into the same error.
+When faced with this error, participants used trial and error to eventually complete the task.
+
+\textbf{Implication 2.} REST API documentation needs to include examples of data types for each API field to satisfy API client developer information need.
+
+\textbf{Observation 3.} API client developers face problem with using the right data format without examples.
+Tasks T3 required the use of a correct format to represent an array of numeric post IDs to be excluded using the API.
+The original documentation mentions the following about this API query parameter: ``exclude: Ensure result set excludes specific IDs.'' Participants attempted to solve this task using multiple different formats for specifying the array.
+For example: they tried with $exclude=[1,4]$, $exclude=1\&4$, $exclude=4\&\&exclude=1$, $id!=1\&id!=4$ and other alternatives before eventually finding the right format as follows $exclude=1,4$.
+Participant P1.6 mentioned the following: \textbf{``It was difficult to figure out whether some of the inputs needed to be arrays or just a comma-separated list.''} 7 out of the 16 participants in G1 faced this error.
+The enhanced documentation showed one example of using multiple IDs and all G2 participants could use the right format.
+
+Task T5 also required the participants to use an ISO8601 formatted date for publishing a post with a specific date.
+Even though the original documentation mentions ISO8601 format, it doesn't provide an example.
+Instead of using ``2016-01-01T00:00:00-00:00'', participants from G1 used date formats as follows: ``January 1, 2016'', ``2016-01-01T12:00:00'', ``20160101'', etc.
+P1.8 mentioned the following: \textbf{``Got hung up trying to figure out the formatting of the date (which doesn't appear to be ISO8601, despite what it says)''}.
+After analyzing the response, we found that P1.8 used the ISO8601 formatted date in the API calls, but didn't provide the $second$ portion of the time as required by the API.
+8 of the 16 participants in G1 had an error in the date formats.
+With an example, 1 of the 10 participants had the same error from G2.
+
+\textbf{Implication 3.} REST API developers need to include examples showing the valid data format for the API elements.
+
+\textbf{Observation 4.} REST API client developers face problem with using request headers without usage examples.
+Tasks T5 and T6 required the use of a HTTP request header named $Content-Type$.
+In the original API documentation, the following was mentioned: ``The API uses JSON exclusively as the request and response format, including error responses.''.
+Experienced software engineers still had problem using this HTTP header as we found from the participants in G1.
+P1.12 mentioned the following: \textbf{``I didn't know how to specify application/json so that I could send POST data via the body instead in the URL parameters...
+Examples would really help - with a sample request and a sample response.''} We were surprised to see that 14 of the 15 participants that attempted T5 and T6 from G1 had the error of not specifying the $Content-Type$ header even though they have years of experience with REST APIs.
+In contrast, 2 out of the 10 participants from G2 failed used this header correctly as indicated in the example in the enhanced documentation.
+Moreover, 1 of the 2 G2 participants that didn't use the header first time, later used it correctly on a subsequent attempt.
+
+\textbf{Implication 4.} If API requests need to use HTTP Request headers, in addition to the request method and body, REST API develpers need to include examples of the HTTP headers.
+
+\textbf{Observation 5.} API client developers face problem with performing API tasks that require multiple API calls.
+Tasks T2 and T6 required the use of multiple API calls to complete successfully.
+Both tasks require an initial API call to find data using the API that is needed make a second API call.
+For example, P1.5 mentioned the following about T6, \textbf{``The post I created is not visible and not editable, though I got 201.
+Not sure why''}.
+After analyzing the response, we observed that P1.5 was able to create a blog post, but failed to publish it because it didn't have the $publish$ status.
+As a result, the post was not returned via the API for T6.
+P2.2 mentioned the following about T2: \textbf{``In case of search posts by author, it's not clear how to find the author\_id.
+Although, I realized that users and authors are same in this case.''}
+
+\textbf{Implication 5.} If there are prerequisites for making an API call, REST API developers need to provide examples showing how to get the prerequisites in the API documentation.
+
+To summarize, we recommend API developers to consider usage examples as an essential requirement for REST API documentation.
+As such, sufficient resources and a high priority should be devoted to generate and maintain the usage examples.
+Based on our observations, we recommend REST API documentation tools to provide first-class support for including usage examples with realistic data.
+
 \section{Threats to Validity}
@@
 used a mixed approach of surveys and in-person interviews with professional software engineers to understand the API learning obstacles \cite{Robillard_what_makes} \cite{Robillard_a_field_study}.
-They found that API documentation was related to one of the most severe obstacles for API learnabilty and recommended considering five factors for API documentation as follows: documentation of intent, code examples, matching APIs with scenarios, penetrability of the API, and format and presentation.
-Gias et al.
-surveyed 323 professional software engineers to understand the factors that fail a API documentation \cite{g_uddin}.
+They found that many obstacles for API learnabilty were related to API documentation and five impactful factors for API documentation as follows: documentation of intent, code examples, matching APIs with scenarios, penetrability of the API, and format and presentation.
+Uddin and Robillard surveyed 323 professional software engineers to understand the different ways in which a piece of documentation can be unfit for purpose \cite{g_uddin}.
 They identified ambiguity, incompleteness, and incorrectness as the three severest problems that lead API documentations to fail to answer the information needs of developers.
-Duala-Ekoko et al.
-performed a study with twenty programmers to understand the types of questions that developers have while using an unfamiliar API \cite{Duala-Ekoko:2012:AAQ:2337223.2337255}.
+Duala-Ekoko and Robillard performed a study with twenty programmers to understand the types of questions that developers have while using an unfamiliar API \cite{Duala-Ekoko:2012:AAQ:2337223.2337255}.
 They categorized the API related information needs into twenty generic questions that can be used to analyze the quality of API documentation.
@@
 
-  In this paper, we focused on the usage examples part of API documentation to further understand the obstacles that API client developers face when the documentation contains relevant information about the API, it's intent, description of the API actions and elements, but lacks code examples.
+Our work fits in the general space of studies of API documentation obstacles by focusing specifically on problems related to the absence of examples, and by expanding into the study of REST APIs.
 
 \subsubsection{Crowd-Sourced API Usage Examples}
-Several existing research has focused on sourcing API usage examples from the Internet.
+Prior research investigated the sourcing API usage examples from the Internet.
 Wang et al.
@@
 They searched the web for usage examples related to the API of five popular Java libraries and found that on an average API examples could be found for 77\% of the 4,637 APIs included in the selected libraries.
-Moreover, they found that the crowd-sourced API documented accounted for 93.7\% of the usage examples compared to 6.13\% that were published on the official documentation.
-Incorporating a corpus of crowd-sourced API examples from the web and using placeholders for commonly related API elements, an evaluation of Jadite found that developers were three times faster to complete API tasks \cite{5295283}.
+Moreover, they found that the crowd-sourcing sites accounted for 93.7\% of the usage examples compared to 6.13\% that were published on the official documentation.
+An evaluation of Jadite showed that developers were three times faster to complete API tasks when given placeholder API elements in combination with usage examples sourced from the web \cite{5295283}.
 Nasehi et al.
@@
 Treude et al.
-proposed a machine learning based approach to find relevant API documentation from StackOverflow using both the text and metadata found from the questions and answers \cite{Treude:2016:AAD:2884781.2884800}.
+proposed a machine learning based approach to find relevant API documentation from StackOverflow using both the text and metadata found from the questions and answers~\cite{Treude:2016:AAD:2884781.2884800}.
 Kim et al.
@@
 Scheller et al.
-presented a framework for automatically measuring the complexity of an API \cite{scheller2015automated}.
+presented a framework for automatically measuring the complexity of an API~\cite{scheller2015automated}.
 They identified a list of measurable API properties and provided formulas to compute complexities related to the interfaces, implementation and setup of an API.
@@
 To improve API usability, Farooq et al.
-proposed using peer reviews of API code in addition to API usability studies to uncover API related bugs and incorporate feedback regarding the API elements \cite{Farooq:2010:AUP:1753326.1753677}.
+proposed using peer reviews of API code in addition to API usability studies to uncover API related bugs and incorporate feedback regarding the API elements~\cite{Farooq:2010:AUP:1753326.1753677}.
 In our work, instead of measuring overall API usability, we have focused specifically on the relationship of REST API usability with respect to usage examples.
@@
 \subsubsection{Controlled Studies}
-Several authors published the results of controlled studies on the topic of API usability.
+Several researchers published the results of controlled studies on the topic of API usability.
 Nasehi et al.
@@
 \section*{Acknowledgment}
+We are grateful to our participants for their valuable time on this study. We thank the University of Calgary for funding this research.
 

### 2017-03-19 controlled_study.tex
lower cased task names in proper English

@@
 
-\textbf{T1: List All Posts Task}.
+\textbf{T1: List all posts task}.
 We ask the participants to use the WordPress REST API to get a list of the blog posts from a given WordPress site.
@@
 
-\textbf{T2: FilterPostsByAuthorTask.} The participants are asked to use the API to filter the list of posts obtained in T1 by an author given the author's user name.
+\textbf{T2: Filter posts by author task.} The participants are asked to use the API to filter the list of posts obtained in T1 by an author given the author's user name.
 To answer this correctly, the participants are required to first make an API call to get the numeric ID of the author given the string based user name.
@@
 
-\textbf{T3: ExcludePostsByIdsTask.} We ask the participants to use the API to get a list of all posts excluding posts with IDs 1 and 4.
+\textbf{T3: Exclude posts by ids task.} We ask the participants to use the API to get a list of all posts excluding posts with IDs 1 and 4.
 Participants need to use the inputs I1-2, and use a desired format on I2 to pass an array of IDs as a parameter.
@@
 
-\textbf{T4: FindTotalPostsTasks.} This task requires the participants to use the API to find a total number of posts.
+\textbf{T4: Find total posts tasks.} This task requires the participants to use the API to find a total number of posts.
 Participants need to use the inputs I1-2 and inspect the HTTP response headers to successfully complete this task.
@@
 
-\textbf{T5: PublishPostTask.} We ask the participants to use the API to publish a blog post with a specific title, content, and a published date.
+\textbf{T5: Publish post task.} We ask the participants to use the API to publish a blog post with a specific title, content, and a published date.
 To successfully complete this task, the participants are required to use all four input types and inspect both the HTTP response header and the response body.
@@
 
-\textbf{T6: UpdatePostTask.} We ask the participants to use the API to update a blog post that they published in T5 with a new excerpt.
+\textbf{T6: Update post task.} We ask the participants to use the API to update a blog post that they published in T5 with a new excerpt.
 Similar to T5, this task requires the use of inputs I1-4, but with different values for the inputs.
@@
 Rama et al.
-presented a set of formulas for computing a measure of API usability based on the API's structural components such as classes, methods, parameters, return values, thread-safety, etc \cite{rama}.
+presented a set of formulas for computing a measure of API usability based on the API's structural components such as classes, methods, parameters, return values, thread-safety, etc~\cite{rama}.
 Scheller et al.

### 2017-03-19 controlled_study.tex
fix typo
add preamble for related work

@@
 
-\textbf{T4: Find total posts tasks.} This task requires the participants to use the API to find a total number of posts.
+\textbf{T4: Find total posts task.} This task requires the participants to use the API to find a total number of posts.
 Participants need to use the inputs I1-2 and inspect the HTTP response headers to successfully complete this task.
@@
 \label{sec:related_work}
+
+In this section, we discuss the existing research on the relationship between API usability and usage examples.
+To this regard, we summarize the related work on APIs from the sub-areas of API learning obstacles, using crowd-sourced API examples, measuring API usability, and controlled studies on the impact of usage examples on API usability.
+
 \subsubsection{API Learning Obstacles}

### 2017-03-19 controlled_study.tex
save space from screenshots

@@
 
-\begin{figure*}[t]
-\begin{center}
-\fbox{\includegraphics[width=\linewidth,height=2.5in,keepaspectratio]{original.png}}
-\end{center}
-\caption{Screenshot of the original WordPress REST API documentation}
+\begin{figure*}[htp]
+\centering
+
+\begin{subfigure}[t]{4in}
+\centering
+\includegraphics[width=4in]{original.png}
+\caption{Original API Documentation}
 \label{fig:original}
-\end{figure*}
-
-\begin{figure*}[tb]
-\begin{center}
-\fbox{\includegraphics[width=\linewidth,height=2.5in,keepaspectratio]{forked.png}}
-\end{center}
-\caption{Screenshot of the enhanced WordPress REST API documentation with a usage example}
+\end{subfigure}
+~
+\begin{subfigure}[t]{3in}
+\centering
+\includegraphics[width=2.7in]{forked.png}
+\caption{Usage Example from the Enhanced API Documentation}
 \label{fig:enhanced}
+\end{subfigure}
+\caption{Screenshots of Original vs. Enhanced API Documentation}
 \end{figure*}

### 2017-03-20 controlled_study.tex
better wording
caption
wording
low fat diet for related work section

@@
 We selected the WordPress REST API V2 for this study.
-WordPress is a blog-like open-source (R2) framework used by over 409 million people to visit 23.6 billion pages each month.
+WordPress is a blog-like open-source (R2) framework used by over 409 million people to visit 23.6 billion pages each month \footnote{\url{https://wordpress.com/activity/}}.
 The API allows programmatic access to list, create, update, and delete WordPress data such as blog posts, comments, users, images, tags, etc.
@@
 Before January 2017, the WordPress REST API V2 was distributed as a plug-in where WordPress users could optionally install the API component.
-The following statistics are for the plug-in installation numbers between May 2015 and October 2016:
+The following statistics are for the plug-in installation numbers between May 2015 and October 2016:\footnote{\url{https://wordpress.org/plugins/rest-api/stats/}}
 
@@
 \includegraphics[width=2.7in]{forked.png}
-\caption{Usage Example from the Enhanced API Documentation}
+\caption{API Documentation Enhanced with an Example}
 \label{fig:enhanced}
@@
 
-Based on the lessons learned from the pilot studies, we designed the actual study process as follows: the original WordPress REST API documentation was enhanced and a total of 3 API usage examples were added to show listing of blog posts with query parameters for filtering, a request to create a blog post and a request to update a blog post.
+We designed the actual study protocol as follows: we enhanced the original WordPress REST API documentation and added a total of 3 API usage examples to show listing of blog posts with query parameters for filtering, a request to create a blog post and a request to update a blog post.
+We used the WordPress API unit tests to find relevant data for these examples.
 Figure \ref{fig:original} shows a screenshot of the original API documentation related to T4 where the API client developers are provided with a reference table describing the different properties that can be used to create a $Post$ object\footnote{\url{http://v2.wp-api.org/reference/posts/}}.
@@
 \subsubsection{API Learning Obstacles}
-  Researchers have published several papers on the topic of API learning obstacles.
 Robillard et al.
 used a mixed approach of surveys and in-person interviews with professional software engineers to understand the API learning obstacles \cite{Robillard_what_makes} \cite{Robillard_a_field_study}.
-They found that many obstacles for API learnabilty were related to API documentation and five impactful factors for API documentation as follows: documentation of intent, code examples, matching APIs with scenarios, penetrability of the API, and format and presentation.
+They found that many obstacles for API learnabilty were related to API documentation.
+They identified five impactful factors for API documentation as follows: documentation of intent, code examples, matching APIs with scenarios, penetrability of the API, and format and presentation.
 Uddin and Robillard surveyed 323 professional software engineers to understand the different ways in which a piece of documentation can be unfit for purpose \cite{g_uddin}.
 They identified ambiguity, incompleteness, and incorrectness as the three severest problems that lead API documentations to fail to answer the information needs of developers.
-Duala-Ekoko and Robillard performed a study with twenty programmers to understand the types of questions that developers have while using an unfamiliar API \cite{Duala-Ekoko:2012:AAQ:2337223.2337255}.
+Duala-Ekoko and Robillard performed a study with 20 programmers to understand the types of questions that developers have while using an unfamiliar API \cite{Duala-Ekoko:2012:AAQ:2337223.2337255}.
 They categorized the API related information needs into twenty generic questions that can be used to analyze the quality of API documentation.
 Myers et al.
-performed a study of enterprise SOA API usability and found that the study participants had limited success finding the relevant API elements to perform an API task \cite{Myers_study}.
-Among other factors, they recommended providing code examples in the API documentation to help API client developers with sample solutions to different patterns of API tasks.
-
+found that participants had limited success finding the relevant API elements to perform an API task using an enterprise API \cite{Myers_study}.
+Among other factors, they recommended providing code examples in the API documentation to help API client developers.
 Our work fits in the general space of studies of API documentation obstacles by focusing specifically on problems related to the absence of examples, and by expanding into the study of REST APIs.
@@
 \subsubsection{Crowd-Sourced API Usage Examples}
-Prior research investigated the sourcing API usage examples from the Internet.
 Wang et al.
-performed an exploratory study to understand the current state of API related knowledge available on the Internet that can be leveraged by API client developers \cite{6462686}.
+performed an exploratory study to understand the current state of API related knowledge available on the Internet \cite{6462686}.
 They searched the web for usage examples related to the API of five popular Java libraries and found that on an average API examples could be found for 77\% of the 4,637 APIs included in the selected libraries.
 Moreover, they found that the crowd-sourcing sites accounted for 93.7\% of the usage examples compared to 6.13\% that were published on the official documentation.
-An evaluation of Jadite showed that developers were three times faster to complete API tasks when given placeholder API elements in combination with usage examples sourced from the web \cite{5295283}.
+An evaluation of Jadite showed that developers were three times faster to complete API tasks with access to auto-referred crowd-sourced usage examples~\cite{5295283}.
 Nasehi et al.
@@
 They found that commonly down-voted API related answers lacked code examples, explanations, and shortcoming of solutions.
-They recommended API developers to evolve the documentation with usage examples to answer API client developer questions that were not anticipated in the existing documentation.
 Treude et al.
@@
 observed a severe inequality within the context of crowd-sourced API usage examples where most of the content were related to the popular APIs \cite{jiau}.
-
-Existing research on crowd-sourced API documentation has mostly focused finding crowd-sourced usage examples of local APIs.
-In this paper, we focused on the impact of usage examples for REST APIs, where the API client development programming language is agnostic to the language used to implement the API.
-This important distinction with local APIs makes it hard to leverage the techniques proposed in the aforementioned papers to find relevant REST API usage examples.
+Existing research on crowd-sourced API documentation has mostly focused finding usage examples of local APIs.
+In this paper, we focused on REST APIs, where the API client development programming language is agnostic to API implementation language.
+This distinction with local APIs makes it hard to leverage the techniques proposed in the aforementioned papers for finding crowd-sourced REST API usage examples.
 
 \subsubsection{Measuring API Usability}
-Several papers have been published on the topic of measuring API usability.
 Rama et al.
@@
 Grill et al.
-proposed an HCI based approach that can be used by API developers to understand and improve on the problem areas related to an API \cite{Grill2012}.
-They suggested using a combination of expert opinion and developer workshops on APIs to identify and collect feedback about API and its documentation related problems.
-To improve API usability, Farooq et al.
-proposed using peer reviews of API code in addition to API usability studies to uncover API related bugs and incorporate feedback regarding the API elements~\cite{Farooq:2010:AUP:1753326.1753677}.
+suggested using a combination of expert opinion and developer workshops to identify and collect feedback about API and its documentation related problems \cite{Grill2012}.
+Farooq et al.
+recommended using peer reviews of API code in addition to API usability studies to uncover API related bugs and incorporate feedback~\cite{Farooq:2010:AUP:1753326.1753677}.
 In our work, instead of measuring overall API usability, we have focused specifically on the relationship of REST API usability with respect to usage examples.
@@
 \subsubsection{Controlled Studies}
-Several researchers published the results of controlled studies on the topic of API usability.
 Nasehi et al.
 performed a controlled study to understand if API unit tests can be used to provide as usage examples to facilitate API client developers \cite{5609553}.
-They grouped participants into two groups, and one of the groups was provided with API unit tests in addition to the documentation.
-The two groups performed the same set of API tasks.
 The researchers found that the examples from the unit tests helped understanding the API concepts better but it was challenging for the participants to locate relevant examples from the corpus of unit tests.
-They recommended automated extraction of relevant high-level API usage examples from API unit tests.
 Endrikat performed a controlled study with four groups of participants to understand the impact of API documentation on APIs that are implemented using programming languages with static and dynamic type systems \cite{Endrikat:2014:ADS:2568225.2568299}.
 Participants were given a set of failing unit tests and were asked to make them pass by writing code using the studied API.
-They defined the API client developer productivity in terms of the time to get the tests passing.
-They found that the participant group using a statically typed API with explicit documentation were more productive than the other groups.
+They found that the participant group using a statically typed API with explicit documentation could get the tests to pass in less time than the other groups.
 Ellis et al.
-performed a controlled study with two groups of Java developers to understand the impact of using a constructor vs.
+performed a controlled study with two groups of developers to understand the impact of using a constructor vs.
 a factory method design pattern on API client development time \cite{Ellis:2007:FPA:1248820.1248863}.
-They found that API client developers needed more time to complete API tasks when the API requires the use a factory method to instantiate objects compared to using a constructor.
-
+They found that API client developers needed more time to complete API tasks using factory method compared to using a constructor.
 Our work in this paper is based on a controlled study and shares part of the setup that were used by the aforementioned API related controlled studies with the following differences: our goal is to understand the obstacles faced by API client developers without usage examples;  we focus on REST APIs instead of a local API; and, the participants are professional software engineers.

### 2017-03-20 controlled_study.tex
down to 8 pages as per guideline

@@
 Participants were divided into two groups and given the same set of API tasks to complete.
-One group was given the official WordPress REST API documentation and another group was given a enhanced version of the documentation where three usage examples were added.
+One group was given the official WordPress REST API documentation and another group was given an enhanced version of the documentation where three usage examples were added.
 Using a novel technique, we collected 539 API calls made by the participants.
@@
 These properties satisfy R1, our requirements for using a representative API.
-
 By selecting an open-source project we are able to access the source-code to inspect the implementation and documentation technique of the WordPress REST API.
@@
 
-In addition to the auto-generated documentation of the the schema and API actions, custom content is added by the contributors to the API documentation project to provide prosaic overviews and usage examples.
+In addition to the auto-generated documentation of the the schema and API actions, contributors add custom content to provide prosaic overviews and usage examples.
 
@@
 We conducted two pilot studies to evaluate and understand an effective process for performing this study.
-The first pilot study involved four participants that were invited to join the first author in-person or using a video conferencing software.
+The first pilot study involved four participants that were invited to join the first author in-person or over video conferencing.
 The study involved tasks using two APIs, the WordPress REST API V2 and the GMail REST API.
-Each participant was given one of the two APIs and a set of tasks to complete using the API within an hour.
-We gave the participants an online answer form to record the answers to the tasks.
-The primary findings from this pilot are as follows: 1) asking participants to use an API to perform the tasks required significant overhead time for them to setup a development environment with the proper API credentials, 2) the intermediate trial attempts of using the API are potentially more valuable than the final answer as it allows us to understand API client developer information needs that may are not answered by the documentation, 3) the number of tasks for the study had to be reduced so the participants could complete the tasks within the one hour limit, and 4) for GMail API, participants used up a large portion of their time on setting up their API credentials that requires understanding of OAuth.
+Each participant was given one of the two APIs and a set of tasks and online answer forms to complete using the API within an hour.
+The primary findings from this pilot are as follows: 1) participants required significant overhead time to setup a development environment with the proper API credentials, 2) the intermediate trial attempts of using the API are potentially more valuable than the final answer as it allows us to study what obstacles the participants face, 3) the number of tasks for the study had to be reduced to fit within the one hour limit, and 4) for GMail API, participants used up a large portion of their time on learning about how to use  OAuth.
 
@@
 The web-base REST API explorer also allowed us to automatically collect all the trial API calls that the participants make for each API task.
-A second pilot study involving seven new participants was performed to understand the features of the web-based REST API explorer in practice and to improve the user interface based on feedback from the participants.
-Participants completed the study on their own without having to meet in-person or over video conferencing.
-Only the WordPress REST API was used to focus on REST APIs without the required learning curve associated with OAuth.
-We found encouraging results from this pilot study as the collected data showed patterns of mistakes that API client developers make that can be reduced by adding usage examples in the API documentation.
+A second pilot study involving seven new participants was performed to collect usability related feedback about the web-based REST API explorer.
+We only used WordPress since it didn't require knowledge about OAuth.
+From this study, we observed patterns of mistakes that API client developers make that can be reduced by adding usage examples in the API documentation.
 
@@
 We analyzed the data artifact qualitatively to understand the obstacles that REST API client developers face when using an API documentation that lacks usage examples.
-The data artifact exported from the web-based REST API explorer included a table with raw data.
+We exported a table of data from the database behind web-based REST API explorer.
 For each row, the table contains the following columns: API task, participant identifier, timestamp, API trial number, I1-4, response headers and body.
@@
 In Figure \ref{graph:productivity} and  \ref{graph:time} we see that, for the simpler tasks T2 and T3, G1 had used more trial API calls and more time on average and reached successful answers with a lower percentage compared to G2.
-For T4, where the participants are required to inspect a HTTP response header, G2 needed more API calls on average, but also yielded 24\% higher success rate.
+For T4, where the participants are required to inspect an HTTP response header, G2 needed more API calls on average, but also yielded 24\% higher success rate.
 For the more complicated tasks T5 and T6, we see that participants in G1 had very low success rate and very high number of average API trials compared to G2.
@@
 From analyzing the participant feedback, we've observed usage examples mentioned as the most frequent and central topic of interest.
-11 of the 16 G2 participants mentioned the lack of examples with HTTP headers, request and response as the primary problem with the API documentation.
+11 of the 16 G1 participants mentioned the lack of examples with HTTP headers, request and response as the primary problem with the API documentation.
 On the other hand, 7 out of the 10 G2 participants mentioned the examples being helpful from the enhanced documentation.
@@
 recommended using peer reviews of API code in addition to API usability studies to uncover API related bugs and incorporate feedback~\cite{Farooq:2010:AUP:1753326.1753677}.
-In our work, instead of measuring overall API usability, we have focused specifically on the relationship of REST API usability with respect to usage examples.
+In our work, instead of measuring API usability, we have focused specifically on the relationship of REST API usability with respect to usage examples.
 
@@
 performed a controlled study to understand if API unit tests can be used to provide as usage examples to facilitate API client developers \cite{5609553}.
-The researchers found that the examples from the unit tests helped understanding the API concepts better but it was challenging for the participants to locate relevant examples from the corpus of unit tests.
+The researchers found that the examples from the unit tests helped understanding the API concepts better but it was challenging for the participants to locate relevant examples.
 Endrikat performed a controlled study with four groups of participants to understand the impact of API documentation on APIs that are implemented using programming languages with static and dynamic type systems \cite{Endrikat:2014:ADS:2568225.2568299}.
@@
 They found that API client developers needed more time to complete API tasks using factory method compared to using a constructor.
-Our work in this paper is based on a controlled study and shares part of the setup that were used by the aforementioned API related controlled studies with the following differences: our goal is to understand the obstacles faced by API client developers without usage examples;  we focus on REST APIs instead of a local API; and, the participants are professional software engineers.
+Our work in this paper is based on a controlled study and shares part of the setup that were used by the aforementioned controlled studies with the following differences: our goal is to understand the obstacles faced by API client developers without usage examples;  we focus on REST APIs instead of a local API; and, the participants are professional software engineers.
 

### 2017-03-20 controlled_study.tex
refactor the qualitative observations to be more specific

@@
 
-\textbf{Observation 1.} REST API client developers want usage examples along with the definition of API elements.
-From analyzing the participant feedback, we've observed usage examples mentioned as the most frequent and central topic of interest.
 11 of the 16 G1 participants mentioned the lack of examples with HTTP headers, request and response as the primary problem with the API documentation.
 On the other hand, 7 out of the 10 G2 participants mentioned the examples being helpful from the enhanced documentation.
-This is also supported by our quantitative analysis as discussed later in this section.
+Based on these feedback, we identify usage examples as the most frequent and central topic of interest from our participants.
+This is also supported by our quantitative analysis results as discussed earlier in this section.
+REST API developers need to provide usage examples showing sample requests and responses with HTTP headers and body for API actions to help API usability.
+In the remainder of this section, we provide our observations related to each task and their implications.
 
-\textbf{Implication 1.} REST API developers need to provide usage examples showing sample requests and responses with HTTP headers and body for API actions to help API usability.
-
-\textbf{Observation 2.} API client developers face problem with using right data type without examples.
-To successfully complete T1 participants had to use an integer ID representing an $Author$ object given the author name.
-In the original API documentation, the following is mentioned about the author parameter: ``author: Limit result set to posts assigned to specific authors.'' The documentation doesn't mention that the data type required is the numeric author ID, not the author name.
-We found 10 of the 16 participants from G1 made API calls with the author name instead of the ID.
+\textbf{Observation 1.} Without usage examples, we observed 10 of the 16 G1 participants used incorrect data type of $String$ instead of $Integer$.
+To successfully complete T1, participants had to use an integer ID representing an $Author$ object given the author name.
+In the original API documentation, the following is mentioned about the author parameter: \textit{``author: Limit result set to posts assigned to specific authors.''} The documentation doesn't mention that the data type required is the numeric author ID, not the author name.
 P1.12 mentioned the following: \textbf{``...to find posts for the author, I had to inspect the response to see that indeed the $author\_id$ was in there''}.
@@
 
-\textbf{Implication 2.} REST API documentation needs to include examples of data types for each API field to satisfy API client developer information need.
+\textbf{Implication 1.} REST API documentation needs to include examples of data types such as Integer, String, and Array for each API field to satisfy API client developer information need.
 
-\textbf{Observation 3.} API client developers face problem with using the right data format without examples.
+\textbf{Observation 2.} We observed participants in G1 faced obstacles using correct data formats for Date and Array type data in API requests without usage examples.
 Tasks T3 required the use of a correct format to represent an array of numeric post IDs to be excluded using the API.
-The original documentation mentions the following about this API query parameter: ``exclude: Ensure result set excludes specific IDs.'' Participants attempted to solve this task using multiple different formats for specifying the array.
+The original documentation mentions the following about this API query parameter: \textit{``exclude: Ensure result set excludes specific IDs.''} Participants attempted to solve this task using multiple different formats for specifying the array.
 For example: they tried with $exclude=[1,4]$, $exclude=1\&4$, $exclude=4\&\&exclude=1$, $id!=1\&id!=4$ and other alternatives before eventually finding the right format as follows $exclude=1,4$.
-Participant P1.6 mentioned the following: \textbf{``It was difficult to figure out whether some of the inputs needed to be arrays or just a comma-separated list.''} 7 out of the 16 participants in G1 faced this error.
+Participant P1.6 mentioned the following: \textbf{``It was difficult to figure out whether some of the inputs needed to be arrays or just a comma-separated list.''}.
+7 out of the 16 participants in G1 faced this error.
 The enhanced documentation showed one example of using multiple IDs and all G2 participants could use the right format.
@@
 
-\textbf{Implication 3.} REST API developers need to include examples showing the valid data format for the API elements.
+\textbf{Implication 2.} REST API developers need to include examples showing the valid data format for the API elements.
 
-\textbf{Observation 4.} REST API client developers face problem with using request headers without usage examples.
+\textbf{Observation 3.} We were surprised to see that 14 of the 15 G1 participants that tried T5 and T6 faced problem using correct request headers without usage examples.
 Tasks T5 and T6 required the use of a HTTP request header named $Content-Type$.
-In the original API documentation, the following was mentioned: ``The API uses JSON exclusively as the request and response format, including error responses.''.
+In the original API documentation, the following was mentioned: \textit{``The API uses JSON exclusively as the request and response format, including error responses.''}.
 Experienced software engineers still had problem using this HTTP header as we found from the participants in G1.
 P1.12 mentioned the following: \textbf{``I didn't know how to specify application/json so that I could send POST data via the body instead in the URL parameters...
-Examples would really help - with a sample request and a sample response.''} We were surprised to see that 14 of the 15 participants that attempted T5 and T6 from G1 had the error of not specifying the $Content-Type$ header even though they have years of experience with REST APIs.
+Examples would really help - with a sample request and a sample response.''}  Even though G1 participants have years of experience with REST APIs, only one of them could use the $Content-Type$ header correctly without example.
 In contrast, 2 out of the 10 participants from G2 failed used this header correctly as indicated in the example in the enhanced documentation.
@@
 
-\textbf{Implication 4.} If API requests need to use HTTP Request headers, in addition to the request method and body, REST API develpers need to include examples of the HTTP headers.
+\textbf{Implication 3.} If API requests need to use HTTP Request headers, in addition to the request method and body, REST API develpers need to include examples of the HTTP headers.
 
-\textbf{Observation 5.} API client developers face problem with performing API tasks that require multiple API calls.
+\textbf{Observation 4.} We observed the participants had problem completing API tasks that require multiple API calls without usage examples.
 Tasks T2 and T6 required the use of multiple API calls to complete successfully.
@@
 
-\textbf{Implication 5.} If there are prerequisites for making an API call, REST API developers need to provide examples showing how to get the prerequisites in the API documentation.
+\textbf{Implication 4.} If there are prerequisites for making an API call, REST API developers need to provide examples showing how to get the prerequisites in the API documentation.
 

### 2017-03-20 controlled_study.tex
+G1
use approx

@@
   \usepackage{tikz,pgfplots,pgfplotstable}
+  \usepackage{mathtools}
   \usepgfplotslibrary{statistics}
@@
 \begin{itemize}
-  \item Total installations: ~248K installs of the plug-in.
-  \item Average daily installations: ~500.
+  \item Total installations: $ \approx 248K$ installs of the plug-in.
+  \item Average daily installations: $ \approx 500$.
 \end{itemize}
@@
 
-\textbf{Observation 4.} We observed the participants had problem completing API tasks that require multiple API calls without usage examples.
+\textbf{Observation 4.} We observed the G1 participants had problem completing API tasks that require multiple API calls without usage examples.
 Tasks T2 and T6 required the use of multiple API calls to complete successfully.

### 2017-03-21 controlled_study.tex
make it clear that the results were suprising

@@
 
-In Figure \ref{graph:productivity} and  \ref{graph:time} we see that, for the simpler tasks T2 and T3, G1 had used more trial API calls and more time on average and reached successful answers with a lower percentage compared to G2.
+Even though we expected G2 to outperform G1, we were surprised to see the difference in effort vs.
+success of the two groups.
+In Figure \ref{graph:productivity} and \ref{graph:time} we see that, for the simpler tasks, T2 and T3, G1 required more time and more trial API calls and reached successful answers with a lower percentage compared to G2.
 For T4, where the participants are required to inspect an HTTP response header, G2 needed more API calls on average, but also yielded 24\% higher success rate.
-For the more complicated tasks T5 and T6, we see that participants in G1 had very low success rate and very high number of average API trials compared to G2.
-The data here clearly depicts that without usage examples REST API client developers spend more time and require more trial and error attempts yet resulting in lower success rates, and the situation gets worse for relatively complicated tasks.
+For the more complex tasks T5 and T6, we see that participants in G1 had very low success rate and very high number of average API trials compared to G2.
+The data here depicts that without usage examples developers spend more time and require more trial and error attempts yet resulting in lower success rates, and the situation gets worse for relatively complicated tasks.
 

### 2017-03-21 controlled_study.tex
titled observations

@@
 
-\subsubsection{Development Environment}
+\subsubsection{Pilot Studies}
 
@@
 
-We designed the actual study protocol as follows: we enhanced the original WordPress REST API documentation and added a total of 3 API usage examples to show listing of blog posts with query parameters for filtering, a request to create a blog post and a request to update a blog post.
+Learning from the pilot studies, we designed the actual study protocol as follows: we enhanced the original WordPress REST API documentation and added a total of 3 API usage examples to show listing of blog posts with query parameters for filtering, a request to create a blog post and a request to update a blog post.
 We used the WordPress API unit tests to find relevant data for these examples.
@@
 
-\begin{subfigure}{0.5\textwidth}
+\begin{subfigure}[b]{0.5\textwidth}
 \centering
@@
         anchor=north,legend columns=-1},
-        ylabel=Average Trial API Calls,
+        ylabel=Average Trial API Calls (Frequency),
         width=2.8in
@@
 \end{subfigure}
-\begin{subfigure}{0.4\textwidth}
+\begin{subfigure}[b]{0.4\textwidth}
 \centering
@@
 Otherwise, it's marked as unsuccessful.
-Based on these annotations, the answers for RQ was computed using the following:
+Based on these annotations, the following formulas were used to compute the quantitative results:
 
-success rate of group = (total number of successful tasks by group) / (number of participants in group * number of API tasks)
+Success Rate = (total number of successful tasks by group) / (number of participants in group * number of API tasks)
 
-API calls per success of group = (number of API calls annotated as successful) / (total number of trial API calls made by group)
+Average Trial API Calls = (number of API calls annotated as successful) / (total number of trial API calls made by group)
 
-average time taken for success = (number of successful API tasks by group) / (total time spent by all participants in the group)
+Average Time Spent = (number of successful API tasks by group) / (total time spent by all participants in the group)
 
@@
 The average number of trial API calls and success rate for each group is computed using the following formulas:
-
-AverageTrialAPICall(task, group) = Total number of API calls made by group on task / Number of participants in group
-
-SuccessRate(task, group) = Number of participants that correctly completed task / Number of participants in group
 
@@
 
-\textbf{Observation 1.} Without usage examples, we observed 10 of the 16 G1 participants used incorrect data type of $String$ instead of $Integer$.
+\textbf{Observation: Use of Data Type.} Without usage examples, we observed 10 of the 16 G1 participants used incorrect data type of $String$ instead of $Integer$.
 To successfully complete T1, participants had to use an integer ID representing an $Author$ object given the author name.
-In the original API documentation, the following is mentioned about the author parameter: \textit{``author: Limit result set to posts assigned to specific authors.''} The documentation doesn't mention that the data type required is the numeric author ID, not the author name.
-P1.12 mentioned the following: \textbf{``...to find posts for the author, I had to inspect the response to see that indeed the $author\_id$ was in there''}.
+In the original API documentation, the following is mentioned about the author parameter: ``author: Limit result set to posts assigned to specific authors.'' The documentation doesn't mention that the data type required is the numeric author ID, not the author name.
+P1.12 mentioned the following: \textit{``...to find posts for the author, I had to inspect the response to see that indeed the $author\_id$ was in there''}.
 The enhanced documentation showed one example of using numeric ID.
@@
 
-\textbf{Observation 2.} We observed participants in G1 faced obstacles using correct data formats for Date and Array type data in API requests without usage examples.
+\textbf{Observation: Use of Data Format.} We observed participants in G1 faced obstacles using correct data formats for Date and Array type data in API requests without usage examples.
 Tasks T3 required the use of a correct format to represent an array of numeric post IDs to be excluded using the API.
-The original documentation mentions the following about this API query parameter: \textit{``exclude: Ensure result set excludes specific IDs.''} Participants attempted to solve this task using multiple different formats for specifying the array.
+The original documentation mentions the following about this API query parameter: ``exclude: Ensure result set excludes specific IDs.'' Participants attempted to solve this task using multiple different formats for specifying the array.
 For example: they tried with $exclude=[1,4]$, $exclude=1\&4$, $exclude=4\&\&exclude=1$, $id!=1\&id!=4$ and other alternatives before eventually finding the right format as follows $exclude=1,4$.
-Participant P1.6 mentioned the following: \textbf{``It was difficult to figure out whether some of the inputs needed to be arrays or just a comma-separated list.''}.
+Participant P1.6 mentioned the following: \textit{``It was difficult to figure out whether some of the inputs needed to be arrays or just a comma-separated list.''}.
 7 out of the 16 participants in G1 faced this error.
@@
 Instead of using ``2016-01-01T00:00:00-00:00'', participants from G1 used date formats as follows: ``January 1, 2016'', ``2016-01-01T12:00:00'', ``20160101'', etc.
-P1.8 mentioned the following: \textbf{``Got hung up trying to figure out the formatting of the date (which doesn't appear to be ISO8601, despite what it says)''}.
+P1.8 mentioned the following: \textit{``Got hung up trying to figure out the formatting of the date (which doesn't appear to be ISO8601, despite what it says)''}.
 After analyzing the response, we found that P1.8 used the ISO8601 formatted date in the API calls, but didn't provide the $second$ portion of the time as required by the API.
@@
 
-\textbf{Observation 3.} We were surprised to see that 14 of the 15 G1 participants that tried T5 and T6 faced problem using correct request headers without usage examples.
+\textbf{Observation: Use of Request Headers.} We were surprised to see that 14 of the 15 G1 participants that tried T5 and T6 faced problem using correct request headers without usage examples.
 Tasks T5 and T6 required the use of a HTTP request header named $Content-Type$.
-In the original API documentation, the following was mentioned: \textit{``The API uses JSON exclusively as the request and response format, including error responses.''}.
+In the original API documentation, the following was mentioned: ``The API uses JSON exclusively as the request and response format, including error responses.''.
 Experienced software engineers still had problem using this HTTP header as we found from the participants in G1.
-P1.12 mentioned the following: \textbf{``I didn't know how to specify application/json so that I could send POST data via the body instead in the URL parameters...
+P1.12 mentioned the following: \textit{``I didn't know how to specify application/json so that I could send POST data via the body instead in the URL parameters...
 Examples would really help - with a sample request and a sample response.''}  Even though G1 participants have years of experience with REST APIs, only one of them could use the $Content-Type$ header correctly without example.
@@
 
-\textbf{Observation 4.} We observed the G1 participants had problem completing API tasks that require multiple API calls without usage examples.
+\textbf{Observation: Use of Interdependent API Calls.} We observed the G1 participants had problem completing API tasks that require multiple API calls without usage examples.
 Tasks T2 and T6 required the use of multiple API calls to complete successfully.
 Both tasks require an initial API call to find data using the API that is needed make a second API call.
-For example, P1.5 mentioned the following about T6, \textbf{``The post I created is not visible and not editable, though I got 201.
+For example, P1.5 mentioned the following about T6, \textit{``The post I created is not visible and not editable, though I got 201.
 Not sure why''}.
@@
 As a result, the post was not returned via the API for T6.
-P2.2 mentioned the following about T2: \textbf{``In case of search posts by author, it's not clear how to find the author\_id.
+P2.2 mentioned the following about T2: \textit{``In case of search posts by author, it's not clear how to find the author\_id.
 Although, I realized that users and authors are same in this case.''}

### 2017-03-22 controlled_study.tex
nearly all comments addressed from Craig

@@
 Generating and maintaining REST API documentation with usage examples can be a time consuming and expensive process for evolving APIs.
-Most REST API documentation tools focus on automating the documentation of the API objects, but require manual effort for capturing realistic usage examples.
+Most REST API documentation tools focus on automating the documentation of the API objects, but require manual effort for capturing usage examples.
 Consequently, REST API developers need to know the cost vs.
-benefit of providing usage examples in the documentation as well as what to include in the usage examples to prioritize the documentation efforts.
-To this end, we have performed a controlled study with 26 experienced software engineers and found that REST API client developers face productivity problems with using correct data types, data formats, required HTTP headers and request body when documentation lacks usage examples.
-By following the REST API documentation guidelines from this paper, practitioners can reduce the errors, improve success rate and satisfaction of REST API client developers.
+benefit of providing usage examples in the documentation to prioritize the documentation efforts.
+To this end, we have performed a controlled study with 26 experienced software engineers to understand problems that REST API client developers face while using an API without usage examples.
+We found that REST API client developers face productivity problems with using correct data types, data formats, required HTTP headers and request body when documentation lacks usage examples.
+By following the REST API documentation guidelines from this paper, REST API developers can reduce the errors, improve success rate and satisfaction of API client developers.
 
@@
 \section{Introduction}
+REST APIs are used as the predominant application integration mechanism over the Internet.
+Generating and maintaining REST API documentation with usage examples can be an expensive process because most API documentation tools do not support effective usage examples.
 Researchers have emphasized API documentation as a key factor that impacts API usability both positively and negatively.
 To improve API documentation, researchers have recommended incorporating usage examples in the API documentation.
-REST APIs are used as the predominant application integration mechanism over the Internet.
-Generating and maintaining REST API documentation with usage examples can be an expensive process because most API documentation tools do not support effective usage examples.
 
 In a resource constrained environment, it is important to understand the value of usage examples on REST API usability to allocate sufficient attention and efforts towards incorporating examples in the API documentation.
-While it is no surprise that examples help API client developers, REST API developers need to know what to include in the examples and why.
+While it is expected that examples help developers, REST API developers need to know what to include in the examples and why.
 
@@
 For example, local API documentation commonly comprises the description of classes and interfaces with their methods.
-In contrast, REST API documentation needs to include information about HTTP headers, request and response body and the data representation formats such as JSON, XML, etc.
+In contrast, REST API documentation needs to include information about HTTP headers, request and response body and the data representation formats such as JSON, XML.
+The existing research on API usability areas have primarily focused on local APIs.
 In our work we have focused on understanding the impact of usage examples within the realm of the distinctive REST API features.
 
-We designed and performed a controlled study with experienced software engineers.
+We designed and performed a controlled study with experienced software engineers to understand how REST API client developers are affected while using an API documentation that describes the API elements but lacks usage examples.
 Participants were divided into two groups and given the same set of API tasks to complete.
@@
 We analyzed the data and observed recurring obstacles faced by the participants while performing the API tasks using the official documentation that lacks usage examples.
-Our contributions on this paper is as follows:
+Our contributions on this paper are as follows:
 
@@
   \item We provide a list of obstacles that REST API client developers face while performing API tasks using documentation that lacks usage examples.
-  \item We provide a list of recommendations for REST developers to be used as a guideline to incorporate usage examples in the API documentation.
+  \item We provide a list of recommendations for REST developers to be used as a guideline to incorporate usage examples in API documentation.
   \item We provide empirical evidence that usage examples in REST API documentation help API client developers perform API tasks with higher developer satisfaction, less time, and better success rate.
@@
 
-The remainder of this paper is organized as follows: In Section \ref{sec:method} we discuss our research goals, requirements, the object of the study and the participant selection, the study setup, data collection and analysis methods.
+The remainder of this paper is organized as follows: In Section \ref{sec:goal} we present our research question.
+In Section \ref{sec:method} we discuss our methodology in terms of the study requirements, the selected API of the study and the participant selection, the study setup, data collection and analysis methods.
 In Section \ref{sec:results} we discuss the results of our analysis.
@@
 
-\section{Methodology}
-\label{sec:method}
-\subsection{Research Goals}
-This research is aimed at answering the following research questions:
+\section{Research Question}
+\label{sec:goal}
+This research is aimed at answering the following research question:
 
@@
 
-\subsection{Requirements} % (fold)
+\section{Methodology} % (fold)
+\label{sec:method}
 
@@
   \item \textbf{R3.
-Time bound.} The total time spent by each participant is required to be time bound to measure productivity.
+Time bound.} The total time spent by each participant is required to be time bound to measure the rate of success within a limited time frame.
 As a result, the study needs to be setup such that participants are able to focus on performing the tasks minimizing any overhead.
   \item \textbf{R4.
-Experienced developers.} Developers with prior experience on REST APIs need to be recruited as study participants to perform the study within a limited amount of time and in a realistic setup.
+Participant Selection.} Developers with prior experience on REST APIs need to be recruited as study participants to perform the study within a limited amount of time and in a realistic setup.
 Furthermore, to reduce a learning bias, only participants with no prior experince of using the WordPress REST API V2 are accepted for this study.
@@
 WordPress is a blog-like open-source (R2) framework used by over 409 million people to visit 23.6 billion pages each month \footnote{\url{https://wordpress.com/activity/}}.
-The API allows programmatic access to list, create, update, and delete WordPress data such as blog posts, comments, users, images, tags, etc.
+The API allows programmatic access to list, create, update, and delete WordPress data such as blog posts, comments, users, images, tags.
 
@@
 
-Starting January 2017, the WordPress REST API V2 is no longer required to be installed as a separate plug-in since it's bundled with every WordPress installation.
+Starting January 2017, the WordPress REST API V2 is no longer required to be installed as a separate plug-in since it is bundled with every WordPress installation.
 According to the code repository on GitHub, there are a total of 99 and 46 contributors that had at least one commit to the code repository of the API and its documentation, respectively.
@@
 On Listing \ref{wordpress_code}, line 4 specifies that this is a schema definition for the API element $Post$.
-On Line 10, it defines $date$, one of the properties of $Post$, followed by a human readable description and and type information.
+On Line 10, it defines $date$, one of the properties of $Post$, followed by a human readable description and type information.
 On line 14, the context of this property is mentioned as $view, edit, embed$, meaning that this property will be returned when the $Post$ object is returned, embedded, or can be used as an input for editing.
 
-The WordPress team leverages the this self-documenting feature to generate and publish the official API documentation as an HTML website.
-Figure \ref{fig:schema} shows a screenshot of the published documentation for the $date$ attribute of the $Post$ API element.\footnote{\url{http://v2.wp-api.org/reference/posts/}}
+The WordPress team leverages this self-documenting feature to generate and publish the official API documentation as an HTML website.
+Fig.
+\ref{fig:schema} shows a screenshot of the published documentation for the $date$ attribute of the $Post$ API element.\footnote{\url{http://v2.wp-api.org/reference/posts/}}
 
@@
 The participants are requested to perform a total of six tasks using the API including one practice task.
-All of the tasks are related to a single API element, $Post$ \footnote{A Post identifies a blog post within WordPress}.
+All of the tasks are related to a single API element, $Post$\footnote{A Post identifies a blog post within WordPress}.
 The tasks get progressively more difficult, and all but the last task can be performed independently of each other.
@@
 In the following paragraphs, we describe each task with its description and the study observation goals against the aforementioned API input and output information.
-For each of the tasks, the participants are required to use the same WordPress REST API and its documentation.
+For each of the tasks, the participants are required to use the same WordPress REST API and accompanying documentation.
 
 \textbf{T1: List all posts task}.
-We ask the participants to use the WordPress REST API to get a list of the blog posts from a given WordPress site.
+We asked the participants to use the WordPress REST API to get a list of the blog posts from a given WordPress site.
 This is the practice task, and the inputs to answer this task are pre-filled for the participants.
@@
 
-\textbf{T3: Exclude posts by ids task.} We ask the participants to use the API to get a list of all posts excluding posts with IDs 1 and 4.
+\textbf{T3: Exclude posts by IDs task.} We ask the participants to use the API to get a list of all posts excluding posts with IDs 1 and 4.
 Participants need to use the inputs I1-2, and use a desired format on I2 to pass an array of IDs as a parameter.
@@
 
-A total 26 participants from sixteen different companies and six different countries (Canada, USA, Germany, Ireland, Brazil, and,
-Bangladesh) were recruited through online announcements posted on Twitter, Facebook, and software developer focused mailing lists.
+\begin{figure*}[t]
+\begin{center}
+\fbox{\includegraphics[width=\linewidth,keepaspectratio]{api_explorer.png}}
+\end{center}
+\caption{Screenshot of the Custom-built Web-based REST API Explorer Used by Study Participants}
+\label{fig:api_explorer}
+\end{figure*}
+
+We recruited a total of 26 participants from sixteen different companies and six different countries (Canada, USA, Germany, Ireland, Brazil, and,
+Bangladesh) through online announcements posted on Twitter, Facebook, and software developer focused mailing lists.
 Table \ref{table:participants} shows a summary of the experience level of the participants in each group.
@@
 The study involved tasks using two APIs, the WordPress REST API V2 and the GMail REST API.
-Each participant was given one of the two APIs and a set of tasks and online answer forms to complete using the API within an hour.
-The primary findings from this pilot are as follows: 1) participants required significant overhead time to setup a development environment with the proper API credentials, 2) the intermediate trial attempts of using the API are potentially more valuable than the final answer as it allows us to study what obstacles the participants face, 3) the number of tasks for the study had to be reduced to fit within the one hour limit, and 4) for GMail API, participants used up a large portion of their time on learning about how to use  OAuth.
+Each participant was given one of the two APIs and the tasks T1-T6 and online answer forms to complete using the API within an hour.
+The primary findings from this pilot are as follows: 1) participants required significant overhead time to setup a development environment with the proper API credentials, 2) the intermediate trial attempts of using the API are potentially more valuable than the final answer as it allows us to study what obstacles the participants face, 3) the number of tasks for the study had to be reduced to fit within the one hour limit, and 4) for the GMail API, participants used up a large portion of their time on learning about how to use  OAuth.
 
-\begin{figure*}[t]
-\begin{center}
-\fbox{\includegraphics[width=\linewidth,keepaspectratio]{api_explorer.png}}
-\end{center}
-\caption{Screenshot of the web-based REST API explorer}
-\label{fig:api_explorer}
-\end{figure*}
-
-To overcome the shortcomings found from the first pilot study, we decided to develop a web-based REST API explorer as shown on Figure \ref{fig:api_explorer} that allows participants to use their browser to make the API calls without setting up any development environment.
+To overcome the shortcomings found from the first pilot study, we decided to develop a web-based REST API explorer as shown on Fig.
+\ref{fig:api_explorer} that allows participants to use their browser to make the API calls without setting up any development environment.
 The web-based API explorer only requires the inputs I1-4, and displays the HTTP response headers and body on the click of a button.
@@
 \subsubsection{Protocol}
-
-Learning from the pilot studies, we designed the actual study protocol as follows: we enhanced the original WordPress REST API documentation and added a total of 3 API usage examples to show listing of blog posts with query parameters for filtering, a request to create a blog post and a request to update a blog post.
+\label{sub:protocol}
+Learning from the pilot studies, we designed the main study protocol as follows: we enhanced the original WordPress REST API documentation and added a total of 3 API usage examples to show listing of blog posts with query parameters for filtering, a request to create a blog post and a request to update a blog post.
 We used the WordPress API unit tests to find relevant data for these examples.
-Figure \ref{fig:original} shows a screenshot of the original API documentation related to T4 where the API client developers are provided with a reference table describing the different properties that can be used to create a $Post$ object\footnote{\url{http://v2.wp-api.org/reference/posts/}}.
-Figure \ref{fig:enhanced} shows a screenshot of the enhanced API documentation with a cURL based usage example.
-cURL is used because it's used by elsewhere in the original API documentation.
+Fig.
+\ref{fig:original} shows a screenshot of the original API documentation related to T4 where the API client developers are provided with a reference table describing the different properties that can be used to create a $Post$ object\footnote{\url{http://v2.wp-api.org/reference/posts/}}.
+Fig.
+\ref{fig:enhanced} shows a screenshot of the enhanced API documentation with a cURL based usage example.
+cURL is used because it is used by elsewhere in the original API documentation.
 In the enhanced API documentation, the example shows one possible API call with realistic values for the data that is described in the reference table and associated API response headers and body.
@@
 All participants were given the same set of API tasks and were requested to limit their participation time to a maximum of one hour.
-No task specific time limit was imposed except an overall limit one hour for the entire study because we wanted participants to spend sufficient time on each task without forcing them to move the next one.
+No task specific time limit was imposed except an overall limit one hour for the entire study because we wanted participants to spend sufficient time on each task without forcing them to move to the next one.
 Participants were allowed to access the internet and external resources alongside the provided API documentation to perform the tasks as they would normally use on a typical work day.
@@
 Task participations are marked successful when I1-4 matches the required values for performing the given task.
-If a participant is able to use the correct I1-4 for one of the two tasks required to complete a single task (T2), we marked it as a partially successful.
+If a participant is able to use the correct I1-4 for one of the two tasks required to complete a single task (T2), we marked it as partially successful.
 Otherwise, it's marked as unsuccessful.
@@
 
-Success Rate = (total number of successful tasks by group) / (number of participants in group * number of API tasks)
+\textbf{Success Rate} = (total number of successful tasks by group) / (number of participants in group * number of API tasks)
 
-Average Trial API Calls = (number of API calls annotated as successful) / (total number of trial API calls made by group)
+\textbf{Average Trial API Calls} = (number of API calls annotated as successful) / (total number of trial API calls made by group)
 
-Average Time Spent = (number of successful API tasks by group) / (total time spent by all participants in the group)
+\textbf{Average Time Spent} = (number of successful API tasks by group) / (total time spent by all participants in the group)
 
@@
 
-We present a summary of the quantitative results as found by analyzing a total of 539 API calls (385 from G1, 152 from G2) made by the participants from this study in Figure \ref{fig:comparison}.
-Figure \ref{graph:productivity} juxtaposes the average number of trial API calls made by each group against their rate of success in performing the given tasks T2-T6.
+We present a summary of the quantitative results as found by analyzing a total of 539 API calls (385 from G1, 152 from G2) made by the participants from this study in Fig.
+\ref{fig:comparison}.
+Fig.
+\ref{graph:productivity} juxtaposes the average number of trial API calls made by each group against their rate of success in performing the given tasks T2-T6.
 The average number of trial API calls and success rate for each group is computed using the following formulas:
@@
 success of the two groups.
-In Figure \ref{graph:productivity} and \ref{graph:time} we see that, for the simpler tasks, T2 and T3, G1 required more time and more trial API calls and reached successful answers with a lower percentage compared to G2.
+In Fig.
+\ref{graph:productivity} and \ref{graph:time} we see that, for the simpler tasks, T2 and T3, G1 required more time and more trial API calls and reached successful answers with a lower percentage compared to G2.
 For T4, where the participants are required to inspect an HTTP response header, G2 needed more API calls on average, but also yielded 24\% higher success rate.
@@
 
-In Figure \ref{fig:rating}, we see that the G2 participants provided higher satisfaction rating compared to G1 participants.
+In Fig.
+\ref{fig:rating}, we see that the G2 participants provided higher satisfaction rating compared to G1 participants.
 P1.1 provided a satisfaction rating of 9, and analyzing the participant feedback against the success, we see a false positive case, where the participant perceived that the tasks were successfully completed even though only 2 of the 5 tasks were correctly answered.
@@
 Examples would really help - with a sample request and a sample response.''}  Even though G1 participants have years of experience with REST APIs, only one of them could use the $Content-Type$ header correctly without example.
-In contrast, 2 out of the 10 participants from G2 failed used this header correctly as indicated in the example in the enhanced documentation.
+In contrast, 2 out of the 10 participants from G2 failed to use this header correctly as indicated in the example in the enhanced documentation.
 Moreover, 1 of the 2 G2 participants that didn't use the header first time, later used it correctly on a subsequent attempt.
@@
   The selection of the API, the tasks, and the participants may introduce selection bias.
-We have developed and followed a set of requirements as discussed earlier on this paper to reduce this bias.
+We have developed and followed a set of requirements as discussed earlier on this paper to reduce this bias (see Section \ref{sub:protocol}).
 This study needs to be replicated with different sets of APIs, tasks, and participants to generalize the results.
@@
 Among other factors, they recommended providing code examples in the API documentation to help API client developers.
-Our work fits in the general space of studies of API documentation obstacles by focusing specifically on problems related to the absence of examples, and by expanding into the study of REST APIs.
+Our work fits in the general space of studies of API documentation obstacles by focusing specifically on problems related to the absence of examples, and by expanding into the area of REST APIs.
 
@@
 performed an exploratory study to understand the current state of API related knowledge available on the Internet \cite{6462686}.
-They searched the web for usage examples related to the API of five popular Java libraries and found that on an average API examples could be found for 77\% of the 4,637 APIs included in the selected libraries.
+They searched the web for usage examples related to the API of five popular Java libraries and found that on average API examples could be found for 77\% of the 4,637 APIs included in the selected libraries.
 Moreover, they found that the crowd-sourcing sites accounted for 93.7\% of the usage examples compared to 6.13\% that were published on the official documentation.
@@
 Rama et al.
-presented a set of formulas for computing a measure of API usability based on the API's structural components such as classes, methods, parameters, return values, thread-safety, etc~\cite{rama}.
+presented a set of formulas for computing a measure of API usability based on the API's structural components such as classes, methods, parameters, return values, thread-safety~\cite{rama}.
 Scheller et al.
@@
 The researchers found that the examples from the unit tests helped understanding the API concepts better but it was challenging for the participants to locate relevant examples.
-Endrikat performed a controlled study with four groups of participants to understand the impact of API documentation on APIs that are implemented using programming languages with static and dynamic type systems \cite{Endrikat:2014:ADS:2568225.2568299}.
+Endrikat et al.
+performed a controlled study with four groups of participants to understand the impact of API documentation on APIs that are implemented using programming languages with static and dynamic type systems \cite{Endrikat:2014:ADS:2568225.2568299}.
 Participants were given a set of failing unit tests and were asked to make them pass by writing code using the studied API.
@@
 They found that API client developers needed more time to complete API tasks using factory method compared to using a constructor.
-Our work in this paper is based on a controlled study and shares part of the setup that were used by the aforementioned controlled studies with the following differences: our goal is to understand the obstacles faced by API client developers without usage examples;  we focus on REST APIs instead of a local API; and, the participants are professional software engineers.
+Our work in this paper is based on a controlled study and shares part of the setup that were used by the aforementioned controlled studies with the following differences: our goal is to understand the obstacles faced by API client developers without usage examples;  we focus on REST APIs instead of local APIs; and, the participants are all professional software engineers.
 
@@
 We identified that, without examples, REST API client developers have trouble using the correct data types, correct data formats, and required HTTP headers and request body.
-We've also presented empirical evidence that adding usage examples it is possible to reduce mistakes, improve success rate and developer satisfaction of using the API.
-Practitioners can follow our guidelines while documenting REST APIs to improve API usability.
+We've also presented empirical evidence that by adding usage examples makes it possible to reduce mistakes, improve success rate and developer satisfaction of using the API.
+Practitioners can follow our implications while documenting REST APIs to improve API usability.
 

### 2017-03-22 controlled_study.tex
remove authors for double-blind
fix-up

@@
 \fi
-\hyphenation{op-tical net-works semi-conduc-tor}
 
@@
 
-\author{\IEEEauthorblockN{S M Sohan, Frank Maurer\IEEEauthorrefmark{1}}
-\IEEEauthorblockA{\IEEEauthorrefmark{1}
-Dept. of Computer Science\\
-University of Calgary\\
-Canada\\
-\{smsohan, frank.maurer\}@ucalgary.ca}
-\and
-\IEEEauthorblockN{Craig Anslow}
-\IEEEauthorblockA{School of Eng. and Computer Science\\
-Victoria University of Wellington\\
-New Zealand\\
-craig@ecs.vuw.ac.nz}
-\and
-\IEEEauthorblockN{Martin Robillard}
+\author{\IEEEauthorblockN{Placeholder Author Name}
 \IEEEauthorblockA{
-School of Computer Science\\
-McGill University\\
-Canada\\
-martin@cs.mcgill.ca}
+Placeholder Department\\
+Placeholder Affiliation\\
+Placeholder Address\\
+Placeholder Email}
 }
+
+% \author{\IEEEauthorblockN{S M Sohan, Frank Maurer\IEEEauthorrefmark{1}}
+% \IEEEauthorblockA{\IEEEauthorrefmark{1}
+% Dept. of Computer Science\\
+% University of Calgary\\
+% Canada\\
+% \{smsohan, frank.maurer\}@ucalgary.ca}
+% \and
+% \IEEEauthorblockN{Craig Anslow}
+% \IEEEauthorblockA{School of Eng. and Computer Science\\
+% Victoria University of Wellington\\
+% New Zealand\\
+% craig@ecs.vuw.ac.nz}
+% \and
+% \IEEEauthorblockN{Martin Robillard}
+% \IEEEauthorblockA{
+% School of Computer Science\\
+% McGill University\\
+% Canada\\
+% martin@cs.mcgill.ca}
+% }
 
@@
 We found that REST API client developers face productivity problems with using correct data types, data formats, required HTTP headers and request body when documentation lacks usage examples.
-By following the REST API documentation guidelines from this paper, REST API developers can reduce the errors, improve success rate and satisfaction of API client developers.
+By following the REST API documentation suggestions from this paper, REST API developers can reduce the errors, improve success rate and satisfaction of API client developers.
 
@@
 \label{sec:goal}
-This research is aimed at answering the following research question:
+This research is aimed at answering the following question:
 
@@
 We've also presented empirical evidence that by adding usage examples makes it possible to reduce mistakes, improve success rate and developer satisfaction of using the API.
-Practitioners can follow our implications while documenting REST APIs to improve API usability.
+Practitioners can follow our implications as a set of guidelines while documenting REST APIs to improve API usability.
 

### 2017-03-24 controlled_study.tex
box around the implications
back to 8+1 pages

@@
 
-Starting January 2017, the WordPress REST API V2 is no longer required to be installed as a separate plug-in since it is bundled with every WordPress installation.
-According to the code repository on GitHub, there are a total of 99 and 46 contributors that had at least one commit to the code repository of the API and its documentation, respectively.
+Starting January 2017, the WordPress REST API V2 is no longer required to be installed as a separate plug-in since it is bundled with WordPress installation.
+According to the code repository, there are a total of 99 and 46 contributors that had at least one commit to the code repository of the API and its documentation, respectively.
 These properties satisfy R1, our requirements for using a representative API.
@@
 \end{center}
-\caption{Screenshot showing auto-generated documentation for Post/date}
+\caption{Screenshot of documentation for Post/date}
 \label{fig:schema}
@@
   \item Currently working as a software engineer.
-  \item At least 1 year of industry experience as a software engineer.
+  \item At least 1 year of industry experience as a developer.
   \item At least 1 year of industry experience with REST APIs.
@@
 
-\textbf{Implication 1.} REST API documentation needs to include examples of data types such as Integer, String, and Array for each API field to satisfy API client developer information need.
+\hspace{-5mm}
+\fbox{
+\begin{minipage}{0.95\linewidth}
+
+\textbf{Recommendation 1.} REST API documentation needs to include examples of data types such as Integer, String, and Array for each API field to satisfy API client developer information need.
+
+\end{minipage}
+}
+\vspace{0.5mm}
 
@@
 
-\textbf{Implication 2.} REST API developers need to include examples showing the valid data format for the API elements.
+\vspace{1mm}
+\hspace{-5mm}
+\fbox{
+\begin{minipage}{0.95\linewidth}
+
+\textbf{Recommendation 2.} REST API developers need to include examples showing the valid data format for the API elements.
+\end{minipage}
+}
+\vspace{0.5mm}
 
@@
 
-\textbf{Implication 3.} If API requests need to use HTTP Request headers, in addition to the request method and body, REST API develpers need to include examples of the HTTP headers.
+\vspace{1mm}
+\hspace{-5mm}
+\fbox{
+\begin{minipage}{0.95\linewidth}
+
+\textbf{Recommendation 3.} If API requests need to use HTTP Request headers, in addition to the request method and body, REST API develpers need to include examples of the HTTP headers.
+\end{minipage}
+}
+\vspace{0.5mm}
 
@@
 
-\textbf{Implication 4.} If there are prerequisites for making an API call, REST API developers need to provide examples showing how to get the prerequisites in the API documentation.
+\vspace{1mm}
+\hspace{-5mm}
+\fbox{
+\begin{minipage}{0.95\linewidth}
+
+\textbf{Recommendation 4.} If there are prerequisites for making an API call, REST API developers need to provide examples showing how to get the prerequisites in the API documentation.
+\end{minipage}
+}
+\vspace{0.5mm}
 
@@
 We've also presented empirical evidence that by adding usage examples makes it possible to reduce mistakes, improve success rate and developer satisfaction of using the API.
-Practitioners can follow our implications as a set of guidelines while documenting REST APIs to improve API usability.
+Practitioners can follow our recommendation as a set of guidelines while documenting REST APIs to improve API usability.
 

### 2017-03-26 controlled_study.tex
use lstinline

@@
   \usepgfplotslibrary{statistics}
+  \usepackage{courier}
 \else
@@
 Selecting a mature REST API for this study reduces the possibility that the obstacles we observe in the study are the results of insignificant accidental problems symptomatic of an immature technology.
-A familiar domain needed to be selected so that participants are able to relate to the API features without requiring upfront training.
-In addition to selecting the API, we had to select tasks that are related to the core features provided by the API to represent reality.
+We selected a familiar domain so that participants are able to relate to the API features without requiring upfront training.
+In addition to selecting the API, we had to select tasks that are related to the core features provided by the API to represent a common usage area of the API.
   \item \textbf{R2.
@@
   \item \textbf{R3.
-Time bound.} The total time spent by each participant is required to be time bound to measure the rate of success within a limited time frame.
+Time bound.} We applied a maximum time constraint for each participant to measure the rate of success within a limited time frame.
 As a result, the study needs to be setup such that participants are able to focus on performing the tasks minimizing any overhead.
@@
 According to the code repository, there are a total of 99 and 46 contributors that had at least one commit to the code repository of the API and its documentation, respectively.
-These properties satisfy R1, our requirements for using a representative API.
+These properties satisfy R1 AND R2, our requirements for using a representative API.
 By selecting an open-source project we are able to access the source-code to inspect the implementation and documentation technique of the WordPress REST API.
@@
 
-\lstset{basicstyle=\footnotesize}
-\begin{lstlisting}[language=php,breaklines=true,showspaces=false,showstringspaces=false,numbers=left,xleftmargin=2em,caption={Example of self-documenting API Code in PHP},label=wordpress_code]
+\lstset{basicstyle=\ttfamily\footnotesize,breaklines=true}
+\begin{lstlisting}[language=php,showspaces=false,showstringspaces=false,numbers=left,xleftmargin=2em,caption={Example of self-documenting API Code in PHP},label=wordpress_code]
 public function get_item_schema() {
@@
 
-On Listing \ref{wordpress_code}, line 4 specifies that this is a schema definition for the API element $Post$.
-On Line 10, it defines $date$, one of the properties of $Post$, followed by a human readable description and type information.
-On line 14, the context of this property is mentioned as $view, edit, embed$, meaning that this property will be returned when the $Post$ object is returned, embedded, or can be used as an input for editing.
+On Listing \ref{wordpress_code}, line 4 specifies that this is a schema definition for the API element \lstinline{Post} \footnote{\url{https://github.com/WP-API/WP-API}}.
+On Line 10, it defines \lstinline{date}, one of the properties of \lstinline{Post}, followed by a human readable description and type information.
+On line 14, the context of this property is mentioned as \lstinline{view, edit, embed}, meaning that this property will be returned when the \lstinline{Post} object is returned, embedded, or can be used as an input for editing.
 
@@
 Fig.
-\ref{fig:schema} shows a screenshot of the published documentation for the $date$ attribute of the $Post$ API element.\footnote{\url{http://v2.wp-api.org/reference/posts/}}
+\ref{fig:schema} shows a screenshot of the published documentation for the \lstinline{date} attribute of the \lstinline{Post} API element.\footnote{\url{http://v2.wp-api.org/reference/posts/}}
 
@@
 The participants are requested to perform a total of six tasks using the API including one practice task.
-All of the tasks are related to a single API element, $Post$\footnote{A Post identifies a blog post within WordPress}.
+All of the tasks are related to a single API element, \lstinline{Post}\footnote{A Post identifies a blog post within WordPress}.
 The tasks get progressively more difficult, and all but the last task can be performed independently of each other.
-Participants are requested to limit the total time on the study to a maximum of one hour.
+Participants are requested to limit the total time on the study to a maximum of one hour (R3).
 Participants are encouraged to proceed to the next task when they are either satisfied with their answer or feel stuck and unable to make progress.
@@
 In the following paragraphs, we describe each task with its description and the study observation goals against the aforementioned API input and output information.
-For each of the tasks, the participants are required to use the same WordPress REST API and accompanying documentation.
+For each of the tasks, the participants are required to the same WordPress REST API and one of the two variants of accompanying documentation.
 
@@
 To answer this correctly, the participants are required to first make an API call to get the numeric ID of the author given the string based user name.
-Then, the ID needs to be used on the Post API to filter posts by the author.
+Then, the ID needs to be used on the \lstinline{Post} API to filter posts by the author.
 This task allows us to understand the impact of usage examples on API client developers when multiple API calls need to be made to perform a task using the API.
@@
 
-\textbf{T6: Update post task.} We ask the participants to use the API to update a blog post that they published in T5 with a new excerpt.
+\textbf{T6: Update post task.} We ask the participants to use the API to update a blog \lstinline{Post} that they published in T5 with a new \lstinline{excerpt}.
 Similar to T5, this task requires the use of inputs I1-4, but with different values for the inputs.
@@
 
-To satisfy the requirement of developers with REST API experience (R3), we used the following criteria for recruiting the study participants:
+To satisfy the requirement of developers with REST API experience (R4), we used the following criteria for recruiting the study participants:
 
@@
 The study involved tasks using two APIs, the WordPress REST API V2 and the GMail REST API.
-Each participant was given one of the two APIs and the tasks T1-T6 and online answer forms to complete using the API within an hour.
+Each participant was given one of the two APIs and and a set of API tasks and online answer forms to complete using the API within an hour.
 The primary findings from this pilot are as follows: 1) participants required significant overhead time to setup a development environment with the proper API credentials, 2) the intermediate trial attempts of using the API are potentially more valuable than the final answer as it allows us to study what obstacles the participants face, 3) the number of tasks for the study had to be reduced to fit within the one hour limit, and 4) for the GMail API, participants used up a large portion of their time on learning about how to use  OAuth.
@@
 Fig.
-\ref{fig:original} shows a screenshot of the original API documentation related to T4 where the API client developers are provided with a reference table describing the different properties that can be used to create a $Post$ object\footnote{\url{http://v2.wp-api.org/reference/posts/}}.
+\ref{fig:original} shows a screenshot of the original API documentation related to T4 where the API client developers are provided with a reference table describing the different properties that can be used to create a \lstinline{Post} object\footnote{\url{http://v2.wp-api.org/reference/posts/}}.
 Fig.
-\ref{fig:enhanced} shows a screenshot of the enhanced API documentation with a cURL based usage example.
-cURL is used because it is used by elsewhere in the original API documentation.
+\ref{fig:enhanced} shows a screenshot of the enhanced API documentation with a cURL\footnote{\url{https://curl.haxx.se/}} based usage example.
+cURL is used because it is used elsewhere in the original API documentation.
 In the enhanced API documentation, the example shows one possible API call with realistic values for the data that is described in the reference table and associated API response headers and body.
@@
 G1 participants were provided with a link to the official API documentation on the web-based API explorer, and G2 participants were provided with a link to the enhanced API documentation with usage examples.
-The web-based API explorer allocated more participants to G1 compared to G2 because we wanted to better understand the impact of the lack of usage examples on API client developer productivity.
+We designed the the web-based API explorer to allocate more participants to G1 compared to G2 because we wanted to better understand the impact of the lack of usage examples on API client developer productivity.
 However, each individual participant was randomly assigned to a group by the web-based API explorer.
 All participants were given the same set of API tasks and were requested to limit their participation time to a maximum of one hour.
-No task specific time limit was imposed except an overall limit one hour for the entire study because we wanted participants to spend sufficient time on each task without forcing them to move to the next one.
+No task specific time limit was imposed except an overall limit of one hour for the entire study because we wanted participants to spend sufficient time on each task without forcing them to move to the next one.
 Participants were allowed to access the internet and external resources alongside the provided API documentation to perform the tasks as they would normally use on a typical work day.
@@
 
-\textbf{Success Rate} = (total number of successful tasks by group) / (number of participants in group * number of API tasks)
+\textbf{Success Rate (Task, Group)} = (No. of participants that successfully completed Task) / (No. of participants in Group)
 
-\textbf{Average Trial API Calls} = (number of API calls annotated as successful) / (total number of trial API calls made by group)
+\textbf{Average Trial API Calls (Task, Group)} = (No. of API calls made by Group on Task) / (No. of participants in Group).
 
-\textbf{Average Time Spent} = (number of successful API tasks by group) / (total time spent by all participants in the group)
+\textbf{Average Time Spent (Task, Group)} = (Time spent by Group on Task) / (No. of participants in Group)
 
@@
 
-Even though we expected G2 to outperform G1, we were surprised to see the difference in effort vs.
+Even though it is expected that G2 would outperform G1 with an enhanced documentation, we were surprised to see the difference in effort vs.
 success of the two groups.
@@
 On the other hand, 7 out of the 10 G2 participants mentioned the examples being helpful from the enhanced documentation.
-Based on these feedback, we identify usage examples as the most frequent and central topic of interest from our participants.
+We identify usage examples as the most frequent and central topic of interest from our participants as per their feedback.
 This is also supported by our quantitative analysis results as discussed earlier in this section.
@@
 
-\textbf{Observation: Use of Data Type.} Without usage examples, we observed 10 of the 16 G1 participants used incorrect data type of $String$ instead of $Integer$.
-To successfully complete T1, participants had to use an integer ID representing an $Author$ object given the author name.
+\textbf{Observation: Use of Data Type.} Without usage examples, we observed 10 of the 16 G1 participants used incorrect data type of \lstinline{String} instead of \lstinline{Integer}.
+To successfully complete T1, participants had to use an integer ID representing an \lstinline{Author} object given the author name.
 In the original API documentation, the following is mentioned about the author parameter: ``author: Limit result set to posts assigned to specific authors.'' The documentation doesn't mention that the data type required is the numeric author ID, not the author name.
@@
 The original documentation mentions the following about this API query parameter: ``exclude: Ensure result set excludes specific IDs.'' Participants attempted to solve this task using multiple different formats for specifying the array.
-For example: they tried with $exclude=[1,4]$, $exclude=1\&4$, $exclude=4\&\&exclude=1$, $id!=1\&id!=4$ and other alternatives before eventually finding the right format as follows $exclude=1,4$.
+For example: they tried with \lstinline{exclude=[1,4], exclude=1&4, exclude=4&&exclude=1, id!=1&id!=4} and other alternatives before eventually finding the right format as follows \lstinline{exclude=1,4}.
 Participant P1.6 mentioned the following: \textit{``It was difficult to figure out whether some of the inputs needed to be arrays or just a comma-separated list.''}.
@@
 Even though the original documentation mentions ISO8601 format, it doesn't provide an example.
-Instead of using ``2016-01-01T00:00:00-00:00'', participants from G1 used date formats as follows: ``January 1, 2016'', ``2016-01-01T12:00:00'', ``20160101'', etc.
+Instead of using ``2016-01-01T00:00:00-00:00'', participants from G1 used date formats as follows: \lstinline{``January 1, 2016'', ``2016-01-01T12:00:00'', ``20160101''}.
 P1.8 mentioned the following: \textit{``Got hung up trying to figure out the formatting of the date (which doesn't appear to be ISO8601, despite what it says)''}.
-After analyzing the response, we found that P1.8 used the ISO8601 formatted date in the API calls, but didn't provide the $second$ portion of the time as required by the API.
+After analyzing the response, we found that P1.8 used the ISO8601 formatted date in the API calls, but didn't provide the required \lstinline{second} portion of the time.
 8 of the 16 participants in G1 had an error in the date formats.
@@
 \textbf{Observation: Use of Request Headers.} We were surprised to see that 14 of the 15 G1 participants that tried T5 and T6 faced problem using correct request headers without usage examples.
-Tasks T5 and T6 required the use of a HTTP request header named $Content-Type$.
+Tasks T5 and T6 required the use of a HTTP request header named \lstinline{Content-Type}.
 In the original API documentation, the following was mentioned: ``The API uses JSON exclusively as the request and response format, including error responses.''.
@@
 P1.12 mentioned the following: \textit{``I didn't know how to specify application/json so that I could send POST data via the body instead in the URL parameters...
-Examples would really help - with a sample request and a sample response.''}  Even though G1 participants have years of experience with REST APIs, only one of them could use the $Content-Type$ header correctly without example.
+Examples would really help - with a sample request and a sample response.''}  Even though G1 participants have years of experience with REST APIs, only one of them could use the \lstinline{Content-Type} header correctly without example.
 In contrast, 2 out of the 10 participants from G2 failed to use this header correctly as indicated in the example in the enhanced documentation.
@@
 Not sure why''}.
-After analyzing the response, we observed that P1.5 was able to create a blog post, but failed to publish it because it didn't have the $publish$ status.
+After analyzing the response, we observed that P1.5 was able to create a blog post, but failed to publish it because it didn't have the \lstinline{publish} status.
 As a result, the post was not returned via the API for T6.

### 2017-03-26 controlled_study.tex
fixes grammar

@@
 Among other factors, they recommended providing code examples in the API documentation to help API client developers.
-Our work fits in the general space of studies of API documentation obstacles by focusing specifically on problems related to the absence of examples, and by expanding into the area of REST APIs.
+Our work fits in the general space of studies of API documentation obstacles.
+We focus specifically on problems related to the absence of examples expanding within the area of REST APIs.
 

### 2017-03-26 controlled_study.tex
add the confidence interval info

@@
 For each row, the table contains the following columns: API task, participant identifier, timestamp, API trial number, I1-4, response headers and body.
-It allowed us to observe and categorize the different input values that the participants used to perform the given API tasks.
 The values for I1-4 used by the participants were manually coded to group the participant responses into categories.
@@
 The analysis started with an empty set of codes and new codes were introduced to describe scenarios that did not fall under  codes that were already applied.
-The first author applied the codes on the raw data artifact and the co-authors verified the codes randomly.
+The first author applied the codes on the raw data artifact and provided a coding scheme to a co-author.
+The co-author applied the codes to 82/539 (10\% confidence interval with 95\% confidence level) API calls made by the participants.
 To resolve disagreements, the authors discussed and updated the codes accordingly.
-The codes were then further categorized to higher level grouping to represent common problems faced by the study participants for each group.
 

### 2017-04-05 controlled_study.tex
back to the correct size + observation number

@@
 The first author applied the codes on the raw data artifact and provided a coding scheme to a co-author.
-The co-author applied the codes to 82/539 (10\% confidence interval with 95\% confidence level) API calls made by the participants.
-To resolve disagreements, the authors discussed and updated the codes accordingly.
+The co-author applied the codes to 82/539 API calls (10\% confidence interval with 95\% confidence level) made by the participants.
+The two authors validated the codes against each other to provide consistency.
 
@@
 
-\textbf{Observation: Use of Data Type.} Without usage examples, we observed 10 of the 16 G1 participants used incorrect data type of \lstinline{String} instead of \lstinline{Integer}.
+\textbf{Observation 1: Use of Data Type.} Without usage examples, we observed 10 of the 16 G1 participants used incorrect data type of \lstinline{String} instead of \lstinline{Integer}.
 To successfully complete T1, participants had to use an integer ID representing an \lstinline{Author} object given the author name.
@@
 
-\textbf{Observation: Use of Data Format.} We observed participants in G1 faced obstacles using correct data formats for Date and Array type data in API requests without usage examples.
+\textbf{Observation 2: Use of Data Format.} We observed participants in G1 faced obstacles using correct data formats for Date and Array type data in API requests without usage examples.
 Tasks T3 required the use of a correct format to represent an array of numeric post IDs to be excluded using the API.
@@
 
-\textbf{Observation: Use of Request Headers.} We were surprised to see that 14 of the 15 G1 participants that tried T5 and T6 faced problem using correct request headers without usage examples.
+\textbf{Observation 3: Use of Request Headers.} We were surprised to see that 14 of the 15 G1 participants that tried T5 and T6 faced problem using correct request headers without usage examples.
 Tasks T5 and T6 required the use of a HTTP request header named \lstinline{Content-Type}.
@@
 
-\textbf{Observation: Use of Interdependent API Calls.} We observed the G1 participants had problem completing API tasks that require multiple API calls without usage examples.
+\textbf{Observation 4: Use of Interdependent API Calls.} We observed the G1 participants had problem completing API tasks that require multiple API calls without usage examples.
 Tasks T2 and T6 required the use of multiple API calls to complete successfully.
@@
 Our work fits in the general space of studies of API documentation obstacles.
-We focus specifically on problems related to the absence of examples expanding within the area of REST APIs.
+We focus on problems related to the absence of examples within the area of REST APIs.
 
@@
 performed an exploratory study to understand the current state of API related knowledge available on the Internet \cite{6462686}.
-They searched the web for usage examples related to the API of five popular Java libraries and found that on average API examples could be found for 77\% of the 4,637 APIs included in the selected libraries.
-Moreover, they found that the crowd-sourcing sites accounted for 93.7\% of the usage examples compared to 6.13\% that were published on the official documentation.
+They searched the web for usage examples related to the API of five popular Java libraries and found that on average API examples could be found for 77\% of the 4,637 APIs from the libraries.
+Moreover, they found that the crowd-sourcing sites accounted for 93.7\% of the usage examples compared to 6.13\% that were officially published.
 An evaluation of Jadite showed that developers were three times faster to complete API tasks with access to auto-referred crowd-sourced usage examples~\cite{5295283}.
 Nasehi et al.
-analyzed StackOverflow threads to understand the characteristics of good and bad code examples based on user provided votes \cite{Nasehi_what_makes}.
+analyzed StackOverflow threads to understand the characteristics of good and bad code examples \cite{Nasehi_what_makes}.
 They found that commonly down-voted API related answers lacked code examples, explanations, and shortcoming of solutions.
@@
 observed a severe inequality within the context of crowd-sourced API usage examples where most of the content were related to the popular APIs \cite{jiau}.
-Existing research on crowd-sourced API documentation has mostly focused finding usage examples of local APIs.
+Existing research on crowd-sourced API documentation mostly focused on finding usage examples of local APIs.
 In this paper, we focused on REST APIs, where the API client development programming language is agnostic to API implementation language.
-This distinction with local APIs makes it hard to leverage the techniques proposed in the aforementioned papers for finding crowd-sourced REST API usage examples.
+This distinction with local APIs makes it hard to use the techniques proposed in the aforementioned papers for finding crowd-sourced REST API usage examples.
 

### 2017-04-05 controlled_study.tex
consistent format

@@
 
-\textbf{Observation 1: Use of Data Type.} Without usage examples, we observed 10 of the 16 G1 participants used incorrect data type of \lstinline{String} instead of \lstinline{Integer}.
+\textbf{Observation 1.
+Use of Data Type.} Without usage examples, we observed 10 of the 16 G1 participants used incorrect data type of \lstinline{String} instead of \lstinline{Integer}.
 To successfully complete T1, participants had to use an integer ID representing an \lstinline{Author} object given the author name.
@@
 
-\textbf{Observation 2: Use of Data Format.} We observed participants in G1 faced obstacles using correct data formats for Date and Array type data in API requests without usage examples.
+\textbf{Observation 2.
+Use of Data Format.} We observed participants in G1 faced obstacles using correct data formats for Date and Array type data in API requests without usage examples.
 Tasks T3 required the use of a correct format to represent an array of numeric post IDs to be excluded using the API.
@@
 
-\textbf{Observation 3: Use of Request Headers.} We were surprised to see that 14 of the 15 G1 participants that tried T5 and T6 faced problem using correct request headers without usage examples.
+\textbf{Observation 3.
+Use of Request Headers.} We were surprised to see that 14 of the 15 G1 participants that tried T5 and T6 faced problem using correct request headers without usage examples.
 Tasks T5 and T6 required the use of a HTTP request header named \lstinline{Content-Type}.
@@
 
-\textbf{Observation 4: Use of Interdependent API Calls.} We observed the G1 participants had problem completing API tasks that require multiple API calls without usage examples.
+\textbf{Observation 4.
+Use of Interdependent API Calls.} We observed the G1 participants had problem completing API tasks that require multiple API calls without usage examples.
 Tasks T2 and T6 required the use of multiple API calls to complete successfully.

### 2017-04-20 controlled_study.tex
adds the suggested text changes by Martin sans the agreement portion

@@
 A second pilot study involving seven new participants was performed to collect usability related feedback about the web-based REST API explorer.
-We only used WordPress since it didn't require knowledge about OAuth.
+We only used WordPress since it did not require knowledge about OAuth.
 From this study, we observed patterns of mistakes that API client developers make that can be reduced by adding usage examples in the API documentation.
@@
 
-\subsection{Quantitative Analysis Results} % (fold)
+\subsection{Quantitative Analysis} % (fold)
 
@@
 \ref{graph:productivity} juxtaposes the average number of trial API calls made by each group against their rate of success in performing the given tasks T2-T6.
-The average number of trial API calls and success rate for each group is computed using the following formulas:
+The average number of trial API calls and success rate for each group is computed using the aforementioned formulas.
 
-Even though it is expected that G2 would outperform G1 with an enhanced documentation, we were surprised to see the difference in effort vs.
+Even though it is expected that G2 would outperform G1 with an enhanced documentation, we were surprised to see the extent of the difference in effort vs.
 success of the two groups.
@@
 \ref{graph:productivity} and \ref{graph:time} we see that, for the simpler tasks, T2 and T3, G1 required more time and more trial API calls and reached successful answers with a lower percentage compared to G2.
-For T4, where the participants are required to inspect an HTTP response header, G2 needed more API calls on average, but also yielded 24\% higher success rate.
+For T4, where the participants are required to inspect an HTTP response header, G2 executed more API calls on average, but also yielded 24\% higher success rate.
 For the more complex tasks T5 and T6, we see that participants in G1 had very low success rate and very high number of average API trials compared to G2.
-The data here depicts that without usage examples developers spend more time and require more trial and error attempts yet resulting in lower success rates, and the situation gets worse for relatively complicated tasks.
+The data here confirms the intuition that without usage examples developers spend more time and execute more trial and error attempts yet have lower success rates.
+And the situation gets worse for relatively complicated tasks.
 
@@
 
-Based on these results, we recommend REST API developers to study the usability of their API documentation with prospective API client developers and provide required usage examples in the documentation to improve API client developer productivity and satisfaction.
+The results provide an initial quantification of the impact of examples for typical REST-API usage tasks.
+This quantification is preliminary evidence that can help justify future investment in API documentation.
 
@@
 
-11 of the 16 G1 participants mentioned the lack of examples with HTTP headers, request and response as the primary problem with the API documentation.
+In their free-form feedback, 11 of the 16 G1 participants mentioned the lack of examples with HTTP headers, request and response as the primary problem with the API documentation.
 On the other hand, 7 out of the 10 G2 participants mentioned the examples being helpful from the enhanced documentation.
 We identify usage examples as the most frequent and central topic of interest from our participants as per their feedback.
-This is also supported by our quantitative analysis results as discussed earlier in this section.
-REST API developers need to provide usage examples showing sample requests and responses with HTTP headers and body for API actions to help API usability.
+This is also supported by our quantitative analysis results.
 In the remainder of this section, we provide our observations related to each task and their implications.
@@
 To successfully complete T1, participants had to use an integer ID representing an \lstinline{Author} object given the author name.
-In the original API documentation, the following is mentioned about the author parameter: ``author: Limit result set to posts assigned to specific authors.'' The documentation doesn't mention that the data type required is the numeric author ID, not the author name.
+In the original API documentation, the following is mentioned about the author parameter: ``author: Limit result set to posts assigned to specific authors.'' The documentation does not mention that the data type required is the numeric author ID, not the author name.
 P1.12 mentioned the following: \textit{``...to find posts for the author, I had to inspect the response to see that indeed the $author\_id$ was in there''}.
@@
 \textbf{Observation 2.
-Use of Data Format.} We observed participants in G1 faced obstacles using correct data formats for Date and Array type data in API requests without usage examples.
+Use of Data Format.} We observed participants in G1 faced obstacles using correct data formats for \lstinline{Date} and \lstinline{Array} type data in API requests without usage examples.
 Tasks T3 required the use of a correct format to represent an array of numeric post IDs to be excluded using the API.
@@
 Participant P1.6 mentioned the following: \textit{``It was difficult to figure out whether some of the inputs needed to be arrays or just a comma-separated list.''}.
-7 out of the 16 participants in G1 faced this error.
+Seven of the 16 participants in G1 faced this error.
 The enhanced documentation showed one example of using multiple IDs and all G2 participants could use the right format.
 
-Task T5 also required the participants to use an ISO8601 formatted date for publishing a post with a specific date.
-Even though the original documentation mentions ISO8601 format, it doesn't provide an example.
-Instead of using ``2016-01-01T00:00:00-00:00'', participants from G1 used date formats as follows: \lstinline{``January 1, 2016'', ``2016-01-01T12:00:00'', ``20160101''}.
+Task T5 also required the participants to use an the ISO8601 formatted date for publishing a post with a specific date.
+Even though the original documentation mentions ISO8601 format, it does not provide an example.
+Instead of using \lstinline{2016-01-01T00:00:00-00:00}, participants from G1 used date formats as follows: \lstinline{January 1, 2016, 2016-01-01T12:00:00, 20160101}.
 P1.8 mentioned the following: \textit{``Got hung up trying to figure out the formatting of the date (which doesn't appear to be ISO8601, despite what it says)''}.
-After analyzing the response, we found that P1.8 used the ISO8601 formatted date in the API calls, but didn't provide the required \lstinline{second} portion of the time.
-8 of the 16 participants in G1 had an error in the date formats.
+After analyzing the response, we found that P1.8 used the ISO8601 formatted date in the API calls, but did not provide the required \lstinline{second} portion of the time.
+Eight the 16 participants in G1 had an error in the date formats.
 With an example, 1 of the 10 participants had the same error from G2.
@@
 In contrast, 2 out of the 10 participants from G2 failed to use this header correctly as indicated in the example in the enhanced documentation.
-Moreover, 1 of the 2 G2 participants that didn't use the header first time, later used it correctly on a subsequent attempt.
+Moreover, 1 of the 2 G2 participants that did not use the header the first time, later used it correctly on a subsequent attempt.
 
@@
 \label{sec:threat}
-  The selection of the API, the tasks, and the participants may introduce selection bias.
-We have developed and followed a set of requirements as discussed earlier on this paper to reduce this bias (see Section \ref{sub:protocol}).
-This study needs to be replicated with different sets of APIs, tasks, and participants to generalize the results.
-The analysis of the raw data may introduce a bias.
-To reduce this, we have involved multiple researchers in the analysis process.
+  The selection of the API, the tasks, and the participants may introduce selection bias
+ that there are obstacles we don't know about that could be more severe, but were not observed because the tasks did not involve anything related to these obstacles.
+  The analysis of the raw data may introduce a bias. To reduce this, we have involved multiple researchers in the analysis process.
 

### 2017-04-20 controlled_study.tex
adds the numbers for coding agreements

@@
 We analyzed the data artifact qualitatively to understand the obstacles that REST API client developers face when using an API documentation that lacks usage examples.
-We exported a table of data from the database behind web-based REST API explorer.
+We exported a table of data from the web-based REST API explorer.
 For each row, the table contains the following columns: API task, participant identifier, timestamp, API trial number, I1-4, response headers and body.
 The values for I1-4 used by the participants were manually coded to group the participant responses into categories.
-The categories help us determine the types of information that API client developers need in the API documentation to perform API tasks successfully.
-The analysis started with an empty set of codes and new codes were introduced to describe scenarios that did not fall under  codes that were already applied.
-The first author applied the codes on the raw data artifact and provided a coding scheme to a co-author.
-The co-author applied the codes to 82/539 API calls (10\% confidence interval with 95\% confidence level) made by the participants.
+The categories help us determine the information type that API client developers need in the API documentation to perform API tasks.
+The analysis started with an empty set of codes and new codes were introduced to describe scenarios that did not fall under existing codes.
+The first author applied the codes on the raw data artifact and provided a coding scheme comprising 11 codes to a co-author.
+The co-author applied the codes to 82/539 API calls (10\% confidence interval with 95\% confidence level).
 The two authors validated the codes against each other to provide consistency.
+The initial coding from the co-author mismatched 4 / 11 codes for 27 / 82 API call samples, which were resolved when the coding was repeated after a discussion.
 
@@
 \label{sec:threat}
-  The selection of the API, the tasks, and the participants may introduce selection bias
- that there are obstacles we don't know about that could be more severe, but were not observed because the tasks did not involve anything related to these obstacles.
-  The analysis of the raw data may introduce a bias. To reduce this, we have involved multiple researchers in the analysis process.
+The selection of the API, the tasks, and the participants may introduce selection bias.
+It is possible that there are obstacles we don't know about that could be more severe, but were not observed because the tasks did not involve anything related to these obstacles.
+The analysis of the raw data may introduce a bias.
+To reduce this, we have involved multiple researchers in the analysis process with a repeatable coding scheme.
 
@@
 We've also presented empirical evidence that by adding usage examples makes it possible to reduce mistakes, improve success rate and developer satisfaction of using the API.
-Practitioners can follow our recommendation as a set of guidelines while documenting REST APIs to improve API usability.
+REST API developers can follow our recommendation as a set of guidelines while documenting REST APIs to improve API usability.
+REST API documentation tool developers can leverage our findings to improve reusable tool support for software developers.
 

### 2017-04-20 controlled_study.tex
more formal

@@
 Not sure why''}.
-After analyzing the response, we observed that P1.5 was able to create a blog post, but failed to publish it because it didn't have the \lstinline{publish} status.
+After analyzing the response, we observed that P1.5 was able to create a blog post, but failed to publish it because it did not have the \lstinline{publish} status.
 As a result, the post was not returned via the API for T6.
@@
 The selection of the API, the tasks, and the participants may introduce selection bias.
-It is possible that there are obstacles we don't know about that could be more severe, but were not observed because the tasks did not involve anything related to these obstacles.
+It is possible that there are obstacles we do not know about that could be more severe, but were not observed because the tasks did not involve anything related to these obstacles.
 The analysis of the raw data may introduce a bias.

### 2017-04-21 controlled_study.tex
fix as per Frank's comment

@@
 In a resource constrained environment, it is important to understand the value of usage examples on REST API usability to allocate sufficient attention and efforts towards incorporating examples in the API documentation.
-While it is expected that examples help developers, REST API developers need to know what to include in the examples and why.
+While it is expected that examples help API client developers, API developers need to know what to include in the examples and why.
 
@@
 
-We conducted two pilot studies to evaluate and understand an effective process for performing this study.
-The first pilot study involved four participants that were invited to join the first author in-person or over video conferencing.
-The study involved tasks using two APIs, the WordPress REST API V2 and the GMail REST API.
+We conducted two pilot studies to develop a process for performing this study.
+The first pilot study involved four participants that joined the first author in-person or over video conferencing.
+The study involved tasks using the WordPress REST API V2 and the GMail REST API.
 Each participant was given one of the two APIs and and a set of API tasks and online answer forms to complete using the API within an hour.
@@
 \ref{fig:api_explorer} that allows participants to use their browser to make the API calls without setting up any development environment.
-The web-based API explorer only requires the inputs I1-4, and displays the HTTP response headers and body on the click of a button.
-Thus the participants could focus on using the right input and verifying the output without having to write any code.
+The web-based API explorer records the inputs I1-4, and displays the API response on the click of a button.
+Thus the participants could focus on using the right input and verifying the output without writing code.
 The web-base REST API explorer also allowed us to automatically collect all the trial API calls that the participants make for each API task.
@@
 We divided the participants into two groups, G1 and G2.
+The participants in the pilot are not counted towards G1 and G2 for the study.
 G1 participants were provided with a link to the official API documentation on the web-based API explorer, and G2 participants were provided with a link to the enhanced API documentation with usage examples.
@@
 The two authors validated the codes against each other to provide consistency.
-The initial coding from the co-author mismatched 4 / 11 codes for 27 / 82 API call samples, which were resolved when the coding was repeated after a discussion.
+The initial coding from the co-author had discrepancies in 4 / 11 codes for 27 / 82 API call samples.
+After discussing the coding scheme, the discrepancies were resolved as the definition of the terms became clearer.
 
@@
 \section*{Acknowledgment}
-We are grateful to our participants for their valuable time on this study. We thank the University of Calgary for funding this research.
+We are grateful to our participants for their valuable time on this study. We thank our university for funding this research.
 

### 2017-07-19 controlled_study.tex
updates the title + adds the stats

@@
 
-\author{\IEEEauthorblockN{Placeholder Author Name}
+ \author{\IEEEauthorblockN{S M Sohan, Frank Maurer\IEEEauthorrefmark{1}}
+ \IEEEauthorblockA{\IEEEauthorrefmark{1}
+ Dept. of Computer Science\\
+ University of Calgary\\
+ Canada\\
+ \{smsohan, frank.maurer\}@ucalgary.ca}
+ \and
+ \IEEEauthorblockN{Craig Anslow}
+ \IEEEauthorblockA{School of Eng. and Computer Science\\
+ Victoria University of Wellington\\
+ New Zealand\\
+ craig@ecs.vuw.ac.nz}
+ \and
+ \IEEEauthorblockN{Martin P. Robillard}
  \IEEEauthorblockA{
-Placeholder Department\\
-Placeholder Affiliation\\
-Placeholder Address\\
-Placeholder Email}
+ School of Computer Science\\
+ McGill University\\
+ Canada\\
+ martin@cs.mcgill.ca}
  }
-
-% \author{\IEEEauthorblockN{S M Sohan, Frank Maurer\IEEEauthorrefmark{1}}
-% \IEEEauthorblockA{\IEEEauthorrefmark{1}
-% Dept. of Computer Science\\
-% University of Calgary\\
-% Canada\\
-% \{smsohan, frank.maurer\}@ucalgary.ca}
-% \and
-% \IEEEauthorblockN{Craig Anslow}
-% \IEEEauthorblockA{School of Eng. and Computer Science\\
-% Victoria University of Wellington\\
-% New Zealand\\
-% craig@ecs.vuw.ac.nz}
-% \and
-% \IEEEauthorblockN{Martin Robillard}
-% \IEEEauthorblockA{
-% School of Computer Science\\
-% McGill University\\
-% Canada\\
-% martin@cs.mcgill.ca}
-% }
 
@@
 \ref{graph:productivity} and \ref{graph:time} we see that, for the simpler tasks, T2 and T3, G1 required more time and more trial API calls and reached successful answers with a lower percentage compared to G2.
-For T4, where the participants are required to inspect an HTTP response header, G2 executed more API calls on average, but also yielded 24\% higher success rate.
+For T4, where the participants are required to inspect an HTTP response header, G2 executed more API calls on average due to P2.7 who made 11 API calls even though reached the successful call on the 3rd attempt.
 For the more complex tasks T5 and T6, we see that participants in G1 had very low success rate and very high number of average API trials compared to G2.
-The data here confirms the intuition that without usage examples developers spend more time and execute more trial and error attempts yet have lower success rates.
-And the situation gets worse for relatively complicated tasks.
+We observed a p-value of 0.0003 for the success rates using Fisher test, and p-values of 0.017  and 0.003 for the trial API calls and time spent using Wilcoxon rank sum test.
+The data here confirms that without usage examples developers spend more time and execute more trial attempts yet have lower success rates.
 
@@
 The selection of the API, the tasks, and the participants may introduce selection bias.
-It is possible that there are obstacles we do not know about that could be more severe, but were not observed because the tasks did not involve anything related to these obstacles.
+It is possible that there are obstacles we do not know about that could be more severe, but were not observed because of our task selection.
 The analysis of the raw data may introduce a bias.
-To reduce this, we have involved multiple researchers in the analysis process with a repeatable coding scheme.
+To reduce this, we have involved multiple researchers a repeatable coding scheme.
+A more complete API documentation without adding usage examples may provide  insights from our findings.
 

### 2017-07-19 controlled_study.tex
all the stats are added

@@
 For the more complex tasks T5 and T6, we see that participants in G1 had very low success rate and very high number of average API trials compared to G2.
-We observed a p-value of 0.0003 for the success rates using Fisher test, and p-values of 0.017  and 0.003 for the trial API calls and time spent using Wilcoxon rank sum test.
+We observed a p-value of 0.0003 for the success rates using Fisher test, and p-values of 0.017  and 0.051 for the trial API calls and time spent using Wilcoxon rank sum test.
 The data here confirms that without usage examples developers spend more time and execute more trial attempts yet have lower success rates.
@@
 In Fig.
-\ref{fig:rating}, we see that the G2 participants provided higher satisfaction rating compared to G1 participants.
+\ref{fig:rating}, we see that the G2 participants provided higher satisfaction rating compared to G1 participants (with a p-value of 0.003 using Wilcoxon rank sum test).
 P1.1 provided a satisfaction rating of 9, and analyzing the participant feedback against the success, we see a false positive case, where the participant perceived that the tasks were successfully completed even though only 2 of the 5 tasks were correctly answered.
@@
 
+These results support using a more qualitative approach due to our limited number of participants.
 The results provide an initial quantification of the impact of examples for typical REST-API usage tasks.
@@
 The analysis of the raw data may introduce a bias.
-To reduce this, we have involved multiple researchers a repeatable coding scheme.
+To reduce this, we have involved multiple researchers and a repeatable coding scheme.
 A more complete API documentation without adding usage examples may provide  insights from our findings.

### 2017-07-20 rebuttal.txt
adds all the required changes

We thank the reviewers for their time and constructive feedback on our paper. We plan to incorporate the suggestions and clarify the concerns as follows:

REVIEW 1

We'd like to clarify that the paper studied 26 developers in two groups. We'll update the title on Table 1 to further clarify this number.

1.1 We agree with the reviewer that a third experiment with an enhanced documentation without adding usage examples may provide different insights. We will mention this as a limitation of this paper in Section V. (Threats to Validity).

1.2 The examples show a single API call for three API actions with their associated responses. The data for the examples are taken from the original API documentation and the unit test code of the API. The original API documentation is located at http://v2.wp-api.org/reference/posts/ and the enhanced documentation is located at https://spyrest.github.io/docs-v2/reference/posts/

1.3 On task T4, average trial API call and time spent by G2 was higher than G1 primarily due to P2.7 who made 11 API calls. The rest of G2 participants had the following numbers of trial API calls (1, 5, 5, 3, 2, 6, 1, 1, 3) on T4. P2.7 reached the correct answer on the third trial, but made 8 subsequent trials to confirm the results.

1.4 In the paper, we performed a qualitative study to understand the types of obstacles faced by REST API client developers without examples with the aim that our results will help the design of API documentation. The findings differ from local APIs as follows: Local APIs are often used by API client developers in the same programming language of the API. As such, information regarding data types and formats can often be inferred using code-completion features. Also, HTTP request and response headers are unique to web APIs that are not used in a local API. We'll clarify these distinctions related to our findings in an the camera ready version.

REVIEW 2

2.1 As we discused in 1.1, we will acknowledge this as a limiation of our work. We have provided links to the original and enhanced API documentation in 1.2. The data presented in the usage examples in the enhanced version is sourced from the original API documentation. For example, in the "Create a Post" section, the original documentation mentions five possible values for the attribute "status". In the enhanced version, we used one of those five values.

2.2 Based on the feedback about quantitative analysis, we performed statistical tests on the four quantiative results and found the following: a) Using Fisher test, we observed a p-value of 0.0003 for the success rates of the two groups, b) p-value of 0.017 for the trial API calls made by the two groups using Wilcoxon rank sum test, c) p-value of 0.003 for the satiscation rating of the two groups using Wilcoxon rank sum test, and d) p-value of 0.051 for the time taken by the two groups using Wilcoxon rank sum test. Based on these results, we see that the observations reported on Figures 4a and 5 are satistically significant, while Figure 4b is not. We will add these findings in the camera-ready version.

2.3 In our study, for three API actions, we added one example API call per action with its request and response data showing the HTTP query parameters, headers and body attributes. We observed these single examples helped developers in performing the API tasks avoiding the obstacles that are faced without the examples. For complicated API actions, it may be necessary to provide multiple examples for a single API action to describe intended usages scenarios. We will mention this as a limitation in our paper.

2.4 We'll update the references in the related work section as suggested and make the grammatical corrections as pointed out by the reviewer.

REVIEW 3

3.1 Please see 1.3

3.2 The participants accessed the web-based REST API explorer used for this study on their own computers and we did not record web search activities. They were told to access the internet search tools as needed only to seek information regarding the given API tasks during their participation.

3.3 We will update the paper to incorporate the suggested corrections in a new version.

----------------------- REVIEW 1 ---------------------
PAPER: 20
TITLE: A Study of the Effectiveness of Usage Examples in REST API Documentation
AUTHORS: S. M. Sohan, Frank Maurer, Craig Anslow and Martin Robillard

Overall evaluation: 2 (accept)
Reviewer's confidence: 4 ((high))

----------- Review -----------
Summary:
This paper evaluates the challenges faced by developers using REST APIs and provides a list of empirically-derived recommendations for addressing those challenges. The paper presents a between-subject study of 16 developers where one treatment got the standard WordPress REST API v2 and the other received an enhanced edition that provided 3 additional usage examples that contained useful additional information not found in the normal API. Participants were given 6 REST API tasks to complete, a together performed a total of 539 API calls. The results showed that participants who received the treatment with additional examples finished more of the tasks than those who didn’t

Review:
This paper was straightforward and well-written. It positioned itself well among the related work and made a compelling argument for how it differed to what’s been done before.

I found the takeaways to be useful, yet I wanted to see more depth from the recommendations beyond “API’s should (1) define expected data types, (2) define data formats, (3) define expected inputs and (4) define prerequisites.” The authors present these takeaways as justifications for including examples in the APIs, but they could also be addressed by having through and complete documentation. It would have been insightful to have the study include a third treatment, one where the existing documentation has the missing information added sans example. I suspect the treatment with examples would still outperform and it would make for a more compelling paper.

I’d like to see all three examples that were provided to participants. It’s unclear in the paper if the three examples were too similar to the six tasks given to participants and thus effectively giving the answers to the participants. I don’t think this is the case, but it would be to verify.

Although the authors already lightly discuss it, I would like to know more about Task 4 where G2 participants spent more time on average than G1 participants. Was it because G1 participants were more likely to not finish that task, whereas G2 participants finished it?

I think one of the difficulties of facing this paper is that it all seems kind of obvious. Add examples to address the questions that participants are likely to have and unsurprisingly, participants do better. This obviousness could have been mitigated by providing some challenges faced by programmers that are unique to REST APIs or implications for design for this unique domain. As far as I can tell, the suggestions provided seem to be similar to the ones in the existing literature for local APIs. However, these are not sufficient reasons to reject the paper, so I would argue for accepting it.

----------------------- REVIEW 2 ---------------------
PAPER: 20
TITLE: A Study of the Effectiveness of Usage Examples in REST API Documentation
AUTHORS: S. M. Sohan, Frank Maurer, Craig Anslow and Martin Robillard

Overall evaluation: -1 (weak reject)
Reviewer's confidence: 5 ((expert))

----------- Review -----------
This paper describes a controlled experiment with 26 experienced
developers using a special-purpose web editor to create REST
calls. Some had usage examples as part of the documentation and some
did not. The people who did not did worse. It is good they they got
"26 participants from sixteen different companies and six different
countries".

It has been known for a long time that examples are helpful and are
preferred by developers, so that is not a new contribution. And the
control condition in this study was clearly-inadequate documentation,
so the fact that these people did worse is no surprise and not
particularly useful to know. It would have been much more valid if the
control condition had well-done, complete documentation. Many, many
previous papers have identified inadequate documentation as a frequent
complaint about APIs, so showing that the official documentation for
these REST APIs is incomplete is not interesting either. I didn't find
any of the Recommendations to be new or non-obvious.

Since you spend a lot of time comparing the two conditions (e.g.,
Figures 4,5), you need to do the statistical analysis to see if there
are actually any statistically significant differences. Any
quantitative comparison is not valid without that.

With respect to Recommdation 4: how many examples are necessary? It
seems like you are requesting that every combination of calls would
need an example. What TYPE of APIs examples are needed?

In related work section 4, you mention: "but it was challenging for
the participants to locate relevant examples" -- did you address this
in your study? If not, no need to mention it. Also, the study in [18]
doesn't seem relevant.

The relate work should include this study, even though it is about
SOAP APIs instead of REST:

Brad A. Myers, Sae Young Jeong, Yingyu Xie, Jack Beaton, Jeff Stylos,
Ralf Ehret, Jan Karstens, Arkin Efeoglu, Daniela K. Busse. "Studying
the Documentation of an API for Enterprise Service-Oriented
Architecture". The Journal of Organizational and End User Computing,
JOEUC, IGI Publishing. vol. 22, no. 1, Jan-Mar, 2010. pp. 23-51. IGI global,
DOI: 10.4018/joeuc.2010101903.

Low level writing issues:

This is a better motivation than what is put in the abstract: "API
developers need to know what to include in the examples and why" - but
you don't actually follow through on doing this.

In R4 on page 2, you say: "with no prior experince of using the
WordPress REST API V2" but you haven't said that is what you are using
yet. Maybe just say "the selected API" here, or better yet - introduce
the target API in the introduction.

typo: documentation of the the

missing word: are required to the same

typo: APIs and and a

Figure 3 - screen shot is too small - can't read.

6) Data Analysis: this section has too much detail, and the analysis
was done wrong anyway (see above).

page 5: "each other to provide consistency" -- better to say "to
evaluate consistency"

Figure 4a - I can't tell which vertical axis goes with which dataset -
are the bars or lines the "success rate"?

----------------------- REVIEW 3 ---------------------
PAPER: 20
TITLE: A Study of the Effectiveness of Usage Examples in REST API Documentation
AUTHORS: S. M. Sohan, Frank Maurer, Craig Anslow and Martin Robillard

Overall evaluation: 2 (accept)
Reviewer's confidence: 3 ((medium))

----------- Review -----------
In this paper, the authors present results from a controlled study they carried out to evaluate the impact of providing usage examples in REST API documentation. While generating and maintaining API documentation together with examples of usage can be expensive, it helps client developers work more efficiently while making fewer mistakes. The authors summarize their findings through:
1) A list of common obstacles client developers who use REST APIs encounter in the absence of usage examples.
2) A list of guidelines on incorporating usage examples in API documentation.
3) Empirical evidence that shows positive impact of usage examples on time taken and success rate for task completion, and overall developer satisfaction.

On page 6, you state "For T4, where the participants...rate." Any insights into why G2 executed more API calls for T4? Given the design of your custom REST API explorer shown in Figure 2, with the response header so prominently displayed, I am curious as to why participants in G2 had to execute more requests to be able to complete the task.

Since participants were allowed to access the internet, did you log those accesses? Any insights into the sort of information G1 and G2 sought through their web searches?

Minor corrections:
- Pg 1, col 2: Change "...obstacles API client..." to "...obstacles do API client..."
- Pg 2, col 1: Change "...images, tags." to "...images, and tags."
- Pg 3, col 1: Change "...requires to the..." to "...required to use the..."
- Pg 7, col 1: Change "Eight the..." to "Eight of the..."

### 2017-07-20 controlled_study.tex
adds all the required changes

@@
 \begin{itemize}
-  \item \textbf{RQ}. What obstacles API client developers face when using a REST API documentation that lacks usage examples?
+  \item \textbf{RQ}. What obstacles do API client developers face when using a REST API documentation that lacks usage examples?
 \end{itemize}
@@
 WordPress is a blog-like open-source (R2) framework used by over 409 million people to visit 23.6 billion pages each month \footnote{\url{https://wordpress.com/activity/}}.
-The API allows programmatic access to list, create, update, and delete WordPress data such as blog posts, comments, users, images, tags.
+The API allows programmatic access to list, create, update, and delete WordPress data such as blog posts, comments, users, images, and tags.
 
@@
 In the following paragraphs, we describe each task with its description and the study observation goals against the aforementioned API input and output information.
-For each of the tasks, the participants are required to the same WordPress REST API and one of the two variants of accompanying documentation.
+For each of the tasks, the participants are required to use the same WordPress REST API and one of the two variants of accompanying documentation.
 
@@
 The co-author applied the codes to 82/539 API calls (10\% confidence interval with 95\% confidence level).
-The two authors validated the codes against each other to provide consistency.
+The two authors validated the codes against each other to evaluate consistency.
 The initial coding from the co-author had discrepancies in 4 / 11 codes for 27 / 82 API call samples.
@@
 After analyzing the response, we found that P1.8 used the ISO8601 formatted date in the API calls, but did not provide the required \lstinline{second} portion of the time.
-Eight the 16 participants in G1 had an error in the date formats.
+Eight of the 16 participants in G1 had an error in the date formats.
 With an example, 1 of the 10 participants had the same error from G2.
@@
 They found that the participant group using a statically typed API with explicit documentation could get the tests to pass in less time than the other groups.
-Ellis et al.
-performed a controlled study with two groups of developers to understand the impact of using a constructor vs.
-a factory method design pattern on API client development time \cite{Ellis:2007:FPA:1248820.1248863}.
-They found that API client developers needed more time to complete API tasks using factory method compared to using a constructor.
 Our work in this paper is based on a controlled study and shares part of the setup that were used by the aforementioned controlled studies with the following differences: our goal is to understand the obstacles faced by API client developers without usage examples;  we focus on REST APIs instead of local APIs; and, the participants are all professional software engineers.

### 2017-07-20 build.sh

@@ -1,0 +1,4 @@
+pdflatex controlled_study.tex
+bibtex controlled_study.aux
+pdflatex controlled_study.tex
+pdflatex controlled_study.tex

### 2017-07-21 controlled_study/controlled_study.tex
adds controlled study to be included as a chapter

--- 2017-07-20 controlled_study.tex
@@
 \fi
+
+\makeatletter
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
+\def\ps@IEEEtitlepagestyle{%
+  \def\@oddfoot{\mycopyrightnotice}%
+  \def\@evenfoot{}%
+}
+\def\mycopyrightnotice{%
+  {\footnotesize 978-1-5386-0443-4/17/\$31.00 ©2017 IEEE\hfill}% <--- Change here
+  \gdef\mycopyrightnotice{}% just in case
+}
 
@@
         anchor=north,legend columns=-1},
-        ylabel=Average Trial API Calls (Frequency),
+        ylabel=Avg. Trial API Calls (Frequency),
         width=2.8in
@@
     \addplot[draw=black] table [col sep=comma, x expr=\coordindex, y=AverageAPICalls] {grouptwo.csv};
-    \legend{G1: Using Original, G2: Using Enhanced}
+    \legend{G1 Avg. Trial API Calls, G2 Avg. Trial API Calls}
     \end{axis}
@@
     \addplot[dashed,mark=*,gray] table [col sep=comma, x expr=\coordindex, y=SuccessRate] {groupone_success.csv};
-    \addlegendentry{G1: Using Original}
+    \addlegendentry{G1 Success Rate}
 
     \addplot[smooth,mark=*,black] table [col sep=comma, x expr=\coordindex, y=SuccessRate] {grouptwo_success.csv};
-      \addlegendentry{G2: Using Enhanced}
+      \addlegendentry{G2 Success Rate}
     \end{axis}
@@
         anchor=north,legend columns=-1},
-        ylabel=Average Time Spent (minutes),
+        ylabel=Avg. Time Spent (minutes),
         width=2.8in
@@
     \addplot[draw=black] table [col sep=comma, x expr=\coordindex, y=Time] {grouptwo_time.csv};
-    \legend{G1: Using Original, G2: Using Enhanced}
+    \legend{G1 Avg. Time Spent, G2 Avg. Time Spent}
     \end{axis}

### 2017-07-21 root.tex
adds controlled study to be included as a chapter

%% Modified documentstyle to documentclass -- compatibility with LaTex2e
%% N. Mancell (98/03/01).

%%  ucalgthes_root.tex        (NM 98/03/01)
%   Modified   92-09-18      Add references to dissertation        D. Teale
%                            Add approval page to toc
%                            Add ref to Title Degree on approval page
%   Modified  2006-09-12     Added geometry package to set up UofC thesis margins
%                            Removed includeprompt option N. Mancell
%
% Modified  2012-04-15  T.Zhang
% 1, the ‘fancy’ package is cancelled. The pagestyle is changed from ‘myheadings’ and ‘headings’ to ‘plain’ to move the page number (footer) from top right to bottom centre.
% 2, The parameter  of using ‘geometry’ package is changed from four different values (top=1in, bottom=1.22in, left=1.40in, right=0.850in) to one single value: 1in (top=1in, bottom= 1in, left= 1in, right= 1in)
% 3, ‘List of Figures’ is modified to ‘List of Figures and Illustrations’
% 4, A new page ‘List of Symbols, Abbreviations and Nomenclature’ is created. Its page number is following ‘List of Figures and Illustrations’ with Roman numerals. A separate file ‘symbols.tex’ is created for students to put content into the list.

\documentclass[12pt]{ucalgthes1}
\usepackage[letterpaper,top=1in, bottom= 1in, left= 1in, right= 1in]{geometry}
%\usepackage{fancyhdr}
%\fancyhead{}
%\fancyfoot{}
%\renewcommand{\headrulewidth}{0pt}
%\fancyhead[RO,LE]{\thepage}
%Define other usepackages here
\usepackage{hyperref}
\usepackage{mathptmx}
\title{Title of Thesis (title case, double spaced, no more than 240 characters. No bold font or symbols, except Greek) \\ \bigskip
Thesis Title Second Line }
%
%            Insert the correct information between the {}
%
\author{Student's Full Name}
\thesisyear{year}
\thesis{thesis}    % the word dissertation can be inserted between {}
\newcommand{\thesistitle}{Title of Thesis}
\monthname{month name}
\dept{GRADUATE PROGRAM IN "NAME OF PROGRAM"}
\degree{"FULL NAME OF DEGREE e.g.: DOCTOR OF PHILOSOPHY"}
%
%                    End of supplied information
%
\begin{document}
\makethesistitle
\pagenumbering{roman}     % resets page counter to one
\setcounter{page}{2}
%\chapter*{UNIVERSITY OF CALGARY \\ FACULTY OF GRADUATE STUDIES}
%\thispagestyle{empty}
%The undersigned certify that they have read, and recommend
%to the Faculty of Graduate Studies for acceptance, a \Thesis\ entitled
%``\thesistitle'' submitted by \Author\
%in partial fulfillment of the requirements for the degree of
%\Degree.\\

%
%                 Substitute  List of Examiners
%
%\begin{signing}{Department of Academic Computing}
%\signline
%Chairman, Dr.~John D.~Doe \\
%Department of Academic Computing \\
%Services  \\
%\signline
%Chairman, Dr.~John D.~Doe \\
%Department of Academic Computing \\
%Services  \\
%\signline
%Chairman, Dr.~John D.~Doe \\
%Department of Academic Computing \\
%Services  \\
%\signline
%Chairman, Dr.~John D.~Doe \\
%Department of Academic Computing \\
%Services  \\
%\newsigncolumn         use this command to start a new column if necessary
%\newsigncolumn
%\signline
%Chairman, Dr.~John D.~Doe \\
%Department of Academic Computing \\
%Services  \\
%\signline
%Dr.~Jane Smith \\
%Department of Academic Computing  \\

%\signline
%Dr.~A.~B.~Brown \\
%Department of Academic Computing  \\
%\end{signing}
%
\newpage
\phantomsection
\altchapter{\bf{Abstract}}
Paragraph 1

Paragraph 2

Paragragh 3

This page is required for all graduate theses.  An abstract is a short paragraph explaining the major points and conclusions of your thesis.  For master's theses, the abstract can be no more than 150 words long, while doctoral abstracts can be no longer than 350 words.

\newpage
\phantomsection
\altchapter{\bf{Acknowledgements}}
Paragraph 1

Paragraph 2

Paragraph 3

\begin{singlespace}
\newpage
\phantomsection
\tableofcontents
\pagestyle{plain}
\newpage
\phantomsection
\listoftables
\pagestyle{plain}
\newpage
\phantomsection
\listoffigures
\pagestyle{plain}
\clearpage
\clearpage          % otherwise tables will be numbered wrong
\end{singlespace}
\newpage
\phantomsection
\chapter*{\bf{List of Symbols, Abbreviations and Nomenclature}\hfill} \addcontentsline{toc}{chapter}{List of Symbols}
\listofsymbols
\pagestyle{plain}
\clearpage

\pagenumbering{arabic}
\include{chapter1}
%\include{chapter2}
%\include{chapter3}
%\include{chapter4}
%\include{chapter5}
\appendix
\include{appendix1}
\end{document}

### 2017-07-21 build.sh
remove unwanted files

@@ -1,4 +1,4 @@
-pdflatex controlled_study.tex
-bibtex controlled_study.aux
-pdflatex controlled_study.tex
-pdflatex controlled_study.tex
+pdflatex root.tex
+bibtex root.aux
+pdflatex root.tex
+pdflatex root.tex

### 2017-07-21 extract_code.rb

@@ -1,11 +1,0 @@
-Dir.glob("*/*/*task*-data.md").each do |file|
-  dir = File.dirname(file)
-  file_name = File.basename(file)
-
-  extrated_code_file = File.join(dir, "codes-from-#{file_name}")
-
-  command =  "grep -i \"Participant\\|\\`CODE\" #{file} > #{extrated_code_file}"
-
-  `#{command}`
-
-end

### 2017-07-21 raw_data_export.rb

@@ -1,16 +1,0 @@
-require 'csv'
-
-codes = ['Confirm Correct Answer', 'Correct Answer', 'Wrong Body', 'Wrong Data Type', 'Wrong Path', 'Missing Headers', 'Partially Correct Answer', 'Repeat Correct Answer', 'Wrong Array Format', 'Wrong Header', 'Wrong Method']
-
-csv_string = CSV.generate do |csv|
-
-  csv << ['Task', 'Participant', 'Method', 'Path', 'Headers', 'Body', 'Confirm Correct Answer', 'Correct Answer', 'Wrong Body', 'Wrong Data Type', 'Wrong Path', 'Missing Headers', 'Partially Correct Answer', 'Repeat Correct Answer', 'Wrong Array Format', 'Wrong Header', 'Wrong Method']
-
-  Task.where.not(id:1).each do |task|
-
-    task.answers.where(default: nil).order("RANDOM()").limit(17).each do |answer|
-
-      csv << [task.id, answer.participant.new_code_name, answer.method, answer.path, answer.request_headers, answer.request_body]
-    end
-  end
-end

### 2017-07-21 rename.rb

@@ -1,17 +1,0 @@
-map = [["P1", "P1.1"], ["P2", "P1.2"], ["P3", "P1.3"], ["P4", "P1.4"], ["P5", "P1.5"], ["P6", "P1.6"], ["P7", "P1.7"], ["P8", "P1.8"], ["P9", "P1.9"], ["P16", "P1.10"], ["P17", "P1.11"], ["P21", "P1.12"], ["P23", "P1.13"], ["P24", "P1.14"], ["P25", "P1.15"], ["P26", "P1.16"]]
-
-
-files = Dir.glob('*/*task-2-data*').each do |file|
-  map.each do |old, new_code|
-    `sed -i '' 's/#{old}\ using\ original/#{new_code}\ using\ original/' #{file}`
-  end
-end
-
-map = [["P10", "P2.1"], ["P11", "P2.2"], ["P13", "P2.4"], ["P14", "P2.5"], ["P15", "P2.6"], ["P18", "P2.7"], ["P19", "P2.8"], ["P20", "P2.9"], ["P22", "P2.10"], ["P12", "P2.3"]]
-
-files = Dir.glob('*/*task-2-data*').each do |file|
-  map.each do |old, new_code|
-    `sed -i '' 's/#{old}\ using\ forked/#{new_code}\ using\ forked/' #{file}`
-  end
-end
-

### 2017-07-21 controlled_study/controlled_study.tex
remove unwanted files

@@
-\documentclass[conference]{IEEEtran}
+%\documentclass[conference]{IEEEtran}
 
-\ifCLASSINFOpdf
-  \usepackage{listings}
-  \usepackage[pdftex]{graphicx}
-  \usepackage{url}
-  % \usepackage{caption}
-  \usepackage{subcaption}
-  \usepackage{tikz,pgfplots,pgfplotstable}
-  \usepackage{mathtools}
-  \usepgfplotslibrary{statistics}
-  \usepackage{courier}
-\else
-\fi
+%\ifCLASSINFOpdf
+  %\usepackage{listings}
+  %\usepackage[pdftex]{graphicx}
+  %\usepackage{url}
+  %\usepackage{subcaption}
+  %\usepackage{tikz,pgfplots,pgfplotstable}
+  %\usepackage{mathtools}
+  %\usepgfplotslibrary{statistics}
+  %\usepackage{courier}
+%\else
+%\fi
 
-\makeatletter
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-\def\ps@IEEEtitlepagestyle{%
-  \def\@oddfoot{\mycopyrightnotice}%
-  \def\@evenfoot{}%
-}
-\def\mycopyrightnotice{%
-  {\footnotesize 978-1-5386-0443-4/17/\$31.00 ©2017 IEEE\hfill}% <--- Change here
-  \gdef\mycopyrightnotice{}% just in case
-}
+%\makeatletter
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
+%\def\ps@IEEEtitlepagestyle{%
+  %\def\@oddfoot{\mycopyrightnotice}%
+  %\def\@evenfoot{}%
+%}
+%\def\mycopyrightnotice{%
+  %{\footnotesize 978-1-5386-0443-4/17/\$31.00 ©2017 IEEE\hfill}% <--- Change here
+  %\gdef\mycopyrightnotice{}% just in case
+%}
+%\begin{document}
+%\title{A Study of the Effectiveness of Usage Examples in REST API Documentation}
+\chapter{A Study of the Effectiveness of Usage Examples in REST API Documentation}
 
-\begin{document}
-\title{A Study of the Effectiveness of Usage Examples in REST API Documentation}
+ %\author{\IEEEauthorblockN{S M Sohan, Frank Maurer\IEEEauthorrefmark{1}}
+ %\IEEEauthorblockA{\IEEEauthorrefmark{1}
+ %Dept. of Computer Science\\
+ %University of Calgary\\
+ %Canada\\
+ %\{smsohan, frank.maurer\}@ucalgary.ca}
+ %\and
+ %\IEEEauthorblockN{Craig Anslow}
+ %\IEEEauthorblockA{School of Eng. and Computer Science\\
+ %Victoria University of Wellington\\
+ %New Zealand\\
+ %craig@ecs.vuw.ac.nz}
+ %\and
+ %\IEEEauthorblockN{Martin P. Robillard}
+ %\IEEEauthorblockA{
+ %School of Computer Science\\
+ %McGill University\\
+ %Canada\\
+ %martin@cs.mcgill.ca}
+ %}
 
- \author{\IEEEauthorblockN{S M Sohan, Frank Maurer\IEEEauthorrefmark{1}}
- \IEEEauthorblockA{\IEEEauthorrefmark{1}
- Dept. of Computer Science\\
- University of Calgary\\
- Canada\\
- \{smsohan, frank.maurer\}@ucalgary.ca}
- \and
- \IEEEauthorblockN{Craig Anslow}
- \IEEEauthorblockA{School of Eng. and Computer Science\\
- Victoria University of Wellington\\
- New Zealand\\
- craig@ecs.vuw.ac.nz}
- \and
- \IEEEauthorblockN{Martin P. Robillard}
- \IEEEauthorblockA{
- School of Computer Science\\
- McGill University\\
- Canada\\
- martin@cs.mcgill.ca}
- }
+%\maketitle
 
-\maketitle
-
-\begin{abstract}
+\section{abstract}
 Generating and maintaining REST API documentation with usage examples can be a time consuming and expensive process for evolving APIs.
@@
 
-\end{abstract}
+%\begin{IEEEkeywords}
+%API; REST; Documentation; Usage Examples; Empirical Study; Controlled Study; Productivity;
 
-\begin{IEEEkeywords}
-API; REST; Documentation; Usage Examples; Empirical Study; Controlled Study; Productivity;
+%\end{IEEEkeywords}
 
-\end{IEEEkeywords}
-
-\IEEEpeerreviewmaketitle
+%\IEEEpeerreviewmaketitle
 
@@
   \begin{center}
-\fbox{\includegraphics[width=\linewidth,keepaspectratio]{schema.png}}
+    \includegraphics[width=\linewidth,keepaspectratio]{schema.png}
   \end{center}
@@
 
-\end{document}
+%\end{document}

### 2017-07-21 build.sh
controlled study paper is wholly included

@@ -1,4 +1,7 @@
+set -e
+git clean -fX
 pdflatex root.tex
 bibtex root.aux
 pdflatex root.tex
 pdflatex root.tex
+skim root.pdf

### 2017-07-21 controlled_study/controlled_study.tex
controlled study paper is wholly included
boxes look good

@@
 
-\section{abstract}
+\section{Abstract}
 Generating and maintaining REST API documentation with usage examples can be a time consuming and expensive process for evolving APIs.
@@
 
-\begin{figure}[h!]
+\begin{figure}
   \begin{center}
-    \includegraphics[width=\linewidth,keepaspectratio]{schema.png}
+    \fbox{\includegraphics[width=\linewidth,keepaspectratio]{schema.png}}
   \end{center}
@@
 \begin{table}[!tbh]
+  \centering
   \caption{Participant Profile}
@@
 
-\begin{tabular}{|p{0.8in} p{.6in} p{0.7in} p{0.7in}|}
+  \begin{tabular}{|p{1.5in} p{1in} p{1.5in} p{1.5in}|}
 
@@
 
-\begin{subfigure}[t]{4in}
+\begin{subfigure}[t]{5in}
 \centering
-\includegraphics[width=4in]{original.png}
+\includegraphics[width=5in]{original.png}
 \caption{Original API Documentation}
@@
 ~
-\begin{subfigure}[t]{3in}
+\begin{subfigure}[t]{5in}
 \centering
-\includegraphics[width=2.7in]{forked.png}
+\includegraphics[width=5in]{forked.png}
 \caption{API Documentation Enhanced with an Example}
@@
 
+\subsubsection{Data Analysis}
+We analyzed the data artifact qualitatively to understand the obstacles that REST API client developers face when using an API documentation that lacks usage examples.
+We exported a table of data from the web-based REST API explorer.
+For each row, the table contains the following columns: API task, participant identifier, timestamp, API trial number, I1-4, response headers and body.
+The values for I1-4 used by the participants were manually coded to group the participant responses into categories.
+The categories help us determine the information type that API client developers need in the API documentation to perform API tasks.
+The analysis started with an empty set of codes and new codes were introduced to describe scenarios that did not fall under existing codes.
+The first author applied the codes on the raw data artifact and provided a coding scheme comprising 11 codes to a co-author.
+The co-author applied the codes to 82/539 API calls (10\% confidence interval with 95\% confidence level).
+The two authors validated the codes against each other to evaluate consistency.
+The initial coding from the co-author had discrepancies in 4 / 11 codes for 27 / 82 API call samples.
+After discussing the coding scheme, the discrepancies were resolved as the definition of the terms became clearer.
+
+For each API task attempted by each participant, we annotated the artifact with one of the following labels: successful, partially successful, and unsuccessful.
+Task participations are marked successful when I1-4 matches the required values for performing the given task.
+If a participant is able to use the correct I1-4 for one of the two tasks required to complete a single task (T2), we marked it as partially successful.
+Otherwise, it's marked as unsuccessful.
+Based on these annotations, the following formulas were used to compute the quantitative results:
+
+\textbf{Success Rate (Task, Group)} = (No. of participants that successfully completed Task) / (No. of participants in Group)
+
+\textbf{Average Trial API Calls (Task, Group)} = (No. of API calls made by Group on Task) / (No. of participants in Group).
+
+\textbf{Average Time Spent (Task, Group)} = (Time spent by Group on Task) / (No. of participants in Group)
+
 \pgfplotsset{compat=1.14}
 \usetikzlibrary{patterns}
-\pgfplotstableread[col sep=comma]{groupone.csv}\groupone
+\pgfplotstableread[col sep=comma]{controlled_study/groupone.csv}\groupone
 
-\begin{figure*}[htb]
+\begin{figure*}[!h]
 
@@
 
-\begin{subfigure}[b]{0.5\textwidth}
+\begin{subfigure}[b]{\textwidth}
 \centering
@@
 \end{subfigure}
-\begin{subfigure}[b]{0.4\textwidth}
+\begin{subfigure}[b]{\textwidth}
 \centering
@@
 \end{figure*}
-
-\subsubsection{Data Analysis}
-We analyzed the data artifact qualitatively to understand the obstacles that REST API client developers face when using an API documentation that lacks usage examples.
-We exported a table of data from the web-based REST API explorer.
-For each row, the table contains the following columns: API task, participant identifier, timestamp, API trial number, I1-4, response headers and body.
-The values for I1-4 used by the participants were manually coded to group the participant responses into categories.
-The categories help us determine the information type that API client developers need in the API documentation to perform API tasks.
-The analysis started with an empty set of codes and new codes were introduced to describe scenarios that did not fall under existing codes.
-The first author applied the codes on the raw data artifact and provided a coding scheme comprising 11 codes to a co-author.
-The co-author applied the codes to 82/539 API calls (10\% confidence interval with 95\% confidence level).
-The two authors validated the codes against each other to evaluate consistency.
-The initial coding from the co-author had discrepancies in 4 / 11 codes for 27 / 82 API call samples.
-After discussing the coding scheme, the discrepancies were resolved as the definition of the terms became clearer.
-
-For each API task attempted by each participant, we annotated the artifact with one of the following labels: successful, partially successful, and unsuccessful.
-Task participations are marked successful when I1-4 matches the required values for performing the given task.
-If a participant is able to use the correct I1-4 for one of the two tasks required to complete a single task (T2), we marked it as partially successful.
-Otherwise, it's marked as unsuccessful.
-Based on these annotations, the following formulas were used to compute the quantitative results:
-
-\textbf{Success Rate (Task, Group)} = (No. of participants that successfully completed Task) / (No. of participants in Group)
-
-\textbf{Average Trial API Calls (Task, Group)} = (No. of API calls made by Group on Task) / (No. of participants in Group).
-
-\textbf{Average Time Spent (Task, Group)} = (Time spent by Group on Task) / (No. of participants in Group)
 
@@
 
+\vspace{10mm}
 \hspace{-5mm}
@@
 }
-\vspace{0.5mm}
+\vspace{10mm}
 
@@
 
-\vspace{1mm}
+\vspace{10mm}
 \hspace{-5mm}
@@
 \bibliographystyle{IEEEtran}
-\bibliography{IEEEabrv,refs}
-
-%\end{document}
+\bibliography{controlled_study/IEEEabrv,controlled_study/refs}

### 2017-07-21 build.sh

@@ -1,7 +1,7 @@
 set -e
 git clean -fX
 pdflatex root.tex
-bibtex root.aux
+bibtex controlled_study/controlled_study.aux
 pdflatex root.tex
 pdflatex root.tex
-skim root.pdf
+/Applications/Skim.app/Contents/MacOS/Skim root.pdf

### 2017-07-21 case-study/change_frequency.tex
DRY up

% \begin{tikzpicture}
%   \begin{axis}[
%     title=\dataSourceName\ API Changes,
%     date coordinates in=x,
%     stack plots=y,
%     area style,
%     xtick = data,
%     xlabel={Release Dates},
%     ylabel={No. of Changes},
%     xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%     enlarge x limits=false,
%     height=0.25\textheight,
%     width=0.9\linewidth,
%     legend style={legend pos=north west,font=\tiny, legend columns=2}
%     ]
    \addplot[fill=green] table[x=Date, y=AddAPIElements]{plotdata/\dataSourceFileName}
    \closedcycle;
    \addplot[fill=red] table[x=Date, y=RemoveAPIElements]{plotdata/\dataSourceFileName}
    \closedcycle;
    \addplot[fill=orange] table[x=Date, y=ChangeAPIElements]{plotdata/\dataSourceFileName}
    \closedcycle;
    \addplot[fill=gray] table[x=Date, y=ChangeFieldDataType]{plotdata/\dataSourceFileName}
    \closedcycle;
    \addplot[fill=violet] table[x=Date, y=ErrorConditionChange]{plotdata/\dataSourceFileName}
    \closedcycle;
    \addplot[fill=yellow] table[x=Date, y=ChangeAuthentication]{plotdata/\dataSourceFileName}
    \closedcycle;
    \addplot[fill=magenta] table[x=Date, y=BehaviorChange]{plotdata/\dataSourceFileName}
    \closedcycle;
    \addplot[fill=blue] table[x=Date, y=ChangeAuthorization]{plotdata/\dataSourceFileName}
    \closedcycle;

    \legend{AddAPIElements, RemoveAPIElements, ChangeAPIElements,
        ChangeFieldDataType, ErrorConditionChange,
        ChangeAuthentication, BehaviorChange, ChangeAuthorization}

%   \end{axis}
% \end{tikzpicture}

### 2017-07-21 case-study/icws_case_study_api_evolution_sohan.tex
DRY up

\documentclass[conference]{IEEEtran}

\ifCLASSINFOpdf
\else
\fi
\usepackage{url}
\usepackage[hypcap]{caption}
\usepackage{listings}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{eurosym}
\usepackage{multirow}
\usepackage{subfigure}

%\usepackage{caption}
%\usepackage{subcaption}

\usetikzlibrary{backgrounds}
\pgfplotsset{compat=1.5}
\usepgfplotslibrary{dateplot}

\begin{document}
\title{A Case Study of Web API Evolution}

\author{\IEEEauthorblockN{S M Sohan, Craig Anslow, Frank Maurer
\IEEEauthorblockA{Department of Computer Science\\
University of Calgary,
Calgary, Alberta T2N 1N4\\ Email: \{smsohan, craig.anslow, frank.maurer\}@ucalgary.ca }
}}

\maketitle

\begin{abstract}
When applications are integrated using web APIs, changes on a web API may break the dependent applications. This problem exists because old versions of the APIs may no longer be supported, a lack of adequate documentation to upgrade to a newer version, and insufficient communication of changes. In this paper we conducted a case study of evolving Web APIs to investigate what changes are made between versions and how the changes are documented and communicated to the API users. The findings are a list of recommendations for practitioners and researchers based on API change profiles, versioning, documentation and communication approaches that are observed in practice. This study will help inform developers of evolving Web APIs to make decision about versioning, documentation and communication methods.
\end{abstract}

\begin{IEEEkeywords}
Case Study; RESTful, SOAP, Web API Evolution; WSDL
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle

\section{Introduction}

Web APIs are used as a key interconnectivity mechanism to access software services over the Internet. Interconnected applications can provide better service at a lower cost to their users. For example, a business search portal can display the businesses near a user's current location on a map by using the Google Maps API. The actual implementation of Web APIs can follow various protocols, such as SOAP\footnote{http://www.w3.org/TR/soap/} and RESTful\footnote{\url{http://www.ics.uci.edu/~fielding/pubs/dissertation/rest_arch_style.htm}} web services. In this paper, the term Web API is used to describe any protocol unless otherwise specified.

Web APIs connect independently maintained applications. As a result, when a Web API evolves, all the integrated applications using the Web API may not be able to evolve on the same schedule. For example, in the case of the aforementioned business search portal, if the Google Maps API changes, it may also require the portal to change. This introduces unique challenges to the evolution of Web APIs compared because the evolution is out of the control of the API users. For example, API users of a local API can use a copy of an older version, but for evolving Web APIs, API users are forced to upgrade if older versions are no longer supported. Documentation of evolving Web APIs is an important topic for case study because of the unique challenges it presents. Web APIs are defined by their HTTP interface, including both HTTP headers as well as request and response data formats. As a result, for an evolving Web API, its documentation needs to indicate changes in any of these fields, unlike the local APIs, where typically no HTTP specific information is involved.

The primary goal of this research is to investigate the challenges that are currently faced by developers and users of evolving Web APIs so that future research can be performed to address the problems. To this regard, we analyzed existing literature on the topic of Web API evolution. We analyzed the evolution of multiple Web APIs to understand the current industry trends and challenges. The juxtaposition of current industry practices with the literature reveals insights that are otherwise unseen when viewed from a single perspective. We use the differences between the literature and industry practices to identify unresolved research problems.

In this case study, multiple Web APIs representing a variety of application areas are selected, analyzed and compared to gain a broader perspective about current industry practices on Web API versioning, documentation and communication of changes. The result of this analysis is a summary of different approaches to Web API evolution, and a list of recommendations for developers of evolving Web APIs.

The remainder of this paper is organized as follows: the next section is related work on Web API evolution. Then, we discuss the case study of multiple Web API evolution. In the following section, we list our findings, followed by a discussion about the findings. Finally, we conclude with the contributions of this work.

\section{Related Work}

We look at related work in the following areas: Empirical Research on Web API evolution, techniques to evolve Web APIs and, tool support for evolving Web APIs.

\subsection{Empirical Research on Web API evolution} % (fold)
\label{sub:empirical_research}
Several case study papers were published on Web API Evolution. Maleshkova et al. analyzed multiple Web APIs to identify the different approaches related to Web API descriptions, data formats, protocols, reusability, granularity, and authentication \cite{maleshkova}. They found that a lack of a standard format to document Web APIs and manual documentation led to API under-specification causing confusions about how to use the APIs for different use cases. They identified a need for automated approaches.

Fokaefs et al. performed an empirical study on Web API evolution involving multiple SOAP web services, Amazon EC2, FedEx Rate, Bing, Paypal, and Fedex Pack., and provided an evolution profile showing the percentage of changes between two consecutive versions of the Web APIs \cite{fokaefs_2011_empirical}. The same Web APIs were studied by Romano et at., but they provided an alternate evolution profile of the Web APIs. Instead of using percentage of change, they profiled the number of operations, parts and XML file elements from WSDL files for each version of the Web APIs \cite{wsdl_diff_2012}. These change profiles point to the ever changing nature of Web APIs.

Jun et al. performed case studies on the evolution of Web APIs and compared the evolution of Web APIs against local library APIs \cite{li_client_2013}. For studying Web APIs, a selection criteria was provided as follows: Web APIs with large number of clients, covering different application areas, owned by different companies in different countries, and well documented API reference and migration guides. They extracted 16 patterns to describe different types of changes on the Web APIs as follows: add or remove parameter, change type of return value, delete method, rename method, rename parameter, change format of parameter, change format of return value, change XML tag, combine methods, split method, expose data, unsupport request method, change default value of parameter, chance default value of parameter, change upper bound of parameter, restrict access to API. They also presented six challenges specific to Web API evolution that are not found with library APIs such as: transformation between JSON and XML, M to N Mapping, delete method, authorization protocol change, rate limit, and authorization of API access.

Espinha et al. looked at the growing pains with Web API evolution from the perspective of Web API users by conducting interviews \cite{espinha}. The interviewees were selected based on the activity of their open-source projects that used Web APIs. They found that when a Web API evolves, maintaining the integration with a Web API took more effort than integrating it the first time since maintenance is carried out over a longer period of time. This makes the evolution of Web APIs an important subject for a case study.

We distinguish our case study from these case studies as follows: we provide up-to-date information and include a diverse set of Web APIs. Our analysis uses new data sources such as API user forums, and question and answer sites. We focused this study on different approaches to versioning, documentation, and communicating changes as an API evolves.
% subsection empirical_research (end)

\subsection{Techniques to evolve Web APIs} % (fold)
\label{sub:techniques}
Kaminski et al. presented a design technique called ``Chain of Adapters'' to implement evolving Web APIs \cite{kaminski2006design}. Using this approach, the source code of a Web API can evolve from a version to the next by introducing an adapter on the original version to implement any required change for new versions. This technique allows for multiple versions to be deployed concurrently since older versions are left unchanged.

Leskey suggests following a consistent and meaningful naming scheme to identify different versions of APIs \cite{laskey2008considerations}. Such a naming scheme can be used to identify if two versions of an API are compatible. Given the URL scheme, Leskey suggests a self documenting Web API, one that exposes API endpoints to describe the details for each version.

Treiber et al. showed a conceptual model to capture the changes of a Web API \cite{treiber2009analyzing}. A graphical representation to display the evolution of SOAP Web Services is presented by Aversano et al. where the differences between versions are highlighted based on their meta information found in the WSDL files \cite{aversano2005visualizing}.

Some papers focused on automatic migration of Web API clients to newer versions. For example, Wilde proposed a conceptual framework where WSDL can be extended so a client can be compatible with potential Web API changes \cite{wilde2004semantically}. A similar approach is proposed by Fang et al. where a fixed set of new XML nodes are used instead of arbitrary extensions \cite{fang2007version}. Juric et al. proposed a number of versioning related extensions to WSDL that can be applied to version SOAP web services in multiple levels of granularity such as, service level and operation level \cite{Juric20091326}. Zuo et al. developed a formal XML based change specification format to describe the evolution of SOAP Web services \cite{zuo6928972}.

For RESTful Web APIs, Mangler et al. introduced an XML based language called RIDDL \cite{mangler2010origin}. RIDDL allows incremental composition of Web API documentation for different versions by adding a changelog to the documentation of an old version.

VRESCo demonstrates an approach where a router between the Web API client and server is used to automatically route Web API calls to new API versions. The router routes API calls to the desired version based on free form version names such as: latest, stable, and fixed version \cite{leitner2008end}. Meng et al. compared different approaches for evolving Web APIs across multiple criteria: granularity of evolution, terminal of evolution, type of evolution, scalability, and maintainability \cite{6835564}. They identified an aspect oriented approach, where an intermediate tier between a Web API and its clients is used similar to VRESCo, can reduce coupling between the two.

Existing techniques to evolve Web APIs focused on finding different techniques for implementing evolving SOAP based Web APIs that cannot be readily applied to RESTful Web APIs. From our case study, we identified future research opportunities on techniques to solve the problems Web API evolution that are also applicable to RESTful APIs.

\subsection{Tool support for evolving Web APIs} % (fold)
\label{sub:tool_development}

%{|p{0.2\linewidth}|p{0.3\linewidth}|p{0.2\line width}|p{0.2\linewidth}|}
%& &  \multicolumn{4}{ c |}{\textbf{Categories}} & & \\ \cline{3-6}
%\multirow{2}{*}{\textbf{Web API}}  & \multirow{2}{*}{\textbf{URL}} &\textbf{Industry} & \textbf{Popularity} &  \textbf{Maturity} & \textbf{Code}& \multirow{2}{*}{\\textbf{Rel.}} & \multirow{2}{*}{\\textbf{Dates}}\\

\begin{table*}[!ht]
\caption{Case Study - Representative Sample of Web APIs.}
\label{web_api_grid}
\begin{tabular} {| l | l  | l  | l | l | l | l  | l |}
\hline
\multirow{2}{*}{\textbf{Web API}} &
\multirow{2}{*}{\textbf{URL}} &
\multicolumn{4}{ c |}{\textbf{Categories}} &
\textbf{Num of} &
\multirow{2}{*}{\textbf{Dates}}\\    \cline{3-6}
  & & Industry & Popularity &  Maturity & Code& \textbf{Releases}& \\
\hline
Facebook Platform API & \tiny{\url{https://developers.facebook.com/docs/apps/versions}} & Social & High & Est & Closed & 11 & 01/2013-07/2014\\
\hline
Twitter REST API & \tiny{\url{https://dev.twitter.com/docs/api}} & Social & High & Est & Closed & 2 & Not Available\\
\hline
Wordpress REST API & \tiny{\url{http://wp-api.org/}} & Business & Growing &  New & Open & 8 & 08/2013-06/2014\\
\hline
Salesforce \& Chatter REST API & \tiny{\url{https://www.salesforce.com/us/developer/docs/api_rest/}} & Business & Popular & Est & Closed & 4 & 01/2013-08/2014\\
\hline
Google Calendar API & \tiny{\url{https://developers.google.com/google-apps/calendar/}} & Business & Popular & Est & Closed & 2 & Not Available\\
\hline
Stripe Payment Processor API & \tiny{\url{https://stripe.com/docs/api}} & Business & Growing & New & Closed & 30 &09/2011-10/2014 \\
\hline
Github API & \tiny{\url{https://developer.github.com/}} & Productivity & Growing & New & Closed  & 59 &09/2012-10/2014\\
\hline
Google Maps API & \tiny{\url{https://developers.google.com/maps/web/}} & Productivity & Popular & Est & Closed & 102 &01/2011-09/2014\\
\hline
OpenStreetMap API &\tiny{\url{http://wiki.openstreetmap.org/wiki/API}} & Productivity & Growing & New & Open & 1 & Not Available\\
\hline
\end{tabular}
\end{table*}

Analyzing the available tool support for evolving Web APIs helped us identify opportunities for future research in this area. WSDarwin introduces a layer of adapters on both the Web API and its client to automatically migrate Web API clients to use a newer version of an API \cite{WSDarwin}. WSDarwin uses WSDL description of the SOAP service to automatically infer the changes between versions and uses preset default values for newly added objects to adapt the client request. VTracker, predecessor of WSDarwin, produces a changelog between versions of WSDL files \cite{fokaefs_2011_empirical}. WSDLDiff extends this by incorporating the semantic meaning of WSDL with an XML Schema Definition into their changelog computation \cite{wsdl_diff_2012}.

To identify if a new version of a Web API has any impact on a specific API client,  Zou et al. show a technique to generate client specific change logs \cite{le2008synchronizing}. A customized per API client change log is generated by filtering the change log based on  recorded API usage data so that only relevant API changes can be communicated effectively to each user.

Existing research on tool support for RESTful Web API evolution is rather limited. hRESTS is an HTML-based specification for describing RESTful Web APIs that can be used by computers to auto-generate client code \cite{4740521}. RESTdesc is another specification for describing RESTful Web API documentation where APIs are described using pre and post conditions \cite{RESTdesc}. However, the input to both hRESTS and RESTDesc require manual effort to produce the desired specification.

In the literature, we found multiple tools exist that rely on WSDL files. WSDL solutions have limited applicability for RESTful Web APIs since WSDL files cannot be used to fully describe RESTful APIs. Based on our case study, we will discuss future research problems on tool development to support the challenges with RESTful Web API evolution.

\section{Case Study}
When companies evolve their Web APIs in practice the new changes can often break applications that rely on these APIs. The aim of this case study is to investigate current industry practices on Web API Evolution to answer the following research questions.

\subsection{Research Questions} % (fold)

\begin{description}
  \item[RQ1] Versioning: How are Web APIs versioned as they evolve?
  \item[RQ2] Documentation: How are Web API changes documented as they evolve?
  \item[RQ3] Communication: How are Web API changes communicated as they evolve?
\end{description}

\subsection{Selection Strategy} % (fold)
\label{sub:selection_strategy}

To gain a broad perspective, we selected evolving Web APIs to represent each of the following categories as shown in Table \ref{web_api_grid}.

\textbf{Industry} - APIs representing different domains: social, business, and productivity.

\textbf{Popularity} - APIs with different number of users.

\textbf{Maturity} - Newer vs. established Web APIs.

\textbf{Source code} - Open source vs. closed source.

\subsection{Research Method - Qualitative Analysis} % (fold)
\label{sub:research_method}

The APIs in this case study are proprietary and open-source evolving Web APIs. For each API, we collected publicly available data from the following sources: API home page, formal API reference, and user forums and question-answer sites. A homepage provides a general introduction to the Web API. The formal API reference described the different elements of the API. The user forums and question-answer sites are used to communicate news and collect feedback about the Web APIs, for example StackOverflow.com.

The first author of this paper read the the API homepages and followed relevant hyperlinks to find answers to RQ1. A summary of the strategies were presented to the co-authors, which was used to categorize Web API versioning strategies into groups. Web API homepages contain changelogs and links to API references which were used to find answers to RQ2.

We used codes to annotate the API changelogs to identify the different types of API changes. The coding scheme started with a set of codes based on the codes from Jun et al. \cite{li_client_2013}. New codes  emerged from our analysis as the existing codes did not describe an API change scenario. An overview of the used codes is provided in Table \ref{coding_scheme}. The codes were applied to the different API elements as described through the following example Web API description:

``A Web API to post a comment on a photo.''

\textbf{Endpoint} - URLs to access the the API.
\textbf{Resource} - An input or output object. For example, comment, photo, etc.
\textbf{Method} - An API operation on a resource. e.g. post a comment.
\textbf{Field} - An attribute of the resource. e.g. timestamps of a comment.
\textbf{FieldValue} - Possible value for a field. For example: a timestamp value of January 1, 2015.
\textbf{Error} - Error cases. e.g. not found error.
\textbf{Authentication} - Authentication methods. e.g. HTTP Basic authentication.

\begin{table*}[!tbh]
\caption{List of Codes}
\label{coding_scheme}
\begin{tabular}{|p{0.2\linewidth}|p{0.7\linewidth}|}
\hline
\textbf{Category} & \textbf{Description}\\
\hline
Add$<$APIElements$>$ & New API elements introduced: Adds a new resource, field, field value, endpoint or method\\
\hline
Remove$<$APIElements$>$ & API elements completely removed: Removes a resource, field, field value, endpoint or method\\
\hline
Change$<$APIElements$>$ & Existing API element is replaced by another: Replaces or renames a resource, field, field value, endpoint or method\\
\hline
ErrorConditionChange & Changes the possible error conditions: New, changed or removed error response \\
\hline
ChangeFieldDataType & Changes the data type of a field: Single item to an array or vice versa, number to strings or vice versa, required to optional or vice versa\\
\hline
BehaviorChange & Changes the outcome or fixes bug: Uses new formula for computing API return values\\
\hline
ChangeAuthentication & Changes how API clients are authenticated: New, changed or deprecated API authentication methods\\
\hline
ChangeAuthorization & Changes who can access which API elements: Add, remove or modify permissions of API user\\
\hline
\end{tabular}
\end{table*}

These codes were applied to the text from the Web API homepages and API changelogs for multiple versions to produce a change profile. The first and second authors of this paper arrived at the same conclusion when they independently applied the  codes on 18 snippets of changelogs randomly sampled from the studied Web APIs. The coding process was manual since the data comprises of unstructured text that are not suitable for automated tagging. Here is an example of a change log from WordPress API with the codes:

\small
``Add post meta endpoints \textbf{(AddEndpoint)}.
Creating, reading, updating and deleting \textbf{(AddMethod * 4)}
post meta \textbf{(AddResource)} is now possible by
using \url{the/posts/<id>/meta} endpoints.''
\normalsize

We analyzed the API user forums and newsboards to find answers to RQ3. Messages in these sources were filtered to focus on the evolution of Web APIs. These data sources provided the perspectives of both the developers and users of Web APIs. To find evolution related messages, based on trial search runs, full text search features on the forums were used to search for these keywords: version, release, migration, breaking, deprecate, change.

\section{Findings} % (fold)
\label{sec:findings}

In this section we present our findings for RQ1-3.

\subsection{RQ1 Versioning} % (fold)
\label{sub:versioning}

\subsubsection{Patterns of Changes}

Based on our coding scheme, we identified the following new change patterns to describe changes between versions of Web APIs in addition to the 16 patterns described by Jun et al. \cite{li_client_2013} as follows:

\textbf{Move API elements} - We found 114 occurrences of element moves, where API elements are moved under new hierarchy or renamed. For example, Stripe API moved two objects under a new API element as found in the following change log from version 2013-08-13:

\small
\begin{quotation}
``Remove fee and fee\_details properties on charge and transfer objects. Instead, fee information is now stored on the corresponding balance transaction.
''\end{quotation}
\normalsize

\textbf{Rename API elements} - We found that 31 API elements were renamed for consistency. The following example from the Facebook API shows a rename \footnote{\url{https://developers.facebook.com/docs/apps/migrations/ads-api-changes-2014-07-02}}:

\small
\begin{quotation}
``Facebook will be renaming the adgroup\_status value ADGROUP\_PAUSED TO PAUSED to be consistent with the rest of the object's APIs.
''\end{quotation}
\normalsize

\textbf{Behavior change} - We have found 247 examples where Web APIs changed the resultant data while keeping the API interfaces intact. These are commonly triggered by bug fixes. For example, Wordpress REST API 1.1 changed as follows \footnote{\url{http://make.wordpress.org/core/2014/06/23/json-rest-api-version-1-1/}}:

\small
\begin{quotation}
``Correct password-protected post handling.
Password-protected posts could previously be exposed to all users, however
could also have broken behavior with excerpts. Password-protected posts are
now hidden to unauthenticated users..
''\end{quotation}
\normalsize

Another example that demonstrates a behavior change can be found from the 2013-10-29 version of Stripe API \footnote{\url{https://stripe.com/docs/upgrades#2014-01-31}}:

\small
\begin{quotation}
``When we apply a \$Y coupon to a \$X dollar invoice, we are no longer applying the remainder of the coupon to the account balance if Y $>$ X. Applications of coupons to \$0 invoices will no longer count as a redemption of the coupon.''
\end{quotation}
\normalsize

\textbf{Post condition change} - In the studied Web APIs we found examples where the immediate result of an API call remained unchanged but new post conditions are imposed requiring additional work for the API users. The following change on Facebook platform API illustrates this \footnote{\url{https://developers.facebook.com/docs/apps/migrations/payments-disputes-realtime-updates-required-2014-05-13}}:

\small
\begin{quotation}
``Starting May 13th, 2014, developers that accept payments will be required to subscribe to and honor Realtime Updates to ensure order fulfillment and appropriate handling of disputes from all payers. If you do not subscribe to and honor these updates, Facebook reserves the right, under our Developer Payment Terms to withhold payouts and/or stop your app from accepting payments.
''\end{quotation}
\normalsize

\textbf{HTTP header change} - Web APIs also change the custom HTTP headers that are used as meta data. For example, the following shows a change when the WordPress Web API introduced two new headers to describe the pagination status of an API call \footnote{\url{http://make.wordpress.org/core/2013/09/12/json-rest-api-version-0-5/}}:

\small
\begin{quotation}
``Send X-WP-Total and X-WP-TotalPages headers for information on post/pagination counts''
\end{quotation}
\normalsize

Similarly, the Github API changed their custom headers about pagination in the following example \footnote{\url{https://developer.github.com/v3/versions/}}:

\small
\begin{quotation}
``No longer using the X-Next or X-Last headers. Pagination info is returned in the Link header instead.''
\end{quotation}
\normalsize

\textbf{Error condition change} - Web APIs change the error conditions that may require the users to adapt their integration. For example, in Salesforce API Version 29 the error codes were changed \footnote{\url{https://na1.salesforce.com/help/pdfs/en/salesforce_winter14_release_notes.pdf#rn_186_api_objects}}:

\small
\begin{quotation}
``...different ExceptionCode and StatusCode values are now returned for some error conditions
when saving user records. We strongly recommend that you test your code and modify it accordingly if your clients rely on specific ExceptionCode and StatusCode values being returned...
''\end{quotation}
\normalsize

These extended list of change patterns need to be addressed by future work on finding techniques and developing tool support for evolving Web APIs.

\subsubsection{Version Identifiers}

We analyzed different approaches to identify the versions of Web APIs to understand if the version identifiers can be used to infer information about their evolution. We can categorize the Web APIs into these groups:

\begin{figure*}[!htp]
\subfigure[Facebook API Changes (11 Releases)]{
\def\dataSourceFileName{facebookx.dat}
\begin{tikzpicture}[show background rectangle,tight background]
  \begin{axis}[
    ybar stacked,
    bar width=5pt,
    %title=Facebook API Changes (11 Releases),
    date coordinates in=x,
    xticklabel = \year-\month,
    xtick = {2013-01-09, 2013-04-01, 2013-07-01, 2013-10-01, 2014-01-01, 2014-04-01, 2014-08-01, 2014-07-02},
    %xlabel={Timeline},
    xlabel style={font=\small},
    ylabel style={font=\small},
    ylabel={No. of Changes},
    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
    enlarge x limits=false,
    height=0.20\textheight,
    width=0.32\linewidth,
    legend style={legend pos=north west,font=\tiny, legend columns=4},
    legend to name=legend
    ]
    \input{change_frequency}
  \end{axis}
\end{tikzpicture}
\label{fig:facebook-chart}
}
\subfigure[Stripe API Changes (30 Releases)]{
\def\dataSourceFileName{stripe.dat}
\begin{tikzpicture}[show background rectangle,tight background]
  \begin{axis}[
    %title=Stripe API Changes (30 Releases),
    date coordinates in=x,
    ybar stacked,
    bar width=5pt,
    xticklabel = \year-\month,
    xtick = {2011-09-15, 2012-01-01, 2013-01-01, 2014-01-01, 2014-10-01},
    %xlabel={Timeline},
    xlabel style={font=\small},
    ylabel style={font=\small},
    ylabel={No. of Changes},
    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
    enlarge x limits=false,
    height=0.20\textheight,
    width=0.32\linewidth,
    legend to name=legend
    ]
    \input{change_frequency}
  \end{axis}
\end{tikzpicture}
\label{fig:stripe-chart}
}
\subfigure[WordPress REST API Changes (8 Releases)]{
\def\dataSourceFileName{wordpress.dat}
\begin{tikzpicture}[show background rectangle,tight background]
  \begin{axis}[
    %title=WordPress REST API Changes (8 Releases),
    date coordinates in=x,
    ybar stacked,
    bar width=5pt,
    xticklabel = \year-\month,
    xtick = {2013-08-07, 2013-11-01, 2014-03-01, 2014-06-26},
    %xlabel={Timeline},
    xlabel style={font=\small},
    ylabel style={font=\small},
    ylabel={No. of Changes},
    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
    enlarge x limits=false,
    height=0.20\textheight,
    width=0.32\linewidth,
    legend to name=legend
    ]
    \input{change_frequency}
  \end{axis}
\end{tikzpicture}
\label{fig:wordpress-chart}
}
\subfigure[GitHub API Changes (59 Releases)]{
\def\dataSourceFileName{github.dat}
\begin{tikzpicture}[show background rectangle,tight background]
  \begin{axis}[
    %title=GitHub API Changes (59 Releases),
    date coordinates in=x,
    ybar stacked,
    bar width=5pt,
    xticklabel = \year-\month,
    xtick = {2012-09-05, 2012-12-01, 2013-03-01, 2013-06-01, 2013-09-01, 2013-12-01, 2014-03-01, 2014-10-06},
    %xlabel={Timeline},
    xlabel style={font=\small},
    ylabel style={font=\small},
    ylabel={No. of Changes},
    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
    enlarge x limits=false,
    height=0.20\textheight,
    width=0.32\linewidth,
    legend to name=legend
    ]
    \input{change_frequency}
  \end{axis}
\end{tikzpicture}
\label{fig:github-chart}
}
\subfigure[Google Maps Web API (102 Releases)]{
\def\dataSourceFileName{gmap.dat}
\begin{tikzpicture}[show background rectangle,tight background]
  \begin{axis}[
    %title=Google Maps Web API (102 Releases),
    date coordinates in=x,
    ybar stacked,
    bar width=5pt,
    xticklabel = \year-\month,
    xtick = {2011-01-06, 2011-07-01, 2012-01-01, 2012-07-01, 2013-01-01, 2013-07-01, 2014-01-01, 2014-07-01, 2014-09-18},
    %xlabel={Timeline},
    xlabel style={font=\small},
    ylabel style={font=\small},
    ylabel={No. of Changes},
    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
    enlarge x limits=false,
    height=0.20\textheight,
    width=0.32\linewidth,
    legend to name=legend
    ]
    \input{change_frequency}
  \end{axis}
\end{tikzpicture}
\label{fig:gmaps-chart}
}
\subfigure[Salesforce REST API (4 Releases)]{
\def\dataSourceName{Salesforce Force.com and Chatter REST}
\def\dataSourceFileName{salesforce.dat}
\begin{tikzpicture}[show background rectangle,tight background]
  \begin{axis}[
    %title=Salesforce REST API (4 Releases),
    date coordinates in=x,
    ybar stacked,
    bar width=5pt,
    xticklabel = \year-\month,
    xtick = {2013-10-04, 2014-01-01, 2014-04-01, 2014-07-01, 2014-10-18},
    %label={Timeline},
    xlabel style={font=\small},
    ylabel style={font=\small},
    ylabel={No. of Changes},
    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
    enlarge x limits=false,
    height=0.20\textheight,
    width=0.32\linewidth,
    legend to name=legend
    ]
    \input{change_frequency}
  \end{axis}
\end{tikzpicture}
\label{fig:salesforce-chart}
}
%% Add the legend here
\begin{center}
\subfigure{
\includegraphics[keepaspectratio, width=10cm]{legend.png}
}
\end{center}
\caption{Case Study - Change Profiles for the Web APIs.}
\label{fig:proflie}
\end{figure*}

\textbf{Numbered identifiers} - Contiguous integers are used to identify new versions by Salesforce and GitHub Web APIs. Such naming scheme clearly communicates the sequence between versions but cannot be used to infer compatibility between versions.

\textbf{Timestamped identifiers} - Stripe Web API uses date as a version identifier. While timestamped versions provide the time frame of a version, they do not provide any compatibility information between versions.

\textbf{Identifiers with major and minor version numbers} - Most of the Web APIs studied use major and minor version numbers to identify a version. Facebook, Twitter, WordPress, Google Calendar, Google Maps and OpenStreetMap follow this approach. However, an explanation of the major and minor version numbers aren't provided and cannot be used to infer compatibility information.

Also, web APIs often evolve with breaking changes without issuing a new version identifier. For example, Facebook, Google Calendar and GitHub APIs often receive breaking changes without using a new version identifier.

\subsubsection{Upgrade Mode} % (fold)

From our case studies, we found the Web APIs followed several different approaches to versioning.

\textbf{Single Version} - Web APIs that force the API clients to migrate either because only one version is available or deprecated older versions will be removed. These include Facebook REST API (90 days), Twitter REST API (no specific time frame), Github API (no specific time frame), OpenStreetMap API (no specific time frame), Google Calendar (6 months), Google Maps (no specific time frame).

\textbf{Multiple Versions} - Multiple versions of the Web API are supported for longer duration. These include Wordpress REST API, Salesforce API, Stripe API.

Both of these categories, single and multiple available versions, have their advantages and disadvantages. Forcing API clients to use a single stable version makes it easy to maintain the Web APIs at the expense of scheduling flexibility for API users. These APIs keep a deprecated version for a limited period of time to provide a buffer time for the API users. For example, Google Calendar API is supporting its deprecated version for 6 months.

Maintaining multiple versions of the Web APIs provide more scheduling flexibility to the API clients but requires more effort from the API publisher. As a result, not all the older versions are maintained. When API clients lag far behind the latest version, migrating to the latest version becomes harder. For example, here is an excerpt of a comment from a Stripe Web API client developer \footnote{\url{https://groups.google.com/a/lists.stripe.com/forum/#!searchin/api-discuss/version/api-discuss/x8TM-yHnhYQ/y9JJKytSiUAJ}}:

\small
``...I just learned that we're about 14 versions behind and would like to work on upgrading but there are lots of breaking changes...''
\normalsize

While both the forced and on demand modes are found in practice, future work needs to be carried out to compare and contrast the benefits of these modes in greater detail.

\subsubsection{Change Profiles}
The Web APIs evolve at different frequencies as well as the actual change that takes place during their evolution varies from one API to another. Fig. \ref{fig:proflie} shows an aggregated view of the count of different categories of changes observed for different Web APIs based on the aforementioned coding scheme.

%
%
%\begin{figure*}[!htp]
%\begin{subfigure}[b]{\textwidth}
%\def\dataSourceFileName{facebookx.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    ybar stacked,
%    bar width=5pt,
%    %title=Facebook API Changes (11 Releases),
%    date coordinates in=x,
%    xticklabel = \year-\month,
%    xtick = {2013-01-09, 2013-04-01, 2013-07-01, 2013-10-01, 2014-01-01, 2014-04-01, 2014-08-01, 2014-07-02},
%    %xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    %width=\linewidth,
%    legend style={legend pos=north west,font=\tiny, legend columns=4},
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%\caption{Facebook API Changes (11 Releases)}
%\end{subfigure}
%
%\begin{subfigure}{\textwidth}
%%\caption{Stripe API Changes (30 Releases)}
%\def\dataSourceFileName{stripe.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    %title=Stripe API Changes (30 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2011-09-15, 2012-01-01, 2013-01-01, 2014-01-01, 2014-10-01},
%    %xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    %width=\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%\end{subfigure}
%
%\begin{subfigure}{\textwidth}
%%\caption{WordPress REST API Changes (8 Releases)}
%\def\dataSourceFileName{wordpress.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    %title=WordPress REST API Changes (8 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2013-08-07, 2013-11-01, 2014-03-01, 2014-06-26},
%    %xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    %width=\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%\end{subfigure}
%
%\begin{subfigure}{\textwidth}
%%\caption{GitHub API Changes (59 Releases)}
%\def\dataSourceFileName{github.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    %title=GitHub API Changes (59 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2012-09-05, 2012-12-01, 2013-03-01, 2013-06-01, 2013-09-01, 2013-12-01, 2014-03-01, 2014-10-06},
%    %xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    %width=\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%\end{subfigure}
%
%\begin{subfigure}{\textwidth}
%%\caption{Google Maps Web API (102 Releases}
%\def\dataSourceFileName{gmap.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    %title=Google Maps Web API (102 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2011-01-06, 2011-07-01, 2012-01-01, 2012-07-01, 2013-01-01, 2013-07-01, 2014-01-01, 2014-07-01, 2014-09-18},
%    %xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    %width=\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%\end{subfigure}
%
%\begin{subfigure}{\textwidth}
%%\caption{[Salesforce REST API (4 Releases)]}
%\def\dataSourceName{Salesforce Force.com and Chatter REST}
%\def\dataSourceFileName{salesforce.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    %title=Salesforce REST API (4 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2013-10-04, 2014-01-01, 2014-04-01, 2014-07-01, 2014-10-18},
%    %label={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    %width=0.9\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%\end{subfigure}
%
%\begin{subfigure}{\textwidth}
%\ref{legend}
%\end{subfigure}
%
%\caption{Case Study - Change Profiles for the Web APIs.}
%\end{figure*}
%

%%%%%%%%%%%%%%%%%%
%
%x.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    ybar stacked,
%    bar width=5pt,
%    title=Facebook API Changes (11 Releases),
%    date coordinates in=x,
%    xticklabel = \year-\month,
%    xtick = {2013-01-09, 2013-04-01, 2013-07-01, 2013-10-01, 2014-01-01, 2014-04-01, 2014-08-01, 2014-07-02},
%    xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    width=\linewidth,
%    legend style={legend pos=north west,font=\tiny, legend columns=4},
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%
%
%\ref{legend}
%
%\def\dataSourceFileName{stripe.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    title=Stripe API Changes (30 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2011-09-15, 2012-01-01, 2013-01-01, 2014-01-01, 2014-10-01},
%    xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    width=\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%
%
%\def\dataSourceFileName{wordpress.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    title=WorPress REST API Changes (8 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2013-08-07, 2013-11-01, 2014-03-01, 2014-06-26},
%    xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    width=\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%
%
%\def\dataSourceFileName{github.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    title=GitHub API Changes (59 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2012-09-05, 2012-12-01, 2013-03-01, 2013-06-01, 2013-09-01, 2013-12-01, 2014-03-01, 2014-10-06},
%    xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    width=\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%
%\def\dataSourceFileName{gmap.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    title=Google Maps Web API (102 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2011-01-06, 2011-07-01, 2012-01-01, 2012-07-01, 2013-01-01, 2013-07-01, 2014-01-01, 2014-07-01, 2014-09-18},
%    xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    width=\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}
%
%\def\dataSourceName{Salesforce Force.com and Chatter REST}
%\def\dataSourceFileName{salesforce.dat}
%\begin{tikzpicture}[show background rectangle,tight background]
%  \begin{axis}[
%    title=Salesforce REST API (4 Releases),
%    date coordinates in=x,
%    ybar stacked,
%    bar width=5pt,
%    xticklabel = \year-\month,
%    xtick = {2013-10-04, 2014-01-01, 2014-04-01, 2014-07-01, 2014-10-18},
%    xlabel={Timeline},
%    xlabel style={font=\small},
%    ylabel style={font=\small},
%    ylabel={No. of Changes},
%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
%    enlarge x limits=false,
%    height=0.20\textheight,
%    width=0.9\linewidth,
%    legend to name=legend
%    ]
%    \input{change_frequency}
%  \end{axis}
%\end{tikzpicture}

The APIs here show very different change trajectories. Facebook API shows an alternating pattern of smaller and larger number of changes between subsequent versions, and each release consists of multiple categories of changes. On the other hand, Stripe and GitHub API changes are more frequent, but on the majority of the releases only include a single category of change. WordPress REST API releases contain multiple categories of changes, but shows only a single API element removal for the studied period. GoogleMaps API changes are also very frequent, dominated by AddAPIElement and BehaviorChange categories in a way that once new API elements are introduced, the subsequent releases primarily focus on bug fixes. The GoogleMaps API shows signs of stability with time. The Salesforce API changelog shows a different picture. Unlike the other APIs studied, Salesforce API evolves quarterly and involves a larger number of API elements with each release. We observed most Web API releases broke backward compatibility as shown by the presence of colors in addition to the green areas on the charts. We consider API changes to be backward compatible when changes are limited to only AddAPIElements category.

Google Calendar API, not shown in a chart, changed the format of all API request and response data from a custom XML to JSON in a single release, effectively forcing all its users to upgrade with a deprecation policy of one year. A plot for the Twitter and OpenStreetMap APIs could not be produced since changelogs for these APIs were no longer available.

These change profiles have little similarity, implying the fact that a common solution to solve their evolution related challenges may not work. Future research on Web API evolution needs to consider these variation of change profiles to find effective solutions for the case in hand.

\subsubsection{Implementation} % (fold)
\label{ssub:implementation}

In our case studies, we included two open-source Web APIs, WordPress and OpenStreetMap, because it gave us access to  their source code to study how versioning is actually implemented. Wordpress REST API is developed as a plugin for Wordpress. As a result, a new version of the Web API is released with a new version of the plugin and it all API clients are affected when a new version is deployed. The source code for OpenStreetMap.org Web API shows the use of configuration files for API Versioning. We found its implementation only allows for a single API version.

To summarize the change profiles, version identifiers and implementation of evolving Web API versions, we observed little convergence and a lack of any common scheme that was followed by the studied Web APIs.

\subsection{RQ2 Documentation} % (fold)
\label{sub:documentation}
Web API documentation is used by API users as the primary source of information for system integration. We analyzed the API reference documentation to understand the contents and the presentation style of their documentation. The documentation is commonly comprised of the following:

\textbf{Web API access information} - Web API documentation includes information about the URL where the API can be reached. The studied APIs used one of the URL, HTTP headers, or a user interface to identify specific API versions.

\textbf{Authentication} - Web API documentation includes information about authentication for their API clients. This is also used to implement client specific rules, such as rate limits and time zones.

\textbf{API Interface definition} - The interface definition is a major part of the Web API documentation where each API element is described with a general overview  and related business rules.

\textbf{Example usage} - Examples that describe different use cases of the Web APIs are also commonly found in the documentation. The examples contain the data and HTTP headers for sample requests and the corresponding API responses.

Some of the studied Web APIs also provide live web browser based API explorers, where API users can make API calls without needing to write any code. However, the studied Web APIs that offered Live API Explorer only did so for the latest API versions.

Web API documentation needs to include additional information compared to local APIs since HTTP is used for communication. As an API evolves lack of documentation for older API versions causes difficulties with the upgrade process as found in the following comment from a Stripe API user\footnote{\url{https://groups.google.com/a/lists.stripe.com/forum/#!searchin/api-discuss/version/api-discuss/li4PyVcweiw/NT9SFTtF-vQJ}}

\small
\begin{quotation}
 ``Does the full API documentation only reflect the current version of the
  API?  Is there a way to access the API docs for outdated versions? ...That would be very helpful. When you are trying to upgrade from one version to another it's impossible to know the implementation differences. We are currently about 4 API versions behind and are stuck behind a version that causes a significant amount of work on our end to support. I'd like to be able to upgrade incrementally through each version.''
\end{quotation}
\normalsize

We also investigated how the API documentation is generated. We found that, Web API documents are manually generated and updated, unlike the documentation for local APIs which are are typically auto-generated. For WordPress and OpenStreetMap, this is apparent from their source code, as we found the use of manually edited wiki and Markdown files to generate the documentation\footnote{\url{http://en.wikipedia.org/wiki/Markdown}}.

In our case studies, we found one example of self-documenting Web API as discussed by Laskey \cite{laskey2008considerations}. The Salesforce Web API has API endpoints that describe all available versions of the API that can be used to programmatically determine the changes between versions.

To summarize, we found that Web API documents are manually generated and widely vary in terms of both their contents and presentation formats.

\subsection{RQ3 Communication} % (fold)
\label{sub:communication}

Communicating updates and changelogs is an important activity for evolving Web APIs. The communication channels used by the studied Web APIs is categorized as follows:

  \textbf{API home page} - Web APIs announce the API changes in their home pages. In practice, the announcements include partial change log with the key changes that are made in a release. These announcements are unstructured text and do not follow any standard format even for the same API.

  \textbf{API response} - Some Web APIs use custom HTTP headers to indicate when a call is made to a deprecated version. For example, Wordpress REST API sends the X-WP-DeprecatedFunction header in response to API calls made to deprecated endpoints.

  \textbf{Email} - Facebook sends customized email alerts to its API clients based on their usage of the API, similar to the idea as presented by Zou et al. \cite{le2008synchronizing}. API calls are recorded to determine if a change has an impact on users. The granularity of the change tracking often results into false-positives, as found in this comment by a Facebook API user \footnote{\url{http://stackoverflow.com/questions/16270446/updating-app-for-breaking-change-non-threaded-comments}}:

\small
\begin{quotation}
...I have received the same notification for one of my apps. My app definitely does not read or create comments on Facebook posts or objects ... Thus Facebook's message is not relevant to every app they send it to.
The change should only affect apps which read or publish comments.
\end{quotation}
\normalsize

  \textbf{Newsfeed} - To keep up-to-date with API changes, the Web API clients are requested to subscribe to newsfeed, typically delivered via mailing lists and Twitter feeds. User feedback is also collected on these platforms. For example, at the time of writing this paper there were 13,239 questions tagged against google-maps-api-3\footnote{\url{https://stackoverflow.com/questions/tagged/google-maps-api-313.239}}.

To summarize, we found both formal and informal channels are used to communicate Web API changes. Most messages are primarily driven by manually edited unstructured text.
% subsection communication (end)

\section{Discussion} % (fold)
\label{sec:discussion}
Our case study findings such as a lack of standard approaches to deal with Web API versioning, documentation and change communication confirms the findings from previous case studies. In addition to the Web API change patterns identified by previous case studies, we have identified six new change patterns: 1) move API elements, 2) rename API elements, 3) behavior change, 4) post condition change, 5) HTTP header change, and 6) error condition change. When creating API changelog and developing related tool support, these new change patterns can be used to communicate the changes using a shared vocabulary. In addition to these new change patterns, from our case study we have compiled a list of recommendations and identified new research problems related to RQ1-3 as discussed in the following paragraphs.

\textbf{RQ1 Versioning} - From our case study, we found that when a new version is released for a Web API, the version identifiers provide little information about backward compatibility and the impact of the new version on existing API users. As a result, the API users are left to follow the free-form newsfeeds to understand the impact for each new API version. To improve this situation, for practitioners, we recommend using Semantic Versioning. Semantic Versioning is a naming technique for versioning software so that it is possible to infer if two versions are backward compatible simply by interpreting the version identifiers\footnote{\url{http://semver.org/}}. Future research can be carried out to automatically generate a Semantic Versioning identifier based on the compatibility between versions.

Our change profiles show that Web APIs often evolve as fast as several times a month and forces API clients to upgrade. Also, releases often combine both bug fixes and other breaking changes, forcing the API users to adapt even though they only need the fixes.  For practitioners, we recommend that bug fixes and breaking changes be released under separate versions to provide more flexibility to the API clients. Future research needs to focus on finding strategies for evolving Web APIs so that bug fixes and other changes can be easily delivered in separate versions and multiple versions can be made available at the same time.

\textbf{RQ2 Documentation} - We observed a that largely manual process is used for generating API documentation for most RESTful APIs. Manually generating and maintaining such documentation is expensive and error-prone. From our analysis, we showed that even when Web APIs are  versioned, documentation for older versions is not always available, and the difference between two versions of a Web API cannot be easily derived from their manually edited API references. We also found that the changelogs and API references are two disconnected documents even though a user needs to read both documents while upgrading. For practitioners, we recommend releasing Web API documentation for each available version that is cross-linked with the changelog. Future research on tool development can be aimed to support automated, version-aware documentation needs of evolving Web APIs.

Live Web API explorers are used by API users to try the APIs with little effort. However, there is a lack of reusable approach to create Live Web API explorers and bespoke implementation of the live API explorers are expensive since they need additional software development. We have seen only two of the nine studied APIs to provide a live API explorer. Moreover, we found that live Web API explorers are only offered for latest versions. Future research can focus on the implementation of a reusable, version-aware live Web API explorer for evolving Web APIs.

Self documenting Web APIs can be used to automatically determine compatibility between two versions of a Web API. We found Salesforce REST API to be the only example from our case studies that implemented self documenting Web APIs. For practitioners, we recommend providing self-documenting Web APIs. We also identify this as an opportunity for future research so that reusable tool support can be provided to create self-documenting Web APIS.

\textbf{RQ3 Communication} - For communicating API changes, we found a number of different channels that are used in practice. These channels need to be compared so that the effective channels can be used for communicating between the API developers and users.

In the change profiles we found the APIs change frequently and this requires manual effort to understand the impact of a change on a specific API user. This can be simplified by alerting users about an API change with a customized changelog. We also found that API related discussions are carried out external to the API documentation sites. API users need to search multiple disconnected information sources when API related questions arise. For practitioners, we recommend publishing a customized changelog for their Web APIs and providing discussion forums with API documentation. Previous research showed a solution based on WSDL files \cite{le2008synchronizing}. Future research is needed to solve this problem for RESTful Web APIs.

The open source Web APIs (Wordpress and OpenStreetMap) had limited documentation and change log information compared to the proprietary Web APIs. This presents an opportunity for researchers and industry practitioners to create reusable open source versioning and documentation tools for evolving Web APIs.

\textbf{Limitations} - We recognize the number of Web APIs studied to be a threat to the construct validity. Although our selection involves a diverse set of Web APIs, the findings may not be generalized for all evolving Web APIs. Our selection criteria included different industry domains, API sizes, popularity and maturity levels to minimize selection bias. The selected APIs represent only publicly available evolving Web APIs because only publicly available data is used for the case study. Future work need to be carried out to compare the applicability of our findings and related implications on privately maintained evolving Web APIs. To mitigate the interpretation bias of the coding scheme, the first two authors of this paper independently annotated randomly selected snippets of change logs from different Web APIs and reached the same conclusion indicating the replicability of the given coding scheme.

Given the state of practice we found a wide variety of ways to evolve Web APIs in terms of versioning, documentation and communication of changes and no consistent way to deal with the changes. This indicates an immature area where cost-effective solutions that are acceptable for both API developers and users are still missing - as discussed. Further research needs to be carried out to validate the usefulness of the aforementioned recommendations and solve the new research problems as identified by this case study.

\section{Conclusion}
Web API evolution poses challenges since a change may break applications that are developed by different teams and organizations. We presented a case study of Web API evolution focusing on the challenges involved with versioning, documentation and communication of API changes. Recommendations for practitioners from our analysis include the use of semantic versioning, separate releases for bug fixes and new features, auto generated API documentation cross-linked with changelogs and providing live API explorers. A list of open research problems are discussed related to Web API evolution that we want to solve in our future work.

\bibliographystyle{./IEEEtran}
\bibliography{IEEEabrv,references}

\end{document}

### 2017-07-22 case_study/case_study.tex
adds the case study paper

--- 2017-07-21 case-study/icws_case_study_api_evolution_sohan.tex
@@
-\documentclass[conference]{IEEEtran}
-
-\ifCLASSINFOpdf
-\else
-\fi
-\usepackage{url}
-\usepackage[hypcap]{caption}
-\usepackage{listings}
-\usepackage{pgfplots}
-\usepackage{pgfplotstable}
-\usepackage{eurosym}
-\usepackage{multirow}
-\usepackage{subfigure}
-
-%\usepackage{caption}
-%\usepackage{subcaption}
-
 \usetikzlibrary{backgrounds}
 \pgfplotsset{compat=1.5}
-\usepgfplotslibrary{dateplot}
 
-\begin{document}
-\title{A Case Study of Web API Evolution}
+\chapter{A Case Study of Web API Evolution}
 
-\author{\IEEEauthorblockN{S M Sohan, Craig Anslow, Frank Maurer
-\IEEEauthorblockA{Department of Computer Science\\
-University of Calgary,
-Calgary, Alberta T2N 1N4\\ Email: \{smsohan, craig.anslow, frank.maurer\}@ucalgary.ca }
-}}
+%\author{\IEEEauthorblockN{S M Sohan, Craig Anslow, Frank Maurer
+%\IEEEauthorblockA{Department of Computer Science\\
+%University of Calgary,
+%Calgary, Alberta T2N 1N4\\ Email: \{smsohan, craig.anslow, frank.maurer\}@ucalgary.ca }
+%}}
 
-\maketitle
+%\maketitle
 
-\begin{abstract}
+\section{Abstract}
 When applications are integrated using web APIs, changes on a web API may break the dependent applications.
@@
 This study will help inform developers of evolving Web APIs to make decision about versioning, documentation and communication methods.
-\end{abstract}
 
-\begin{IEEEkeywords}
-Case Study; RESTful, SOAP, Web API Evolution; WSDL
-\end{IEEEkeywords}
+%\begin{IEEEkeywords}
+%Case Study; RESTful, SOAP, Web API Evolution; WSDL
+%\end{IEEEkeywords}
 
-\IEEEpeerreviewmaketitle
+%\IEEEpeerreviewmaketitle
 
@@
 
-%{|p{0.2\linewidth}|p{0.3\linewidth}|p{0.2\line width}|p{0.2\linewidth}|}
-%& &  \multicolumn{4}{ c |}{\textbf{Categories}} & & \\ \cline{3-6}
-%\multirow{2}{*}{\textbf{Web API}}  & \multirow{2}{*}{\textbf{URL}} &\textbf{Industry} & \textbf{Popularity} &  \textbf{Maturity} & \textbf{Code}& \multirow{2}{*}{\\textbf{Rel.}} & \multirow{2}{*}{\\textbf{Dates}}\\
-
-\begin{table*}[!ht]
-\caption{Case Study - Representative Sample of Web APIs.}
-\label{web_api_grid}
-\begin{tabular} {| l | l  | l  | l | l | l | l  | l |}
-\hline
-\multirow{2}{*}{\textbf{Web API}} &
-\multirow{2}{*}{\textbf{URL}} &
-\multicolumn{4}{ c |}{\textbf{Categories}} &
-\textbf{Num of} &
-\multirow{2}{*}{\textbf{Dates}}\\    \cline{3-6}
-  & & Industry & Popularity &  Maturity & Code& \textbf{Releases}& \\
-\hline
-Facebook Platform API & \tiny{\url{https://developers.facebook.com/docs/apps/versions}} & Social & High & Est & Closed & 11 & 01/2013-07/2014\\
-\hline
-Twitter REST API & \tiny{\url{https://dev.twitter.com/docs/api}} & Social & High & Est & Closed & 2 & Not Available\\
-\hline
-Wordpress REST API & \tiny{\url{http://wp-api.org/}} & Business & Growing &  New & Open & 8 & 08/2013-06/2014\\
-\hline
-Salesforce \& Chatter REST API & \tiny{\url{https://www.salesforce.com/us/developer/docs/api_rest/}} & Business & Popular & Est & Closed & 4 & 01/2013-08/2014\\
-\hline
-Google Calendar API & \tiny{\url{https://developers.google.com/google-apps/calendar/}} & Business & Popular & Est & Closed & 2 & Not Available\\
-\hline
-Stripe Payment Processor API & \tiny{\url{https://stripe.com/docs/api}} & Business & Growing & New & Closed & 30 &09/2011-10/2014 \\
-\hline
-Github API & \tiny{\url{https://developer.github.com/}} & Productivity & Growing & New & Closed  & 59 &09/2012-10/2014\\
-\hline
-Google Maps API & \tiny{\url{https://developers.google.com/maps/web/}} & Productivity & Popular & Est & Closed & 102 &01/2011-09/2014\\
-\hline
-OpenStreetMap API &\tiny{\url{http://wiki.openstreetmap.org/wiki/API}} & Productivity & Growing & New & Open & 1 & Not Available\\
-\hline
-\end{tabular}
-\end{table*}
-
 Analyzing the available tool support for evolving Web APIs helped us identify opportunities for future research in this area.
@@
 \label{sub:selection_strategy}
+\begin{table*}[!htbp]
+\caption{Case Study - Representative Sample of Web APIs.}
+\label{web_api_grid}
+\rotatebox{-90}{
+  \begin{tabular} {| p{1.5in}| p{2in}| l  | l | l | l | l  | l |}
+    \hline
+    \multirow{2}{*}{\textbf{Web API}} &
+    \multirow{2}{*}{\textbf{URL}} &
+    \multicolumn{4}{ c |}{\textbf{Categories}} &
+    \textbf{Num of} &
+    \multirow{2}{*}{\textbf{Dates}}\\    \cline{3-6}
+    & & Industry & Popularity &  Maturity & Code& \textbf{Releases}& \\
+    \hline
+    Facebook Platform API & \tiny{\url{https://developers.facebook.com/docs/apps/versions}} & Social & High & Est & Closed & 11 & 01/2013-07/2014\\
+    \hline
+    Twitter REST API & \tiny{\url{https://dev.twitter.com/docs/api}} & Social & High & Est & Closed & 2 & Not Available\\
+    \hline
+    Wordpress REST API & \tiny{\url{http://wp-api.org/}} & Business & Growing &  New & Open & 8 & 08/2013-06/2014\\
+    \hline
+    Salesforce \& Chatter REST API & \tiny{\url{https://www.salesforce.com/us/developer/docs/api_rest/}} & Business & Popular & Est & Closed & 4 & 01/2013-08/2014\\
+    \hline
+    Google Calendar API & \tiny{\url{https://developers.google.com/google-apps/calendar/}} & Business & Popular & Est & Closed & 2 & Not Available\\
+    \hline
+    Stripe Payment Processor API & \tiny{\url{https://stripe.com/docs/api}} & Business & Growing & New & Closed & 30 &09/2011-10/2014 \\
+    \hline
+    Github API & \tiny{\url{https://developer.github.com/}} & Productivity & Growing & New & Closed  & 59 &09/2012-10/2014\\
+    \hline
+    Google Maps API & \tiny{\url{https://developers.google.com/maps/web/}} & Productivity & Popular & Est & Closed & 102 &01/2011-09/2014\\
+    \hline
+    OpenStreetMap API &\tiny{\url{http://wiki.openstreetmap.org/wiki/API}} & Productivity & Growing & New & Open & 1 & Not Available\\
+    \hline
+  \end{tabular}
+}
+\end{table*}
 
@@
 
-\begin{table*}[!tbh]
+\begin{table*}[!htbp]
 \caption{List of Codes}
 \label{coding_scheme}
-\begin{tabular}{|p{0.2\linewidth}|p{0.7\linewidth}|}
+\begin{tabular}{|p{0.3\linewidth}|p{0.6\linewidth}|}
 \hline
@@
 
-Another example that demonstrates a behavior change can be found from the 2013-10-29 version of Stripe API \footnote{\url{https://stripe.com/docs/upgrades#2014-01-31}}:
+Another example that demonstrates a behavior change can be found from the 2013-10-29 version of Stripe API \footnote{\url{https://stripe.com/docs/upgrades\#2014-01-31}}:
 
@@
 \textbf{Error condition change} - Web APIs change the error conditions that may require the users to adapt their integration.
-For example, in Salesforce API Version 29 the error codes were changed \footnote{\url{https://na1.salesforce.com/help/pdfs/en/salesforce_winter14_release_notes.pdf#rn_186_api_objects}}:
+For example, in Salesforce API Version 29 the error codes were changed \footnote{\url{https://na1.salesforce.com/help/pdfs/en/salesforce_winter14_release_notes.pdf\#rn_186_api_objects}}:
 
@@
 We can categorize the Web APIs into these groups:
+\textbf{Numbered identifiers} - Contiguous integers are used to identify new versions by Salesforce and GitHub Web APIs.
+Such naming scheme clearly communicates the sequence between versions but cannot be used to infer compatibility between versions.
 
-\begin{figure*}[!htp]
-\subfigure[Facebook API Changes (11 Releases)]{
+\textbf{Timestamped identifiers} - Stripe Web API uses date as a version identifier.
+While timestamped versions provide the time frame of a version, they do not provide any compatibility information between versions.
+
+\textbf{Identifiers with major and minor version numbers} - Most of the Web APIs studied use major and minor version numbers to identify a version.
+Facebook, Twitter, WordPress, Google Calendar, Google Maps and OpenStreetMap follow this approach.
+However, an explanation of the major and minor version numbers aren't provided and cannot be used to infer compatibility information.
+
+Also, web APIs often evolve with breaking changes without issuing a new version identifier.
+For example, Facebook, Google Calendar and GitHub APIs often receive breaking changes without using a new version identifier.
+
+\begin{figure*}[!hbtp]
+  \begin{subfigure}[b]{\textwidth}
     \def\dataSourceFileName{facebookx.dat}
@@
           bar width=5pt,
-    %title=Facebook API Changes (11 Releases),
           date coordinates in=x,
@@
           xtick = {2013-01-09, 2013-04-01, 2013-07-01, 2013-10-01, 2014-01-01, 2014-04-01, 2014-08-01, 2014-07-02},
-    %xlabel={Timeline},
           xlabel style={font=\small},
@@
           height=0.20\textheight,
-    width=0.32\linewidth,
+          width=\textwidth,
           legend style={legend pos=north west,font=\tiny, legend columns=4},
@@
         ]
-    \input{change_frequency}
+        \input{case_study/change_frequency}
       \end{axis}
@@
     \label{fig:facebook-chart}
-}
-\subfigure[Stripe API Changes (30 Releases)]{
+    \caption{Facebook API Changes (11 Releases)}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{\textwidth}
     \def\dataSourceFileName{stripe.dat}
@@
       \begin{axis}[
-    %title=Stripe API Changes (30 Releases),
           date coordinates in=x,
@@
           height=0.20\textheight,
-    width=0.32\linewidth,
+          width=\textwidth,
           legend to name=legend
         ]
-    \input{change_frequency}
+        \input{case_study/change_frequency}
       \end{axis}
@@
     \label{fig:stripe-chart}
-}
-\subfigure[WordPress REST API Changes (8 Releases)]{
+    \caption{Stripe API Changes (30 Releases)}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{\textwidth}
     \def\dataSourceFileName{wordpress.dat}
@@
     \label{fig:wordpress-chart}
-}
-\subfigure[GitHub API Changes (59 Releases)]{
+    \caption{WordPress REST API Changes (8 Releases)}
+  \end{subfigure}
+  \begin{subfigure}[b]{\textwidth}
     \def\dataSourceFileName{github.dat}
@@
     \label{fig:github-chart}
-}
-\subfigure[Google Maps Web API (102 Releases)]{
+    \caption{GitHub API Changes (59 Releases)}
+  \end{subfigure}
+
+\end{figure*}
+\begin{figure*}[!h]\ContinuedFloat
+
+  \begin{subfigure}[b]{\textwidth}
     \def\dataSourceFileName{gmap.dat}
@@
     \label{fig:gmaps-chart}
-}
-\subfigure[Salesforce REST API (4 Releases)]{
+    \caption{Google Maps Web API (102 Releases)}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{\textwidth}
     \def\dataSourceName{Salesforce Force.com and Chatter REST}
@@
     \label{fig:salesforce-chart}
-}
+    \caption{Salesforce REST API (4 Releases)}
+  \end{subfigure}
+
 %% Add the legend here
   \begin{center}
-\subfigure{
+    \begin{subfigure}[b]{10cm}
       \includegraphics[keepaspectratio, width=10cm]{legend.png}
-}
+      \caption{Legend}
+    \end{subfigure}
   \end{center}
@@
 \end{figure*}
-
-\textbf{Numbered identifiers} - Contiguous integers are used to identify new versions by Salesforce and GitHub Web APIs.
-Such naming scheme clearly communicates the sequence between versions but cannot be used to infer compatibility between versions.
-
-\textbf{Timestamped identifiers} - Stripe Web API uses date as a version identifier.
-While timestamped versions provide the time frame of a version, they do not provide any compatibility information between versions.
-
-\textbf{Identifiers with major and minor version numbers} - Most of the Web APIs studied use major and minor version numbers to identify a version.
-Facebook, Twitter, WordPress, Google Calendar, Google Maps and OpenStreetMap follow this approach.
-However, an explanation of the major and minor version numbers aren't provided and cannot be used to infer compatibility information.
-
-Also, web APIs often evolve with breaking changes without issuing a new version identifier.
-For example, Facebook, Google Calendar and GitHub APIs often receive breaking changes without using a new version identifier.
 
@@
 When API clients lag far behind the latest version, migrating to the latest version becomes harder.
-For example, here is an excerpt of a comment from a Stripe Web API client developer \footnote{\url{https://groups.google.com/a/lists.stripe.com/forum/#!searchin/api-discuss/version/api-discuss/x8TM-yHnhYQ/y9JJKytSiUAJ}}:
+For example, here is an excerpt of a comment from a Stripe Web API client developer \footnote{\url{https://groups.google.com/a/lists.stripe.com/forum/\#!searchin/api-discuss/version/api-discuss/x8TM-yHnhYQ/y9JJKytSiUAJ}}:
 
@@
 \ref{fig:proflie} shows an aggregated view of the count of different categories of changes observed for different Web APIs based on the aforementioned coding scheme.
-
-%
-%
-%\begin{figure*}[!htp]
-%\begin{subfigure}[b]{\textwidth}
-%\def\dataSourceFileName{facebookx.dat}
-%\begin{tikzpicture}[show background rectangle,tight background]
-%  \begin{axis}[
-%    ybar stacked,
-%    bar width=5pt,
-%    %title=Facebook API Changes (11 Releases),
-%    date coordinates in=x,
-%    xticklabel = \year-\month,
-%    xtick = {2013-01-09, 2013-04-01, 2013-07-01, 2013-10-01, 2014-01-01, 2014-04-01, 2014-08-01, 2014-07-02},
-%    %xlabel={Timeline},
-%    xlabel style={font=\small},
-%    ylabel style={font=\small},
-%    ylabel={No. of Changes},
-%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
-%    enlarge x limits=false,
-%    height=0.20\textheight,
-%    %width=\linewidth,
-%    legend style={legend pos=north west,font=\tiny, legend columns=4},
-%    legend to name=legend
-%    ]
-%    \input{change_frequency}
-%  \end{axis}
-%\end{tikzpicture}
-%\caption{Facebook API Changes (11 Releases)}
-%\end{subfigure}
-%
-%\begin{subfigure}{\textwidth}
-%%\caption{Stripe API Changes (30 Releases)}
-%\def\dataSourceFileName{stripe.dat}
-%\begin{tikzpicture}[show background rectangle,tight background]
-%  \begin{axis}[
-%    %title=Stripe API Changes (30 Releases),
-%    date coordinates in=x,
-%    ybar stacked,
-%    bar width=5pt,
-%    xticklabel = \year-\month,
-%    xtick = {2011-09-15, 2012-01-01, 2013-01-01, 2014-01-01, 2014-10-01},
-%    %xlabel={Timeline},
-%    xlabel style={font=\small},
-%    ylabel style={font=\small},
-%    ylabel={No. of Changes},
-%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
-%    enlarge x limits=false,
-%    height=0.20\textheight,
-%    %width=\linewidth,
-%    legend to name=legend
-%    ]
-%    \input{change_frequency}
-%  \end{axis}
-%\end{tikzpicture}
-%\end{subfigure}
-%
-%\begin{subfigure}{\textwidth}
-%%\caption{WordPress REST API Changes (8 Releases)}
-%\def\dataSourceFileName{wordpress.dat}
-%\begin{tikzpicture}[show background rectangle,tight background]
-%  \begin{axis}[
-%    %title=WordPress REST API Changes (8 Releases),
-%    date coordinates in=x,
-%    ybar stacked,
-%    bar width=5pt,
-%    xticklabel = \year-\month,
-%    xtick = {2013-08-07, 2013-11-01, 2014-03-01, 2014-06-26},
-%    %xlabel={Timeline},
-%    xlabel style={font=\small},
-%    ylabel style={font=\small},
-%    ylabel={No. of Changes},
-%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
-%    enlarge x limits=false,
-%    height=0.20\textheight,
-%    %width=\linewidth,
-%    legend to name=legend
-%    ]
-%    \input{change_frequency}
-%  \end{axis}
-%\end{tikzpicture}
-%\end{subfigure}
-%
-%\begin{subfigure}{\textwidth}
-%%\caption{GitHub API Changes (59 Releases)}
-%\def\dataSourceFileName{github.dat}
-%\begin{tikzpicture}[show background rectangle,tight background]
-%  \begin{axis}[
-%    %title=GitHub API Changes (59 Releases),
-%    date coordinates in=x,
-%    ybar stacked,
-%    bar width=5pt,
-%    xticklabel = \year-\month,
-%    xtick = {2012-09-05, 2012-12-01, 2013-03-01, 2013-06-01, 2013-09-01, 2013-12-01, 2014-03-01, 2014-10-06},
-%    %xlabel={Timeline},
-%    xlabel style={font=\small},
-%    ylabel style={font=\small},
-%    ylabel={No. of Changes},
-%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
-%    enlarge x limits=false,
-%    height=0.20\textheight,
-%    %width=\linewidth,
-%    legend to name=legend
-%    ]
-%    \input{change_frequency}
-%  \end{axis}
-%\end{tikzpicture}
-%\end{subfigure}
-%
-%\begin{subfigure}{\textwidth}
-%%\caption{Google Maps Web API (102 Releases}
-%\def\dataSourceFileName{gmap.dat}
-%\begin{tikzpicture}[show background rectangle,tight background]
-%  \begin{axis}[
-%    %title=Google Maps Web API (102 Releases),
-%    date coordinates in=x,
-%    ybar stacked,
-%    bar width=5pt,
-%    xticklabel = \year-\month,
-%    xtick = {2011-01-06, 2011-07-01, 2012-01-01, 2012-07-01, 2013-01-01, 2013-07-01, 2014-01-01, 2014-07-01, 2014-09-18},
-%    %xlabel={Timeline},
-%    xlabel style={font=\small},
-%    ylabel style={font=\small},
-%    ylabel={No. of Changes},
-%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
-%    enlarge x limits=false,
-%    height=0.20\textheight,
-%    %width=\linewidth,
-%    legend to name=legend
-%    ]
-%    \input{change_frequency}
-%  \end{axis}
-%\end{tikzpicture}
-%\end{subfigure}
-%
-%\begin{subfigure}{\textwidth}
-%%\caption{[Salesforce REST API (4 Releases)]}
-%\def\dataSourceName{Salesforce Force.com and Chatter REST}
-%\def\dataSourceFileName{salesforce.dat}
-%\begin{tikzpicture}[show background rectangle,tight background]
-%  \begin{axis}[
-%    %title=Salesforce REST API (4 Releases),
-%    date coordinates in=x,
-%    ybar stacked,
-%    bar width=5pt,
-%    xticklabel = \year-\month,
-%    xtick = {2013-10-04, 2014-01-01, 2014-04-01, 2014-07-01, 2014-10-18},
-%    %label={Timeline},
-%    xlabel style={font=\small},
-%    ylabel style={font=\small},
-%    ylabel={No. of Changes},
-%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
-%    enlarge x limits=false,
-%    height=0.20\textheight,
-%    %width=0.9\linewidth,
-%    legend to name=legend
-%    ]
-%    \input{change_frequency}
-%  \end{axis}
-%\end{tikzpicture}
-%\end{subfigure}
-%
-%\begin{subfigure}{\textwidth}
-%\ref{legend}
-%\end{subfigure}
-%
-%\caption{Case Study - Change Profiles for the Web APIs.}
-%\end{figure*}
-%
-
-%%%%%%%%%%%%%%%%%%
-%
-%x.dat}
-%\begin{tikzpicture}[show background rectangle,tight background]
-%  \begin{axis}[
-%    ybar stacked,
-%    bar width=5pt,
-%    title=Facebook API Changes (11 Releases),
-%    date coordinates in=x,
-%    xticklabel = \year-\month,
-%    xtick = {2013-01-09, 2013-04-01, 2013-07-01, 2013-10-01, 2014-01-01, 2014-04-01, 2014-08-01, 2014-07-02},
-%    xlabel={Timeline},
-%    xlabel style={font=\small},
-%    ylabel style={font=\small},
-%    ylabel={No. of Changes},
-%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
-%    enlarge x limits=false,
-%    height=0.20\textheight,
-%    width=\linewidth,
-%    legend style={legend pos=north west,font=\tiny, legend columns=4},
-%    legend to name=legend
-%    ]
-%    \input{change_frequency}
-%  \end{axis}
-%\end{tikzpicture}
-%
-%
-%\ref{legend}
-%
-%\def\dataSourceFileName{stripe.dat}
-%\begin{tikzpicture}[show background rectangle,tight background]
-%  \begin{axis}[
-%    title=Stripe API Changes (30 Releases),
-%    date coordinates in=x,
-%    ybar stacked,
-%    bar width=5pt,
-%    xticklabel = \year-\month,
-%    xtick = {2011-09-15, 2012-01-01, 2013-01-01, 2014-01-01, 2014-10-01},
-%    xlabel={Timeline},
-%    xlabel style={font=\small},
-%    ylabel style={font=\small},
-%    ylabel={No. of Changes},
-%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
-%    enlarge x limits=false,
-%    height=0.20\textheight,
-%    width=\linewidth,
-%    legend to name=legend
-%    ]
-%    \input{change_frequency}
-%  \end{axis}
-%\end{tikzpicture}
-%
-%
-%\def\dataSourceFileName{wordpress.dat}
-%\begin{tikzpicture}[show background rectangle,tight background]
-%  \begin{axis}[
-%    title=WorPress REST API Changes (8 Releases),
-%    date coordinates in=x,
-%    ybar stacked,
-%    bar width=5pt,
-%    xticklabel = \year-\month,
-%    xtick = {2013-08-07, 2013-11-01, 2014-03-01, 2014-06-26},
-%    xlabel={Timeline},
-%    xlabel style={font=\small},
-%    ylabel style={font=\small},
-%    ylabel={No. of Changes},
-%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
-%    enlarge x limits=false,
-%    height=0.20\textheight,
-%    width=\linewidth,
-%    legend to name=legend
-%    ]
-%    \input{change_frequency}
-%  \end{axis}
-%\end{tikzpicture}
-%
-%
-%\def\dataSourceFileName{github.dat}
-%\begin{tikzpicture}[show background rectangle,tight background]
-%  \begin{axis}[
-%    title=GitHub API Changes (59 Releases),
-%    date coordinates in=x,
-%    ybar stacked,
-%    bar width=5pt,
-%    xticklabel = \year-\month,
-%    xtick = {2012-09-05, 2012-12-01, 2013-03-01, 2013-06-01, 2013-09-01, 2013-12-01, 2014-03-01, 2014-10-06},
-%    xlabel={Timeline},
-%    xlabel style={font=\small},
-%    ylabel style={font=\small},
-%    ylabel={No. of Changes},
-%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
-%    enlarge x limits=false,
-%    height=0.20\textheight,
-%    width=\linewidth,
-%    legend to name=legend
-%    ]
-%    \input{change_frequency}
-%  \end{axis}
-%\end{tikzpicture}
-%
-%\def\dataSourceFileName{gmap.dat}
-%\begin{tikzpicture}[show background rectangle,tight background]
-%  \begin{axis}[
-%    title=Google Maps Web API (102 Releases),
-%    date coordinates in=x,
-%    ybar stacked,
-%    bar width=5pt,
-%    xticklabel = \year-\month,
-%    xtick = {2011-01-06, 2011-07-01, 2012-01-01, 2012-07-01, 2013-01-01, 2013-07-01, 2014-01-01, 2014-07-01, 2014-09-18},
-%    xlabel={Timeline},
-%    xlabel style={font=\small},
-%    ylabel style={font=\small},
-%    ylabel={No. of Changes},
-%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
-%    enlarge x limits=false,
-%    height=0.20\textheight,
-%    width=\linewidth,
-%    legend to name=legend
-%    ]
-%    \input{change_frequency}
-%  \end{axis}
-%\end{tikzpicture}
-%
-%\def\dataSourceName{Salesforce Force.com and Chatter REST}
-%\def\dataSourceFileName{salesforce.dat}
-%\begin{tikzpicture}[show background rectangle,tight background]
-%  \begin{axis}[
-%    title=Salesforce REST API (4 Releases),
-%    date coordinates in=x,
-%    ybar stacked,
-%    bar width=5pt,
-%    xticklabel = \year-\month,
-%    xtick = {2013-10-04, 2014-01-01, 2014-04-01, 2014-07-01, 2014-10-18},
-%    xlabel={Timeline},
-%    xlabel style={font=\small},
-%    ylabel style={font=\small},
-%    ylabel={No. of Changes},
-%    xticklabel style= {rotate=90,anchor=near xticklabel,font=\tiny},
-%    enlarge x limits=false,
-%    height=0.20\textheight,
-%    width=0.9\linewidth,
-%    legend to name=legend
-%    ]
-%    \input{change_frequency}
-%  \end{axis}
-%\end{tikzpicture}
 
@@
 Web API documentation needs to include additional information compared to local APIs since HTTP is used for communication.
-As an API evolves lack of documentation for older API versions causes difficulties with the upgrade process as found in the following comment from a Stripe API user\footnote{\url{https://groups.google.com/a/lists.stripe.com/forum/#!searchin/api-discuss/version/api-discuss/li4PyVcweiw/NT9SFTtF-vQJ}}
+As an API evolves lack of documentation for older API versions causes difficulties with the upgrade process as found in the following comment from a Stripe API user\footnote{\url{https://groups.google.com/a/lists.stripe.com/forum/\#!searchin/api-discuss/version/api-discuss/li4PyVcweiw/NT9SFTtF-vQJ}}
 
@@
 
-\bibliographystyle{./IEEEtran}
-\bibliography{IEEEabrv,references}
-
-\end{document}
+\bibliographystyle{IEEEtran}
+\bibliography{IEEEabrv,case_study/references}

### 2017-07-22 build.sh

@@ -2,6 +2,7 @@ set -e
 git clean -fX
 pdflatex root.tex
 bibtex controlled_study/controlled_study.aux
+bibtex case_study/case_study.aux
 pdflatex root.tex
 pdflatex root.tex
 /Applications/Skim.app/Contents/MacOS/Skim root.pdf

### 2017-07-22 cisco_study/cisco_study.tex
adds the cisco study paper

\section{Automated Example Oriented REST API Documentation at Cisco}

%\author{\IEEEauthorblockN{S M Sohan}
%\IEEEauthorblockA{
%Security Group\\
%Cisco Systems Ltd.\\
%Calgary, Canada\\
%sosohan@cisco.com}
%\and
%\IEEEauthorblockN{Craig Anslow}
%\IEEEauthorblockA{
%Department of Computer Science\\
%Middlesex University\\
%London, United Kingdom\\
%c.anslow@mdx.ac.uk}
%\and
%\IEEEauthorblockN{Frank Maurer}
%\IEEEauthorblockA{
%Department of Computer Science\\
%University of Calgary\\
%Calgary, Canada\\
%frank.maurer@ucalgary.ca}
%}

\section{Abstract}
Generating and maintaining an up-to-date API documentation is a challenging problem for evolving REST APIs. At Cisco, we've used SpyREST, an automated REST API documentation tool, via our functional tests to solve this problem with one of our APIs for a cyber security application over the past eighteen months. Using this approach, we've avoided the need for extensive manual effort by leveraging our test code to also generate a continuously updated API documentation as the API evolved. Our always-updated API documentation has helped creating a fast feedback loop between the developers and QA engineers. The findings from this paper can be used by practitioners to introduce automation to reduce the manual effort associated to their REST API documentation process.

%\begin{IEEEkeywords}
%API; REST; Documentation; Tool; Case study; Test; Automation; HTTP; Web API;

%\end{IEEEkeywords}

\section{Introduction}

Researchers identified the documentation of APIs as both the primary source of information as well as the key obstacle for API usability \cite{Robillard_what_makes}. To this regard, researchers have identified the qualities of ``good API documentation'' as follows: complete, correct, includes thorough explanations and code examples, provides consistent presentation and organization \cite{Robillard_what_makes,Myers_study}. In our previous work, we introduced a novel technique and SpyREST, an implementation, based on an HTTP proxy server to automatically intercept example REST API calls and synthesize the data to produce REST API documentation to meet the aforementioned qualities \cite{DBLP:conf/kbse/SohanAM15}.

The primary motivation of this paper is to share the results of an evaluation of SpyREST and it's underlying REST API documentation technique based on its production use at Cisco. While REST APIs are published and documented by a large number of today's Internet based software companies, we see a lack of published literature on the topic of reusable and automated documentation techniques for evolving APIs. By sharing our lessons learned from a production use of a new technique in an industry setting we aim to help practitioners and researchers with a practical approach to generate and maintain example oriented REST API documentation.

The team at Cisco was tasked to develop a REST API such that it could frequently evolve to support incremental releases and addition of new features with backward compatibility as needed by the business. We compared several existing REST API documentation tools to evaluate the features against the documentation requirements for this evolving API. SpyREST was selected because the team found it to provide better automation support to achieve the requirements compared to the alternatives. We expect the information on this paper to be of use for other evolving REST APIs that need to maintain an up-to-date documentation without duplicating developer effort.

SpyREST is being used in production at Cisco for the documentation of a commercial REST API of a cloud based Cyber security product. The first author of this paper is affiliated with the team behind the product. It provides us with a unique opportunity to analyze the impact of the industry adoption of a tool developed in research. Production usage over an eighteen month period also allows us to understand the problem and opportunities presented by SpyREST in depth.

Our core contributions from this case study are as follows:

\begin{itemize}
  \item \textbf{Test driven REST API documentation}. For practitioners, we discuss a reusable technique for producing example oriented REST API documentation as a byproduct from automated API test code.
  \item \textbf{Evolution of API documentation}. For practitioners, we discuss a viable process to support evolution of API documentation as the API evolves without duplicating effort.
  \item \textbf{Evaluation of SpyREST at Cisco}. We present an industrial evaluation of SpyREST and its underlying technique of generating API documentation by intercepting API examples based on our case study at Cisco. Using this as evidence, researchers can extend the interception based the API documentation technique to improve tool support for the documentation of other forms of APIs beyond REST.
\end{itemize}

The remainder of this paper is organized as follows: in the following section we present  related work to discuss the current state of research on REST API documentation. Then, we provide a brief overview of our REST API documentation technique and the tool, SpyREST, followed by a case study of using SpyREST at Cisco. Then, we discuss our lessons learned and the limitations of this case study.

\section{Related Work}

\subsection{API Usability and Documentation} % (fold)

Several papers in the existing literature have focused on identifying the characteristics that make an API usable based on case studies. Robillard studied API usability by surveying 83 software developers at Microsoft \cite{Robillard_a_field_study}. They found that 78\% of the survey participants read API documentation to learn the APIs, 55\% used code examples, 34\% experimented with the APIs, 30\% read articles, and 29\% asked colleagues. Robillard et al. found that the most severe API learning obstacles are related to the API documentation. They suggested the following requirements as must-have for API documentation: include good examples, be complete, support many example usage scenarios, be conveniently organized, and include relevant design elements. Myers et al. also recognized the documentation as a key component for API usability and suggested using examples in the documentation to answer API related questions \cite{Myers_improving_api}. Zibran et al. found that 27.3\% of the reported bugs are API documentation bugs studying repositories for 562 API usability related bugs from five different projects \cite{zibran2011useful}. Scheller et al. provided a framework for measuring API usability based on the number and types of different objects and methods that the API provides \cite{scheller2015automated}.

Kuhn et al. performed a user study with 19 professional software developers to understand requirements for tool development to support API learnability \cite{Kuhn_on_designing}. They recommended the following as requirements for API documentation: trustworthiness, confidentiality, lack of information overload and the need for code examples as first-class documentation artifacts. Shi et al. recommended API documentation tools to support editors for custom content to provide simple ways to include API usage examples without syntax errors \cite{shi2011empirical}. Ko et al. found that thorough introductions to the concepts, standards and ideas in API documentation are a prerequisite for API usability \cite{ko2011role}.

Because REST APIs are a category of APIs, the strong relationship between the documentation and usability of the API as discussed in the aforementioned papers also applies to the context of REST APIs.

\subsection{Usage Examples in API Documentation} % (fold)

% subsection tool_support (end)

Several authors introduced tool support for including usage examples with API documentation. Hoffman et al. recommended using executable examples in API documentation \cite{Hoffman_api_documentation}. They introduced the Roast test as tool support to combine prosaic descriptions of Java APIs along with executable code examples. Montandon et al. developed APIMiner as a search tool for Java APIs and recommended providing production-like API usage examples in the API documentation \cite{montandon2013documenting}. Zhu et al. developed UsETeC to extract API usage examples by automatically synthesizing JUnit test code of the APIs \cite{zhu2014mining}. Stylos et al. presented Jadeite to provide developers with faster access to relevant API documentation by using placeholders for API methods \cite{Jadeite}.

Several authors presented techniques for linking official API documentation with crowd-sourced API usage examples that is otherwise fragmented. Nasehi et al. recommended mining knowledge repositories such as StackOverflow and developer forums should be considered for retrieving useful code examples \cite{Nasehi_what_makes}. Parnin et al. found that examples of 87.9\% of all jQuery API methods are found by searching software development blogs and forums \cite{Parnin_measuring}. Wu et al. presented an Eclipse plugin called CoDocent that can automatically find code examples using various online code search engines and link with the relevant official API documentation \cite{wu2010codocent}. Chen et al. presented a technique to automatically link official documentation with crowd-sourced documentation by recording the API related web searches that are performed by developers \cite{Chen_who_asked}. Dagenais et al. presented a tool called RecoDoc to link code-like elements from API mailing lists and developer forums with their corresponding code elements \cite{dagenais2012recovering}. Treude et al. presented a machine learning based technique called SISE to augment useful information from StackOverflow to API documentation by using text similarity of API elements and StackOverflow content \cite{Treude:2016:AAD:2884781.2884800}.

We observed that the research on usage examples in API documentation related tools have focused on local APIs such as Java library APIs. While local APIs are defined by programming language specific constructs such as classes, and methods, REST APIS are defined by HTTP terminology such as HTTP request, response, and headers. We found a lack of published work on the tool support for including usage examples with REST API documentation.

\subsection{REST API Documentation} % (fold)
The concept behind RESTful API was introduced by Fielding \cite{Fielding_rest}. Maleshkova found that most REST APIs are manually documented which results in API underspecification, and a lack of support for common tasks and reusable tools \cite{Maleshkova_investigating}. Myers et al. performed a user study on the usability of a complex API for enterprise SOA \cite{Myers_study}. They recommended providing a consistent look-and-feel with explanation for the starting points and an overall map comprising of both text and diagrams, providing a browsing experience with breadcrumb trail following a hierarchy, an effective search interface, providing example code and a way to exercise the examples online without writing code. In a previous case study, we found the documentation of REST APIs are generated manually or using bespoke tools \cite{sohan2015case}.

Several authors have suggested machine readable specification languages for REST APIs that can be used to transform into API documentation and auto generated API client code such as WADL (Web Application Description Language), RIDDL, hRESTS, RESTdesc, and WIfL (Web Interface Language), OmniVoke \cite{hadley2006web, mangler2010origin,Kopecky_hrests,RESTdesc,Danielsen_validation,Ning_omnivoke}. Polak proposed a specification format for REST API using the Model-Driven Architectural principle \cite{Polak:2015:RAM:2790798.2790820}. In addition to the existing literature, several REST API description languages have been proposed by industry practitioners such as RAML, API Blueprint, and Swagger\footnote{ \url{http://raml.org/} \url{https://apiblueprint.org/} \url{http://swagger.io/}}. Listing \ref{listing_swagger} shows a fragment of REST API specification following swagger \footnote{\url{https://github.com/OAI/OpenAPI-Specification}}.

\lstset{basicstyle=\footnotesize}
\begin{lstlisting}[language=Ruby,breaklines=true,showspaces=false,showstringspaces=false,numbers=left,xleftmargin=2em,caption={Example Swagger REST API Specification},label=listing_swagger]
basePath: /v1
schemes:
  - http
consumes:
  - application/json
produces:
  - application/json
paths:
  /pets:
    get:
      summary: List all pets
      operationId: listPets
      tags:
        - pets
      parameters:
        - name: limit
          in: query
\end{lstlisting}

The primary advantage of these specification languages is code generation and automatic transformation into REST API documentation. For some strongly typed programming languages, such as Java, it is possible to extract a schema representation of the API objects from the code. However, for dynamically typed languages, such as Ruby, the lack of type information prevents automated API specification. As a result, manual effort required from the developers to maintain a separate API specification. Even when API schema automatically extracted, API developers need to manually maintain usage examples because the examples don't exist in the API code.

\begin{table}[!tbh]
  \caption{REST API Documentation}
  \label{table:rest_apis}
\begin{tabular}{|p{2in} | p{4in}|}
\hline
\textbf{Desirable Property} & \textbf{Current State} \\
\hline
Detailed introduction &
Manually edited contents are commonly used.
 \\
\hline
Includes Examples &
Commonly include manually generated API examples.
 \\
\hline
Executable Examples &
Bespoke tooling is used to provide API explorers.
 \\
\hline
Automated &
Tools rely on manually written specifications.\\
\hline
Consistent Presentation &
Includes access information, resources, actions, request and response structures and API examples.\\
\hline
\end{tabular}
\end{table}

Table \ref{table:rest_apis} contrasts the current state of tool support for REST API documentation against a set of properties that researchers identified as required for API usability. In summary, practitioners and researchers have attempted to solve the problem of REST API documentation by proposing specifications for describing REST APIs. Manual work is needed by REST API developers to generate and maintain the API specifications. These specification formats describe the structure of different API elements (the syntax), but lacks support for automatically generated executable API usage examples with realistic data (the semantics). We found a lack of published papers on the effectiveness of the aforementioned specification languages in an industry setting. In this research, we aim to fill this gap by sharing our experience of using an automated REST API documentation technique in the industry.

\section{REST API Documentation Tool Selection}

\subsection{Requirements}

Tulach summarized developer feelings about writing API documentation as follows: \textbf{``Everyone who ever tried to convince a developer to write documentation knows how hard that is.
As a result, the general feeling is that developers don't like and are not able to provide documentation for their own code.''} \cite{Tulach2008}. At Cisco, a focused group was formed involving five API developers and two QA engineers to select an appropriate REST API documentation tool that they would actually use. The group collected the following requirements: \textbf{R1.} support multiple versions, \textbf{R2.} support multiple environments, \textbf{R3.} easy to maintain, \textbf{R4.} include executable API usage examples, \textbf{R5.} use familiar tools, and \textbf{R6.} allow customization.

\subsection{Evaluation of Tools}
A list of REST API documentation tools were selected for ``spikes'', a practice used by the software engineers to better understand how a tool fits a problem in context. The API developers and QA engineers shared opinions based on their spikes about three different tools, Swagger, API Blueprint, and SpyREST. Swagger and Blueprint were researched because of their maturity and industry adoption information as found through online searches. The authors of this paper previously developed SpyREST based on their past experience outside the context of this specific Cisco API. The focused group participated in the ``spike'' and discussed the pros and cons of each selected tool. Table \ref{table:compare} summarizes the findings:

\begin{table}[!tbh]
\caption{Comparison of Tools against the Requirements}
  \label{table:compare}
  \begin{tabular}{|p{1in} | p{1.5in}| p{1.7in}|p{1.5in}|}

\hline
\textbf{Req}. & \textbf{Swagger} & \textbf{API Blueprint} & \textbf{SpyREST} \\
\hline
R1 (version) & Separate file / version & Separate file / version & API example / version \\
\hline
R2 (env.)& Separate file / environment & Separate file / environment & Automated \\
\hline
R3 (maint.)& Manually maintain separate artifact from code & Manually maintain separate artifact from code & Derived from automated tests \\
\hline
R4 (exe. exampl.)& Automated (Web UI and cURL) & Manual & Automated (cURL) \\
\hline
R5 (familiarity)& Swagger API Spec and code inspection & Markdown & Automated tests \\
\hline
R6 (custom.)& Manual & Manual & Automated tests and Manual \\
\hline
\end{tabular}
\end{table}

As shown in Table \ref{table:compare}, the verdict to use SpyREST was based on the observation that using Swagger or API blueprint required us to manually maintain the REST API documentation as a separate static artifact from the API related code. SpyREST was found to provide better support for R3, R4, and R5 over Swagger or API Blueprint as the developers could auto-generate the documentation from the functional test suite instead of ``writing separate documentation'' that are required to be written and maintained with the code. A round of informal feedback on SpyREST was collected after the first production release of the API and the team decided to continue using SpyREST.

\section{SpyREST -- Overview} % (fold)
\label{sec:overview}
We provide a brief overview of SpyREST, the REST API documentation tool and the underlying technique used in this case study \cite{DBLP:conf/kbse/SohanAM15, DBLP:conf/kbse/SohanAM15a}. At the heart of the technique is a pass-through HTTP proxy server which acts as an interceptor between an API client and the API. This allows the proxy server to inspect the raw HTTP request and response data from the example API calls. However, for usable API documentation the raw HTTP data needs to be further processed and enriched with meta data. For example, given the following HTTP request and response data from an example API call to create a Blog post:
\begin{lstlisting}
  Request Verb: POST
  Request URL: /v2/posts
  Request Headers:
    Content-Type: application/json
    Authorization: Basic dXNlcjpwYXNzd29yZA==
  Request Body:
    {
      "title":"My New Blog post",
      "content": "This is a new blog post"
    }

  Response Headers:
    Location: "/v2/posts/1"
    host: "blog.example.com"

\end{lstlisting}

A series of transformation needs to take place to produce usable API documentation. The proxy server used by SpyREST is customized to record and synthesize such example API calls. The transformation process involves the following analyzers:

\begin{itemize}[leftmargin=*]
  \item \textbf{API version analyzer}. The version analyzer automatically infers the API version used by the example API call. From the aforementioned example, the version analyzer auto-detects the API version as ``v2'' based on the URL.

  \item \textbf{API resource analyzer}. To generate a hierarchical representation of the API elements, it's important to group multiple API actions that correspond to a single API resource under a single hierarchy. API resource analyzer automatically detects $post$ as the API resource for this example.

  \item \textbf{API action analyzer}. The API action analyzer automatically infers the API action on a resource from example API calls. For the given example, the action analyzer identifies $POST /v2/posts$ as the API action.

  \item \textbf{API query parameter analyzer:}. The query parameter analyzer records each query parameter that is used by the example API calls and automatically infers the data types such as integer, string, timestamp, etc.

  \item \textbf{API request header analyzer}: The request header analyzer removes  standard authorization credentials from the API documentation. For the given example, the API requests header analyzer transforms the $Authorization$ header's value as $Authorization: Basic FILTERED$.

  \item \textbf{API body analyzer}: The body analyzer captures the request and response bodies and infers the structure of the body as an object with auto-detected field names and data types. The request body analyzers produces the following information from the given example for API documentation:\\

  \begin{tabular}{|l | l | l|}
    \hline
    Field & Data Type & Example \\
    \hline
    title & String & My New Blog post \\
    \hline
    content & String & This is a new blog post \\
    \hline
  \end{tabular}\\

  \item \textbf{API response header analyzer}: The response header analyzer automatically removes noisy headers such as (Server, Pragma, etc.) from the documentation.

  \item \textbf{Custom content analyzer}: SpyREST automatically infers a human readable description for each API action. For the given example, the auto detected description is as follows: $Create a Post$. The custom content analyzer allows API developers to override the auto-detected attributes such as the API version, resource, action and description by using a set of SpyREST specific HTTP headers.
\end{itemize}

\begin{figure*}
  \begin{subfigure}[b]{\textwidth}
    \centering{
      \includegraphics[keepaspectratio, height=0.4\textheight]{overview.png}
    }
    \label{fig:screenshot_overview}
    \caption{API Overview in SpyREST}
  \end{subfigure}

  \begin{subfigure}[b]{\textwidth}
    \centering{
      \includegraphics[keepaspectratio, height=0.4\textheight]{fields.png}
    }
    \label{fig:screenshot_fields}
    \caption{SpyREST Analyzed API Structure}
  \end{subfigure}
\end{figure*}
\begin{figure*}\ContinuedFloat
  \begin{subfigure}[t]{0.5\textwidth}
    \centering{
      \includegraphics[keepaspectratio, width=\textwidth]{request.png}
    }
    \label{fig:screenshot_request}
    \caption{SpyREST Recorded API Request}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\textwidth}
    \centering{
      \includegraphics[keepaspectratio, width=\textwidth]{response.png}
    }
    \label{fig:screenshot_response}
    \caption{SpyREST Recorded API Response}
  \end{subfigure}

  \caption{Screenshots of the REST API documentation from Cisco using SpyREST}
  \label{fig:screenshot}
\end{figure*}

Fig. \ref{fig:screenshot} shows SpyREST screenshots of the production REST API documentation at Cisco. Fig. \ref{fig:screenshot_overview} shows a fragment of the manually written overview information juxtaposed to the auto-generated index of the API resources. Fig. \ref{fig:screenshot_fields} shows the output of the different analyzers within SpyREST proxy server that automatically detects the version (v1), API resource (Computer), action (PATCH /v1/compouters/...), and the structure of the response fields from the single example API call. Fig. \ref{fig:screenshot_request} shows the transformed example API request headers, body, and an executable cURL command that can be used to exercise the API call. Fig. \ref{fig:screenshot_response} shows the API response headers and body for the example.

To summarize, SpyREST provides tool support to improve the process of REST API documentation with usage examples, automatic updates, executable examples and a consistent hierarchical representation of the API.

\section{Case Study: SpyREST at Cisco}

\subsection{Methodology}
For this case study, we use a specific REST API within Cisco. SpyREST has been used over the past eighteen months and is being used at the time of writing this paper to document the REST API for a cloud delivered cyber security application at Cisco named Advanced Malware Protection for Endpoints. The API allows the customers of the cyber security application to extract and modify security related data specific to their businesses from the cloud to facilitate automation and custom third-party integration. For example, some customers use the API to automatically create an investigation ticket on their ticketing system for each malware detection event.

The REST API is documented using SpyREST via automated tests. To evaluate the outcome of the documentation process using SpyREST, we've incorporated stakeholder feedback as well as collected metrics from the various log files associated with the development and deployment of API docs and the actual API changes. 25 API stakeholders from Cisco were invited via email to provide free form feedback about the API documentation and its underlying process via emails to the first author of this paper.

\subsection{Context}
\subsubsection{The Stakeholders} % (fold)
The REST API has several stakeholders as follows: customers, API client developers, customer support team, product management, API developers, and quality assurance engineers. Customers and potential customers are interested in an API so that custom tooling and business specific integration can be performed. API client developers are often employed by the customers to implement API integrations. At present, over a hundred unique customers use the API at least once a month as per the API usage logs.

Inside Cisco, the stakeholders are the customer support team, product management team, API developers (including the first author) and quality assurance engineers. At present, a total of 25 people are involved as the Cisco stakeholders in this case study.

\subsubsection{Participants}
We received a total of 9 feedback emails comprising of the following: one project manager (PM), one customer support engineer (SE), five API developers (D1-5) and two QA engineers (QA1-2). Developer D2 has 5 years, and the rest of the participants each have more than 10 years of software industry experience.

\subsubsection{API Properties} % (fold)
At present, the API is currently available under two different versions, v0 and v1. The API is also served from three different Cisco managed cloud environments for North America, Europe, and Asia Pacific. Additionally, the API is shipped with a private cloud appliance, where customers can host the software on-premise. Even though the environments are designed to eventually serve the API with identical features, the deployment of the API to each environment is independently managed. The underlying application, introduced in 2010, uses Ruby on Rails \footnote{\url{http://rubyonrails.org}} web framework, and has a total of ~120K lines of custom code with a code to test ratio of 1:1.3.

In the latest version, v1, the API has nine different API resources, and a total of twenty API actions, with multiple variants per action, to be performed on those resources. For example, the API has a resource named ``Computer,'' and an action $PATCH /v1/computers/{:connector_guid}$ for updating a computer. The older version, v0, has 6 resources and 10 API actions. Currently, v1 is backward compatible with v0, but backward compatibility is not guaranteed with future versions.

\subsubsection{API Evolution}
The API was first launched with version v0 in February 2015, and a fully compatible version v1 was released in May, 2016. The monthly release notes have mentioned at least one API related change for 12 times out of a total of 18 publicly announced releases of the software. Internally, the automated functional tests for the API documentation have changed 95 times by 6 contributors (as per the version control logs). The actual number of internal API evolution before they are released is unknown since they often happen several times a day, and don't always involve a change in the functional tests. The version numbers v0 and v1 were left unchanged to release backward compatible new features and bug fixes. At present, the API serves 80,000 API calls on an average per day (as per the API call logs) and is  used by over a hundred enterprise customers that are globally distributed. The customers use the SpyREST generated documentation as the sole information source for learning the API features. The actual number of API client developers is not known at this time since multiple API client developers can be involved in writing an API client application for the customers.

\subsection{The API Development Process} % (fold)
\begin{figure*}[t]
\begin{center}
\includegraphics[width=18cm,keepaspectratio]{spyrestflow.png}
\end{center}
\caption{API Development workflow using SpyREST at Cisco}
\label{fig:spyrest_workflow}
\end{figure*}

\subsubsection{Documentation of a new API Action}

The REST API developers and quality assurance engineers of the cyber security product at Cisco use SpyREST to generate the API documentation throughout the lifecycle of an API from the developer workstation to a production instance as shown in Fig \ref{fig:spyrest_workflow}. When a new API action is introduced, the API developers write one or more functional tests against the API to show an intended usage example. If the functional test is run using SpyREST proxy server, it automatically generates the API documentation. Listing \ref{listing_test} provides an example of an automated test fragment from the case study that renders part of the documentation as shown in Fig. \ref{fig:screenshot_request} \ref{fig:screenshot_response}.

\lstset{basicstyle=\footnotesize}
\begin{lstlisting}[language=Ruby,breaklines=true,showspaces=false,showstringspaces=false,numbers=left,xleftmargin=2em,caption={Functional Test Code},label=listing_test]
context 'v1' do
  describe 'Computer' do
   it 'Moves computer to a group with given connector_guid and group_guid' do
      response = API.patch("/v1/computers/#{connector_guid}", {
        headers: 'x-spy-rest-action' => '/v1/computers/{:connector_guid}',
        body: { group_guid: group_guid }.to_json
      })
      expect(response.code).to eql(202)
      #more assertions
    end
  end
end
\end{lstlisting}

This example code is written using RSpec\footnote{\url{http://rspec.info}}, a Ruby based test framework. Line 1 mentions the API version, line 2 mentions the API resource of interest, and line 3 shows a human readable description of the test. On lines 5-7, an example HTTP patch API request is made using the method $API.patch$. Then, an example assertion is added on line 8. the $API$ class has capabilities to make the $patch$ HTTP request over an HTTP Proxy server. While using the proxy server, the $API$ class also sends the test description ``Moves computer to a group with given $connector_guid$ and $group_guid$'' with a SpyREST specific header to automatically map test description as the description of the API usage example. As discussed earlier, the SpyREST analyzers can inspect the HTTP request and response information with the custom headers to produce usable REST API documentation from this automated test code.

This practice of using tests for documentation was mentioned as an advantage by 4 of the 9 participants. For example, QA2 mentioned the following: \textbf{``SpyREST was a useful tool for us... as the (API) endpoints were documented using unit tests, (API) developers would write the documentation for a similar audience (API client developers).''} The PM mentioned the following feedback about the executable API examples in the documentation: \textbf{``The embedded examples are great for allowing users to try API calls quickly and easily to get a feel for how the API works.''}

The next step in the lifecycle of the API documentation is in the automated build server where the test suite is executed against a deployed API on a staging server through a SpyREST proxy. This promotes the REST API documentation from the developer workstation to a shared instance that is used by the quality assurance engineers, product management, and the peer developers to verify against the acceptance criteria. This often results in several loops between the developer workstations and the staging servers until the API and its documentation meets the acceptance criteria. During the team retrospective following the API production release, the following comment was captured in the meeting notes by one of the QA engineers (QA1):  \textbf{``Up-to-date documentation using SpyREST helped developing and collecting lots of input about the API in small pieces during the weekly meetings before production release''}. The API developers and quality assurance engineers add any custom content as needed to describe complex API concepts and prerequisites that aren't captured in the automated examples.

The next step in the lifecycle of the API is a release to one or more of the production environments that customers use. With the production release of an updated API, its documentation is automatically promoted from the staging environment to production. The promotion of the documentation is a three-step process. First, the data from the staging server is exported in a portable archive file. Then, a transformation is performed to update staging specific data such as URLs, email domains, and hostnames within the archive to match the desired production environment. Finally, the transformed archive is published to production as a read-only artifact.

\subsubsection{Documentation of a new Version for an Existing API Action}

When API version v1 is launched, all the functional tests for v0 are run on a loop, once per version. This code reuse minimizes the effort required to document multiple versions of the same API action when the API versions are backward compatible. The rest of the lifecycle steps follow similar process as the documentation of a new API action.

\subsubsection{Documentation of an updated API Action}
When an API update requires a new example, a new automated test case is written. For example, an API action was updated to receive a new query parameter to support an additional operation.

When an API update doesn't require a new API example to describe the change, the API documentation is automatically updated by the build server. For example, in the case study an API action was updated to include a new field with the response body and the API documentation was automatically updated without needing a change in the test code.

\subsubsection{REST API Evolution} % (fold)
\label{sub:rest_api_evolution}
In the case study we have observed the evolution of the API and it's documentation is triggered by the following: new requirements, and internal reviews of the API. For example, the primary difference between v0 and v1 of the API is the addition of a new requirement to allow API client developers the features to modify API objects in v1 that only allowed read-only access on v0.

While the new requirements are developed, before they are released to production, the API undergoes a more frequent evolution, several times a day, triggered by internal reviews of the API by the peer API developers and the quality assurance team. The auto-generated API documentation using SpyREST is used in the internal reviews to suggest alternatives and verify API acceptance criteria.

API evolution before and after the production release of the API has two opposing forces, yet one helps achieve the other. After an API is published to production, we are unable to make any breaking changes without affecting customers. To support this feature, we found it important to be able to evolve the API frequently before the API is published. The always up-to-date API documentation with SpyREST has helped establish a quick feedback loop between the API development and QA teams throughout the life-cycle of the API.

\section{Discussion}

\subsection{REST API Documentation from Test Code}
API developers are required to write the tests for the APIs even if not used for documentation. At Cisco, we have learned that the documentation of the REST API from its functional test code to be a welcome side-effect. Developer D3 mentioned the following feedback: \textbf{``Using a set of functional tests to both ensure correctness of the API, as well as a source of documentation is an excellent strategy''}. When the data from the tests are intercepted to generate API documentation, API developers need to actively think about the API usage scenarios against a realistic setup. For example, one of the Cisco developers updated the test name for an API from ``Finds computers that have connected with an IP for v0'' to ``Fetches list of computers that have connected to a given IP address'' to better reflect an API action for documentation based on internal review. This enforces the need for a stable test suite since a failing or incomplete test suite also breaks the API documentation. Eight of the nine study participants mentioned the executable API examples as a key feature of SpyREST that helped them to review the details of the API without writing code.

Continuously updated documentation improves its correctness and verification of version compatibility as an API evolves. A breaking change in the API fails the underlying tests. For non-breaking changes, such as addition of new REST API objects, the test suite helps proving backward compatibility.

One limitation of this approach is that not all API tests are suitable to be used in the REST API documentation. For example, even though unhappy paths are commonly written in tests, they are seldom included in the documentation for brevity. API developers may need to select a subset of the tests that are run through SpyREST proxy server that are meaningful for documentation.

Based on our experience at Cisco, we recommend REST API developers to utilize the automated test suite to drive API usage examples in the API documentation using SpyREST. SpyREST can intercept REST API calls irrespective of the technology used to implement the API and the API test client because it works as an HTTP proxy. The approach shared in this paper can be applied to document REST APIs within other organizations where the technology behind the API implementation may differ.

\subsection{Maintaining Custom Content}
Based on the feedback on SpyREST by Cisco developers, we've learned both the benefits and drawbacks of using a wiki-like editor for custom content with the API documentation. The wiki within SpyREST allows the API developers to add rich content comprising of proses, images, tables, code fragments, and web links, etc. to explain concepts that are required to understand the API. One developer feedback (D1) mentioned the following about the wiki: \textbf{``Good customization on the overview of each resource.''} Developer D3 mentioned the following: \textbf{``Being able to edit the documentation in place, similar to a wiki works really well, effectively allowing an organization to crowd-source the job.''} For example, at Cisco, we've added an overview explaining API access information, rate-limit, and common approaches to perform pagination within the API actions as custom content.

The custom content using a wiki has a drawback as it's maintained as a separate artifact within SpyREST. Developer D4 mentioned the following feedback: \textbf{``Tests are great for examples and showing lists of available endpoints and parameters, but there also needs to be some explanation of how to use the API in general.''} Since the wiki-edited custom content lives outside the code repository, manual effort is needed to maintain the evolution of custom content when needed. One of the Cisco developers (D2) mentioned the following in the feedback about SpyREST: \textbf{``It is helpful for deployment if the content in the overview section on the home page can be auto-generated through SpyREST code.''}

We identify this as an opportunity to continue further research to improve the collaborative editing of rich customized content with auto-generated API documentation such that the custom content can be put into the version control systems.

\subsection{Handling Flexible API Elements}
The API elements within REST are not restricted to follow a strict structure. For example, in the case study, we have a flexible API element of type ``Event'' to denote a malware detection event. The actual structure of the ``Event'' API element can be widely different based on the type of the event. For example, a file detection event contains several file specific data such as the fingerprint, name, path on disk, size. On the other hand, a network detection event contains network specific data such as the remote host, URL, IP, protocol. A single API call may return a list of ``Events'' of such different types. Incomplete list of API response attributes caused a confusion to a customer and the SE mentioned the following feedback: \textbf{``The only criticism I have is making sure that all of the returnable data that is available in the Response Fields, because I have found some instances where this data was missing''}. We've incorporated this feedback and changed SpyREST so that the ``Event'' API element is documented as a union of all the different attributes of the different event types (fingerprint, name, path on disk, size file, remote host, URL, IP, protocol) from the intercepted example API calls.

We have learned the main advantage is, merging the attributes of different types of API elements in an Array provides a list of all possible attributes that are observed in an API element. An API client developer reading the API documentation gets a complete picture of the API element. On the other hand, it may not clearly communicate the fact that only a subset of the documented attributes may be returned by an API call for each API element. In our case-study at Cisco, the API usage examples captured by SpyREST shows specific examples with different ``Event'' types to provide the context around the event types. We recommend practitioners to follow this pattern since it allows API client developers to understand the implied structure of flexible API elements with specific use cases.

\subsection{API Documentation Life-cycle}
We've learned that it is a critical feature that the API documentation follows the API throughout its life-cycle stages such as development, staging, and production environments for each available API version. In our case study at Cisco, we've achieved this requirement from the test suite that is maintained under the same version control as the REST API code. As a result, for any version of the API, corresponding documentation can be generated from test code. The PM provided the following feedback about the customer impact of this auto-updated API documentation: \textbf{``One of the great values of a self-documented API is that it's always current and up to date. This simple, but valuable fact is not lost on our customers who feel like companies often forsake documentation and leave it as an afterthought.''}

We've also learned that when REST API documentation is published from a staging environment to production, the data from the API examples may need to be obfuscated and transformed. This need was mentioned on the feedback from developer D5: \textbf{``It would be cool if you could dynamically re-write the hostname within the rails app (SpyREST), so you could hit an internal server but show docs for a public one.''}. This idea was incorporated within SpyREST to automatically rewrite all URLs and email domains within the documentation during the production release process. However, there may be other confidential information auto-captured by SpyREST that may need to be obfuscated. For example, at Cisco, one of our API examples show the user login name on a computer as an API response. To prevent the leakage of such confidential data, we run the tests against a sandbox API environment that serves dummy data. We suggest practitioners to incorporate this idea of continuous API documentation while working on APIs to improve collaboration and feedback.

\subsection{API Changelog}
We found the manual process of generating API changelog to be both error-prone and time consuming. For an evolving API, a changelog is the primary information source for existing API client developers to learn about the API changes. To verify the changes during rapid development, we've found a changelog to be a necessary communication artifact between stakeholders. SpyREST can automatically detect API changes when API objects such as query parameters, headers, request, and response objects are introduced or removed between versions because it records the API objects for each version. We have used the data from SpyREST and version control history of the tests as interim changelogs before a production release.

For production release, we manually inspect the API documentation and the interim changelogs to write a new changelog for customers. The manually written changelogs show the high level API changes leaving out the details. For example, the v1 changelog includes the following ``Move a computer from one group to another'', but further details were absent in the changelog. The underlying data model within SpyREST can be utilized to bridge the API reference documentation with relevant changelog for each version. Future work needs to be done to extend automation support for API changelog generation for evolving REST APIs.

\subsection{Cross-Referencing API Elements}
The primary navigation offered by SpyREST automatically presents a hierarchical view of the API elements comprising of API version, resources, actions on resources, and specific examples for each action. While this allows API client developers to get a quick index into the API elements, it may not provide the conceptual cross-references between API elements. Developer D2 suggested following enhancement about the SpyREST generated API Index: \textbf{``Is it possible to manually order the resources on the API doc site? Right now, the ordering is ensured in the SpyREST code.''}

We've learned that cross-referencing the API elements may provide useful alternative navigation experiences. For example, the API in the case study has the API resources ``Group'' and ``Computer,'' where a ``Computer'' belongs to a ``Group''. This nesting relationship is not captured in the SpyREST generated navigation. Similarly, the ``Group'' resource is related to a ``Policy'' resource since a ``Policy'' is applied on a ``Group.'' In the API documentation for both ``Group'' and ``Policy'' resources there are references to one another in the API response example. This dependency relationship is not captured within the SpyREST generated API navigation. Future work needs to be carried out to automatically surface such dependency relationships among API elements to help the API client developers.

\subsection{Extending beyond REST API Documentation}
From this case study, we've learned that it is feasible to use interception to generate REST API documentation with examples. While an HTTP proxy server cannot intercept non-HTTP APIs, the core concept of interception can be applied to such APIs. For example, to document a local API, an interceptor can be written to intercept example API calls in memory to generate API documentation with examples.

We identify several benefits of this approach over publishing the existing unit test code as documentation. First, organizations may not allow publishing their test code to external API users due to intellectual property issues. Second, even if unit tests are published as API documentation, the API client developers may not be familiar with the unit test framework and external dependencies. Third, unit tests often use test specific code such as complex setup, tear down, stubs and mocks, that may not be useful in the documentation. Fourth, the use of interception technique transforms task of documentation to writing executable code, a welcome change for developers. Researchers can extend the interception based documentation technique on non-REST APIs to document such APIs with executable examples.

\subsection{Limitations} % (fold)

The first author on this paper is a member of the team at Cisco where the case study is performed. While it provides us with a unique insights, it also introduces a confirmation bias. Our results may include a selection bias since the results are based on a single single set of REST APIs. While the aforementioned feedback shows that REST API documentation process is effective for us, we have a lack understanding of the impact of the documentation on the API client developers. This case study needs to be replicated against REST APIs that are implemented by other organizations using different technologies to reduce our limitations.

\section{Conclusion}
In this research, we presented the results of using SpyREST at Cisco to maintain an up-to-date documentation with usage examples for an evolving REST API. Our primary findings provide an evidence that SpyREST can be used as a practical REST API documentation tool. By leveraging automated functional tests to document REST APIs, we have reduced human efforts with automation in the API documentation process while improving the test suite. We discussed techniques to deal with custom content, flexible API elements and including the documentation of the REST API with every life-cycle step to establish a quick feedback loop. We presented our case for extending the interception based API documentation technique beyond REST APIs. In future, we plan to improve SpyREST by incorporating the ideas from this case study. We will perform an experiment to evaluate the impact of usage examples on REST API client developers.

\section*{Acknowledgments}
\footnotesize
We express our gratitude to Cisco Systems Ltd. and the University of Calgary.
\normalsize

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,cisco_study/references}

### 2017-07-22 build.sh

@@ -3,6 +3,7 @@ git clean -fX
 pdflatex root.tex
 bibtex controlled_study/controlled_study.aux
 bibtex case_study/case_study.aux
+bibtex cisco_study/cisco_study.aux
 pdflatex root.tex
 pdflatex root.tex
 /Applications/Skim.app/Contents/MacOS/Skim root.pdf

### 2017-07-22 demo_paper/demo_paper.tex
demo paper is added

%\documentclass[conference]{IEEEtran}

%\usepackage{cite}
%\usepackage{listings}
%\usepackage{amsmath}
%\lstset{showstringspaces=false, basicstyle=\footnotesize, numbers=left, numbersep=1pt }
%\usepackage[tableposition=top]{caption}
%\ifCLASSINFOpdf
%\usepackage[pdftex]{graphicx}
%\usepackage[pdftex]{color}
%\usepackage{subcaption}
%\else
%\fi
%\usepackage{url}
%\usepackage{upquote}
%\usepackage{mdframed}
%\usepackage{balance}
%\hyphenation{op-tical net-works semi-conduc-tor}

\chapter{SpyREST in Action: An Automated RESTful API Documentation Tool}

%\author{\IEEEauthorblockN{S M Sohan, Craig Anslow, and Frank Maurer}
%\IEEEauthorblockA{Department of Computer Science\\
%University of Calgary\\
%Calgary, Alberta T2N 1N4\\
%Email: \{smsohan, craig.anslow, frank.maurer\}@ucalgary.ca  }
%}
%\maketitle

\section{Abstract}
RESTful APIs are often manually documented. As a result, the process of maintaining the documentation of RESTful APIs is both expensive and error-prone. In this demonstration paper, we present SpyREST as an automated software as a service tool that can be used to document RESTful APIs. SpyREST leverages an HTTP Proxy server to intercept real API calls to automatically collect and generate RESTful API documentation by processing HTTP traffic involved in API calls. SpyREST provides an automated yet customizable example based documentation solution for RESTful APIs. RESTful API developers can use SpyREST to automatically generate and maintain updated API documentation.

%\begin{IEEEkeywords}
%RESTful API, Web API, Documentation, Automation, Example based documentation
%\end{IEEEkeywords}

%\IEEEpeerreviewmaketitle

\section{Introduction}
Fielding introduced RESTful APIs as a versatile mechanism for connecting internet based applications \cite{Fielding_rest}. For example, a hotel web site may use the Google Maps API to provide driving directions to the hotel. RESTful API developers need to provide and maintain API documentations when releasing the APIs so that the APIs can be used. Maintaining API documentation manually is less efficient and often less reliable than automated documentation \cite{Espinha_web}.

Automated solutions exist for documentation of local APIs within the context of classes, and methods. For example, JavaDoc can automatically generate the documentation for Java APIs. For RESTful APIs, instead of classes and objects, the context of APIs are defined by API endpoints, resources, actions, and example API calls in terms of the HTTP headers, request and response payload formats \cite{Danielsen_validation}. The existing tools for documenting local APIs do not natively support the different context for RESTful APIs and cannot be readily used to document RESTful APIs.

In this paper we present SpyREST which provides the necessary tool support to address the unique needs for RESTful API documentation. To this regard, in the next section we have discussed the design goals followed by a HTTP proxy server based prototype implementation to demonstrate an automated RESTful API documentation tool. We then present two use cases to show how API developers can use SpyREST to automatically generate and maintain RESTful API documentation. In the discussion, we compare SpyREST against other tools. We provide an analysis of how the requirements for SpyREST are derived in the related work section.

\section{SpyREST}

\subsection{Design} % (fold)
\label{sub:how_it_works}

SpyREST is a web-based software as a service tool\footnote{\url{http://SpyREST.com}} \cite{Sohan_spyrest}. The design goals of SpyREST is to provide an automated, technology agnostic, and example driven solution for documenting RESTful APIs so that SpyREST can be used a reusable tool to document multiple RESTful APIs . To auto-generate documentation for RESTful APIs, SpyREST relies on HTTP traffic information captured while example API calls are made using a HTTP proxy. The high level process can be described as a three step process: API developers make example API calls using a HTTP proxy server, the HTTP proxy server captures the HTTP headers, request and response data, the collected data is synthesized and presented on a web application.

The proxy server used in SpyREST is a specialized server for RESTful API documentation. A typical HTTP proxy server can intercept and record the HTTP traffic for example API calls but further processing is required to generate a usable RESTful API documentation out of the raw HTTP data. For example, API reference documentations need to be version aware so that the documentation can clearly refer to the relevant API versions. To provide human readable descriptions to the example API calls, relevant information needs to be captured in addition to the raw HTTP traffic. Post processing is required to generate the structure of request and response payloads by parsing and consolidating raw HTTP data from multiple example API calls since the optional fields may not be obvious when looked at individual example calls.

To generate usable documentation, SpyREST automatically parses the captured HTTP request and response data and presents summary tables showing the structure for HTTP query parameters, headers, request and response data. For each field, SpyREST displays the name of the field, example values and automatically inferred data type information such as integer, date time, boolean, and string. To reduce verbosity from large API responses, SpyREST automatically strips large responses to only show representative samples from repeated values in arrays. SpyREST does not capture and strip off any authorization headers used in the example API calls to prevent confidential information from being part of the documentation.

\begin{figure}[tbh]
  \centering
  \includegraphics[width=\linewidth]{spyrest_components.png}
  \caption{SpyREST Design}
  \label{fig:components}
\end{figure}

By using an HTTP proxy server, SpyREST enables the documentation of different RESTful APIs independent of the implementation technology behind the APIs. SpyREST can be used both as a software as a service tool or as a self-hosted tool when isolation is desired. SpyREST is released as an open-source product and for self-hosted solutions it can be customized to fit unique requirements.

The web interface of SpyREST provides a hierarchical navigation for RESTful API documentation as follows: each API host has one or more API versions, each API version has one or more resources, each API resource has one or more API actions, each API action may have a structure of HTTP query parameters, headers, request and response payloads as well as many API examples, each API example may have HTTP query parameters, headers, request and response payloads. The web interface also provides a wiki-like editor to allow API developers to override auto generated API documentation. On each page, the web interface includes collaboration features so that API developers can discuss questions and collect feedback about specific API documentation pages. For each API example, SpyREST generates a cURL command that can be run to execute the API examples without writing any code.

\subsection{Implementation} % (fold)
\label{sub:implementation}

SpyREST is composed of three main components as shown in Fig. \ref{fig:components}. The ``Spy'' component is the customized HTTP proxy server. This component is written using NodeJS. The Spy has internal logic to decode HTTP request URL and headers and auto-detect API versions for the commonly used version identifier patterns. Table \ref{table:versions} shows examples of auto detected versions from accept HTTP header and URLs.

\begin{table}[!tbh]
  \caption{Examples of auto detected versions}
  \begin{tabular}{|p{0.2\linewidth}|p{0.5\linewidth}|p{0.2\linewidth}|}
    \hline
    \textbf{Type} & \textbf{Example Value} & \textbf{Detected version}\\
    \hline
    ``accept'' header & application/vnd.ex.ca.v3+json & v3\\
    \hline
    ``accept'' header & application/vnd.ex.ca.v3.1+json & v3.1\\
    \hline
    URL & \url{/v2/x} & v2\\
    \hline
    URL & \url{/v2.1/x} & v2.1\\
    \hline
    URL & \url{/v2.1-pre/users} & v2.1-pre\\
    \hline
  \end{tabular}
  \label{table:versions}
\end{table}

The Spy also parses the URL to auto detect the API resources, API actions and query parameters for the example API calls. For example, given an example API call made to \texttt{GET \url{https://api.github.com/repositories?since=100}}, the Spy automatically detects the API resource as \texttt{repositories}, the API action as \texttt{GET /repositories}, and a query parameter \texttt{since} with an example value of \texttt{100}.

Even though the auto detected API version, resource and actions cover the commonly observed patterns, API developers can override the auto-detections of any of these fields by using custom Spy specific HTTP headers when making the example API calls. The HTTP headers \texttt{x-spy-rest-version}, \texttt{x-spy-rest-resource}, and \texttt{x-spy-rest-action} can be used to override auto-detection of these respective fields. Additionally, API developers can use \texttt{x-spy-rest-desc} header to attach human readable descriptions for each API example so that the web interface can tag the examples against meaningful descriptions.

The ``Web App'' component is implemented using Ruby on Rails. To allow RESTful API developers to edit auto generated documentation, the Web App uses the Markdown Syntax. To facilitate collaboration across all the API documentation pages, the Web App uses Disqus, a popular third-party collaboration service, for commenting.

The Spy writes the captured and synthesized API example data into a MongoDB ``Database''. The Web App reads and saves custom edits on the same database.

All three SpyREST components are released as Docker containers. Docker is a lightweight virtualization solution that can either be hosted in-house or using many of the popular cloud hosting providers. The public instance of SpyREST has been tested with the three components in three separate docker containers running on a single Linux server with 512MB memory, 1 core processor, and 20GB disk space.

\begin{figure*}[!tbh]
  \begin{mdframed}
    \centering
    \begin{subfigure}[t]{0.4\textwidth}
      \includegraphics[width=\linewidth]{summary.png}
      \caption{API documentation summary section}
      \label{fig:summary}
    \end{subfigure}
    \begin{subfigure}[t]{0.5\textwidth}
      \includegraphics[width=\linewidth]{examples.png}
      \caption{API documentation examples section}
      \label{fig:examples}
    \end{subfigure}
    \caption{SpyREST Screen shots showing auto generated API documentation}
    \label{fig:spyrest_screenshots}
  \end{mdframed}
\end{figure*}

\section{Use Cases}
In this section, we describe two different use cases from the perspective of initially generating a RESTful API documentation for the first time and  maintaining the documentation as the API evolves over time.

\subsection{Initial API Documentation} % (fold)
When releasing a RESTful API, API designers need to publish the API documentation so that other developers can use the APIs. For demonstration, consider generating documentation for the ``GET /notifications (List your notifications)'' API that is exposed by GitHub\footnote{\url{https://developer.github.com/v3/activity/notifications/\#list-your-notifications}}. Using this API a developer can paginate through the list of all public repositories within GitHub. This API action takes an optional parameter ``since'' that can be used to specify a time filter.

To manually generate documentation for the \texttt{GET /notifications} API, a developer needs to follow these steps: S1) make example authenticated API calls to $GET /notifications$, S2) record the HTTP headers with request and response data, S3) remove duplicate information, S4) remove authentication information, S5) format and stylize the reduced information, S6) add custom content, S7) organize multiple formatted documents together on a web site with navigation, and S8) publish the documentation website. Steps S1-S7 need to be repeated for each of the API endpoints. Since these steps are manually performed, there is also a room for human errors in the aforementioned steps. SpyREST aims to minimize the potential for human errors by automating steps S2-S5 and S7-S8.

To use SpyREST for auto-generating documentation of an API, a developer only needs to make the example API calls via the Spy proxy server with custom HTTP headers for human readable descriptions for the API example. Custom edits to the auto generated documentation may be made by using the Markdown editor. Fig \ref{fig:spyrest_screenshots} shows a partial screen shot of SpyREST generated documentation for this API given four example API calls. SpyREST fully automates the steps S2-S5, and S7-S8 as follows: S2) the Spy proxy server records the HTTP traffic, S3) the Spy automatically strips off repeated items from the arrays in the response body, S4) the Spy does not save any \texttt{Authorization} header, S5) the Web App renders the documentation for all APIs following a consistent look and feel, S7) the Web App provides breadcrumbs for hierarchical navigation of the API objects, and S8) the recorded data is automatically published by the Web App in real-time. The API developers only need to focus on finding good example API calls for S1 and describing the complex rules about APIs that are not otherwise derivable from looking at the request and response data alone for S7. In addition to automating the RESTful API documentation process, SpyREST also displays a cURL command for each recorded API example so that developers can try the API examples without having to write any code. Using in page collaboration developers can discuss API related questions and find past conversation all in one place. In this specific example, we see the official documentation of $GET /notifications$ provided by GitHub was outdated and missing several fields (e.g. $subscription\_url$, $repository.releases\_url$, $repository.labels\_url$, $repository.notifications\_url$ and 28 more) that are automatically documented by SpyREST. API developers can also customize the auto generated summary information using a wiki-like editor on SpyREST. This allows API developers to explain behaviors that cannot be derived from analyzing the individual example API calls. For example, users can edit the table of Query Parameters to explain any parameter that is not used in the example API calls or explain validation rules (such as required, interdependent fields) about the query parameters.

\subsection{Maintaining API Documentation for evolving APIs} % (fold)
Maintenance of RESTful API documentation is another use case of interest since the documentation for RESTful APIs need to evolve with the APIs to reflect the updated information. The 8 steps mentioned previously need to be repeated every time any of the API objects changes when a manual process is followed. Since APIs continue to evolve throughout their lifetime, manually maintaining their documentation incurs further costs or documentation can quickly become outdated.

SpyREST automatically updates the recorded information for each example API call, so rerunning the API examples automatically updates the auto-generated documentation. To uniquely identify each example API call, the Spy computes a digital signature of each example API call based on its API version, resource, URL and custom description. To replay the example API calls, API developers can use automated scripts so that once an API example is scripted, it can be run repeatedly. Moreover, the automated scripts can be written as functional tests where the generation of API documentation becomes a side-benefit of running the tests. Running the automated scripts to exercise the API examples frequently can prevent out of date API documentation as shown on the previous use case.

\begin{lstlisting}[language=ruby, breaklines=true, caption={}, label=list:ex, float,floatplacement=H, caption=Example API call using SpyREST, label={lst:notifications}]
class Github

  include HTTParty

  base_uri 'https://api.github.com'
  headers('Accept' => 'application/vnd.github.v3+json',
    'User-Agent' => 'curl/7.37.1',
    'content-type' => 'application/json',
    'Authorization' => "token GITHUB_API_TOKEN'"
  )

  host, port = 'spyrest.com', 9081
  http_proxy host, port

end

describe 'Notifications' do

  it 'List all notifications for the current user, where they are participating, since a time' do

    response = Github.get '/notifications?all=true&participating=true&since=2014-01-01T00:00:00Z'
    assert_equal 200, response.code
  end

end\end{lstlisting}

Listing \ref{lst:notifications} shows an example script written using the Ruby based test framework Minitest. In this example script, on line 1 the class \texttt{Github} is setup so that the Spy proxy server is used to make example API calls. \texttt{HTTParty} used on line 3 is a third-party library for making RESTful API calls. Then the required headers are setup for using GitHub API on line 6. Finally, we setup the proxy connection to SpyREST server. Then, on lines 17-25, we define an example API call as an automated test script to generate documentation for the \texttt{GET /notifications} API action with three query parameters \texttt{all, participating}, and \texttt{since}. The results of the API call can be used to test against expected results. Using this script SpyREST will generate the documentation for this API and auto-update the documentation by rerunning the script anytime the API changes while still providing test coverage. If custom edits are made to the summary section of the SpyREST generated documentation, then a re-run of the same API example will only update the example section of the documentation with new data leaving the summary section unchanged. Thus, the custom edited content sustains over multiple runs of the API examples.

\section{Discussion}
With SpyREST we have aimed to provide a solution to RESTful API documentation. A manual approach is both expensive and susceptible to human errors. SpyREST offers several benefits over existing API documentation tools. We performed a preliminary evaluation of SpyREST by generating documentation of 25 RESTful API actions from 3 different APIs providers (Github.com, KISSMetrics.com, LiquidPlanner.com) using 272 lines of test code\footnote{\url{https://github.com/smsohan/spyrest_examples}}. Comparing SpyREST generated documents against the official documentation (manually generated) we found that the official documentation for 5 of the 25 API actions had outdated response information that did not match with the actual API responses as captured by SpyREST. This is an expected advantage of automation, since updated documentation can be produced by replaying existing scripts instead of requiring any manual efforts.

Using automated tests to generate documentation serves a dual purpose for maintaining a test suite since the documentation of the RESTful API becomes an artifact from the execution of tests. This is a unique advantage of SpyREST over other API documentation tools that translate manually edited API descriptions into API documents instead of executing actual API calls.

Most general purpose API documentation tools such as JavaDoc extract documentation based on formatted comments from the source code. Since comments are not executable code, manual effort is required to keep the comments updated as the underlying API changes. RESTful API documentation typically includes additional information about HTTP headers, request and response payloads as well as example API calls that are not natively supported by general purpose API documentation tools. Also, API documentation tools for library APIs are often programming language specific. SpyREST overcomes these limitations of code comments based API documentation tools by leveraging a HTTP proxy server where executable API examples are converted into API documentation that can be used to document RESTful APIs independent of independent of what programming languages are used to implement the RESTful APIs..

SpyREST offers unique features compared to other RESTful API documentation services. Swagger\footnote{https://github.com/swagger-api/swagger-spec/blob/master/versions/2.0.md} and Blueprint\footnote{https://github.com/apiaryio/api-blueprint} are two software as a service tools for RESTful API documentation. The key difference between SpyREST and these tools is in the underlying approach used. These tools require the API developers to follow custom vendor specific API specification formats to describe the API objects. Once the APIs are described, the tools can auto-generate the documentations for the APIs. There is no automated tool support to generate the required API description in the vendor specific formats. Thus manual effort is required to produce the required API specifications and maintain it over time. SpyREST does not require any intermediate API specification as data is sourced for the documentation based on live API calls. Swagger and Blueprint do not support a self-hosted service to document RESTful APIs whereas SpyREST allows a self-hosted service to document RESTful APIs that cannot be externally shared for confidentiality issues.

The technology choice behind for our implementation of SpyREST components is based on our past experience of using these tools but the concepts behind SpyREST is not dependent on the aforementioned technology choice. As a proof of concept, SpyREST only analyzes JSON based API request and response payloads.

SpyREST is a proof of concept implementation for a novel RESTful API documentation technique. We identify the following limitations. SpyREST has yet to be evaluated. We plan to evaluate SpyREST with a large number of RESTful APIs as well as conduct a user evaluation involving API developers to understand the strengths and weaknesses of our tool. The current implementation of SpyREST only works on JSON based RESTful APIs. SpyREST needs to be enhanced to support other data types such as XML, and CSV. SpyREST, like other HTTP proxy servers, performs an intentional ``man in the middle attack'' to intercept HTTP traffic for example API calls over SSL/TLS. We consider the security impact of this approach to be low since the data is commonly meant to be included in documentation for public use. Otherwise, API developers can use the self-hosted mode and provide SpyREST with the required SSL keys. Web APIs often include a set of commonly used error responses across a variety of API endpoints. Additional work is required on SpyREST to automatically provide a summary of error responses based on example API calls. SpyREST also needs to be extended to support auto-linking between different API endpoints so that API developers can navigate between related API documentations.

\section{Related Work}
Several papers have studied API documentation to understand and recommend requirements for effective API documentation tools. Robillard recommended the following requirements for API documentation: include good examples, be complete, support many complex usage scenarios, be conveniently organized, and include relevant design elements \cite{Robillard_what_makes, Robillard_a_field_study}. Kuhn et al. discussed the importance of examples in API documentation as a key recommendation based on a survey of software developers using APIs \cite{Kuhn_on_designing}. They identified trustworthiness, confidentiality, and limiting information overload as other key recommendations for API documentation. Hoffman et al. recommended providing executable test cases with API documentation \cite{Hoffman_api_documentation}. Nasehi et al. recommended the use of wiki-like editing features for online API documentation to foster collaboration \cite{Nasehi_what_makes}. Parnin et al. and Chen et al. also recommended incorporating collaborative features with API documentations \cite{Parnin_measuring, Chen_who_asked}. Stepalina identified several benefits of using software as a service model for API documentations as follows: cost effective yet powerful, platform agnostic and high accessibility, improved document quality, content re-use, automated tools, and organization of robust and scalable documentation process \cite{Stepalina_saas}.

Several papers have discussed the topic of RESTful API documentation. Espinha et al. found most RESTful API documentations to be less reliable because they are manually generated \cite{Espinha_web}. Danielsen et al. proposed a vocabulary called Web Interface Language (WIfL) for documenting RESTful APIs so that different APIs can be described using a standard terminology \cite{Danielsen_validation}. RESTdesc, RDDL and hRESTS have been proposed as custom specifications to describe RESTful APIs \cite{Verborgh_functional, Mangler_rddl, Kopecky_hrests}. For RESTful API documentation, Myers et al. recommended providing a consistent look-and-feel and an overall map comprising of both text and diagrams, providing a browsing experience with breadcrumb trail following a hierarchy, an effective search interface, providing example code and a way to exercise the examples without writing code \cite{Myers_study}.

Our work on SpyREST is a result of following the recommendations as found on the aforementioned related work as well as critically analyzing existing API documentation tools such as JavaDoc, Swagger, and Blueprint, to identify the unmet requirements. Based on our analysis, we incorporated the following recommendations as requirements for SpyREST: automated RESTful API documentation, executable examples, consistent navigation and look and feel, software as a service model, wiki-like editing, automatic stripping of confidential data, reducing information overload, and in place collaboration.

\section{Conclusion}
In this paper we have presented SpyREST as a prototype tool to demonstrate an automated solution to RESTful API documentation. Our main contribution is a novel tool that leverages a HTTP proxy server to intercept and automatically extract API documentation from example API calls. SpyREST provides an integrated tool for RESTful API documentation by providing features for automated generation, customization, maintenance, collaboration and executable examples under a single cloud based software as a service platform. By automating the repeated parts of the RESTful API documentation process, SpyREST provides a more effective alternative to the manual process. API developers can use SpyREST to save time and costs for generating and maintaining RESTful API documentation. In the future we plan to conduct a quantitative evaluation by using SpyREST to auto generate documentation for a large set of RESTful APIs. We will also conduct a qualitative user study involving API developers to collect feedback about the strengths and weaknesses of SpyREST.

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,demo_paper/references}

### 2017-07-22 build.sh

@@ -2,6 +2,7 @@ set -e
 git clean -fX
 pdflatex root.tex
 bibtex controlled_study/controlled_study.aux
+bibtex demo_paper/demo_paper.aux
 bibtex case_study/case_study.aux
 bibtex cisco_study/cisco_study.aux
 pdflatex root.tex

### 2017-07-22 spy_rest/spy_rest.tex
adds all 5 papers, reasonable ok formatted

\chapter{SpyREST: Automated RESTful API Documentation using an HTTP Proxy Server}

%\author{\IEEEauthorblockN{S M Sohan, Craig Anslow, and Frank Maurer}
%\IEEEauthorblockA{Department of Computer Science\\
%University of Calgary\\
%Calgary, Alberta T2N 1N4\\
%Email: \{smsohan, craig.anslow, frank.maurer\}@ucalgary.ca  }
%}
%\maketitle

\section{Abstract}
RESTful API documentation is expensive to produce and maintain due to the lack of reusable tools and automated solutions. Most RESTful APIs are documented manually and the API developers are responsible for keeping the documentation up to date as the API evolves making the process both costly and error-prone. In this paper we introduce a novel technique using an HTTP proxy server that can be used to automatically generate RESTful API documentation and demonstrate SpyREST, an example implementation of the proposed technique. SpyREST uses a proxy to intercept example API calls and intelligently produces API documentation for RESTful Web APIs by processing the request and response data. Using the proposed HTTP proxy server based technique, RESTful API developers can significantly reduce the cost of producing and maintaining API documentation by replacing a large manual process with an automated process.

%\IEEEpeerreviewmaketitle

%\begin{IEEEkeywords}
%RESTful API, Web API, Documentation, Automation, Example based documentation
%\end{IEEEkeywords}

\section{Introduction}
RESTful APIs, introduced by Fielding, are used as a primary interconnection mechanism among modern day web based systems \cite{Fielding_rest}. For example, the website of a restaurant can use the RESTful API from Twitter to show the latest tweets mentioning the restaurant so that prospective customers can read the experience shared by others. To allow others to use their APIs, Twitter and other RESTful API developers publish documentation describing different features of their RESTful API. The documentation of such RESTful APIs are often produced and maintained using a manual process that is expensive and error-prone.

API documentation for library APIs, such as Java Standard Edition, commonly leverage reusable tools such as JavaDoc. The documentation produced by such tools include description of objects and methods, with custom texts primarily sourced via comments in the source code. On the other hand, RESTful API documentation includes additional information such as HTTP headers, request parameters, request and response data in serialized formats (e.g. JSON, XML). Using comments for these additional information requires significant manual effort. There is a lack of reusable tools to automate the documentation process. This makes the task of producing RESTful API documentation a costly and error-prone one. API developers also need to publish and often maintain the documentation for multiple versions as the RESTful API evolves. This requires further manual effort.

To produce RESTful API documentation, the manual process used can be described as a six-step process as follows: 1) craft an example call to an API endpoint with required headers, URL parameters and request body, 2) make the call, 3) capture the response headers and data, 4) strip any confidential data from the captured information, 5) add custom descriptions to the captured data and 6) publish the API documentation. This six-step process is essentially repeated for all API endpoints that are documented. With \emph{SpyREST}, we have implemented an innovative solution to largely automate the aforementioned manual process of RESTful API documentation so that all but steps 1 and 5 are automated. Steps 1 and 5 are left to a manual process to allow for a pragmatic solution. Our solution relies on an HTTP proxy server to intercept example API calls to automatically capture all HTTP traffic. The collected data is then synthesized to present the documentation for RESTful APIs. SpyREST can generate documentation for all RESTful APIs irrespective of the technology used to implement the API since it leverages an HTTP proxy server.

The key contributions of our research are as follows: we discuss a list of requirements for tool development by analyzing related work to automate the RESTful API documentation process and present a new technique based on an HTTP proxy server to meet the requirements. We demonstrate an example implementation of the proposed technique that generates automated yet customizable, version-aware, collaboration enabled and reusable API documentation software as a service platform that can be used to generate and maintain documentation for any RESTful API. We present a case study of using SpyREST to compare the advantages of our proposed technique over existing solutions for RESTful API documentation.

The remainder of this paper is organized as follows: in the next section we discuss related work on API documentation. The requirements and implementation details of SpyREST is provided next. Then, we present a case study to demonstrate SpyREST in action. Finally, we discuss the implications and limitations of our work in the discussion section.

\section{Related Work} % (fold)
\label{sec:related_work}
\subsection{General API Documentation} % (fold)
\label{ssub:api_learnability}
Several papers have studied the documentation of APIs to recommend best practices that are also applicable to RESTful API documentation. Robillard et al. found that  developers faced severe obstacles learning new APIs due to inappropriate documentation and other learning resources \cite{Robillard_what_makes, Robillard_a_field_study}. Robillard recommended the following for API documentation: include good examples, be complete, support many complex usage scenarios, be conveniently organized, and include relevant design elements. Kuhn et al. discussed the importance of examples in API documentation \cite{Kuhn_on_designing}. Hoffman et al. recommended making the API example scenarios to be executable test cases so that a user can execute an API \cite{Hoffman_api_documentation}.

Nasehi et al. performed a case study based on StackOverflow discussions to find out what makes good code examples \cite{Nasehi_what_makes}. They recommended API developers to include examples in the API documentation and the use of wiki-like collaborative tools with online API documentation. Parnin et al. found that using social media gives API documentation readers a chance to engage with authors of the APIs \cite{Parnin_measuring}. Chen et al. recommended integrating crowdsourced frequently asked questions into API documents so that users can easily find relevant discussions when questions arise \cite{Chen_who_asked}. Subramanian et al. presented an automated approach to link API documentation of different Java and JavaScript libraries with code examples that are shared on StackOverflow \cite{Subramanian_live}.

Stepalina identified several advantages of software as a service (SaaS) based solutions for API documentation systems such as, cost effective yet powerful, platform agnostic and high accessibility, improved document quality, content re-use, automated tools, and organization of robust and scalable documentation process \cite{Stepalina_saas}. Several tools exist that help automatic generation of API documentation for local APIs such as JavaDoc\footnote{http://www.oracle.com/technetwork/java/javase/documentation/index-jsp-135444.html}, RDoc\footnote{http://rdoc.sourceforge.net/}, Jadeite \cite{Jadeite}. While these tools have been proven to work for local library APIs, they have limited applicability for documenting RESTful APIs because HTTP specific information are not natively supported by these tools.

\subsection{RESTful API Documentation} % (fold)
\label{sub:restful_api_documentation}

Several related work proposed specifications for RESTful APIs. Maleshkova et al. found that a lack of a standard format to document Web APIs and manual documentation causes confusions about how to use the APIs for different use cases \cite{Maleshkova_investigating}. Espinha et al. observed that most RESTful APIs are documented manually by API developers making the documentation a less reliable one \cite{Espinha_web}. Danielsen et al. presented a vocabulary for documenting RESTful Web APIs called Web Interface Language (WIfL) \cite{Danielsen_validation}. Verborgh et at. presented RESTdesc, a Resource oriented and Hyper-link based specification for describing RESTful APIs \cite{Verborgh_functional}. Mangler et al. presented RDDL, an XML based specification for describing RESTful APIs \cite{Mangler_rddl}. Kopecky´ et al. presented hRESTS, a machine readable micro-format to describe RESTful APIs that use an alternate representation compared to WIfL \cite{Kopecky_hrests}. Ning et al. presented OmniVoke, a RESTful API based invocation engine that provides an abstraction layer for RESTful API calls to multiple APIs that follow different conventions \cite{Ning_omnivoke}. Manual configuration is required to generate these aforementioned specifications for RESTful APIs. Myers et al. recommended providing a consistent look-and-feel with explanation for the starting points and an overall map comprising of both text and diagrams, providing a browsing experience with breadcrumb trail following a hierarchy, an effective search interface, providing example code and a way to exercise the examples online without writing code \cite{Myers_study}.

In addition to the research community, there are several custom RESTful API specification formats that are observed in the industry such as, Swagger\footnote{https://github.com/swagger-api/swagger-spec/blob/master/versions/2.0.md}, RAML\footnote{http://raml.org/spec.html}, and Blueprint\footnote{https://github.com/apiaryio/api-blueprint}. In addition to producing RESTful API documentation, there are SaaS based tools providers that can be used to publish the documentation and auto generate API client code for RESTful APIs that are described using one of these formats. To use these tools to generate HTML based RESTful API documentation, API developers need to manually construct the intermediate documentation format such as Swagger, Blueprint, RAML, etc. since there is no automated tool to produce this.
% subsubsection restful_api_specification (end)

\section{SpyREST} % (fold)
\label{sec:spyrest}

\subsection{Requirements} % (fold)
\label{sub:spyrest_requirements}

The following list of requirements for SpyREST, R1-7, in priority order, is derived from analyzing the aforementioned related work and current API documentation practices as observed in the industry:

  \textbf{R1 - Automated RESTful API documentation}: RESTful APIs are either documented manually or using custom tools to partially automate the process. The primary requirement for SpyREST is to find a cost-effective approach to automate RESTful API documentation.

  \textbf{R2 - Example based}: As discussed in the related work section, several authors have emphasized including example scenarios with API documentation can help users understand how to use an API \cite{Robillard_what_makes, Kuhn_on_designing, Hoffman_api_documentation,Nasehi_what_makes}. SpyREST generated API documentation needs to include example scenarios.

  \textbf{R3 - Executable documentation}: In addition to including examples, SpyREST also needs to allow users to execute the example scenarios so that they can try the API features without having to write code \cite{Hoffman_api_documentation, Myers_study}.

  \textbf{R4 - Version awareness}: SpyREST needs to allow API developers to publish documentation for multiple versions of a Web API as it evolves. The following comment from an API client developer of Stripe, an online payment processing company, shows the importance of a version-aware documentation tool\footnote{\url{https://groups.google.com/a/lists.stripe.com/forum/\#!searchin/api-discuss/version/api-discuss/li4PyVcweiw/NT9SFTtF-vQJ}}
:
  \small
  \begin{quotation}
   ``Does the full API documentation only reflect the current version of the
    API?  Is there a way to access the API docs for outdated versions? ...That would be very helpful. When you are trying to upgrade from one version to another it's impossible to know the implementation differences...''
  \end{quotation}
  \normalsize

  \textbf{R5 - Customizable}: It is important to customize the overview information about how to use an API to reduce API learning obstacles and describe complicated business rules\cite{Robillard_what_makes}. SpyREST needs to allow API developers to add customized content to auto generated RESTful API documentation.

  \textbf{R6 - Reusable}: SpyREST needs to work as a reusable platform so that multiple REST API documentations can be generated and published on a single platform to get the advantages of a SaaS platform \cite{Stepalina_saas}.

  \textbf{R7 - Collaborative}: SpyREST needs to allow people to collaborate on the API documentation so that questions and answers about APIs can coexist with the API documentation to to overcome the knowledge fragmentation \cite{Chen_who_asked}.

\subsection{Design} % (fold)
From our research, we identified an HTTP proxy server based solution as a novel approach for generating RESTful API documentation meeting requirements R1-7. Our approach works as follows: API developers make example API calls using a proxy $\longrightarrow$ the proxy records and processes HTTP traffic $\longrightarrow$ a Web App generates RESTful API documentation. SpyREST, an example implementation of this approach\footnote{\url{https://github.com/smsohan/demo_paper/raw/master/SpyREST_tool_demo.pdf}}, is composed of three main components as follows:

\textbf{The Spy:} The Spy component in SpyREST is an HTTP proxy server. The Spy is named so because it records the HTTP traffic for example API calls that are made using it as a proxy server. For example, when using Spy as an HTTP proxy to make the following HTTP request:

\footnotesize
\begin{description}
  \item[Method] GET
  \item[URL] https://api.github.com/repositories?since=100
  \item[Headers] accept: application/vnd.github.v3+json
\end{description}
\normalsize
that produces the following response:
\footnotesize
\begin{description}
  \item[Headers] status: 200 OK \\
content-type: application/json; charset=utf-8 \\
...
  \item[Body]
\begin{lstlisting}
[
  {
    "id": 1,
    "name": "grit",
    "full_name": "mojombo/grit",
    ...
  }
]\end{lstlisting}
\end{description}
\normalsize
Spy automatically saves the raw request and response data with HTTP headers. Additionally, the Spy synthesizes the data and saves the following meta data about this example API call:

\footnotesize
\begin{tabular}{r p{0.55\linewidth}}
  version & v3 \\
  resource & repositories \\
  action & GET /repositories \\
  query & since: 100 \\
  strippedResponseBody & a subset of the response body \\
  description & blank \\
  digest & base64 hash value of the version, resource, url and description \\
  requiresAuth & false \\
  apiToken & blank, used when a SpyREST API token is provided \\
  userId & blank, used when a user is found for given apiToken \\
\end{tabular}
\normalsize

As shown in this example, the version ``v3'' is automatically detected by parsing the accept request header. Next, the resource field is also auto detected as ``repositories'' by parsing the request URL. The action ``GET /repositories'' is automatically detected by combining the request HTTP method with request path. The strippedResponseBody field automatically saves a shorter version of the actual response where large response body is truncated to a smaller one to reduce noise from the generated documentation. These meta fields allow SpyREST to structure the example API calls for a given API host in a hierarchical model as follows: an APIHost (e.g. GitHub.com) has many APIVersions (e.g. v3), each APIVersion has many APIResources (e.g. repositories), each APIResource has many APIActions (e.g. GET /repositories), and each APIAction has many APIExamples.

\begin{figure*}[!tbh]
  \centering
  \begin{subfigure}[t]{0.4\textwidth}
    \includegraphics[width=\linewidth]{spyrest_header.png}
    \caption{API documentation summary section}
    \label{fig:summary}
  \end{subfigure}
  \begin{subfigure}[t]{0.5\textwidth}
    \includegraphics[width=\linewidth]{spyrest_examples.png}
    \caption{API documentation examples section}
    \label{fig:examples}
  \end{subfigure}
  \caption{SpyREST Screen shots showing auto generated API documentation}
  \label{fig:spyrest_screenshots}
\end{figure*}

\textbf{The Database:} The database component saves the data that is captured by the Spy and the Web App. The Spy saves the recorded data about the API examples and the auto-computed meta data in the database. The Web App saves data about SpyREST users and custom modifications on auto-generated API documentation on the same database.

\textbf{The Web App:} The Web App component further processes and presents auto generated documentation that is captured by Spy in the database. The Web App also allows the API developers to edit or add custom content to the auto generated documentation. The source code for both the Spy and Web App components are released as open source projects.

\subsection{Features} % (fold)

Now that we have explained the design and implementation of SpyREST, we discuss how SpyREST meets the aforementioned requirements R1-7.

\textbf{R1 - Automated RESTful API documentation:} The work flow for automated RESTful API documentation can be explained by the following steps: 1) API developer uses the Spy HTTP proxy to run example API calls, 2) The Spy records example API calls, 3) The Spy extracts meta information about the API call, 4) The Web App displays the auto generated documentation, 5) API developers can optionally customize the auto generated documentation.

Fig \ref{fig:spyrest_screenshots} shows screen shots (shortened) from SpyREST auto generated documentation that is solely based on an example API call to URL \url{https://api.github.com/search/repositories?q=tetris+language:assembly&sort=stars&order=desc} with a custom header ``x-spy-rest-description'' to provide the short description for the example.

The auto generated API documentation features two sections, a summary section as shown in Fig. \ref{fig:summary} and an examples section as shown in Fig. \ref{fig:examples}. The summary section includes breadcrumbs to show hierarchy of the API objects related to each API action to help API client developers easily navigate the API documentation. This section also includes three tables that display the structure of query parameters, and request and response payloads. In addition to the structure, automated type detection is used to show the data type and example values for each field in these tables. The examples section on the API documentation shows all the recorded API examples for a given API action. For each example, it shows a description, the request URL, query parameters, and request and response headers and bodies. Spy filters out the ``authorization'' request header before saving the examples in the database and the documentation rendered by the Web App displays this header with a placeholder text as ``FILTERED''.

\textbf{R2 - Example based:} SpyREST generated RESTful API documentation includes both the structure of the API objects and concrete examples for different use cases. The API examples are annotated with user provided descriptions through a custom HTTP request header ``x-spy-rest-description'' that is otherwise hard to automatically infer. To record the examples in SpyREST, any REST API client can be used as long as it supports an HTTP proxy to call the API.

\textbf{R3 -  Executable documentation:} SpyREST keeps a copy of the API request information including URL, headers and request body. As a result, SpyREST can also recreate the example API calls that can be executed by users of the generated API documentation. SpyREST automatically generates executable test cases that can be run using cURL\footnote{http://curl.haxx.se/}.

\textbf{R4 - Version awareness:} SpyREST allows API developers to publish the documentation for multiple versions of their RESTful APIs. To organize API documentation and examples under multiple versions, SpyREST has an automatic version detection algorithm that parses the ``accept'' request header or the URL. The auto version detection algorithm can be suppressed by specifying the version in the custom ``x-spy-rest-version'' header when running example API calls through the Spy.

\textbf{R5 - Customizable:} SpyREST allows API developers to modify and add custom free-form contents to the auto generated summary section of the documentation using Markdown\footnote{\url{http://en.wikipedia.org/wiki/Markdown}} syntax. Custom content can be used to provide overview information and explain complicated business rules about the APIs that are not derivable from simply synthesizing the examples. We consider this as a pragmatic solution to augment manual effort with the largely automated solution to the RESTful API documentation effort. The custom edits are persisted in the database and are not overridden when the API examples are replayed unless the user decides to revert back to automated summaries.

\textbf{R6 - Reusable:} SpyREST relies on an HTTP proxy to auto generate RESTful API documentation. To feed SpyREST with data, any REST API client can be used as long as it supports using a proxy server. This technology agnostic feature allows SpyREST to be a reusable RESTful API documentation tool. SpyREST is offered as a SaaS tool at \url{http://spyrest.com} that can be used to auto-generate and publish documentation of multiple RESTful APIs. For APIs where a SaaS solution is not acceptable, SpyREST can also be installed and used as a self-hosted solution. Because SpyREST is an open-source application, users can modify the source to support any unique requirements that are not supported by SpyREST.

\textbf{R7 - Collaborative:} SpyREST allows the API developers to comment and discuss API related questions with API client developers right next to the documentation so that crowdsourced documentations can coexist with the officially published API documentation. Each API documentation page on SpyREST features its own discussion thread to help users easily locate relevant information.

\section{SpyREST Case Study}
\label{sec:case_study}
In this section we discuss a case study of using SpyREST to auto generate documentation for 25 API actions randomly sampled from three RESTful API providers. These 25 API APIs are documented using 272 lines of code\footnote{\url{https://github.com/smsohan/spyrest_examples}}. These three API providers are: GitHub.com, KISSMetrics.com (Online analytics tool), and LiquidPlanner.com (Online project management tool).

\subsection{SpyREST Documentation vs. Official Documentation} % (fold)
\label{sub:spyrest_vs_official_documentation}

We found SpyREST generated documentation for 5 of the 25 API actions from the case study included fields that are found from actual API responses but not included in their official documentation. For example, the SpyREST generated documentation for Github \texttt{GET /notifications} API action included 34 additional fields, such as: \texttt{forks\_url, keys\_url, collaborators\_url}, and 29 more that were not mentioned in the official API documentation even though actually returned as API response. Similarly, the official documentation for GitHub.com \texttt{GET /search/code} API action did not include the \texttt{releases\_url} API response field. Official documentation of KISSMetrics.com \texttt{GET core/accounts} and \texttt{GET core/accounts/:account\_id} API actions did not mention the \texttt{data} field that is found in the actual response as documented by SpyREST. The official documentation for \texttt{GET /api/account} action on LiquidPlanner.com did not mention the fields \texttt{workspaces, last\_workspace\_id, and disabled\_workspaces\_count} that were included in the SpyREST documentation. These examples show the error-prone nature of a manual process that requires API developers to ensure the documentation is updated to reflect any change in the API. We consider the proxy based solution to solve this problem since the API documentation can be updated by replaying the example API calls.

The official documentation of these three API providers did not include integrated collaboration features. We found 172 unanswered questions out of 662 questions with the tag \texttt{github-api} on StackOverflow.com. We found 123 questions about LiquidPlanner API on their developer forum that are not linked to officially generated documentation published as a PDF file. On StackOverflow.com, we found 4 unanswered questions out of 9 questions on KISSMetrics API. SpyREST provides the collaboration features with auto generated API documentation so that API developers and users can discuss and locate related discussions when browsing RESTful API documentation on a single web interface. SpyREST includes executable API examples, support for multiple versions and presents the auto generated documentation for different RESTful APIs using a consistent look and feel that are not provided by these studied APIs.

\section{Discussion}
RESTful API documentation is expensive, error-prone, and often incomplete because of the manual effort involved in the process. Our core contribution is a novel approach where an HTTP proxy server is used to largely automate the process for RESTful API documentation. There is a lack of tool support for automatically generating documentation of RESTful APIs. We presented SpyREST as an example implementation of the HTTP proxy server based solution to fill this need. The requirements for SpyREST are derived from analyzing the existing literature and industry practices as well as our own professional experiences developing RESTful APIs and their documentation. In the literature, several papers provided lists of recommendations for API documentation based on studying the existing tools and techniques and feedback collected from API developers. We found a lack of available tool support to generate RESTful API documentation following those recommendations and designed SpyREST to provide a solution. As a result, SpyREST features automated documentation with executable examples, customizable contents, collaboration, version awareness and reusability to provide a ready to use tool support for RESTful API documentation following the recommendations.

The novel approach of using an HTTP proxy sever offers some unique benefits over existing tools from academia and industry. For example, tools that rely on user provided comments on source code and code inspection, such as JavaDoc, require the user to write formatted comments that are often applicable to a single programming language. The comments are not executable, and manual effort is required to ensure the comments and the API that it describes are kept in sync. SpyREST generated documentation can reflect the latest information since the documentation is generated after executing the example API calls. The verbose nature of RESTful API documentation makes it difficult to write and maintain the documentation as comments in source code. Using an HTTP proxy server, SpyREST offers a language agnostic solution so that RESTful APIs written in any programming language can be documented using a single tool.

Another key feature of SpyREST is the shared platform for publishing RESTful API documentation. A similar shared platform is provided by several other tools such as Swagger, Blueprint. We identify the key advantage of SpyREST over these tools to be the fact that SpyREST relies on example API calls instead of relying on a custom API specification that are required by these tools. The suggested API specifications can be used to define the structure of API objects but do not capture examples of API usage. The work flow for using these API specification formats can be described as follows: (1) execute example API calls $\longrightarrow$ (2) manually generate API specification $\longrightarrow$ (3) automatically generate API documentation. SpyREST does not rely on any custom API specification, so the associated manual effort can be largely avoided resulting in a smaller work flow as follows: (1) execute example API calls $\longrightarrow$ (2) automatically generate API documentation. We consider this process to be a cost effective approach for APIs with several endpoints or APIs that evolve. Unlike other available SaaS only solutions such as Swagger, and Blueprint, SpyREST can be used both as a SaaS and a self-hosted platform. The self-hosted mode can be used for documenting internal RESTful APIs that cannot be released on the internet or modifying the open-sourced code of SpyREST to fit unique API specific needs. We have demonstrated using SpyREST with automated test code. Using the test code helps automatically testing different use-case scenarios of the API actions in addition to generating the API documentation.

While SpyREST offers an automated solution for documenting RESTful APIs, API developers can customize the auto generated content. Several researchers have identified the need for rich content comprising of text and diagrams to explain complex business rules about APIs. To achieve this goal, we consider augmenting auto generated documentation with user contributed documentation, when necessary, to provide a pragmatic solution. To summarize, from our example implementation of SpyREST and the case-study, we have shown the advantages of the proposed HTTP proxy server based solution over the existing techniques for RESTful API documentation.

\subsection{Threats to Validity} % (fold)
\label{sub:threats_to_validity}

\textbf{Internal Threats.} Our proposed technique uses an HTTP proxy server that can only intercept and record the data when it is either in clear text or trusted to be decrypted by SpyREST. Because SpyREST can be used as a shared platform, any secrets used in the example API calls get decrypted in memory, even though not saved by SpyREST. To overcome these threat, we recommend API developers to use disposable secrets so that the secrets used to run an API call are valid for a single API call. For use cases where this is unacceptable, the self-hosted mode can be used as an alternative. SpyREST currently only supports JSON based APIs for their widespread use among popular RESTful APIs such as Twitter, Facebook, Google Maps.

\textbf{External Threats.} We used SpyREST to auto generate documentation for APIs randomly sampled from three RESTful API providers as a proof of concept and found the proxy server based solution met the requirements R1-7. Further evaluation is needed involving a larger set of RESTful APIs to evaluate the effectiveness of the proposed technique of using an HTTP proxy server for RESTful API documentation.

\section{Conclusion}
In this paper, we have presented a novel technique of using an HTTP proxy server to automate the RESTful API documentation process that otherwise requires a largely manual process. The proxy server based solution supports integrated features such as automated RESTful API documentation with executable example API calls, support for multiple versions, customization and collaboration that are offered both as a SaaS and self-hosted platforms. These features have been recommended by existing research on the documentation of APIs. Overall, we conclude that a proxy server based approach shows a pragmatic solution to the RESTful API documentation problem. In the future, we will perform a qualitative user study involving RESTful API developers to evaluate the effectiveness about the aforementioned HTTP proxy based RESTful API documentation approach.
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,spy_rest/references}

### 2017-07-22 build.sh

@@ -2,6 +2,7 @@ set -e
 git clean -fX
 pdflatex root.tex
 bibtex controlled_study/controlled_study.aux
+bibtex spy_rest/spy_rest.aux
 bibtex demo_paper/demo_paper.aux
 bibtex case_study/case_study.aux
 bibtex cisco_study/cisco_study.aux

### 2017-07-22 case_study/case_study.tex
fix the label/caption order
structurally complete, sans abstract / introduction / conclusion

@@
 \usetikzlibrary{backgrounds}
+    \label{fig:screenshot_response}
 \pgfplotsset{compat=1.5}
@@
 
-\subsection{Empirical Research on Web API evolution} % (fold)
+\subsection{Empirical Research on Web API Evolution} % (fold)
 \label{sub:empirical_research}
@@
 
-\subsection{Techniques to evolve Web APIs} % (fold)
+\subsection{Techniques to Evolve Web APIs} % (fold)
 \label{sub:techniques}
@@
 
-\subsection{Tool support for evolving Web APIs} % (fold)
+\subsection{Tool Support for Evolving Web APIs} % (fold)
 \label{sub:tool_development}
@@
     \end{tikzpicture}
+    \caption{Facebook API Changes (11 Releases)}
     \label{fig:facebook-chart}
-    \caption{Facebook API Changes (11 Releases)}
   \end{subfigure}
@@
     \end{tikzpicture}
+    \caption{Stripe API Changes (30 Releases)}
     \label{fig:stripe-chart}
-    \caption{Stripe API Changes (30 Releases)}
   \end{subfigure}
@@
     \end{tikzpicture}
+    \caption{WordPress REST API Changes (8 Releases)}
     \label{fig:wordpress-chart}
-    \caption{WordPress REST API Changes (8 Releases)}
   \end{subfigure}
@@
     \end{tikzpicture}
+    \caption{GitHub API Changes (59 Releases)}
     \label{fig:github-chart}
-    \caption{GitHub API Changes (59 Releases)}
   \end{subfigure}
@@
     \end{tikzpicture}
+    \caption{Google Maps Web API (102 Releases)}
     \label{fig:gmaps-chart}
-    \caption{Google Maps Web API (102 Releases)}
   \end{subfigure}
@@
     \end{tikzpicture}
+    \caption{Salesforce REST API (4 Releases)}
     \label{fig:salesforce-chart}
-    \caption{Salesforce REST API (4 Releases)}
   \end{subfigure}

### 2017-07-22 demo_paper/demo_paper.tex
structurally complete, sans abstract / introduction / conclusion

@@
 \begin{table}[!tbh]
-  \caption{Examples of auto detected versions}
+  \caption{Examples of Auto Detected Versions}
   \begin{tabular}{|p{0.2\linewidth}|p{0.5\linewidth}|p{0.2\linewidth}|}
@@
     \end{subfigure}
-    \caption{SpyREST Screen shots showing auto generated API documentation}
+    \caption{SpyREST Screenshots Showing Auto Generated API Documentation}
     \label{fig:spyrest_screenshots}
@@
 
-\subsection{Maintaining API Documentation for evolving APIs} % (fold)
+\subsection{Maintaining API Documentation for Evolving APIs} % (fold)
 Maintenance of RESTful API documentation is another use case of interest since the documentation for RESTful APIs need to evolve with the APIs to reflect the updated information.

### 2017-07-22 root.tex
remove unwanted files
controlled study paper is wholly included
boxes look good
adds the case study paper
adds the cisco study paper
demo paper is added
adds all 5 papers, reasonable ok formatted
structurally complete, sans abstract / introduction / conclusion

@@
-%% Modified documentstyle to documentclass -- compatibility with LaTex2e
-%% N. Mancell (98/03/01).
-
-%%  ucalgthes_root.tex        (NM 98/03/01)
-%   Modified   92-09-18      Add references to dissertation        D. Teale
-%                            Add approval page to toc
-%                            Add ref to Title Degree on approval page
-%   Modified  2006-09-12     Added geometry package to set up UofC thesis margins
-%                            Removed includeprompt option N. Mancell
-%
-% Modified  2012-04-15  T.Zhang
-% 1, the ‘fancy’ package is cancelled. The pagestyle is changed from ‘myheadings’ and ‘headings’ to ‘plain’ to move the page number (footer) from top right to bottom centre.
-% 2, The parameter  of using ‘geometry’ package is changed from four different values (top=1in, bottom=1.22in, left=1.40in, right=0.850in) to one single value: 1in (top=1in, bottom= 1in, left= 1in, right= 1in)
-% 3, ‘List of Figures’ is modified to ‘List of Figures and Illustrations’
-% 4, A new page ‘List of Symbols, Abbreviations and Nomenclature’ is created.
-Its page number is following ‘List of Figures and Illustrations’ with Roman numerals.
-A separate file ‘symbols.tex’ is created for students to put content into the list.
-
 \documentclass[12pt]{ucalgthes1}
+\usepackage[pdftex]{graphicx}
+\usepackage[pdftex]{color}
 \usepackage[letterpaper,top=1in, bottom= 1in, left= 1in, right= 1in]{geometry}
-%\usepackage{fancyhdr}
-%\fancyhead{}
-%\fancyfoot{}
-%\renewcommand{\headrulewidth}{0pt}
-%\fancyhead[RO,LE]{\thepage}
-%Define other usepackages here
 \usepackage{hyperref}
 \usepackage{mathptmx}
-\title{Title of Thesis (title case, double spaced, no more than 240 characters. No bold font or symbols, except Greek) \\ \bigskip
-Thesis Title Second Line }
+\usepackage{listings}
+\usepackage{url}
+\usepackage[tableposition=top]{caption}
+\usepackage{subcaption}
+\usepackage{tikz,pgfplots,pgfplotstable}
+\usepackage{mathtools}
+\usepgfplotslibrary{dateplot}
+\usepgfplotslibrary{statistics}
+\usepackage{courier}
+\usepackage[square,numbers,sectionbib]{natbib}
+\usepackage[sectionbib]{chapterbib}
+\usepackage[numbib]{tocbibind}
+\newcommand{\newblock}{}
+\usepackage{eurosym}
+\usepackage{multirow}
+\usepackage[T1]{fontenc}
+\usepackage{textcomp}
+\lstset{upquote=true}
+\usepackage{float}
+\floatstyle{boxed}
+\usepackage{blindtext}
+\let\labelindent\relax
+\usepackage{upquote}
+\usepackage{mdframed}
+%\usepackage{balance}
+\usepackage{enumitem}
+
+\title{Automated Example Oriented REST API Documenation}
 %
@@
 %
-\author{Student's Full Name}
-\thesisyear{year}
+\author{S M Sohan}
+\thesisyear{2017}
 \thesis{thesis}    % the word dissertation can be inserted between {}
-\newcommand{\thesistitle}{Title of Thesis}
-\monthname{month name}
-\dept{GRADUATE PROGRAM IN "NAME OF PROGRAM"}
-\degree{"FULL NAME OF DEGREE e.g.: DOCTOR OF PHILOSOPHY"}
+\newcommand{\thesistitle}{Automated Example Oriented REST API Documenation}
+\monthname{September}
+\dept{GRADUATE PROGRAM IN COMPUTER SCIENCE}
+\degree{DOCTOR OF PHILOSOPHY}
 %
@@
 \setcounter{page}{2}
-%\chapter*{UNIVERSITY OF CALGARY \\ FACULTY OF GRADUATE STUDIES}
-%\thispagestyle{empty}
-%The undersigned certify that they have read, and recommend
-%to the Faculty of Graduate Studies for acceptance, a \Thesis\ entitled
-%``\thesistitle'' submitted by \Author\
-%in partial fulfillment of the requirements for the degree of
-%\Degree.\\
+\chapter*{UNIVERSITY OF CALGARY \\ FACULTY OF GRADUATE STUDIES}
+\thispagestyle{empty}
+The undersigned certify that they have read, and recommend
+to the Faculty of Graduate Studies for acceptance, a \Thesis\ entitled
+``\thesistitle'' submitted by \Author\
+in partial fulfillment of the requirements for the degree of
+\Degree.\\
 
@@
 %
-%\begin{signing}{Department of Academic Computing}
-%\signline
-%Chairman, Dr.~John D.~Doe \\
-%Department of Academic Computing \\
-%Services  \\
-%\signline
-%Chairman, Dr.~John D.~Doe \\
-%Department of Academic Computing \\
-%Services  \\
-%\signline
-%Chairman, Dr.~John D.~Doe \\
-%Department of Academic Computing \\
-%Services  \\
-%\signline
-%Chairman, Dr.~John D.~Doe \\
-%Department of Academic Computing \\
-%Services  \\
-%\newsigncolumn         use this command to start a new column if necessary
-%\newsigncolumn
-%\signline
-%Chairman, Dr.~John D.~Doe \\
-%Department of Academic Computing \\
-%Services  \\
-%\signline
-%Dr.~Jane Smith \\
-%Department of Academic Computing  \\
-
-%\signline
-%Dr.~A.~B.~Brown \\
-%Department of Academic Computing  \\
-%\end{signing}
+\begin{signing}{Department of Computer Science}
+\signline
+Dr. Frank Maurer \\
+Supervisor\\
+Department of Computer Science \\
+University of Calgary \\
+Canada\\
+\signline
+Dr. Mea Wang\\
+Supervisory Committee Member\\
+Department of Computer Science \\
+University of Calgary \\
+Canada\\
+\signline
+Dr. Craig Anslow\\
+Supervisory Committee Member\\
+School of Eng. and Computer Science \\
+Victoria University of Wellington\\
+New Zealand\\
+\end{signing}
 %
@@
 \altchapter{\bf{Abstract}}
-Paragraph 1
-
-Paragraph 2
-
-Paragragh 3
-
-This page is required for all graduate theses.
-An abstract is a short paragraph explaining the major points and conclusions of your thesis.
-For master's theses, the abstract can be no more than 150 words long, while doctoral abstracts can be no longer than 350 words.
-
 \newpage
@@
 \altchapter{\bf{Acknowledgements}}
-Paragraph 1
-
-Paragraph 2
-
-Paragraph 3
 
@@
 \pagenumbering{arabic}
-\include{chapter1}
-%\include{chapter2}
-%\include{chapter3}
-%\include{chapter4}
-%\include{chapter5}
+\graphicspath{{controlled_study/}, {case_study/}, {cisco_study/}, {spy_rest/}, {demo_paper/}}
+\pgfplotsset{table/search path={controlled_study,case_study}}
+\renewcommand\bibname{References}
+\chapter{Introduction}
+
+\include{case_study/case_study}
+\include{spy_rest/spy_rest}
+\include{demo_paper/demo_paper}
+\include{cisco_study/cisco_study}
+\include{controlled_study/controlled_study}
+\chapter{Conclusion}
 \appendix
-\include{appendix1}
 \end{document}

### 2017-07-22 spy_rest/spy_rest.tex
structurally complete, sans abstract / introduction / conclusion

@@
   \end{subfigure}
-  \caption{SpyREST Screen shots showing auto generated API documentation}
+  \caption{SpyREST Screenshots Showing Auto Generated API Documentation}
   \label{fig:spyrest_screenshots}
@@
 
-Fig \ref{fig:spyrest_screenshots} shows screen shots (shortened) from SpyREST auto generated documentation that is solely based on an example API call to URL \url{https://api.github.com/search/repositories?q=tetris+language:assembly&sort=stars&order=desc} with a custom header ``x-spy-rest-description'' to provide the short description for the example.
+Fig \ref{fig:spyrest_screenshots} shows screenshots (shortened) from SpyREST auto generated documentation that is solely based on an example API call to URL \url{https://api.github.com/search/repositories?q=tetris+language:assembly&sort=stars&order=desc} with a custom header ``x-spy-rest-description'' to provide the short description for the example.
 

### 2017-07-23 root.tex
abstract v1

@@
 \altchapter{\bf{Abstract}}
+API documentation presents both a problem and an opportunity for API usability.
+Most developers learn to use an API based on its documentation and at the same time face obstacles using the APIs due to lack of necessary information in the documentation \cite{Robillard_what_makes}.
+REST APIs provide interconnectivity between applications over HTTP.
+Documentation of a REST API is a key information source for software developers that use these types of API.
+As a REST API evolves, the documentation needs to be maintained alongside.
+This implies that REST API developers also need to efficiently generate and maintain the documentation of their APIs with the qualities that make the API usable.
+This introduces the problem of interest in this research: how to document REST APIs?
+
+This is a problem because there is a lack of scientifically established guidelines and automated tool support for REST API documentation with usage examples.
+Existing local API documentation techniques and tools are not readily usable for the documentation of REST APIs because REST APIs have HTTP protocol specific parameters, and API clients are agonistic to the programming languages used to implement the APIs.
+Existing REST API documentation techniques rely on custom API definition languages that require manual effort to generate and maintain REST API documentation as a separate artifact as the API evolves.
+
+In this reasearch, I proposed and evaluated a novel technique to solve REST API documentaiton requirements.
+I identified the requirements by studying the existing literature and the current state of REST API documentation in practice.
+The technique involves the interception of example REST API calls using an HTTP proxy server to auto-generate an accurate and updated REST API documentation.
+
+Practioners can use the presented technique to automatically generate and maintain an updated documentation of REST APIs including the API structure and usage examples from their API test code.
+Researchers can apply the novel concept of using interception to automatically document API usage examples from test scripts to other forms of APIs.
+
 \newpage

### 2017-07-23 root.tex
all chapters added

@@
 \renewcommand\bibname{References}
+
 \chapter{Introduction}
+Application Programming Interfaces, commonly known as APIs, are used to express a software component in terms of its operations, inputs, outputs, and their types\footnote{\url{https://en.wikipedia.org/wiki/Application_programming_interface}}.
+Robillard defines an API as follows: An API is the interface to implement functionality that developers can access to perform various tasks \cite{Robillard_a_field_study} \cite{Robillard_what_makes}.
+APIs enable multiple software components to interact with each other.
+
+REST APIs are a type of APIs that are used to integrate software components using web technologies.
+Fielding defined Representational State Transfer or REST as an architectural style for developing distributed hypermedia systems \cite{Fielding_rest}.
+For example, the GitHub REST API\footnote{\url{https://developer.github.com/v3/repos/\#create}} has a resource called \texttt{Repository} to denote a code repository that can be identified by the URL \url{/user/repos}.
+To create a new \texttt{Repository} for a user, the GitHub API can be invoked via HTTP $POST$ at \url{/user/repos} using the following JSON representation of a \texttt{Repository}:
+
+\begin{verbatim}
+{
+  "name": "Hello-World",
+  "description": "This is your first repository",
+  "homepage": "https://github.com",
+  "private": false,
+  "has_issues": true,
+  "has_wiki": true,
+  "has_downloads": true
+}
+\end{verbatim}
+
+In today's world of technology, REST APIs have become ubiquitous and the primary choice for integrating Internet enabled applications due to its simplicity and similarity with HTTP \cite{mangler2010origin}.
+For example, a real estate listing website uses a REST API to collect ``walk score'' and another REST API to show a map view of a property listing.
+By incorporating the map and walk score using REST APIs, the real estate listng site provides important information to their users.
+Most REST APIs, including these two examples, are often documented manually or using custom implemented tools specific to the APIs that are not publicly available.
+This requires effort to generate and maintain the documentation of REST APIs over time since there is a lack of reusable tool support.
+
+Previous research in the area of API usability mostly focused on local APIs such as Java libraries.
+Researchers identified the documentation of APIs as both the primary source of information as well as the key obstacle for API usability \cite{Robillard_what_makes}.
+Hence the quality of the API documentation plays an important role in API usability.
+To this regard, researchers have identified the qualities of ``good API documentation'' as follows: complete, correct, includes thorough explanations and code examples, provides consistent presentation and organization \cite{Robillard_what_makes,Myers_study}.
+Today, there are several tools such as Junit\footnote{\url{http://junit.org/}}, UsETeC \cite{zhu2014mining}, Jadeite \cite{Jadeite}, APIMiner \cite{montandon2013documenting}, Roast \cite{Hoffman_api_documentation} that can be used to document local APIs with the aforementioned qualities .
+
+While there are overlaps between the documentation requirements of local and REST APIs, there are significant requirements that are unique to each.
+For example, REST API documentation needs to include information about its API endpoints, HTTP headers, and request and response payloads.
+On the other hand, documentation of local APIs need to explain the classes and methods.
+Due to such differences between local and REST APIs, the existing technique and tools for local API documentation cannot be readily used to generate REST API documentation.
+
+\begin{figure}[htb]
+  \includegraphics[width=\linewidth]{manual_workflow.png}
+  \caption{Manual REST API Documentation Steps}
+  \label{fig:manual}
+\end{figure}
+
+Currently, the process of documenting REST APIs are largely manual.
+Conceptually, the high level steps involved in the manual process can be identified as shown in Figure \ref{fig:manual}.
+These steps S1-S7 can be described as follows: S1 - API developer prepares an example API with the required HTTP parameters and request body, S2 - API developer uses a REST API client to make the example API call, S3 - the API developer records the HTTP data , S4 - API developer then edits the recorded data so that only relevant content is selected for documentation, S5 - API developers use any custom content to describe the API example, S6 - API developers combine the custom content with the edited content from the HTTP traffic into HTML to publish to the web, and S7 - API developer adds custom overview information to explain API concepts and general rules, and publishes the final documentation so other developers can learn the API.
+The steps S1-S6 are repeated for every API action that needs to be documented, and S1-S7 need to repeated for every version of the REST API as it evolves.
+As depicted here, this manual process can be both time consuming and error-prone.
+
+This thesis is ogranized as follows.
+In Chapter~\ref{chapter:case_study}, I discuss a case study to summarize the state of existing literature and industry practices related to the versioning, documentation, and change communication patterns of evolving Web APIs.
+The primary findings from this study are as follows: Web APIs evolve frequently, often several times a week, including both compatible and incompatible API changes, using manual or bespoke methods to version, document, and communicate the changes as they evolve.
+I found a lack of reusable techniques in the existing literature to automatically generate and maintain the documentation of multi-version evolving REST APIs with usage examples.
+I persued further research to solve this problem as discussed in the following paragraphs.
+
+In Chapter~\ref{chapter:spy_rest}, I present the following list of REST API documentation requirements by studying the literature and applying the findings from Chapter~\ref{chapter:case_study}: 1) automated, 2) example based, 3) executable, 4) version aware, 5) customizable, 6) reusable, and 7) collaborative.
+I present, SpyREST, a novel technique to intercept example REST API calls using an HTTP proxy server to collect and autmatically synthesize API traffic to generate and update API documentation to satisfy the aforementioned requirements.
+The primary advantage of the presented technique is that the documentation is generated from executable code.
+Existing techniques that rely on formatted comments or manually written API descriptions following custom languages require manual effort to find and update the API documentation everytime an API evolves.
+
+In Chapter~\ref{chapter:demo_paper}, I present a prototype implementation of the aformentioned technique to demonstrate a reusable tool and compare it against the existing REST API documentation tools that rely on custom API description languages.
+I also present an example where automated test code can be used to generate and maintain REST API documentation using the developed tool.
+
+I discuss an industrial evaluation of the proposed technique and the tool based on an eigtheen month period of production use at Cisco in Chapter~\ref{chapter:cisco_study}.
+From the study I found that it is feasible to use the interception technique and automated tests to document REST APIs satisfying the aforementioned requirements for an evolving API.
+Automated always-updated REST API documentation was found to help establish a quick feedback cycle among the stakeholders.
+I also discuss several refinements and limitations of the proposed technique based on the lessons learned from this case study.
+
+In Chapter~\ref{chapter:controlled_study}, I discuss a controlled study performed with 26 professional software engineers to undestand the impact of usage examples on REST API client developers.
+I provided an enahanced REST API documentation with usage examples containing data from API test code to participants of one of two study groups to perform the same set of API tasks.
+I found that without usage examples participants faced obstacles related to using appropriate data types, data formats, and required elements in the HTTP headers, query parameters, and request body.
+I observed that providing usage examples with data taken from API test code improves API client developer success rate, reduces required trial attempts, and improves satisfaction ratings.
 

### 2017-07-23 build.sh

@@ -8,4 +8,17 @@ bibtex case_study/case_study.aux
 bibtex cisco_study/cisco_study.aux
 pdflatex root.tex
 pdflatex root.tex
-/Applications/Skim.app/Contents/MacOS/Skim root.pdf
+
+
+pdffile=root.pdf
+/usr/bin/osascript << EOF
+  set theFile to POSIX file "${pdffile}" as alias
+  tell application "Skim"
+  # uncomment the folloing line if you want Skim to get the focus
+  activate
+  set theDocs to get documents whose path is (get POSIX path of theFile)
+  if (count of theDocs) > 0 then revert theDocs
+  open theFile
+  end tell
+EOF
+

### 2017-07-24 root.tex
introduction first pass is complete

@@
 
-I discuss an industrial evaluation of the proposed technique and the tool based on an eigtheen month period of production use at Cisco in Chapter~\ref{chapter:cisco_study}.
-From the study I found that it is feasible to use the interception technique and automated tests to document REST APIs satisfying the aforementioned requirements for an evolving API.
+I discuss an industrial evaluation of the proposed technique and the tool in Chapter~\ref{chapter:cisco_study}.
+The evaluation is carried out based on the data collected from an eighteen month of production use of SpyREST at Cisco.
+From the study I found that it is feasible to use the proposed interception technique and automated tests to document evolving REST APIs satisfying the aforementioned requirements.
 Automated always-updated REST API documentation was found to help establish a quick feedback cycle among the stakeholders.
@@
 
-In Chapter~\ref{chapter:controlled_study}, I discuss a controlled study performed with 26 professional software engineers to undestand the impact of usage examples on REST API client developers.
-I provided an enahanced REST API documentation with usage examples containing data from API test code to participants of one of two study groups to perform the same set of API tasks.
-I found that without usage examples participants faced obstacles related to using appropriate data types, data formats, and required elements in the HTTP headers, query parameters, and request body.
-I observed that providing usage examples with data taken from API test code improves API client developer success rate, reduces required trial attempts, and improves satisfaction ratings.
+In Chapter~\ref{chapter:controlled_study}, I discuss a controlled study performed with 26 professional software engineers to undestand the impact of usage examples in API documentation on REST API client developers.
+I enahanced an existing REST API documentation with usage examples containing data from the API's test code and provided it to the participants of one of two study groups.
+I found that participants using the original documentation without usage examples faced obstacles related to using appropriate data types, data formats, and required elements in the HTTP input.
+I observed that providing usage examples with data taken from API test code reduces the obstacles and improves API client developer success rate, reduces required trial attempts, and improves satisfaction ratings.
+
+In summary, this thesis makes the following key contributions to the body of research in software engineering:
+
+\begin{enumerate}
+  \item A list of REST API documentation requirements derived from the existing literature and the current state of industry practices to be used as a guideline by practitioners and researchers.
+  \item SpyREST, a novel technique and a reusable tool that uses interception to automatically generate and maintain evolving REST API documentation with usage examples from executable code.
+  \item An industrial evaluation of SpyREST to demonstrate the feasibility of the proposed technique that practitioners can follow to improve their REST API documentation experience.
+  \item An empricial evidence showing the common obstacles faced by REST API client developers that can be reduced with usage examples to help the prioritization of REST API documentation efforts.
+\end{enumerate}
 

### 2017-07-24 appendix.tex
almost done with conclusion

\chapter{Source Code Fragments of the Spy from Chapter~\ref{chapter:demo_paper}}

\lstdefinelanguage{JavaScript}{
  keywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break},
  keywordstyle=\color{blue}\bfseries,
  ndkeywords={class, export, boolean, throw, implements, import, this},
  ndkeywordstyle=\color{darkgray}\bfseries,
  identifierstyle=\color{black},
  sensitive=false,
  comment=[l]{//},
  morecomment=[s]{/*}{*/},
  commentstyle=\color{purple}\ttfamily,
  stringstyle=\color{red}\ttfamily,
  morestring=[b]',
  morestring=[b]"
}

\lstset{
   language=JavaScript,
   %backgroundcolor=\color{lightgray},
   extendedchars=true,
   basicstyle=\footnotesize\ttfamily,
   showstringspaces=false,
   showspaces=false,
   numbers=left,
   numberstyle=\footnotesize,
   numbersep=9pt,
   tabsize=2,
   breaklines=true,
   showtabs=false,
   captionpos=b
}

\begin{lstlisting}[caption=index.js]
(function() {
  var ApiThough;

  ApiThough = require('./lib/api_through');

  process.env.NODE_TLS_REJECT_UNAUTHORIZED = "0";

  process.on('uncaughtException', function(err) {
    console.error('uncaughtException');
    console.error(err);
    return console.error(err.stack);
  });

  new ApiThough().start();

}).call(this);
\end{lstlisting}

\begin{lstlisting}[caption=lib/api\_through.js]
(function() {
  var ApiThrough, PassThrough;

  module.exports = function() {
    return new ApiThrough();
  };

  PassThrough = require('stream').PassThrough;

  ApiThrough = (function() {
    var Proxy, SSLCert, fs;

    Proxy = require('http-mitm-proxy');

    SSLCert = require('./ssl_cert');

    fs = require('fs');

    function ApiThrough() {
      var host, mongooose;
      mongooose = require('mongoose');
      host = process.env["MONGODB_PORT_27017_TCP_ADDR"] || 'localhost';
      console.log("Connecting to mongo at " + host);
      mongooose.connect("mongodb://" + host + ":27017/apis");
    }

    ApiThrough.prototype.start = function() {
      var proxy;
      proxy = new Proxy();
      proxy.use(this);
      proxy.onCertificateMissing = (function(_this) {
        return function(ctx, files, callback) {
          return _this.onCertificateMissing(ctx, files, callback);
        };
      })(this);
      proxy.listen({
        port: process.env['PROXY_PORT'] || 9081,
        sslCertCacheDir: './scripts/certs/http-mitm-proxy'
      });
      return this.proxy = proxy;
    };

    ApiThrough.prototype.onCertificateMissing = function(ctx, files, callback) {
      var sslCert;
      console.log('Looking for "%s" certificates', ctx.hostname);
      console.log('"%s" missing', ctx.files.keyFile);
      console.log('"%s" missing', ctx.files.certFile);
      sslCert = new SSLCert(ctx.hostname);
      return sslCert.create((function(_this) {
        return function() {
          return callback(null, {
            keyFileData: fs.readFileSync(ctx.files.keyFile),
            certFileData: fs.readFileSync(ctx.files.certFile)
          });
        };
      })(this));
    };

    ApiThrough.prototype.onError = function(ctx, err) {
      console.error('proxy error:', err);
      return console.error('proxy error stack:', err.stack);
    };

    ApiThrough.prototype.onRequest = function(ctx, callback) {
      var ApiExample, User, apiExample, responseAggregator, responseBody;
      ApiExample = require('./api_example');
      User = require('./user');
      ctx.onError((function(_this) {
        return function(ctx, err) {
          return _this.onError(ctx, err);
        };
      })(this));
      apiExample = new ApiExample();
      apiExample.populateFromRequest(ctx.clientToProxyRequest);
      apiExample.setFullUrl(ctx.isSSL, ctx.proxyToServerRequestOptions);
      responseBody = '';
      responseAggregator = new PassThrough();
      responseAggregator.on('finish', function() {
        apiExample.responseBody = responseBody;
        apiExample.stripResponseBody();
        return User.findOne({
          api_token: apiExample.apiToken
        }, function(err, user) {
          var apiExampleRaw;
          console.log('findOne', err, user);
          if (user) {
            apiExample.userId = user.id;
          }
          apiExampleRaw = apiExample.toObject();
          delete apiExampleRaw._id;
          return ApiExample.findOneAndUpdate({
            digest: apiExample.digest
          }, apiExampleRaw, {
            upsert: true
          }, function(error) {
            if (error != null) {
              return console.log("Failed to save due to error", error);
            }
          });
        });
      });
      ctx.addResponseFilter(responseAggregator);
      ctx.onRequestData(function(ctx, chunk, callback) {
        apiExample.requestBody += chunk.toString('utf8');
        return callback(null, chunk);
      });
      ctx.onResponse(function(ctx, callback) {
        apiExample.responseHeaders = ctx.serverToProxyResponse.headers;
        apiExample.statusCode = ctx.serverToProxyResponse.statusCode;
        ctx.serverToProxyResponse.on("finish", function() {
          return console.log("FINISH");
        });
        return callback();
      });
      ctx.onResponseData(function(ctx, chunk, callback) {
        responseBody += chunk.toString('utf8');
        return callback(null, chunk);
      });
      return callback();
    };

    return ApiThrough;

  })();

}).call(this);

\end{lstlisting}

\begin{lstlisting}[caption=lib/api\_example.js]

(function() {
  var ApiExample, ApiExamplesSchema, CUSTOM_HEADERS, VERSION_IN_HEADER, VERSION_IN_URL, crypto, inflector, mongoose, url, validator, _u;

  mongoose = require('mongoose');

  url = require('url');

  _u = require('underscore');

  crypto = require('crypto');

  validator = require('validator');

  inflector = require('i')();

  VERSION_IN_HEADER = /v(\d|\.\d)+/;

  VERSION_IN_URL = /\/(v\d[^\/]*)/;

  CUSTOM_HEADERS = {
    DESC_HEADER: "x-spy-rest-desc",
    VERSION_HEADER: "x-spy-rest-version",
    RESOURCE_HEADER: "x-spy-rest-resource",
    API_TOKEN_HEADER: "x-spy-rest-api-token",
    HOST_HEADER: 'x-spy-rest-host',
    NO_STRIP_HEADER: 'x-spy-rest-no-strip',
    ACTION_HEADER: 'x-spy-rest-action'
  };

  ApiExamplesSchema = new mongoose.Schema({
    description: {
      type: String
    },
    version: {
      type: String,
      "default": 'Default'
    },
    resource: {
      type: String
    },
    action: {
      type: String
    },
    url: {
      type: String
    },
    templatedURL: {
      type: String
    },
    query: {
      type: Object
    },
    host: {
      type: String
    },
    http_method: {
      type: String
    },
    requestBody: {
      type: String,
      "default": ''
    },
    requestHeaders: {
      type: Object
    },
    statusCode: {
      type: Number
    },
    responseHeaders: {
      type: Object
    },
    responseBody: {
      type: String,
      "default": ''
    },
    strippedResponseBody: {
      type: String
    },
    recordedAt: {
      type: Date,
      "default": function() {
        return new Date();
      }
    },
    fullURL: {
      type: String,
      "default": ''
    },
    digest: {
      type: String,
      "default": ''
    },
    requiresAuth: {
      type: Boolean,
      "default": false
    },
    apiToken: {
      type: String,
      "default": ''
    },
    userId: {
      type: mongoose.Schema.Types.ObjectId,
      "default": null
    },
    authType: {
      type: String
    }
  }, {
    collection: 'api_examples'
  });

  ApiExamplesSchema.index({
    host: 1,
    version: 1,
    resource: 1,
    action: 1
  });

  ApiExample = mongoose.model('ApiExample', ApiExamplesSchema);

  ApiExample.prototype.populateFromRequest = function(request) {
    this.host = request.headers[CUSTOM_HEADERS.HOST_HEADER] || request.headers.host;
    this.url = this.filteredUrl(request.url);
    this.http_method = request.method;
    this.requestHeaders = request.headers;
    this.description = request.headers[CUSTOM_HEADERS.DESC_HEADER];
    this.version = request.headers[CUSTOM_HEADERS.VERSION_HEADER] || this.guessedVersion() || 'Default';
    this.resource = request.headers[CUSTOM_HEADERS.RESOURCE_HEADER] || this.guessedResource();
    this.query = this.parsedUrl().query;
    this.apiToken = this.requestHeaders[CUSTOM_HEADERS.API_TOKEN_HEADER];
    this.filterAuthHeaders();
    this.action = this.computedAction();
    return this.digest = this.computeDigest();
  };

  ApiExample.prototype.stripResponseBody = function() {
    var StrippedObject, strippedObject;
    if (!this.shouldStrip()) {
      return;
    }
    StrippedObject = require('./stripped_object');
    strippedObject = new StrippedObject();
    try {
      return this.strippedResponseBody = JSON.stringify(strippedObject.strip(JSON.parse(this.responseBody)));
    } catch (_error) {
      return this.strippedResponseBody = this.responseBody;
    }
  };

  ApiExample.prototype.saveWithErrorLog = function() {
    return this.save(function(error) {
      if (error) {
        return console.error("Failed to save because of error", error);
      }
    });
  };

  ApiExample.prototype.guessedVersion = function() {
    return this.guessedVersionFromURL() || this.guessedVersionAcceptHeader();
  };

  ApiExample.prototype.guessedVersionFromURL = function() {
    if (this.url == null) {
      return null;
    }
    if (this.url.match(VERSION_IN_URL)) {
      return RegExp.$1;
    }
  };

  ApiExample.prototype.guessedVersionAcceptHeader = function() {
    var matches;
    if (this.requestHeaders['accept'] == null) {
      return null;
    }
    matches = this.requestHeaders['accept'].match(VERSION_IN_HEADER);
    if (matches != null) {
      return matches[0];
    }
  };

  ApiExample.prototype.parsedUrl = function() {
    return url.parse(this.url, true);
  };

  ApiExample.prototype.guessedResource = function() {
    var parts, possibleResource;
    parts = this.parsedUrl().pathname.split('/').reverse();
    possibleResource = _u.find(parts, function(part) {
      return part.match(/^[^\d].*$/);
    });
    if (possibleResource != null) {
      return possibleResource.split('.')[0];
    }
  };

  ApiExample.prototype.computedAction = function() {
    if (this.requestHeaders[CUSTOM_HEADERS.ACTION_HEADER] != null) {
      return "" + this.http_method + " " + this.requestHeaders[CUSTOM_HEADERS.ACTION_HEADER];
    } else {
      this.templatizeURL();
      return "" + this.http_method + " " + this.templatedURL;
    }
  };

  ApiExample.prototype.setFullUrl = function(isSSL, hostPort) {
    var host, port, scheme;
    scheme = isSSL ? "https" : "http";
    host = hostPort.host;
    port = '';
    if (hostPort.port) {
      if (scheme === "http" && hostPort.port !== 80) {
        port = ":" + hostPort.port;
      } else if (scheme === "https" && hostPort.port !== 443) {
        port = ":" + hostPort.port;
      }
    }
    return this.fullURL = "" + scheme + "://" + host + port + this.url;
  };

  ApiExample.prototype.filterAuthHeaders = function() {
    return this.requestHeaders = _u.reduce(this.requestHeaders, (function(_this) {
      return function(filteredHeaders, value, key) {
        if (key === 'authorization') {
          console.log("value");
          console.log(value);
          if (value.match(/Basic/)) {
            console.log(value);
            _this.authType = 'Basic';
            filteredHeaders[key] = 'Basic FILTERED';
          } else {
            filteredHeaders[key] = 'FILTERED';
          }
          _this.requiresAuth = true;
        } else {
          filteredHeaders[key] = value;
        }
        return filteredHeaders;
      };
    })(this), {});
  };

  ApiExample.prototype.filteredUrl = function(rawUrl) {
    var api_key_param, parasedRawUrl;
    if (!rawUrl) {
      return rawUrl;
    }
    parasedRawUrl = url.parse(rawUrl, true);
    api_key_param = _u.find(_u.keys(parasedRawUrl.query), function(param) {
      return param.toLowerCase() === 'api_key';
    });
    if (!api_key_param) {
      return rawUrl;
    }
    return rawUrl.replace("" + api_key_param + "=" + parasedRawUrl.query[api_key_param], "" + api_key_param + "=FILTERED");
  };

  ApiExample.prototype.computeDigest = function() {
    var hash, text;
    text = "__VERSION__" + this.version + "__RESOURCE__" + this.resource + "__URL__" + this.action + "__DESC__" + this.description;
    hash = crypto.createHash('sha');
    console.log("digest for text " + text);
    hash.update(text);
    return hash.digest('base64');
  };

  ApiExample.prototype.shouldStrip = function() {
    var strip_header_value;
    if (!this.requestHeaders) {
      return true;
    }
    console.log("HERE");
    strip_header_value = this.requestHeaders[CUSTOM_HEADERS.NO_STRIP_HEADER];
    return strip_header_value !== 'true';
  };

  ApiExample.prototype.templatizeURL = function() {
    var parts, path, templatedParts;
    path = this.parsedUrl().pathname;
    parts = path.split("/");
    templatedParts = _u.map(parts, function(part, index) {
      var previousPart;
      previousPart = index > 0 ? parts[index - 1] : '';
      if (validator.isNumeric(previousPart)) {
        return part;
      }
      if (validator.isNumeric(part)) {
        return "{:" + (inflector.singularize(previousPart)) + "-id}";
      } else if (validator.isUUID(part)) {
        return "{:" + (inflector.singularize(previousPart)) + "-guid}";
      } else {
        return part;
      }
    });
    return this.templatedURL = templatedParts.join('/');
  };

  module.exports = ApiExample;

}).call(this);
\end{lstlisting}

### 2017-07-24 root.tex
almost done with conclusion

@@
 \chapter{Conclusion}
+
+In this thesis, I presented a novel idea of using interception to automate the process of REST API documentation with usage examples from executable code.
+I started by investigating the current industry practices and challenges associated with developing and maintaining evolving Web APIs.
+I found modern Web APIs predominently embraced REST architectural style, but the existing literature had primarily focused on solving problems related to the SOAP APIs.
+I also found that most REST APIs are documented using manual approaches that are time consuming and error- prone to maintain when an API evolves several times a month.
+Through my research, I aimed to fill this gap between the REST API documentation challenges and the existing literature by finding a reusable solution to the common problems.
+
+In my second study, I investigated the existing literature on API usablity to derive a set of re- quirements that are applicable to REST API documentation and proposed an idea of automatically generating and maintaining REST API documentation from executable code.
+I found that existing techniques for REST API documentation systems relied on custom API description languagees that are manually maintained as a parallel artifact to the API code, causing potential drift between the actual API and it’s documentation as it evolves over time.
+By using interception, I have demon- strated that it is possible to automatically record and synthesize the recorded API traffic to generate pragmatic REST API documentation from its test code.
+Additionally, the technique allows API de- velopers to collaboratively document the API with realistic and executable examples for multiple versions as the API evolves.
+
+In my third study, I evaluated the proposed technique in an industrial setup and found it to be a pragmatic solution for production use.
+The concept of leveraging API tests to maintain an always-updated API documentation has helped the organization establish a quick feedback loop during the API development process.
+Practitioners can incorporate the findings from this study to improve their own REST API documentation practice.
+
+To understand the impact of usage examples that can be derived from API tests, I performed the last study with professional software engineers and found that the examples help reduce the common obstacles faced by REST API client developers.
+I derived a list of recommendations for documenting usage examples to satisfy the common information needs when REST API client developers use an API.
+Practitiones can follow the recommendations as guideline to allocate ap- propriate priority while documenting their REST APIs in a resource constrained environment.
+
 \appendix
+\include{appendix}
 \end{document}

### 2017-07-24 root.tex
adds the future work

@@
 
+\section{Future Work}
+Based on my findings from the this research, I identified the following list of potential future research problems.
+
+\begin{itemize}
+  \item \textbf{Extension to Support Complex Usage Scenarios.} Future research needs to be carried out to find techniques to automate the process of documenting usage scenarios that require multiple API calls to perform complex API tasks.
+Future research is also needed to investigate techniques for automatically cross-referencing API elements to allow API client developers with an improved navigation experience while reading the REST API documentation.
+  \item \textbf{Automated Changelog Generation.} For frequently evolving APIs, future research needs to extend the proposed technique in this thesis to automatically compute human readable changelogs and detect version compatibility to follow a semantic versioning scheme.
+This would help API client developers with an accurate information about the potential impact of an API update on their own application.
+  \item \textbf{Extensions of the Interception Based Documentation to non-REST APIs.}
+\end{itemize}
+
 \appendix

### 2017-07-24 root.tex
future work is done

@@
 This would help API client developers with an accurate information about the potential impact of an API update on their own application.
-  \item \textbf{Extensions of the Interception Based Documentation to non-REST APIs.}
+  \item \textbf{Extensions of the Interception Based Documentation to non-REST APIs.} The proposed technique can be extended to document usage examples for other stateless APIs outside the realm of REST APIs.
+The structure of REST APIs are commonly documented using programming language features and formatted comments.
+Future research can investigate the feasibility of leveraging the interception technique to source realistic usage examples for local stateless APIs from existing API test code.
 \end{itemize}

### 2017-07-24 root.tex
future work is done

@@
 This would help API client developers with an accurate information about the potential impact of an API update on their own application.
+  \item \textbf{Automated API Client Code Generation.} Several tools exist that can convert custom REST API description files to API client code in multiple programming languages.
+Future research can be carried out to map the SpyREST data model with the custom API description languages to leverage the interception technique to reduce the amount of manual effort required to maintain the custom formatted API descriptions as the API evolves.
   \item \textbf{Extensions of the Interception Based Documentation to non-REST APIs.} The proposed technique can be extended to document usage examples for other stateless APIs outside the realm of REST APIs.

### 2017-07-25 introduction.tex
adds refs to introduction

--- 2017-07-24 root.tex
@@
-\documentclass[12pt]{ucalgthes1}
-\usepackage[pdftex]{graphicx}
-\usepackage[pdftex]{color}
-\usepackage[letterpaper,top=1in, bottom= 1in, left= 1in, right= 1in]{geometry}
-\usepackage{hyperref}
-\usepackage{mathptmx}
-\usepackage{listings}
-\usepackage{url}
-\usepackage[tableposition=top]{caption}
-\usepackage{subcaption}
-\usepackage{tikz,pgfplots,pgfplotstable}
-\usepackage{mathtools}
-\usepgfplotslibrary{dateplot}
-\usepgfplotslibrary{statistics}
-\usepackage{courier}
-\usepackage[square,numbers,sectionbib]{natbib}
-\usepackage[sectionbib]{chapterbib}
-\usepackage[numbib]{tocbibind}
-\newcommand{\newblock}{}
-\usepackage{eurosym}
-\usepackage{multirow}
-\usepackage[T1]{fontenc}
-\usepackage{textcomp}
-\lstset{upquote=true}
-\usepackage{float}
-\floatstyle{boxed}
-\usepackage{blindtext}
-\let\labelindent\relax
-\usepackage{upquote}
-\usepackage{mdframed}
-%\usepackage{balance}
-\usepackage{enumitem}
-
-\title{Automated Example Oriented REST API Documenation}
-%
-%            Insert the correct information between the {}
-%
-\author{S M Sohan}
-\thesisyear{2017}
-\thesis{thesis}    % the word dissertation can be inserted between {}
-\newcommand{\thesistitle}{Automated Example Oriented REST API Documenation}
-\monthname{September}
-\dept{GRADUATE PROGRAM IN COMPUTER SCIENCE}
-\degree{DOCTOR OF PHILOSOPHY}
-%
-%                    End of supplied information
-%
-\begin{document}
-\makethesistitle
-\pagenumbering{roman}     % resets page counter to one
-\setcounter{page}{2}
-\chapter*{UNIVERSITY OF CALGARY \\ FACULTY OF GRADUATE STUDIES}
-\thispagestyle{empty}
-The undersigned certify that they have read, and recommend
-to the Faculty of Graduate Studies for acceptance, a \Thesis\ entitled
-``\thesistitle'' submitted by \Author\
-in partial fulfillment of the requirements for the degree of
-\Degree.\\
-
-%
-%                 Substitute  List of Examiners
-%
-\begin{signing}{Department of Computer Science}
-\signline
-Dr. Frank Maurer \\
-Supervisor\\
-Department of Computer Science \\
-University of Calgary \\
-Canada\\
-\signline
-Dr. Mea Wang\\
-Supervisory Committee Member\\
-Department of Computer Science \\
-University of Calgary \\
-Canada\\
-\signline
-Dr. Craig Anslow\\
-Supervisory Committee Member\\
-School of Eng. and Computer Science \\
-Victoria University of Wellington\\
-New Zealand\\
-\end{signing}
-%
-\newpage
-\phantomsection
-\altchapter{\bf{Abstract}}
-API documentation presents both a problem and an opportunity for API usability.
-Most developers learn to use an API based on its documentation and at the same time face obstacles using the APIs due to lack of necessary information in the documentation \cite{Robillard_what_makes}.
-REST APIs provide interconnectivity between applications over HTTP.
-Documentation of a REST API is a key information source for software developers that use these types of API.
-As a REST API evolves, the documentation needs to be maintained alongside.
-This implies that REST API developers also need to efficiently generate and maintain the documentation of their APIs with the qualities that make the API usable.
-This introduces the problem of interest in this research: how to document REST APIs?
-
-This is a problem because there is a lack of scientifically established guidelines and automated tool support for REST API documentation with usage examples.
-Existing local API documentation techniques and tools are not readily usable for the documentation of REST APIs because REST APIs have HTTP protocol specific parameters, and API clients are agonistic to the programming languages used to implement the APIs.
-Existing REST API documentation techniques rely on custom API definition languages that require manual effort to generate and maintain REST API documentation as a separate artifact as the API evolves.
-
-In this reasearch, I proposed and evaluated a novel technique to solve REST API documentaiton requirements.
-I identified the requirements by studying the existing literature and the current state of REST API documentation in practice.
-The technique involves the interception of example REST API calls using an HTTP proxy server to auto-generate an accurate and updated REST API documentation.
-
-Practioners can use the presented technique to automatically generate and maintain an updated documentation of REST APIs including the API structure and usage examples from their API test code.
-Researchers can apply the novel concept of using interception to automatically document API usage examples from test scripts to other forms of APIs.
-
-\newpage
-\phantomsection
-\altchapter{\bf{Acknowledgements}}
-
-\begin{singlespace}
-\newpage
-\phantomsection
-\tableofcontents
-\pagestyle{plain}
-\newpage
-\phantomsection
-\listoftables
-\pagestyle{plain}
-\newpage
-\phantomsection
-\listoffigures
-\pagestyle{plain}
-\clearpage
-\clearpage          % otherwise tables will be numbered wrong
-\end{singlespace}
-\newpage
-\phantomsection
-\chapter*{\bf{List of Symbols, Abbreviations and Nomenclature}\hfill} \addcontentsline{toc}{chapter}{List of Symbols}
-\listofsymbols
-\pagestyle{plain}
-\clearpage
-
-\pagenumbering{arabic}
-\graphicspath{{controlled_study/}, {case_study/}, {cisco_study/}, {spy_rest/}, {demo_paper/}}
-\pgfplotsset{table/search path={controlled_study,case_study}}
-\renewcommand\bibname{References}
-
 \chapter{Introduction}
@@
 
-\include{case_study/case_study}
-\include{spy_rest/spy_rest}
-\include{demo_paper/demo_paper}
-\include{cisco_study/cisco_study}
-\include{controlled_study/controlled_study}
-\chapter{Conclusion}
-
-In this thesis, I presented a novel idea of using interception to automate the process of REST API documentation with usage examples from executable code.
-I started by investigating the current industry practices and challenges associated with developing and maintaining evolving Web APIs.
-I found modern Web APIs predominently embraced REST architectural style, but the existing literature had primarily focused on solving problems related to the SOAP APIs.
-I also found that most REST APIs are documented using manual approaches that are time consuming and error- prone to maintain when an API evolves several times a month.
-Through my research, I aimed to fill this gap between the REST API documentation challenges and the existing literature by finding a reusable solution to the common problems.
-
-In my second study, I investigated the existing literature on API usablity to derive a set of re- quirements that are applicable to REST API documentation and proposed an idea of automatically generating and maintaining REST API documentation from executable code.
-I found that existing techniques for REST API documentation systems relied on custom API description languagees that are manually maintained as a parallel artifact to the API code, causing potential drift between the actual API and it’s documentation as it evolves over time.
-By using interception, I have demon- strated that it is possible to automatically record and synthesize the recorded API traffic to generate pragmatic REST API documentation from its test code.
-Additionally, the technique allows API de- velopers to collaboratively document the API with realistic and executable examples for multiple versions as the API evolves.
-
-In my third study, I evaluated the proposed technique in an industrial setup and found it to be a pragmatic solution for production use.
-The concept of leveraging API tests to maintain an always-updated API documentation has helped the organization establish a quick feedback loop during the API development process.
-Practitioners can incorporate the findings from this study to improve their own REST API documentation practice.
-
-To understand the impact of usage examples that can be derived from API tests, I performed the last study with professional software engineers and found that the examples help reduce the common obstacles faced by REST API client developers.
-I derived a list of recommendations for documenting usage examples to satisfy the common information needs when REST API client developers use an API.
-Practitiones can follow the recommendations as guideline to allocate ap- propriate priority while documenting their REST APIs in a resource constrained environment.
-
-\section{Future Work}
-Based on my findings from the this research, I identified the following list of potential future research problems.
-
-\begin{itemize}
-  \item \textbf{Extension to Support Complex Usage Scenarios.} Future research needs to be carried out to find techniques to automate the process of documenting usage scenarios that require multiple API calls to perform complex API tasks.
-Future research is also needed to investigate techniques for automatically cross-referencing API elements to allow API client developers with an improved navigation experience while reading the REST API documentation.
-  \item \textbf{Automated Changelog Generation.} For frequently evolving APIs, future research needs to extend the proposed technique in this thesis to automatically compute human readable changelogs and detect version compatibility to follow a semantic versioning scheme.
-This would help API client developers with an accurate information about the potential impact of an API update on their own application.
-  \item \textbf{Extensions of the Interception Based Documentation to non-REST APIs.}
-\end{itemize}
-
-\appendix
-\include{appendix}
-\end{document}
+\bibliographystyle{IEEEtran}
+\bibliography{IEEEabrv,references}

### 2017-07-25 root.tex
adds refs to introduction

@@
 \altchapter{\bf{Acknowledgements}}
+\begin{singlespace}
+I am grateful to Shahana, my wife, for supporting me and the family while I took the time away to work on this research.
 
-\begin{singlespace}
+I am thankful to my supervisor Dr. Frank Maurer for giving me the opportunity and valuable guidance to pursue this research.
+
+I am also thankful to Dr. Craig Anslow for his continuous encouragement and the countless number of hours that he spent to provide feedback about my work.
+\newpage
+\phantomsection
+\altchapter{\bf{Dedication}}
+\begin{center}
+  To Shahana
+\end{center}
 \newpage
@@
 
-\chapter{Introduction}
-Application Programming Interfaces, commonly known as APIs, are used to express a software component in terms of its operations, inputs, outputs, and their types\footnote{\url{https://en.wikipedia.org/wiki/Application_programming_interface}}.
-Robillard defines an API as follows: An API is the interface to implement functionality that developers can access to perform various tasks \cite{Robillard_a_field_study} \cite{Robillard_what_makes}.
-APIs enable multiple software components to interact with each other.
-
-REST APIs are a type of APIs that are used to integrate software components using web technologies.
-Fielding defined Representational State Transfer or REST as an architectural style for developing distributed hypermedia systems \cite{Fielding_rest}.
-For example, the GitHub REST API\footnote{\url{https://developer.github.com/v3/repos/\#create}} has a resource called \texttt{Repository} to denote a code repository that can be identified by the URL \url{/user/repos}.
-To create a new \texttt{Repository} for a user, the GitHub API can be invoked via HTTP $POST$ at \url{/user/repos} using the following JSON representation of a \texttt{Repository}:
-
-\begin{verbatim}
-{
-  "name": "Hello-World",
-  "description": "This is your first repository",
-  "homepage": "https://github.com",
-  "private": false,
-  "has_issues": true,
-  "has_wiki": true,
-  "has_downloads": true
-}
-\end{verbatim}
-
-In today's world of technology, REST APIs have become ubiquitous and the primary choice for integrating Internet enabled applications due to its simplicity and similarity with HTTP \cite{mangler2010origin}.
-For example, a real estate listing website uses a REST API to collect ``walk score'' and another REST API to show a map view of a property listing.
-By incorporating the map and walk score using REST APIs, the real estate listng site provides important information to their users.
-Most REST APIs, including these two examples, are often documented manually or using custom implemented tools specific to the APIs that are not publicly available.
-This requires effort to generate and maintain the documentation of REST APIs over time since there is a lack of reusable tool support.
-
-Previous research in the area of API usability mostly focused on local APIs such as Java libraries.
-Researchers identified the documentation of APIs as both the primary source of information as well as the key obstacle for API usability \cite{Robillard_what_makes}.
-Hence the quality of the API documentation plays an important role in API usability.
-To this regard, researchers have identified the qualities of ``good API documentation'' as follows: complete, correct, includes thorough explanations and code examples, provides consistent presentation and organization \cite{Robillard_what_makes,Myers_study}.
-Today, there are several tools such as Junit\footnote{\url{http://junit.org/}}, UsETeC \cite{zhu2014mining}, Jadeite \cite{Jadeite}, APIMiner \cite{montandon2013documenting}, Roast \cite{Hoffman_api_documentation} that can be used to document local APIs with the aforementioned qualities .
-
-While there are overlaps between the documentation requirements of local and REST APIs, there are significant requirements that are unique to each.
-For example, REST API documentation needs to include information about its API endpoints, HTTP headers, and request and response payloads.
-On the other hand, documentation of local APIs need to explain the classes and methods.
-Due to such differences between local and REST APIs, the existing technique and tools for local API documentation cannot be readily used to generate REST API documentation.
-
-\begin{figure}[htb]
-  \includegraphics[width=\linewidth]{manual_workflow.png}
-  \caption{Manual REST API Documentation Steps}
-  \label{fig:manual}
-\end{figure}
-
-Currently, the process of documenting REST APIs are largely manual.
-Conceptually, the high level steps involved in the manual process can be identified as shown in Figure \ref{fig:manual}.
-These steps S1-S7 can be described as follows: S1 - API developer prepares an example API with the required HTTP parameters and request body, S2 - API developer uses a REST API client to make the example API call, S3 - the API developer records the HTTP data , S4 - API developer then edits the recorded data so that only relevant content is selected for documentation, S5 - API developers use any custom content to describe the API example, S6 - API developers combine the custom content with the edited content from the HTTP traffic into HTML to publish to the web, and S7 - API developer adds custom overview information to explain API concepts and general rules, and publishes the final documentation so other developers can learn the API.
-The steps S1-S6 are repeated for every API action that needs to be documented, and S1-S7 need to repeated for every version of the REST API as it evolves.
-As depicted here, this manual process can be both time consuming and error-prone.
-
-This thesis is ogranized as follows.
-In Chapter~\ref{chapter:case_study}, I discuss a case study to summarize the state of existing literature and industry practices related to the versioning, documentation, and change communication patterns of evolving Web APIs.
-The primary findings from this study are as follows: Web APIs evolve frequently, often several times a week, including both compatible and incompatible API changes, using manual or bespoke methods to version, document, and communicate the changes as they evolve.
-I found a lack of reusable techniques in the existing literature to automatically generate and maintain the documentation of multi-version evolving REST APIs with usage examples.
-I persued further research to solve this problem as discussed in the following paragraphs.
-
-In Chapter~\ref{chapter:spy_rest}, I present the following list of REST API documentation requirements by studying the literature and applying the findings from Chapter~\ref{chapter:case_study}: 1) automated, 2) example based, 3) executable, 4) version aware, 5) customizable, 6) reusable, and 7) collaborative.
-I present, SpyREST, a novel technique to intercept example REST API calls using an HTTP proxy server to collect and autmatically synthesize API traffic to generate and update API documentation to satisfy the aforementioned requirements.
-The primary advantage of the presented technique is that the documentation is generated from executable code.
-Existing techniques that rely on formatted comments or manually written API descriptions following custom languages require manual effort to find and update the API documentation everytime an API evolves.
-
-In Chapter~\ref{chapter:demo_paper}, I present a prototype implementation of the aformentioned technique to demonstrate a reusable tool and compare it against the existing REST API documentation tools that rely on custom API description languages.
-I also present an example where automated test code can be used to generate and maintain REST API documentation using the developed tool.
-
-I discuss an industrial evaluation of the proposed technique and the tool in Chapter~\ref{chapter:cisco_study}.
-The evaluation is carried out based on the data collected from an eighteen month of production use of SpyREST at Cisco.
-From the study I found that it is feasible to use the proposed interception technique and automated tests to document evolving REST APIs satisfying the aforementioned requirements.
-Automated always-updated REST API documentation was found to help establish a quick feedback cycle among the stakeholders.
-I also discuss several refinements and limitations of the proposed technique based on the lessons learned from this case study.
-
-In Chapter~\ref{chapter:controlled_study}, I discuss a controlled study performed with 26 professional software engineers to undestand the impact of usage examples in API documentation on REST API client developers.
-I enahanced an existing REST API documentation with usage examples containing data from the API's test code and provided it to the participants of one of two study groups.
-I found that participants using the original documentation without usage examples faced obstacles related to using appropriate data types, data formats, and required elements in the HTTP input.
-I observed that providing usage examples with data taken from API test code reduces the obstacles and improves API client developer success rate, reduces required trial attempts, and improves satisfaction ratings.
-
-In summary, this thesis makes the following key contributions to the body of research in software engineering:
-
-\begin{enumerate}
-  \item A list of REST API documentation requirements derived from the existing literature and the current state of industry practices to be used as a guideline by practitioners and researchers.
-  \item SpyREST, a novel technique and a reusable tool that uses interception to automatically generate and maintain evolving REST API documentation with usage examples from executable code.
-  \item An industrial evaluation of SpyREST to demonstrate the feasibility of the proposed technique that practitioners can follow to improve their REST API documentation experience.
-  \item An empricial evidence showing the common obstacles faced by REST API client developers that can be reduced with usage examples to help the prioritization of REST API documentation efforts.
-\end{enumerate}
-
+\include{introduction}
 \include{case_study/case_study}

### 2017-07-25 build.sh

@@ -1,6 +1,7 @@
 set -e
 git clean -fX
 pdflatex root.tex
+bibtex introduction.aux
 bibtex controlled_study/controlled_study.aux
 bibtex spy_rest/spy_rest.aux
 bibtex demo_paper/demo_paper.aux

### 2017-08-01 introduction.tex
adds the new intro with a greater detail

@@
 Application Programming Interfaces, commonly known as APIs, are used to express a software component in terms of its operations, inputs, outputs, and their types\footnote{\url{https://en.wikipedia.org/wiki/Application_programming_interface}}.
-Robillard defines an API as follows: An API is the interface to implement functionality that developers can access to perform various tasks \cite{Robillard_a_field_study} \cite{Robillard_what_makes}.
+Robillard defines an API as follows: An API is the interface to implement functionality that developers can access to perform various tasks \cite{Robillard_a_field_study, Robillard_what_makes}.
 APIs enable multiple software components to interact with each other.
@@
 Fielding defined Representational State Transfer or REST as an architectural style for developing distributed hypermedia systems \cite{Fielding_rest}.
-For example, the GitHub REST API\footnote{\url{https://developer.github.com/v3/repos/\#create}} has a resource called \texttt{Repository} to denote a code repository that can be identified by the URL \url{/user/repos}.
+For example, GitHub, an online code repostory has a REST API that can be used by third-party applications to integrate with GitHub features.
+The GitHub REST API\footnote{\url{https://developer.github.com/v3/repos/\#create}} has a resource called \texttt{Repository} to denote a code repository that can be identified by the URL \url{/user/repos}.
 To create a new \texttt{Repository} for a user, the GitHub API can be invoked via HTTP $POST$ at \url{/user/repos} using the following JSON representation of a \texttt{Repository}:
@@
 To this regard, researchers have identified the qualities of ``good API documentation'' as follows: complete, correct, includes thorough explanations and code examples, provides consistent presentation and organization \cite{Robillard_what_makes,Myers_study}.
-Today, there are several tools such as Junit\footnote{\url{http://junit.org/}}, UsETeC \cite{zhu2014mining}, Jadeite \cite{Jadeite}, APIMiner \cite{montandon2013documenting}, Roast \cite{Hoffman_api_documentation} that can be used to document local APIs with the aforementioned qualities .
+Today, there are several tools such as Javadoc\footnote{\url{http://www.oracle.com/technetwork/java/javase/documentation/index-jsp-135444.html}}, UsETeC \cite{zhu2014mining}, Jadeite \cite{Jadeite}, APIMiner \cite{montandon2013documenting}, Roast \cite{Hoffman_api_documentation} that can be used to document local APIs with the aforementioned qualities .
 
 While there are overlaps between the documentation requirements of local and REST APIs, there are significant requirements that are unique to each.
-For example, REST API documentation needs to include information about its API endpoints, HTTP headers, and request and response payloads.
-On the other hand, documentation of local APIs need to explain the classes and methods.
+The documentation of most local APIs contain a structrual definition of the API classses and methods.
+API usage examples for local APIs are documented using the same programming language of the API.
+On the other hand, REST APIs are defined using HTTP specific parameters such as API endpoints, query parameters, HTTP headers and body.
+REST API clients can use any programming language agnostic to the API implementation language.
 Due to such differences between local and REST APIs, the existing technique and tools for local API documentation cannot be readily used to generate REST API documentation.
@@
 As depicted here, this manual process can be both time consuming and error-prone.
+In my research, I have aimed to automate the repeating manual efforts involved in the aforementioned workflow.
 
@@
 In Chapter~\ref{chapter:case_study}, I discuss a case study to summarize the state of existing literature and industry practices related to the versioning, documentation, and change communication patterns of evolving Web APIs.
+The results are based on a study of nine industrical APIs selected based on their diverse application domain, maturity, and popularity including both open and closed source implementations.
 The primary findings from this study are as follows: Web APIs evolve frequently, often several times a week, including both compatible and incompatible API changes, using manual or bespoke methods to version, document, and communicate the changes as they evolve.
@@
 In Chapter~\ref{chapter:spy_rest}, I present the following list of REST API documentation requirements by studying the literature and applying the findings from Chapter~\ref{chapter:case_study}: 1) automated, 2) example based, 3) executable, 4) version aware, 5) customizable, 6) reusable, and 7) collaborative.
+I discuss the available REST API documentation techniques that require repeatitive manual efforts to generate and maintain custom formatted API descriptions that fail to satisfy these requirements.
 I present, SpyREST, a novel technique to intercept example REST API calls using an HTTP proxy server to collect and autmatically synthesize API traffic to generate and update API documentation to satisfy the aforementioned requirements.
 The primary advantage of the presented technique is that the documentation is generated from executable code.
-Existing techniques that rely on formatted comments or manually written API descriptions following custom languages require manual effort to find and update the API documentation everytime an API evolves.
+As a result, a continuously updated REST API documentation with realistic usage examples is achievable with a reduced manual effort
 
@@
 I discuss an industrial evaluation of the proposed technique and the tool in Chapter~\ref{chapter:cisco_study}.
-The evaluation is carried out based on the data collected from an eighteen month of production use of SpyREST at Cisco.
-From the study I found that it is feasible to use the proposed interception technique and automated tests to document evolving REST APIs satisfying the aforementioned requirements.
+The evaluation is carried out based on the data collected from an eighteen month of production use of SpyREST at Cisco, my current employer.
+From the study I found that it is feasible to use the proposed interception technique and automated tests to document evolving REST APIs satisfying the previously preseted REST API documentation requirements.
 Automated always-updated REST API documentation was found to help establish a quick feedback cycle among the stakeholders.
@@
 In Chapter~\ref{chapter:controlled_study}, I discuss a controlled study performed with 26 professional software engineers to undestand the impact of usage examples in API documentation on REST API client developers.
-I enahanced an existing REST API documentation with usage examples containing data from the API's test code and provided it to the participants of one of two study groups.
-I found that participants using the original documentation without usage examples faced obstacles related to using appropriate data types, data formats, and required elements in the HTTP input.
+I mapped the participants into two groups and in API troduced usage examples as the control variable.
+I enhanced an existing REST API documentation with new usage examples.
+I gave the enhanced API documentation to participants from one of the study groups.
+I found that participants using the original documentation without usage examples faced common obstacles related to using appropriate data types, data formats, and required elements in the HTTP headers and request body.
 I observed that providing usage examples with data taken from API test code reduces the obstacles and improves API client developer success rate, reduces required trial attempts, and improves satisfaction ratings.
+The findings from this study provides an empirical evidence that by using the proposed technique of generating REST API documentation from API test code can help API client developers to achive a higher level of success with a reduced effort.
 

### 2017-08-01 introduction.tex
revised tool paper intro

@@
 In Chapter~\ref{chapter:demo_paper}, I present a prototype implementation of the aformentioned technique to demonstrate a reusable tool and compare it against the existing REST API documentation tools that rely on custom API description languages.
-I also present an example where automated test code can be used to generate and maintain REST API documentation using the developed tool.
+In this prototype implementation, I discuss the process that transforms a number of example HTTP requests and responses into a usable REST API documentation.
+I also present an example where automated test code is used to generate and maintain a continuously updated REST API documentation using the developed tool.
 

### 2017-08-01 root.tex
adds the new intro with a greater detail
revised tool paper intro

@@
 
-I am thankful to my supervisor Dr. Frank Maurer for giving me the opportunity and valuable guidance to pursue this research.
+I am thankful to my supervisor Dr.
+Frank Maurer for giving me the opportunity and valuable guidance to pursue this research.
+I am also thankful to Dr.
+Craig Anslow for his continuous encouragement and the countless number of hours that he spent to provide feedback about my work.
 
-I am also thankful to Dr. Craig Anslow for his continuous encouragement and the countless number of hours that he spent to provide feedback about my work.
+I am grateful to Cisco for allowing me to pursue this research alongside my job. I am grateful to the study participants in my case studies for sharing their time with me on the experiments.
 \newpage
@@
 \begin{center}
-  To Shahana
+  To my dear wife.
 \end{center}
@@
 In my second study, I investigated the existing literature on API usablity to derive a set of re- quirements that are applicable to REST API documentation and proposed an idea of automatically generating and maintaining REST API documentation from executable code.
-I found that existing techniques for REST API documentation systems relied on custom API description languagees that are manually maintained as a parallel artifact to the API code, causing potential drift between the actual API and it’s documentation as it evolves over time.
+I found that existing techniques for REST API documentation systems relied on custom API description languagees that are manually maintained as a parallel artifact to the API code, causing potential drift between the actual API and its documentation as it evolves over time.
 By using interception, I have demon- strated that it is possible to automatically record and synthesize the recorded API traffic to generate pragmatic REST API documentation from its test code.

### 2017-08-01 controlled_study/controlled_study.tex
DRY up
structurally complete, sans abstract / introduction / conclusion
all chapters added
first reasonably ok pass

@@
 \chapter{A Study of the Effectiveness of Usage Examples in REST API Documentation}
+\label{chapter:controlled_study}
 
@@
 According to the code repository, there are a total of 99 and 46 contributors that had at least one commit to the code repository of the API and its documentation, respectively.
-These properties satisfy R1 AND R2, our requirements for using a representative API.
+These properties satisfy R1 and R2, our requirements for using a representative API.
 By selecting an open-source project we are able to access the source-code to inspect the implementation and documentation technique of the WordPress REST API.
@@
   \end{center}
-  \caption{Screenshot of documentation for Post/date}
+  \caption{Screenshot of Documentation for Post/date}
   \label{fig:schema}
@@
 \bibliographystyle{IEEEtran}
-\bibliography{controlled_study/IEEEabrv,controlled_study/refs}
+\bibliography{IEEEabrv,controlled_study/refs}

### 2017-08-01 symbols.tex
first reasonably ok pass

%%%%%%%%%%%%%%%%%%%%%%%
%Sample List of Symbols
%%%%%%%%%%%%%%%%%%%%%%%

\begin{tabbing}
Symbol~~~~~\= \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \  \parbox{5in}{Definition}\\

\addsymbol \mbox{API}: {Application Programming Interface}
\addsymbol \mbox{HTTP}: {Hyper Text Transfer Protocol}
\addsymbol \mbox{JSON}: {JavaScript Object Notation}
\addsymbol \mbox{REST}: {Representational State Transfer, used interchangeably with RESTful}
\addsymbol \mbox{SOAP}: {Simple Object Protocol}
\addsymbol \mbox{WSDL}: {Web Service Definition Language}
\addsymbol \mbox{XML}: {Extensible Markup Language}

% .
% .
% .
% ALWAYS KEEP THE FOLLOWING LINE
\end{tabbing}

### 2017-08-01 introduction.tex
adds the self refs

@@
 Fielding defined Representational State Transfer or REST as an architectural style for developing distributed hypermedia systems \cite{Fielding_rest}.
-For example, GitHub, an online code repostory has a REST API that can be used by third-party applications to integrate with GitHub features.
+For example, GitHub, an online code repostory hosting tool, has a REST API that can be used by third-party applications to integrate with GitHub features.
 The GitHub REST API\footnote{\url{https://developer.github.com/v3/repos/\#create}} has a resource called \texttt{Repository} to denote a code repository that can be identified by the URL \url{/user/repos}.
@@
 This thesis is ogranized as follows.
-In Chapter~\ref{chapter:case_study}, I discuss a case study to summarize the state of existing literature and industry practices related to the versioning, documentation, and change communication patterns of evolving Web APIs.
-The results are based on a study of nine industrical APIs selected based on their diverse application domain, maturity, and popularity including both open and closed source implementations.
+In Chapter~\ref{chapter:case_study}, I discuss a case study to summarize the state of existing literature and industry practices related to the versioning, documentation, and change communication patterns of evolving Web APIs \cite{sohan2015case}.
+The results are based on a study of nine industrical APIs selected based on their diverse application domains, maturity level, and popularity including both open and closed source implementations.
 The primary findings from this study are as follows: Web APIs evolve frequently, often several times a week, including both compatible and incompatible API changes, using manual or bespoke methods to version, document, and communicate the changes as they evolve.
@@
 
-In Chapter~\ref{chapter:spy_rest}, I present the following list of REST API documentation requirements by studying the literature and applying the findings from Chapter~\ref{chapter:case_study}: 1) automated, 2) example based, 3) executable, 4) version aware, 5) customizable, 6) reusable, and 7) collaborative.
+In Chapter~\ref{chapter:spy_rest}, I present the following list of REST API documentation requirements by studying the literature and applying the findings from Chapter~\ref{chapter:case_study}: 1) automated, 2) example based, 3) executable, 4) version aware, 5) customizable, 6) reusable, and 7) collaborative \cite{sohan2015spyrest}.
 I discuss the available REST API documentation techniques that require repeatitive manual efforts to generate and maintain custom formatted API descriptions that fail to satisfy these requirements.
@@
 
-In Chapter~\ref{chapter:demo_paper}, I present a prototype implementation of the aformentioned technique to demonstrate a reusable tool and compare it against the existing REST API documentation tools that rely on custom API description languages.
+In Chapter~\ref{chapter:demo_paper}, I present a prototype implementation of the aformentioned technique to demonstrate a reusable tool and compare it against the existing REST API documentation tools that rely on custom API description languages \cite{sohan2015spyrest_tool}.
 In this prototype implementation, I discuss the process that transforms a number of example HTTP requests and responses into a usable REST API documentation.
@@
 
-I discuss an industrial evaluation of the proposed technique and the tool in Chapter~\ref{chapter:cisco_study}.
+I discuss an industrial evaluation of the proposed technique and the tool in Chapter~\ref{chapter:cisco_study} \cite{sohan_cisco}.
 The evaluation is carried out based on the data collected from an eighteen month of production use of SpyREST at Cisco, my current employer.
-From the study I found that it is feasible to use the proposed interception technique and automated tests to document evolving REST APIs satisfying the previously preseted REST API documentation requirements.
+From the study I found that it is feasible to use the proposed interception technique and automated tests to document evolving REST APIs satisfying the previously presented REST API documentation requirements.
 Automated always-updated REST API documentation was found to help establish a quick feedback cycle among the stakeholders.
@@
 
-In Chapter~\ref{chapter:controlled_study}, I discuss a controlled study performed with 26 professional software engineers to undestand the impact of usage examples in API documentation on REST API client developers.
-I mapped the participants into two groups and in API troduced usage examples as the control variable.
+In Chapter~\ref{chapter:controlled_study}, I discuss a controlled study performed with 26 professional software engineers to undestand the impact of usage examples in API documentation on REST API client developers \cite{sohan_vlhcc}.
+I mapped the participants into two groups and introduced usage examples as the control variable.
 I enhanced an existing REST API documentation with new usage examples.

### 2017-08-01 root.tex
first reasonably ok pass
adds the self refs

@@
 Supervisory Committee Member\\
-School of Eng. and Computer Science \\
+School of Eng. and Comp. Science \\
 Victoria University of Wellington\\
@@
 API documentation presents both a problem and an opportunity for API usability.
-Most developers learn to use an API based on its documentation and at the same time face obstacles using the APIs due to lack of necessary information in the documentation \cite{Robillard_what_makes}.
+Most developers learn to use an API based on its documentation and at the same time face obstacles using the APIs due to lack of necessary information in the documentation.
 REST APIs provide interconnectivity between applications over HTTP.

### 2017-08-16 root.tex
better abstract

@@
 
-\title{Automated Example Oriented REST API Documenation}
+\title{Automated Example Oriented REST API Documentation}
 %
@@
 API documentation presents both a problem and an opportunity for API usability.
-Most developers learn to use an API based on its documentation and at the same time face obstacles using the APIs due to lack of necessary information in the documentation.
 REST APIs provide interconnectivity between applications over HTTP.
-Documentation of a REST API is a key information source for software developers that use these types of API.
-As a REST API evolves, the documentation needs to be maintained alongside.
-This implies that REST API developers also need to efficiently generate and maintain the documentation of their APIs with the qualities that make the API usable.
-This introduces the problem of interest in this research: how to document REST APIs?
+Documentation of a REST API is a key information source for API client developers.
+Most REST APIs are documented using a manual approach, which can be time consuming and error-prone.
+REST API developers need to efficiently document their APIs with the qualities that make the API usable.
+In this research I focused on the topic of automated REST API documentation to satisfy this need.
 
-This is a problem because there is a lack of scientifically established guidelines and automated tool support for REST API documentation with usage examples.
-Existing local API documentation techniques and tools are not readily usable for the documentation of REST APIs because REST APIs have HTTP protocol specific parameters, and API clients are agonistic to the programming languages used to implement the APIs.
-Existing REST API documentation techniques rely on custom API definition languages that require manual effort to generate and maintain REST API documentation as a separate artifact as the API evolves.
+In this thesis, I present and evaluate of a novel technique to solve REST API documentaiton requirements.
+I present a set of REST API documentation requirements by studying the existing literature and the current industry practice.
+From this study, I observed that REST APIs evolve frequently, but used a manual or bespoke approach for generating and maintaining their documentation.
+I present a reusable technique with a prototype implementation to automate the REST API documentation process with executable API usage examples.
+The technique involves the interception of example REST API calls using an HTTP proxy server to auto-generate an accurate and updated REST API documentation.
+I present an industrial evaluation of the proposed technique based on a period of eighteen months of production use.
+From this study, I found that it was feasible to leverage API test code to automatically generate an always-updated REST API documentation for evolving REST APIs with usage examples.
+To evaluate the impact of usage examples on REST API client developers, I performed a controlled study with experienced REST API developers.
+From this case study, I found that REST API client developers faced patterns of obstacles using REST APIs that can be reduced by including API usage examples in the API documentation, as suggested in the proposed technique.
 
-In this reasearch, I proposed and evaluated a novel technique to solve REST API documentaiton requirements.
-I identified the requirements by studying the existing literature and the current state of REST API documentation in practice.
-The technique involves the interception of example REST API calls using an HTTP proxy server to auto-generate an accurate and updated REST API documentation.
-
-Practioners can use the presented technique to automatically generate and maintain an updated documentation of REST APIs including the API structure and usage examples from their API test code.
-Researchers can apply the novel concept of using interception to automatically document API usage examples from test scripts to other forms of APIs.
+The findings of this research can be used as a guideline by practitioners to automatically generate an updated REST API documentation with usage examples.
+Researchers can extend the novel concept of using interception to automatically document API usage examples from test scripts to other forms of APIs.
 

### 2017-08-16 introduction.tex
better intro

@@
 REST API clients can use any programming language agnostic to the API implementation language.
-Due to such differences between local and REST APIs, the existing technique and tools for local API documentation cannot be readily used to generate REST API documentation.
+Due to such structural and implemenation differences between local and REST APIs, the existing technique and tools for local API documentation cannot be readily used to generate REST API documentation.
 
@@
 
-Currently, the process of documenting REST APIs are largely manual.
+Currently, the process of documenting REST APIs is largely manual.
 Conceptually, the high level steps involved in the manual process can be identified as shown in Figure \ref{fig:manual}.
-These steps S1-S7 can be described as follows: S1 - API developer prepares an example API with the required HTTP parameters and request body, S2 - API developer uses a REST API client to make the example API call, S3 - the API developer records the HTTP data , S4 - API developer then edits the recorded data so that only relevant content is selected for documentation, S5 - API developers use any custom content to describe the API example, S6 - API developers combine the custom content with the edited content from the HTTP traffic into HTML to publish to the web, and S7 - API developer adds custom overview information to explain API concepts and general rules, and publishes the final documentation so other developers can learn the API.
+These steps S1-S7 can be described as follows:
+
+\begin{itemize}
+  \item S1 - API developer prepares an example API with the required HTTP parameters and request body.
+  \item S2 - API developer uses a REST API client to make the example API call.
+  \item S3 - the API developer records the HTTP data.
+  \item S4 - API developer then edits the recorded data so that only relevant content is selected for documentation.
+  \item S5 - API developers use any custom content to describe the API example.
+  \item S6 - API developers combine the custom content with the edited content from the HTTP traffic into HTML to publish to the web.
+  \item S7 - API developer adds custom overview information to explain API concepts and general rules, and publishes the final documentation so other developers can learn the API
+\end{itemize}
+
 The steps S1-S6 are repeated for every API action that needs to be documented, and S1-S7 need to repeated for every version of the REST API as it evolves.
 As depicted here, this manual process can be both time consuming and error-prone.
-In my research, I have aimed to automate the repeating manual efforts involved in the aforementioned workflow.
 
-This thesis is ogranized as follows.
-In Chapter~\ref{chapter:case_study}, I discuss a case study to summarize the state of existing literature and industry practices related to the versioning, documentation, and change communication patterns of evolving Web APIs \cite{sohan2015case}.
+The overarching goal of my research was to improve the current state of REST API documentation process by automating the repeating manual steps in the aforementioned workflow.
+More specifically, I aimed to solve the following research goals:
+
+\begin{itemize}
+  \item G1: To identify a set of REST API documentation requirements.
+  \item G2: To develop and evaluate a technique to overcome the commonly observed but unresolved challenges related to REST API documentation.
+\end{itemize}
+
+To reach these goals, I have followed a strategy involving five smaller research projects as follows.
+The first research project, related to G1, is described in Chapter~\ref{chapter:case_study}, where I discuss a case study to summarize the state of existing literature and industry practices related to the versioning, documentation, and change communication patterns of evolving Web APIs \cite{sohan2015case}.
 The results are based on a study of nine industrical APIs selected based on their diverse application domains, maturity level, and popularity including both open and closed source implementations.
@@
 
-In Chapter~\ref{chapter:spy_rest}, I present the following list of REST API documentation requirements by studying the literature and applying the findings from Chapter~\ref{chapter:case_study}: 1) automated, 2) example based, 3) executable, 4) version aware, 5) customizable, 6) reusable, and 7) collaborative \cite{sohan2015spyrest}.
+In Chapter~\ref{chapter:spy_rest}, I discuss my second research project, related to both G1 and G2, where I derived a set of REST API documentation requirements and proposed a new technique to satisfy the requirements.
+Here, I present the following list of REST API documentation requirements by studying the literature and applying the findings from Chapter~\ref{chapter:case_study}: 1) automated, 2) example based, 3) executable, 4) version aware, 5) customizable, 6) reusable, and 7) collaborative \cite{sohan2015spyrest}.
 I discuss the available REST API documentation techniques that require repeatitive manual efforts to generate and maintain custom formatted API descriptions that fail to satisfy these requirements.
@@
 The primary advantage of the presented technique is that the documentation is generated from executable code.
-As a result, a continuously updated REST API documentation with realistic usage examples is achievable with a reduced manual effort
+As a result, a continuously updated REST API documentation with realistic usage examples is achievable with a reduced manual effort.
 
-In Chapter~\ref{chapter:demo_paper}, I present a prototype implementation of the aformentioned technique to demonstrate a reusable tool and compare it against the existing REST API documentation tools that rely on custom API description languages \cite{sohan2015spyrest_tool}.
+I present my third research project, related to G2, in Chapter~\ref{chapter:demo_paper}, where I present a prototype implementation of the aformentioned technique \cite{sohan2015spyrest_tool}.
+Here, I demonstrate a reusable tool and compare it against the existing REST API documentation tools that rely on custom API description languages.
 In this prototype implementation, I discuss the process that transforms a number of example HTTP requests and responses into a usable REST API documentation.
@@
 
-I discuss an industrial evaluation of the proposed technique and the tool in Chapter~\ref{chapter:cisco_study} \cite{sohan_cisco}.
+I discuss my fourth research project, related to G2, involving an industrial evaluation of the proposed technique and the tool in Chapter~\ref{chapter:cisco_study} \cite{sohan_cisco}.
 The evaluation is carried out based on the data collected from an eighteen month of production use of SpyREST at Cisco, my current employer.
@@
 
-In Chapter~\ref{chapter:controlled_study}, I discuss a controlled study performed with 26 professional software engineers to undestand the impact of usage examples in API documentation on REST API client developers \cite{sohan_vlhcc}.
+I present my fifth research project, related to G2, in Chapter~\ref{chapter:controlled_study}, where I discuss a controlled study to undestand the impact of usage examples in API documentation on REST API client developers \cite{sohan_vlhcc}.
+I recruited a total of 26 professional software engineers with REST API experience for this study.
 I mapped the participants into two groups and introduced usage examples as the control variable.

### 2017-08-16 root.tex
better intro

@@
 REST API developers need to efficiently document their APIs with the qualities that make the API usable.
-In this research I focused on the topic of automated REST API documentation to satisfy this need.
+In this research, I focused on the topic of automated REST API documentation to satisfy this need.
 
@@
 
-The findings of this research can be used as a guideline by practitioners to automatically generate an updated REST API documentation with usage examples.
-Researchers can extend the novel concept of using interception to automatically document API usage examples from test scripts to other forms of APIs.
+The findings of this research can be used as a guideline by practitioners to automatically generate REST API documentation with usage examples.
+Researchers can extend the novel concept of using interception to automatically document API usage examples for other forms of APIs.
 

### 2017-08-16 introduction.tex
revised intro

@@
 
+\section{Research Goals}
 The overarching goal of my research was to improve the current state of REST API documentation process by automating the repeating manual steps in the aforementioned workflow.
@@
 \begin{itemize}
-  \item G1: To identify a set of REST API documentation requirements.
-  \item G2: To develop and evaluate a technique to overcome the commonly observed but unresolved challenges related to REST API documentation.
+  \item RG1: To identify a set of REST API documentation requirements.
+Existing research on API usability has derived a number of API documentation requirements primarily focusing on local APIs.
+Due to the structural and implementation differences between local and REST APIs, I aimed to identity REST API specific documentation requirements in my research.
+This has several implicatons as follows: having access to a scientifically derived list of REST API documentation requirements allows REST API developers to make an informed decision related to planing and execution of their documentation efforts.
+The requirements provide a background research that can be used by practitioners to compare and contrast existing REST API documentation tools while selecting an appropriate tool for their API.
+REST API documentation tool developers can also use the requirements to build necessary features.
+  \item RG2: To develop and evaluate a technique to overcome the commonly observed but unresolved challenges related to REST API documentation.
+I aimed to devise a reusable technique so that REST APIs can be documented following the identified REST API documentation requirements without following manual or bespoke methods.
+It has the following implications: REST API developers can utilize the technique to overcome the common challenges with the generation and maintenance of their API documentation.
+REST API client developers can access a usable API documentation.
 \end{itemize}
 
-To reach these goals, I have followed a strategy involving five smaller research projects as follows.
-The first research project, related to G1, is described in Chapter~\ref{chapter:case_study}, where I discuss a case study to summarize the state of existing literature and industry practices related to the versioning, documentation, and change communication patterns of evolving Web APIs \cite{sohan2015case}.
+\section{Research Objectives}
+Related to RG1 and RG2, I aimed to address the following research objectives:
+
+\begin{itemize}
+  \item RO1: How are evolving REST APIs documented?
+This question is aimed to analyze the gap between current industry practice around the documentation of evolving REST APIs and the proposed solutions from existing literature.
+This objective helped me to identify the research opportunity to work on unresolved problems related to REST API documentation.
+  \item RO2: What are the requirements for usable REST API documentation? This objective provides a neccessary background work for improving the state of REST API documentation process.
+  \item RO3: What technique can be used to document REST APIs to meet the requirements?
+This is the central research objective.
+I researched to develop a new technique that can fill the gap between existing REST API documentation methods and the identified requirements.
+  \item RO4: Is the proposed technique feasible for REST API developers?
+An evaliation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
+  \item RO5: How does a documentation using the proposed technique impact REST API client developers?
+This question investigates the impact of the proposed technique from the perspective of the REST API client developers.
+\end{itemize}
+
+\section{Cohesion and Unitary Report}
+To address the aformentioned research goals and objectives, my research strategy included five sub-projects, result of which have been published as papers in peer-reviewed international venues.
+Table~\ref{table:chapter_summary} shows a high level overview of the flow of arguments between the sub-projects to meet the research goals.
+
+\begin{table}
+  \caption{Summary and Flow of Arguments of the Sub-projects}
+  \begin{tabular}{|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{4.5cm}|p{6cm}|}
+  \hline
+  Chapter & Goal & Objective & Research Project & Primary Outcome \\
+  \hline
+  Chapter~\ref{chapter:case_study}  & RG1 & RO1 & A case study of Web API evolution \cite{sohan2015case}.
+& A set of challenging problems related to evolving Web APIs.
+Selected REST API documentation as the problem of interest for this research.
+\\
+  \hline
+  Chapter~\ref{chapter:spy_rest} & RG1 and RG2 & RO3 and RO3 & SpyREST: Automated RESTful API documentation using an HTTP proxy server \cite{sohan2015spyrest}.
+& A list of requirements and the concept of using interception to generate REST API documentation to satisfy the requirements.
+\\
+  \hline
+  Chapter~\ref{chapter:demo_paper} & RG2 & RO3 & SpyREST in action: An automated RESTful API documentation tool \cite{sohan2015spyrest_tool}.
+& A reusable tool and a demonstration of the proposed REST API documentation technique.
+\\
+  \hline
+  Chapter~\ref{chapter:cisco_study} & RG2 & RO4 & Automated example oriented REST API documentation at cisco \cite{sohan_cisco}.
+& An empirical evidence that the proposed technique is feasible for practical use, and a list of problems for potential future research.\\
+  \hline
+  Chapter~\ref{chapter:controlled_study} & RG2 & RO5 & A study of the effectiveness of usage examples in REST API documentation \cite{sohan_vlhcc}.
+& An empirical evidence that REST API client developers can overcome common obstacles when the documentation contains usage examples, as suggested in the proposed technique.\\
+  \hline
+\end{tabular}
+\label{table:chapter_summary}
+\end{table}
+
+The first research project, related to RG1, is described in Chapter~\ref{chapter:case_study}, where I discuss a case study to summarize the state of existing literature and industry practices related to the versioning, documentation, and change communication patterns of evolving Web APIs \cite{sohan2015case}.
 The results are based on a study of nine industrical APIs selected based on their diverse application domains, maturity level, and popularity including both open and closed source implementations.
@@
 
-In Chapter~\ref{chapter:spy_rest}, I discuss my second research project, related to both G1 and G2, where I derived a set of REST API documentation requirements and proposed a new technique to satisfy the requirements.
+In Chapter~\ref{chapter:spy_rest}, I discuss my second research project, related to both RG1 and RG2, where I derived a set of REST API documentation requirements and proposed a new technique to satisfy the requirements.
 Here, I present the following list of REST API documentation requirements by studying the literature and applying the findings from Chapter~\ref{chapter:case_study}: 1) automated, 2) example based, 3) executable, 4) version aware, 5) customizable, 6) reusable, and 7) collaborative \cite{sohan2015spyrest}.
@@
 
-I present my third research project, related to G2, in Chapter~\ref{chapter:demo_paper}, where I present a prototype implementation of the aformentioned technique \cite{sohan2015spyrest_tool}.
+I present my third research project, related to RG2, in Chapter~\ref{chapter:demo_paper}, where I present a prototype implementation of the aformentioned technique \cite{sohan2015spyrest_tool}.
 Here, I demonstrate a reusable tool and compare it against the existing REST API documentation tools that rely on custom API description languages.
@@
 
-I discuss my fourth research project, related to G2, involving an industrial evaluation of the proposed technique and the tool in Chapter~\ref{chapter:cisco_study} \cite{sohan_cisco}.
+I discuss my fourth research project, related to RG2, involving an industrial evaluation of the proposed technique and the tool in Chapter~\ref{chapter:cisco_study} \cite{sohan_cisco}.
 The evaluation is carried out based on the data collected from an eighteen month of production use of SpyREST at Cisco, my current employer.
@@
 
-I present my fifth research project, related to G2, in Chapter~\ref{chapter:controlled_study}, where I discuss a controlled study to undestand the impact of usage examples in API documentation on REST API client developers \cite{sohan_vlhcc}.
+I present my fifth research project, related to RG2, in Chapter~\ref{chapter:controlled_study}, where I discuss a controlled study to undestand the impact of usage examples in API documentation on REST API client developers \cite{sohan_vlhcc}.
 I recruited a total of 26 professional software engineers with REST API experience for this study.
@@
 The findings from this study provides an empirical evidence that by using the proposed technique of generating REST API documentation from API test code can help API client developers to achive a higher level of success with a reduced effort.
+
+In Chatper~\ref{chapter:conclusion}, I conclude by discussing how the research goals are achieved and provide a list of problems for future research.
 

### 2017-08-16 introduction.tex
more introduction tidy up

@@
   \hline
-  Chapter~\ref{chapter:spy_rest} & RG1 and RG2 & RO3 and RO3 & SpyREST: Automated RESTful API documentation using an HTTP proxy server \cite{sohan2015spyrest}.
+  Chapter~\ref{chapter:spy_rest} & RG1 and RG2 & RO2 and RO3 & SpyREST: Automated RESTful API documentation using an HTTP proxy server \cite{sohan2015spyrest}.
 & A list of requirements and the concept of using interception to generate REST API documentation to satisfy the requirements.
@@
 
-The first research project, related to RG1, is described in Chapter~\ref{chapter:case_study}, where I discuss a case study to summarize the state of existing literature and industry practices related to the versioning, documentation, and change communication patterns of evolving Web APIs \cite{sohan2015case}.
-The results are based on a study of nine industrical APIs selected based on their diverse application domains, maturity level, and popularity including both open and closed source implementations.
-The primary findings from this study are as follows: Web APIs evolve frequently, often several times a week, including both compatible and incompatible API changes, using manual or bespoke methods to version, document, and communicate the changes as they evolve.
+The first research project is described in Chapter~\ref{chapter:case_study}, where I discuss a case study to summarize the state of existing literature and industry practices related to the versioning, documentation, and change communication patterns of evolving Web APIs \cite{sohan2015case}.
+The results are based on a study of nine industrical APIs.
 I found a lack of reusable techniques in the existing literature to automatically generate and maintain the documentation of multi-version evolving REST APIs with usage examples.
@@
 
-In Chapter~\ref{chapter:spy_rest}, I discuss my second research project, related to both RG1 and RG2, where I derived a set of REST API documentation requirements and proposed a new technique to satisfy the requirements.
-Here, I present the following list of REST API documentation requirements by studying the literature and applying the findings from Chapter~\ref{chapter:case_study}: 1) automated, 2) example based, 3) executable, 4) version aware, 5) customizable, 6) reusable, and 7) collaborative \cite{sohan2015spyrest}.
-I discuss the available REST API documentation techniques that require repeatitive manual efforts to generate and maintain custom formatted API descriptions that fail to satisfy these requirements.
-I present, SpyREST, a novel technique to intercept example REST API calls using an HTTP proxy server to collect and autmatically synthesize API traffic to generate and update API documentation to satisfy the aforementioned requirements.
+In Chapter~\ref{chapter:spy_rest}, I discuss my second research project where I derived a set of REST API documentation requirements and proposed a new technique to satisfy the requirements \cite{sohan2015spyrest}.
+Here, I present a list of REST API documentation requirements by studying the literature and applying the findings from Chapter~\ref{chapter:case_study}.
+I discuss the available REST API documentation techniques and compare against the requirements.
+I present, SpyREST, a novel technique to intercept example REST API calls using an HTTP proxy server to collect and autmatically synthesize API traffic to generate and update API documentation to satisfy the identified requirements.
 The primary advantage of the presented technique is that the documentation is generated from executable code.
@@
 I present my third research project, related to RG2, in Chapter~\ref{chapter:demo_paper}, where I present a prototype implementation of the aformentioned technique \cite{sohan2015spyrest_tool}.
-Here, I demonstrate a reusable tool and compare it against the existing REST API documentation tools that rely on custom API description languages.
 In this prototype implementation, I discuss the process that transforms a number of example HTTP requests and responses into a usable REST API documentation.
@@
 I present my fifth research project, related to RG2, in Chapter~\ref{chapter:controlled_study}, where I discuss a controlled study to undestand the impact of usage examples in API documentation on REST API client developers \cite{sohan_vlhcc}.
-I recruited a total of 26 professional software engineers with REST API experience for this study.
-I mapped the participants into two groups and introduced usage examples as the control variable.
-I enhanced an existing REST API documentation with new usage examples.
-I gave the enhanced API documentation to participants from one of the study groups.
-I found that participants using the original documentation without usage examples faced common obstacles related to using appropriate data types, data formats, and required elements in the HTTP headers and request body.
 I observed that providing usage examples with data taken from API test code reduces the obstacles and improves API client developer success rate, reduces required trial attempts, and improves satisfaction ratings.
@@
 
-In Chatper~\ref{chapter:conclusion}, I conclude by discussing how the research goals are achieved and provide a list of problems for future research.
+In Chatper~\ref{chapter:conclusion}, I conclude by discussing how the research objectives are achieved and provide a list of problems for future research.
 

### 2017-08-16 introduction.tex
contribution summary is a section

@@
 \begin{itemize}
-  \item RG1: To identify a set of REST API documentation requirements.
-Existing research on API usability has derived a number of API documentation requirements primarily focusing on local APIs.
+  \item \textit{RG1: To identify a set of REST API documentation requirements.} Existing research on API usability has derived a number of API documentation requirements primarily focusing on local APIs.
 Due to the structural and implementation differences between local and REST APIs, I aimed to identity REST API specific documentation requirements in my research.
-This has several implicatons as follows: having access to a scientifically derived list of REST API documentation requirements allows REST API developers to make an informed decision related to planing and execution of their documentation efforts.
+This has several implicatons as follows: having access to a scientifically derived list of REST API documentation requirements allows REST API developers to make an informed decision related to planning and execution of their documentation efforts.
 The requirements provide a background research that can be used by practitioners to compare and contrast existing REST API documentation tools while selecting an appropriate tool for their API.
 REST API documentation tool developers can also use the requirements to build necessary features.
-  \item RG2: To develop and evaluate a technique to overcome the commonly observed but unresolved challenges related to REST API documentation.
-I aimed to devise a reusable technique so that REST APIs can be documented following the identified REST API documentation requirements without following manual or bespoke methods.
+  \item \textit{RG2: To develop and evaluate a technique to overcome the commonly observed but unresolved challenges related to REST API documentation.} I aimed to devise a reusable technique so that REST APIs can be documented following the identified REST API documentation requirements without following manual or bespoke methods.
 It has the following implications: REST API developers can utilize the technique to overcome the common challenges with the generation and maintenance of their API documentation.
@@
 \begin{itemize}
-  \item RO1: How are evolving REST APIs documented?
-This question is aimed to analyze the gap between current industry practice around the documentation of evolving REST APIs and the proposed solutions from existing literature.
+  \item \textit{RO1: How are evolving REST APIs documented?} This question is aimed to analyze the gap between current industry practice around the documentation of evolving REST APIs and the proposed solutions from existing literature.
 This objective helped me to identify the research opportunity to work on unresolved problems related to REST API documentation.
-  \item RO2: What are the requirements for usable REST API documentation? This objective provides a neccessary background work for improving the state of REST API documentation process.
-  \item RO3: What technique can be used to document REST APIs to meet the requirements?
-This is the central research objective.
+  \item \textit{RO2: What are the requirements for usable REST API documentation?} This objective provides a neccessary background work for improving the state of REST API documentation process.
+  \item \textit{RO3: What technique can be used to document REST APIs to meet the requirements?} This is the central research objective.
 I researched to develop a new technique that can fill the gap between existing REST API documentation methods and the identified requirements.
-  \item RO4: Is the proposed technique feasible for REST API developers?
-An evaliation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
-  \item RO5: How does a documentation using the proposed technique impact REST API client developers?
-This question investigates the impact of the proposed technique from the perspective of the REST API client developers.
+  \item \textit{RO4: Is the proposed technique feasible for REST API developers?} An evaliation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
+  \item \textit{RO5: How does a documentation using the proposed technique impact REST API client developers?} This question investigates the impact of the proposed technique from the perspective of the REST API client developers.
 \end{itemize}
@@
 
+\section{Contributions}
 In summary, this thesis makes the following key contributions to the body of research in software engineering:

### 2017-08-16 root.tex
revised intro
conclusion is in

@@
 \usepackage{tikz,pgfplots,pgfplotstable}
+\usetikzlibrary{shapes,arrows}
 \usepackage{mathtools}
@@
 \chapter{Conclusion}
+\label{chapter:conclusion}
 
-In this thesis, I presented a novel idea of using interception to automate the process of REST API documentation with usage examples from executable code.
-I started by investigating the current industry practices and challenges associated with developing and maintaining evolving Web APIs.
-I found modern Web APIs predominently embraced REST architectural style, but the existing literature had primarily focused on solving problems related to the SOAP APIs.
-I also found that most REST APIs are documented using manual approaches that are time consuming and error- prone to maintain when an API evolves several times a month.
-Through my research, I aimed to fill this gap between the REST API documentation challenges and the existing literature by finding a reusable solution to the common problems.
+In this thesis, I presented a set of five primary research objectives related to the topic of REST API documentation.
+I have addressed the first research objective, RO1, of finding the gap between the current industry practices and existing research on REST API documentation by performing a case study in my first research sub-project.
+I started by reviewing the existing literature on Web API evolution.
+From the existing empirical research on Web API evolution, I found that researchers have focused on profiling SOAP based web services to identify the common change patterns.
+They suggested focusing on the maintenance efforts related to Web API evolution because of its lingering effects.
+In the existing literature, I found several techniques proposed by researchers related to supporting the evolution of SOAP APIs.
+Analyzing the existing literature on tool support for Web API evolution, I found tools developed by researchers that primarly rely on custom formatted API specification documents that require a manual effort to generate and maintain.
+In this sub-project, I have also analyzed the evolution of nine Web APIs in production, selected based on their industry, popularity, maturity, and source code access to understand the current state of practice around Web API versioning, documentation, and change comminucation patterns.
+By comparing and contrasting the findings from the literature review and the industry practices, I observed that it was important to carry out further research to find techniques to support versioning, documentation, and change communication needs for evolving REST APIs.
 
-In my second study, I investigated the existing literature on API usablity to derive a set of re- quirements that are applicable to REST API documentation and proposed an idea of automatically generating and maintaining REST API documentation from executable code.
-I found that existing techniques for REST API documentation systems relied on custom API description languagees that are manually maintained as a parallel artifact to the API code, causing potential drift between the actual API and its documentation as it evolves over time.
-By using interception, I have demon- strated that it is possible to automatically record and synthesize the recorded API traffic to generate pragmatic REST API documentation from its test code.
-Additionally, the technique allows API de- velopers to collaboratively document the API with realistic and executable examples for multiple versions as the API evolves.
+I addressed RO2, finding RESAPI T documentation requirements, in my second research sub-project.
+I started by analyzing the existing literature on API usability to summarize a list of general API documentation requirements as derived by other researchers.
+Then, I analyzed the existing literature on REST API documentation.
+Based on these two analyses, I derived the following list of documentation requirements specific for evolving REST APIs: automated, example based, executable, version aware, customizable, reusable, and collaborative.
+These requirements are used as a background work to address my subsequent research objectives.
 
-In my third study, I evaluated the proposed technique in an industrial setup and found it to be a pragmatic solution for production use.
-The concept of leveraging API tests to maintain an always-updated API documentation has helped the organization establish a quick feedback loop during the API development process.
-Practitioners can incorporate the findings from this study to improve their own REST API documentation practice.
+RO3, the primary objective of finding a new technique to satisfy the documentation requirements of evolving REST APIs, is addressed in my third and fourth sub-projects.
+I observed that most existing techniques for REST API documentation relied on custom API specification formats as an input.
+This approach requires manual effort to generate and maintain the API specification as a separate artifact with the API as the API evolves and fails to satisfy the identified REST API documentation requirements.
+I proposed a novel technique called SpyREST to automatically intercept and transform example REST API calls into usable API documentation following the identified requirements.
+An HTTP proxy server is used to collect data from example API calls.
+The collected data is then synthesized to automatically infer the structural elements of the API such as the version, API resource, HTTP verb, API action, query parameters, request and response body, HTTP headers.
+In addition to the structure, executable API examples are automatically constructed with realistic data.
+I developed a prototype tool to demonstrate a concrete implementation of the proposed concept.
+In a preliminary case study with three evolving REST APIs, I have shown that it is possible to use the proposed technique to generate more a accurate documentation of REST APIs compared to their official documentation that are  manually maintained.
 
-To understand the impact of usage examples that can be derived from API tests, I performed the last study with professional software engineers and found that the examples help reduce the common obstacles faced by REST API client developers.
-I derived a list of recommendations for documenting usage examples to satisfy the common information needs when REST API client developers use an API.
-Practitiones can follow the recommendations as guideline to allocate ap- propriate priority while documenting their REST APIs in a resource constrained environment.
+I addressed RO4, an evaluation of the proposed technique from the perspective of REST API developers, in my fourth sub-project using a case study.
+I presented the proposed technique and the prototype tool to a team of software engineers at Cisco when we were in the process of selecting a documentation method for a suit of new REST APIs to be built.
+I discussed the selection process behind choosing SpyREST as documentation tool for this project.
+In this case study, I discussed how the proposed technique of using test code to drive REST API documentation has been leveraged in Cisco to create and maintain an always-updated and accurate documentation for an evolving REST API in production for a period of eighteen months.
+I gathered feedback from the stakeholders including the API developers, QA engineers, project managers, and customer support engineers to evaluate the merits and shortcomings of the proposed techqniue based on their production use.
+The evaluation results show that it is feasible to use API test code to generate and maintain REST API documentation for an evolving API with the identified REST API documentation requirements.
+From this evaluation, I also found that different stakeholders are able to establish a fast and continuous feedback loop about the API and its documentation while developing the API features when the API documentation is auto updated from executable code.
+Additionally, from this evaluation, I found a list of limitations of the proposed technique related to the lack of support for using version control features for custom edited API documentation content and automatically generated changelogs showing the differences between API elements of different API versions.
+Future research can be carried out to address these limitations.
+
+My last research objective, RO5, to evaluate the impact of the proposed technique on REST API client developers, was addressed in my fifth sub-project.
+I performed a controlled study with 26 professional software engineers with prior REST API experience.
+The participants were asked to perform tasks using a REST API following its accopmpanying documentation.
+I enhanced an existing API documentation with usage examples leveraging the API's test code.
+This enhancement mimics the outcome of the proposed technique but does not use the proposed tool.
+Participants were mapped into two groups.
+One of the groups used the original API documentation and the other group used the enhanced documentation.
+I provided a web based API explorer so that participants could only focus on the required API inputs instead of writing a full API client application to perform the tasks.
+I observed that participants using the original API documentation faced similar patterns of obstacles involving the use of correct data types, data formats, HTTP headers, and request body.
+Participants using the enhanced documentation were able to reduce these obstacles to peform API tasks with a higher success rate, requiring less trial API calls and effort, and a higher satisfaction rating.
+This provides empirical evidence that the proposed technique has a positive impact on the REST API client developers.
+I also derived a list of recommentations related to the content of usage examples for REST API documentation that can be used by practitioners as a guideline to prioritize their documentation efforts.
+From this study, I also observed that participants used more trial API calls for API tasks that involve multiple dependent API calls even while using the enhanced documentation.
+Future research is needed to generate API documentation with examples for complex API usage scenarios to reduce this obstacle.
+
+In summary, through the sub-projects, I have derived a set of REST API documentation requirements, and proposed a novel technique of using interception to automatically document REST APIs with usage examples from executable code.
+I have presented evaluations of the proposed technique from the perspectives of both the developers that write the APIs and the developers that use the APIs.
+Practitioners and researchers can leverage the findings from this thesis to improve the current state of REST API documentation process.
 

### 2017-08-16 introduction.tex
spell corrections

@@
 Fielding defined Representational State Transfer or REST as an architectural style for developing distributed hypermedia systems \cite{Fielding_rest}.
-For example, GitHub, an online code repostory hosting tool, has a REST API that can be used by third-party applications to integrate with GitHub features.
+For example, GitHub, an online code repository hosting tool, has a REST API that can be used by third-party applications to integrate with GitHub features.
 The GitHub REST API\footnote{\url{https://developer.github.com/v3/repos/\#create}} has a resource called \texttt{Repository} to denote a code repository that can be identified by the URL \url{/user/repos}.
@@
 For example, a real estate listing website uses a REST API to collect ``walk score'' and another REST API to show a map view of a property listing.
-By incorporating the map and walk score using REST APIs, the real estate listng site provides important information to their users.
+By incorporating the map and walk score using REST APIs, the real estate listing site provides important information to their users.
 Most REST APIs, including these two examples, are often documented manually or using custom implemented tools specific to the APIs that are not publicly available.
@@
 While there are overlaps between the documentation requirements of local and REST APIs, there are significant requirements that are unique to each.
-The documentation of most local APIs contain a structrual definition of the API classses and methods.
+The documentation of most local APIs contain a structural definition of the API classes and methods.
 API usage examples for local APIs are documented using the same programming language of the API.
@@
 REST API clients can use any programming language agnostic to the API implementation language.
-Due to such structural and implemenation differences between local and REST APIs, the existing technique and tools for local API documentation cannot be readily used to generate REST API documentation.
+Due to such structural and implementation differences between local and REST APIs, the existing technique and tools for local API documentation cannot be readily used to generate REST API documentation.
 
@@
 Due to the structural and implementation differences between local and REST APIs, I aimed to identity REST API specific documentation requirements in my research.
-This has several implicatons as follows: having access to a scientifically derived list of REST API documentation requirements allows REST API developers to make an informed decision related to planning and execution of their documentation efforts.
+This has several implications as follows: having access to a scientifically derived list of REST API documentation requirements allows REST API developers to make an informed decision related to planning and execution of their documentation efforts.
 The requirements provide a background research that can be used by practitioners to compare and contrast existing REST API documentation tools while selecting an appropriate tool for their API.
@@
 This objective helped me to identify the research opportunity to work on unresolved problems related to REST API documentation.
-  \item \textit{RO2: What are the requirements for usable REST API documentation?} This objective provides a neccessary background work for improving the state of REST API documentation process.
+  \item \textit{RO2: What are the requirements for usable REST API documentation?} This objective provides a necessary background work for improving the state of REST API documentation process.
   \item \textit{RO3: What technique can be used to document REST APIs to meet the requirements?} This is the central research objective.
 I researched to develop a new technique that can fill the gap between existing REST API documentation methods and the identified requirements.
-  \item \textit{RO4: Is the proposed technique feasible for REST API developers?} An evaliation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
+  \item \textit{RO4: Is the proposed technique feasible for REST API developers?} An evaluation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
   \item \textit{RO5: How does a documentation using the proposed technique impact REST API client developers?} This question investigates the impact of the proposed technique from the perspective of the REST API client developers.
@@
 \section{Cohesion and Unitary Report}
-To address the aformentioned research goals and objectives, my research strategy included five sub-projects, result of which have been published as papers in peer-reviewed international venues.
+To address the aforementioned research goals and objectives, my research strategy included five sub-projects, result of which have been published as papers in peer-reviewed international venues.
 Table~\ref{table:chapter_summary} shows a high level overview of the flow of arguments between the sub-projects to meet the research goals.
@@
   \hline
-  Chapter~\ref{chapter:cisco_study} & RG2 & RO4 & Automated example oriented REST API documentation at cisco \cite{sohan_cisco}.
+  Chapter~\ref{chapter:cisco_study} & RG2 & RO4 & Automated example oriented REST API documentation at Cisco \cite{sohan_cisco}.
 & An empirical evidence that the proposed technique is feasible for practical use, and a list of problems for potential future research.\\
@@
 The first research project is described in Chapter~\ref{chapter:case_study}, where I discuss a case study to summarize the state of existing literature and industry practices related to the versioning, documentation, and change communication patterns of evolving Web APIs \cite{sohan2015case}.
-The results are based on a study of nine industrical APIs.
+The results are based on a study of nine industrial APIs.
 I found a lack of reusable techniques in the existing literature to automatically generate and maintain the documentation of multi-version evolving REST APIs with usage examples.
-I persued further research to solve this problem as discussed in the following paragraphs.
+I pursued further research to solve this problem as discussed in the following paragraphs.
 
@@
 I discuss the available REST API documentation techniques and compare against the requirements.
-I present, SpyREST, a novel technique to intercept example REST API calls using an HTTP proxy server to collect and autmatically synthesize API traffic to generate and update API documentation to satisfy the identified requirements.
+I present, SpyREST, a novel technique to intercept example REST API calls using an HTTP proxy server to collect and automatically synthesize API traffic to generate and update API documentation to satisfy the identified requirements.
 The primary advantage of the presented technique is that the documentation is generated from executable code.
@@
 
-I present my third research project, related to RG2, in Chapter~\ref{chapter:demo_paper}, where I present a prototype implementation of the aformentioned technique \cite{sohan2015spyrest_tool}.
+I present my third research project, related to RG2, in Chapter~\ref{chapter:demo_paper}, where I present a prototype implementation of the aforementioned technique \cite{sohan2015spyrest_tool}.
 In this prototype implementation, I discuss the process that transforms a number of example HTTP requests and responses into a usable REST API documentation.
@@
 
-I present my fifth research project, related to RG2, in Chapter~\ref{chapter:controlled_study}, where I discuss a controlled study to undestand the impact of usage examples in API documentation on REST API client developers \cite{sohan_vlhcc}.
+I present my fifth research project, related to RG2, in Chapter~\ref{chapter:controlled_study}, where I discuss a controlled study to understand the impact of usage examples in API documentation on REST API client developers \cite{sohan_vlhcc}.
 I observed that providing usage examples with data taken from API test code reduces the obstacles and improves API client developer success rate, reduces required trial attempts, and improves satisfaction ratings.
-The findings from this study provides an empirical evidence that by using the proposed technique of generating REST API documentation from API test code can help API client developers to achive a higher level of success with a reduced effort.
+The findings from this study provides an empirical evidence that by using the proposed technique of generating REST API documentation from API test code can help API client developers to achieve a higher level of success with a reduced effort.
 
-In Chatper~\ref{chapter:conclusion}, I conclude by discussing how the research objectives are achieved and provide a list of problems for future research.
+In Chapter~\ref{chapter:conclusion}, I conclude by discussing how the research objectives are achieved and provide a list of problems for future research.
 
@@
   \item An industrial evaluation of SpyREST to demonstrate the feasibility of the proposed technique that practitioners can follow to improve their REST API documentation experience.
-  \item An empricial evidence showing the common obstacles faced by REST API client developers that can be reduced with usage examples to help the prioritization of REST API documentation efforts.
+  \item An empirical evidence showing the common obstacles faced by REST API client developers that can be reduced with usage examples to help the prioritization of REST API documentation efforts.
 \end{enumerate}

### 2017-08-16 root.tex
spell corrections

@@
 
-In this thesis, I present and evaluate of a novel technique to solve REST API documentaiton requirements.
+In this thesis, I present and evaluate of a novel technique to solve REST API documentation requirements.
 I present a set of REST API documentation requirements by studying the existing literature and the current industry practice.
@@
 In the existing literature, I found several techniques proposed by researchers related to supporting the evolution of SOAP APIs.
-Analyzing the existing literature on tool support for Web API evolution, I found tools developed by researchers that primarly rely on custom formatted API specification documents that require a manual effort to generate and maintain.
-In this sub-project, I have also analyzed the evolution of nine Web APIs in production, selected based on their industry, popularity, maturity, and source code access to understand the current state of practice around Web API versioning, documentation, and change comminucation patterns.
+Analyzing the existing literature on tool support for Web API evolution, I found tools developed by researchers that primarily rely on custom formatted API specification documents that require a manual effort to generate and maintain.
+In this sub-project, I have also analyzed the evolution of nine Web APIs in production, selected based on their industry, popularity, maturity, and source code access, to understand the current state of practice around Web API versioning, documentation, and change communication patterns.
 By comparing and contrasting the findings from the literature review and the industry practices, I observed that it was important to carry out further research to find techniques to support versioning, documentation, and change communication needs for evolving REST APIs.
 
-I addressed RO2, finding RESAPI T documentation requirements, in my second research sub-project.
+I addressed RO2, finding REST API documentation requirements, in my second research sub-project.
 I started by analyzing the existing literature on API usability to summarize a list of general API documentation requirements as derived by other researchers.
@@
 An HTTP proxy server is used to collect data from example API calls.
-The collected data is then synthesized to automatically infer the structural elements of the API such as the version, API resource, HTTP verb, API action, query parameters, request and response body, HTTP headers.
-In addition to the structure, executable API examples are automatically constructed with realistic data.
+The collected data is then synthesized to automatically infer the structural elements of the API such as the version, API resource, HTTP verb, API action, query parameters, request and response body, and HTTP headers.
+In addition to the structure, executable API examples are automatically re-constructed with realistic data from the recorded API calls.
 I developed a prototype tool to demonstrate a concrete implementation of the proposed concept.
-In a preliminary case study with three evolving REST APIs, I have shown that it is possible to use the proposed technique to generate more a accurate documentation of REST APIs compared to their official documentation that are  manually maintained.
+In a preliminary case study with three evolving REST APIs, I have shown that it is possible to use the proposed technique to generate a more accurate documentation of REST APIs compared to their official documentation that are  manually maintained.
 
-I addressed RO4, an evaluation of the proposed technique from the perspective of REST API developers, in my fourth sub-project using a case study.
-I presented the proposed technique and the prototype tool to a team of software engineers at Cisco when we were in the process of selecting a documentation method for a suit of new REST APIs to be built.
+I addressed RO4, an evaluation of the proposed technique from the perspective of REST API developers, in my fourth sub-project using an industrial case study.
+I presented the proposed technique and the prototype tool to a team of software engineers at Cisco when we were in the process of selecting a documentation method for a suite of new REST APIs to be built.
 I discussed the selection process behind choosing SpyREST as documentation tool for this project.
 In this case study, I discussed how the proposed technique of using test code to drive REST API documentation has been leveraged in Cisco to create and maintain an always-updated and accurate documentation for an evolving REST API in production for a period of eighteen months.
-I gathered feedback from the stakeholders including the API developers, QA engineers, project managers, and customer support engineers to evaluate the merits and shortcomings of the proposed techqniue based on their production use.
+I gathered feedback from the stakeholders including the API developers, QA engineers, project managers, and customer support engineers to evaluate the merits and shortcomings of the proposed technique based on their production use.
 The evaluation results show that it is feasible to use API test code to generate and maintain REST API documentation for an evolving API with the identified REST API documentation requirements.
@@
 I performed a controlled study with 26 professional software engineers with prior REST API experience.
-The participants were asked to perform tasks using a REST API following its accopmpanying documentation.
+The participants were asked to perform tasks using a REST API following its accompanying documentation.
 I enhanced an existing API documentation with usage examples leveraging the API's test code.
@@
 I observed that participants using the original API documentation faced similar patterns of obstacles involving the use of correct data types, data formats, HTTP headers, and request body.
-Participants using the enhanced documentation were able to reduce these obstacles to peform API tasks with a higher success rate, requiring less trial API calls and effort, and a higher satisfaction rating.
+Participants using the enhanced documentation were able to reduce these obstacles to perform API tasks with a higher success rate, requiring less trial API calls and effort, and a higher satisfaction rating.
 This provides empirical evidence that the proposed technique has a positive impact on the REST API client developers.
-I also derived a list of recommentations related to the content of usage examples for REST API documentation that can be used by practitioners as a guideline to prioritize their documentation efforts.
+I also derived a list of recommendations related to the content of usage examples for REST API documentation that can be used by practitioners as a guideline to prioritize their documentation efforts.
 From this study, I also observed that participants used more trial API calls for API tasks that involve multiple dependent API calls even while using the enhanced documentation.

### 2017-08-16 case_study/case_study.tex
all chapters added
more spell checks

@@
 \chapter{A Case Study of Web API Evolution}
-
-%\author{\IEEEauthorblockN{S M Sohan, Craig Anslow, Frank Maurer
-%\IEEEauthorblockA{Department of Computer Science\\
-%University of Calgary,
-%Calgary, Alberta T2N 1N4\\ Email: \{smsohan, craig.anslow, frank.maurer\}@ucalgary.ca }
-%}}
-
-%\maketitle
+\label{chapter:case_study}
 
@@
 This study will help inform developers of evolving Web APIs to make decision about versioning, documentation and communication methods.
-
-%\begin{IEEEkeywords}
-%Case Study; RESTful, SOAP, Web API Evolution; WSDL
-%\end{IEEEkeywords}
-
-%\IEEEpeerreviewmaketitle
 
@@
 
-The first author of this paper read the the API homepages and followed relevant hyperlinks to find answers to RQ1.
+The first author of this paper read the API homepages and followed relevant hyperlinks to find answers to RQ1.
 A summary of the strategies were presented to the co-authors, which was used to categorize Web API versioning strategies into groups.
@@
 
-\textbf{Endpoint} - URLs to access the the API.
+\textbf{Endpoint} - URLs to access the API.
 \textbf{Resource} - An input or output object. For example, comment, photo, etc.
@@
 
-We analyzed the API user forums and newsboards to find answers to RQ3.
+We analyzed the API user forums and news boards to find answers to RQ3.
 Messages in these sources were filtered to focus on the evolution of Web APIs.

### 2017-08-16 cisco_study/cisco_study.tex
fix the label/caption order
adds all 5 papers, reasonable ok formatted
structurally complete, sans abstract / introduction / conclusion
all chapters added
more spell checks

@@
-\section{Automated Example Oriented REST API Documentation at Cisco}
-
-%\author{\IEEEauthorblockN{S M Sohan}
-%\IEEEauthorblockA{
-%Security Group\\
-%Cisco Systems Ltd.\\
-%Calgary, Canada\\
-%sosohan@cisco.com}
-%\and
-%\IEEEauthorblockN{Craig Anslow}
-%\IEEEauthorblockA{
-%Department of Computer Science\\
-%Middlesex University\\
-%London, United Kingdom\\
-%c.anslow@mdx.ac.uk}
-%\and
-%\IEEEauthorblockN{Frank Maurer}
-%\IEEEauthorblockA{
-%Department of Computer Science\\
-%University of Calgary\\
-%Calgary, Canada\\
-%frank.maurer@ucalgary.ca}
-%}
+\chapter{Automated Example Oriented REST API Documentation at Cisco}
+\label{chapter:cisco_study}
 
@@
 The findings from this paper can be used by practitioners to introduce automation to reduce the manual effort associated to their REST API documentation process.
-
-%\begin{IEEEkeywords}
-%API; REST; Documentation; Tool; Case study; Test; Automation; HTTP; Web API;
-
-%\end{IEEEkeywords}
 
@@
     }
+    \caption{API Overview in SpyREST}
     \label{fig:screenshot_overview}
-    \caption{API Overview in SpyREST}
   \end{subfigure}
@@
     }
+    \caption{SpyREST Analyzed API Structure}
     \label{fig:screenshot_fields}
-    \caption{SpyREST Analyzed API Structure}
   \end{subfigure}
@@
     }
+    \caption{SpyREST Recorded API Request}
     \label{fig:screenshot_request}
-    \caption{SpyREST Recorded API Request}
   \end{subfigure}
@@
     }
+    \caption{SpyREST Recorded API Response}
     \label{fig:screenshot_response}
-    \caption{SpyREST Recorded API Response}
   \end{subfigure}
 
-  \caption{Screenshots of the REST API documentation from Cisco using SpyREST}
+  \caption{Screenshots of the REST API Documentation from Cisco using SpyREST}
   \label{fig:screenshot}
@@
 
-The next step in the lifecycle of the API documentation is in the automated build server where the test suite is executed against a deployed API on a staging server through a SpyREST proxy.
+The next step in the life cycle of the API documentation is in the automated build server where the test suite is executed against a deployed API on a staging server through a SpyREST proxy.
 This promotes the REST API documentation from the developer workstation to a shared instance that is used by the quality assurance engineers, product management, and the peer developers to verify against the acceptance criteria.
@@
 
-The next step in the lifecycle of the API is a release to one or more of the production environments that customers use.
+The next step in the life cycle of the API is a release to one or more of the production environments that customers use.
 With the production release of an updated API, its documentation is automatically promoted from the staging environment to production.
@@
 First, the data from the staging server is exported in a portable archive file.
-Then, a transformation is performed to update staging specific data such as URLs, email domains, and hostnames within the archive to match the desired production environment.
+Then, a transformation is performed to update staging specific data such as URLs, email domains, and host names within the archive to match the desired production environment.
 Finally, the transformed archive is published to production as a read-only artifact.
@@
 This code reuse minimizes the effort required to document multiple versions of the same API action when the API versions are backward compatible.
-The rest of the lifecycle steps follow similar process as the documentation of a new API action.
+The rest of the life cycle steps follow similar process as the documentation of a new API action.
 

### 2017-09-24 introduction.tex
add changes based on feedback from craig + mea

@@
 \chapter{Introduction}
+
+Robillard defines an API as follows: An API is the interface to implement functionality that developers can access to perform various tasks \cite{Robillard_a_field_study, Robillard_what_makes}.
 Application Programming Interfaces, commonly known as APIs, are used to express a software component in terms of its operations, inputs, outputs, and their types\footnote{\url{https://en.wikipedia.org/wiki/Application_programming_interface}}.
-Robillard defines an API as follows: An API is the interface to implement functionality that developers can access to perform various tasks \cite{Robillard_a_field_study, Robillard_what_makes}.
 APIs enable multiple software components to interact with each other.
+Web APIs allow software-to-software interconnectivity using web technologies as the transport mechanism of information between systems.
 
-REST APIs are a type of APIs that are used to integrate software components using web technologies.
+REST APIs are a type of API that are used to integrate software components using web technologies.
 Fielding defined Representational State Transfer or REST as an architectural style for developing distributed hypermedia systems \cite{Fielding_rest}.
+He described the design of modern Web as an example of a REST-style architecture.
+REST-style architecture allows software components to satisfy the data transfer needs of the modern Internet-scale applications by levergading a generic interface based on URIs and Internet media types that allow intemediate processing such as caching to improve system performance.
+
+REST APIs allow access to application resources over HTTP.
+While REST describes a generic protocol for accessing application resources, the applications are responsible for the specific representation of the resources.
+For example, a REST API may support multiple Internet media types such as JSON, XML, CSV for a single application resource.
+Similarly, two diffenent REST APIs may use the same media type but contain entirely different formats to represent their objects.
+
 For example, GitHub, an online code repository hosting tool, has a REST API that can be used by third-party applications to integrate with GitHub features.
@@
 On the other hand, REST APIs are defined using HTTP specific parameters such as API endpoints, query parameters, HTTP headers and body.
-REST API clients can use any programming language agnostic to the API implementation language.
+REST API clients can be written in any programming language agnostic to the API implementation language.
 Due to such structural and implementation differences between local and REST APIs, the existing technique and tools for local API documentation cannot be readily used to generate REST API documentation.
@@
   \item S5 - API developers use any custom content to describe the API example.
-  \item S6 - API developers combine the custom content with the edited content from the HTTP traffic into HTML to publish to the web.
-  \item S7 - API developer adds custom overview information to explain API concepts and general rules, and publishes the final documentation so other developers can learn the API
+  \item S6 - API developers combine the custom content with the edited content from the HTTP traffic into HTML to publish on the web.
+  \item S7 - API developer adds custom overview information to explain API concepts and general rules, and publishes the final documentation so other developers can learn the API.
 \end{itemize}
@@
 \section{Research Goals}
-The overarching goal of my research was to improve the current state of REST API documentation process by automating the repeating manual steps in the aforementioned workflow.
+The overarching goal of my research is to improve the current state of the proces of REST API documentation process by automating the repeating manual steps in the aforementioned workflow.
 More specifically, I aimed to solve the following research goals:
@@
 REST API documentation tool developers can also use the requirements to build necessary features.
-  \item \textit{RG2: To develop and evaluate a technique to overcome the commonly observed but unresolved challenges related to REST API documentation.} I aimed to devise a reusable technique so that REST APIs can be documented following the identified REST API documentation requirements without following manual or bespoke methods.
-It has the following implications: REST API developers can utilize the technique to overcome the common challenges with the generation and maintenance of their API documentation.
-REST API client developers can access a usable API documentation.
+  \item \textit{RG2: To develop and evaluate a technique to overcome the commonly observed challenges related to REST API documentation.} I aimed to devise a reusable technique so that REST APIs can be documented following the identified REST API documentation requirements without following manual or bespoke methods.
+It has the following implications: REST API developers can utilize the technique to overcome the common challenges with the generation and maintenance of their API documentation, and REST API client developers can access an API documentation.
 \end{itemize}
@@
 \section{Research Objectives}
-Related to RG1 and RG2, I aimed to address the following research objectives:
+Related to RG1 and RG2, I aimed to address the following research questions as my objectives:
 
@@
 This objective helped me to identify the research opportunity to work on unresolved problems related to REST API documentation.
-  \item \textit{RO2: What are the requirements for usable REST API documentation?} This objective provides a necessary background work for improving the state of REST API documentation process.
+  \item \textit{RO2: What are the requirements for usable REST API documentation?} This objective provides a necessary background for improving the state of REST API documentation process.
   \item \textit{RO3: What technique can be used to document REST APIs to meet the requirements?} This is the central research objective.
 I researched to develop a new technique that can fill the gap between existing REST API documentation methods and the identified requirements.
-  \item \textit{RO4: Is the proposed technique feasible for REST API developers?} An evaluation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
-  \item \textit{RO5: How does a documentation using the proposed technique impact REST API client developers?} This question investigates the impact of the proposed technique from the perspective of the REST API client developers.
+  \item \textit{RO4: Is the proposed technique a practical one for REST API developers?} A qualitiative evaluation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
+  \item \textit{RO5: How does a documentation using the proposed technique impact REST API client developers?} This question investigates the impact of the proposed technique from the perspective of the REST API client developer effort and success rate while performing API tasks.
 \end{itemize}
@@
 \section{Cohesion and Unitary Report}
-To address the aforementioned research goals and objectives, my research strategy included five sub-projects, result of which have been published as papers in peer-reviewed international venues.
+To address the aforementioned research goals and objectives, my research strategy included five sub-projects, have resulted in publications at peer-reviewed international conferences.
 Table~\ref{table:chapter_summary} shows a high level overview of the flow of arguments between the sub-projects to meet the research goals.
@@
   Chapter~\ref{chapter:spy_rest} & RG1 and RG2 & RO2 and RO3 & SpyREST: Automated RESTful API documentation using an HTTP proxy server \cite{sohan2015spyrest}.
-& A list of requirements and the concept of using interception to generate REST API documentation to satisfy the requirements.
+& A list of requirements.
+A conceptual design using interception to generate REST API documentation to meet the requirements.
 \\
@@
   Chapter~\ref{chapter:demo_paper} & RG2 & RO3 & SpyREST in action: An automated RESTful API documentation tool \cite{sohan2015spyrest_tool}.
-& A reusable tool and a demonstration of the proposed REST API documentation technique.
-\\
+& A reusable tool that implements the conceptual design and shows how the identified REST API documentation requirements are met.\\
   \hline
   Chapter~\ref{chapter:cisco_study} & RG2 & RO4 & Automated example oriented REST API documentation at Cisco \cite{sohan_cisco}.
-& An empirical evidence that the proposed technique is feasible for practical use, and a list of problems for potential future research.\\
+& Empirical evidence that the proposed technique provides a practical documentation solution for REST API developers.\\
   \hline
   Chapter~\ref{chapter:controlled_study} & RG2 & RO5 & A study of the effectiveness of usage examples in REST API documentation \cite{sohan_vlhcc}.
-& An empirical evidence that REST API client developers can overcome common obstacles when the documentation contains usage examples, as suggested in the proposed technique.\\
+& Empirical evidence that the proposed technique allows REST API client developers to perform API tasks with less effort, higher success and satisfaction ratings.\\
   \hline
@@
 
-In Chapter~\ref{chapter:spy_rest}, I discuss my second research project where I derived a set of REST API documentation requirements and proposed a new technique to satisfy the requirements \cite{sohan2015spyrest}.
+In Chapter~\ref{chapter:spy_rest}, I discuss my second research project where I derived a set of REST API documentation requirements and proposed a new technique to meet the requirements \cite{sohan2015spyrest}.
 Here, I present a list of REST API documentation requirements by studying the literature and applying the findings from Chapter~\ref{chapter:case_study}.
 I discuss the available REST API documentation techniques and compare against the requirements.
-I present, SpyREST, a novel technique to intercept example REST API calls using an HTTP proxy server to collect and automatically synthesize API traffic to generate and update API documentation to satisfy the identified requirements.
+I present, \textit{SpyREST}, a novel technique to intercept example REST API calls using an HTTP proxy server to collect and automatically synthesize API traffic to generate and update API documentation to satisfy the identified requirements.
 The primary advantage of the presented technique is that the documentation is generated from executable code.
 As a result, a continuously updated REST API documentation with realistic usage examples is achievable with a reduced manual effort.
+The remaining chapters provide evidence that the proposed technique can satisfy the identified requirements.
 
 I present my third research project, related to RG2, in Chapter~\ref{chapter:demo_paper}, where I present a prototype implementation of the aforementioned technique \cite{sohan2015spyrest_tool}.
-In this prototype implementation, I discuss the process that transforms a number of example HTTP requests and responses into a usable REST API documentation.
+In this prototype implementation, I discuss the process that transforms a number of example HTTP requests and responses into usable REST API documentation.
 I also present an example where automated test code is used to generate and maintain a continuously updated REST API documentation using the developed tool.

### 2017-09-24 root.tex
add changes based on feedback from craig + mea

@@
 Most REST APIs are documented using a manual approach, which can be time consuming and error-prone.
-REST API developers need to efficiently document their APIs with the qualities that make the API usable.
+REST API developers need to efficiently document their APIs with qualities that make the API usable.
 In this research, I focused on the topic of automated REST API documentation to satisfy this need.
@@
 From this study, I observed that REST APIs evolve frequently, but used a manual or bespoke approach for generating and maintaining their documentation.
-I present a reusable technique with a prototype implementation to automate the REST API documentation process with executable API usage examples.
+I present, SpyREST, a reusable technique with a prototype implementation to automate the REST API documentation process with executable API usage examples.
 The technique involves the interception of example REST API calls using an HTTP proxy server to auto-generate an accurate and updated REST API documentation.
@@
 \begin{singlespace}
-I am grateful to Shahana, my wife, for supporting me and the family while I took the time away to work on this research.
+I am grateful to Shahana, my wife, for supporting me and the family while I took an unacceptable amount of the time and attention away from them to work on this research.
 

### 2017-09-24 introduction.tex
added changes based on feedback

@@
 APIs enable multiple software components to interact with each other.
-Web APIs allow software-to-software interconnectivity using web technologies as the transport mechanism of information between systems.
+Web APIs allow software-to-software interconnectivity using web technologies as the transport mechanism between systems.
 
@@
 He described the design of modern Web as an example of a REST-style architecture.
-REST-style architecture allows software components to satisfy the data transfer needs of the modern Internet-scale applications by levergading a generic interface based on URIs and Internet media types that allow intemediate processing such as caching to improve system performance.
+REST-style architecture allows software components to satisfy the data transfer needs of the modern Internet-scale applications by leveraging a generic interface based on URIs and Internet media types that allow intemediate processing such as caching to improve system performance.
 
@@
 For example, a REST API may support multiple Internet media types such as JSON, XML, CSV for a single application resource.
-Similarly, two diffenent REST APIs may use the same media type but contain entirely different formats to represent their objects.
+Similarly, two different REST APIs may use the same media type but contain entirely different formats to represent their objects.
 
@@
 \section{Research Goals}
-The overarching goal of my research is to improve the current state of the proces of REST API documentation process by automating the repeating manual steps in the aforementioned workflow.
+The overarching goal of my research is to improve the current state of the process of REST API documentation process by automating the repeating manual steps in the aforementioned workflow.
 More specifically, I aimed to solve the following research goals:
@@
 I researched to develop a new technique that can fill the gap between existing REST API documentation methods and the identified requirements.
-  \item \textit{RO4: Is the proposed technique a practical one for REST API developers?} A qualitiative evaluation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
+  \item \textit{RO4: Is the proposed technique a practical one for REST API developers?} A qualitative evaluation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
   \item \textit{RO5: How does a documentation using the proposed technique impact REST API client developers?} This question investigates the impact of the proposed technique from the perspective of the REST API client developer effort and success rate while performing API tasks.
@@
   \item SpyREST, a novel technique and a reusable tool that uses interception to automatically generate and maintain evolving REST API documentation with usage examples from executable code.
-  \item An industrial evaluation of SpyREST to demonstrate the feasibility of the proposed technique that practitioners can follow to improve their REST API documentation experience.
+  \item An industrial evaluation of SpyREST to demonstrate the feasibility of the proposed technique that REST API developers can follow to improve their REST API documentation experience.
   \item An empirical evidence showing the common obstacles faced by REST API client developers that can be reduced with usage examples to help the prioritization of REST API documentation efforts.

### 2017-09-24 root.tex
added changes based on feedback

@@
 Department of Computer Science \\
-University of Calgary \\
-Canada\\
+University of Calgary, Canada \\
 \signline
@@
 Supervisory Committee Member\\
-School of Eng. and Comp. Science \\
-Victoria University of Wellington\\
-New Zealand\\
+School of Engineering and Computer Science \\
+Victoria University of Wellington, New Zealand\\
+\signline
+Dr. Nils Daniel Forkert\\
+Internal Examiner\\
+Department of Radiology and Hotchkiss Brain Institute\\
+University of Calgary, Canada\\
+\signline
+Dr. Yves Lucet\\
+External Examiner\\
+Computer Science\\
+University of British Columbia (Okanagan), Canada\\
 \end{signing}
@@
 From this study, I observed that REST APIs evolve frequently, but used a manual or bespoke approach for generating and maintaining their documentation.
-I present, SpyREST, a reusable technique with a prototype implementation to automate the REST API documentation process with executable API usage examples.
+I present, \textit{SpyREST}, a reusable technique with a prototype implementation to automate the REST API documentation process with executable API usage examples.
 The technique involves the interception of example REST API calls using an HTTP proxy server to auto-generate an accurate and updated REST API documentation.

### 2017-11-02 cisco_study/cisco_study.tex
footnotes added + removed the dashes

@@
-\chapter{Automated Example Oriented REST API Documentation at Cisco}
+\chapter{\texorpdfstring{Automated Example Oriented REST API Documentation at Cisco \protect\footnote{S.Sohan, C.Anslow, and F.Maurer,``Automated example oriented REST API documentation at cisco,'' in Proc.
+of IEEE/ACM International Conference on Software Engineering, 2017, pp.
+213-222.
+}}{Automated Example Oriented REST API Documentation at Cisco}}
 \label{chapter:cisco_study}

### 2017-11-02 demo_paper/demo_paper.tex
all chapters added
more spell checks
footnotes added + removed the dashes

@@
-%\documentclass[conference]{IEEEtran}
-
-%\usepackage{cite}
-%\usepackage{listings}
-%\usepackage{amsmath}
-%\lstset{showstringspaces=false, basicstyle=\footnotesize, numbers=left, numbersep=1pt }
-%\usepackage[tableposition=top]{caption}
-%\ifCLASSINFOpdf
-%\usepackage[pdftex]{graphicx}
-%\usepackage[pdftex]{color}
-%\usepackage{subcaption}
-%\else
-%\fi
-%\usepackage{url}
-%\usepackage{upquote}
-%\usepackage{mdframed}
-%\usepackage{balance}
-%\hyphenation{op-tical net-works semi-conduc-tor}
-
-\chapter{SpyREST in Action: An Automated RESTful API Documentation Tool}
-
-%\author{\IEEEauthorblockN{S M Sohan, Craig Anslow, and Frank Maurer}
-%\IEEEauthorblockA{Department of Computer Science\\
-%University of Calgary\\
-%Calgary, Alberta T2N 1N4\\
-%Email: \{smsohan, craig.anslow, frank.maurer\}@ucalgary.ca  }
-%}
-%\maketitle
+\chapter{\texorpdfstring{SpyREST in Action: An Automated RESTful API Documentation Tool \protect\footnote{S.
+Sohan, C.
+Anslow, and F.
+Maurer, ``Spyrest in action: An automated RESTful API documentation tool,'' in Proc.
+of IEEE/ACM International Conference on Automated Software Engineering, 2015, pp.
+813-818.}}{SpyREST in Action: An Automated RESTful API Documentation Tool}}
+\label{chapter:demo_paper}
 
@@
 RESTful API developers can use SpyREST to automatically generate and maintain updated API documentation.
-
-%\begin{IEEEkeywords}
-%RESTful API, Web API, Documentation, Automation, Example based documentation
-%\end{IEEEkeywords}
-
-%\IEEEpeerreviewmaketitle
 

### 2017-11-02 spy_rest/spy_rest.tex
all chapters added
first reasonably ok pass
more spell checks
footnotes added + removed the dashes

@@
-\chapter{SpyREST: Automated RESTful API Documentation using an HTTP Proxy Server}
-
-%\author{\IEEEauthorblockN{S M Sohan, Craig Anslow, and Frank Maurer}
-%\IEEEauthorblockA{Department of Computer Science\\
-%University of Calgary\\
-%Calgary, Alberta T2N 1N4\\
-%Email: \{smsohan, craig.anslow, frank.maurer\}@ucalgary.ca  }
-%}
-%\maketitle
+\chapter{\texorpdfstring{SpyREST: Automated RESTful API Documentation using an HTTP Proxy Server \protect\footnote{S.
+Sohan, C.
+Anslow, and F.
+Maurer, Spyrest: ``Automated RESTful API documentation using an HTTP proxy server,'' in Proc.
+of IEEE/ACM International Conference on Automated Software Engineering, 2015, pp.
+271-276.}}{SpyREST: Automated RESTful API Documentation using an HTTP Proxy Server}}
+\label{chapter:spy_rest}
 
@@
 Using the proposed HTTP proxy server based technique, RESTful API developers can significantly reduce the cost of producing and maintaining API documentation by replacing a large manual process with an automated process.
-
-%\IEEEpeerreviewmaketitle
-
-%\begin{IEEEkeywords}
-%RESTful API, Web API, Documentation, Automation, Example based documentation
-%\end{IEEEkeywords}
 
@@
 presented a vocabulary for documenting RESTful Web APIs called Web Interface Language (WIfL) \cite{Danielsen_validation}.
-Verborgh et at.
+Verborgh et al.
 presented RESTdesc, a Resource oriented and Hyper-link based specification for describing RESTful APIs \cite{Verborgh_functional}.
@@
 presented RDDL, an XML based specification for describing RESTful APIs \cite{Mangler_rddl}.
-Kopecky´ et al.
+Kopecky et al.
 presented hRESTS, a machine readable micro-format to describe RESTful APIs that use an alternate representation compared to WIfL \cite{Kopecky_hrests}.

### 2017-11-03 introduction.tex
updates the intro to add the images for local and REST API doc

@@
 
-While there are overlaps between the documentation requirements of local and REST APIs, there are significant requirements that are unique to each.
-The documentation of most local APIs contain a structural definition of the API classes and methods.
-API usage examples for local APIs are documented using the same programming language of the API.
-On the other hand, REST APIs are defined using HTTP specific parameters such as API endpoints, query parameters, HTTP headers and body.
-REST API clients can be written in any programming language agnostic to the API implementation language.
-Due to such structural and implementation differences between local and REST APIs, the existing technique and tools for local API documentation cannot be readily used to generate REST API documentation.
+While there are overlaps between the documentation requirements of local and REST APIs, there are significant requirements that are unique to each. The differences between the documentation
+of local vs. REST APIs can be depicted by an example as follows:
+\newcommand*{\captionsource}[2]{%
+  \caption[{#1}]{%
+    \textbf{Source:} #2%
+    \\\hspace{\linewidth}%
+    #1%
+  }%
+}
+\begin{figure}[htb]
+  \includegraphics[width=\linewidth]{javadoc.png}
+  \captionsource{Example Local API Documentation from Formatted Comments Using JavaDoc}{Log4J}
+  \label{fig:javadoc}
+\end{figure}
+
+Figure~\ref{fig:javadoc} shows an example of how JavaDoc is used to automatically extract formatted comments from the API source code to
+generate human readable documentation. Here the developers of the \texttt{debug} API method provided comments describing the intended way to use the method. JavaDoc can
+automatically parse such comments to render the API documentation. Since the comments are not executable code, the API developers need to ensure the correctness of the
+comments, and maintain it manually when the API evolves.
+
+\begin{figure}[htb]
+  \includegraphics[width=\linewidth]{wordpress_doc.png}
+  \captionsource{Difference between a REST API Code and its Documentation}{WordPress REST API}
+  \label{fig:wordpress_doc}
+\end{figure}
+
+Figure~\ref{fig:wordpress_doc} shows an example of a REST API source code and a fragment of its API documentation
+taken from WordPress REST API. The developers of the \texttt{get\_items} API provided formatted comments to describe the intended use of this method similar to the example
+in Figure~\ref{fig:javadoc}. However, there is a stark difference between the API source code and its documentation. This is because even though the API is implemented
+using a specific language, PHP in this case, the documentation is agnostic to this choice of technology because REST APIs are described by their HTTP interface elements such as
+HTTP headers, URLs, and body. As shown in this example, using existing local API documentaiton tools such as JavaDoc that converts formatted comments into API documentation would require extensive
+manuall effort for REST APIs for its verbosity and difference between the source code and the documentation constructs.
+
+The documentation of most local APIs
+contain a structural definition of the API classes and methods. API usage examples for local APIs are documented using the same programming language of the API. On the other hand, REST APIs
+are defined using HTTP specific parameters such as API endpoints, query parameters, HTTP headers and body. REST API clients can be written in any programming language agnostic to the API
+implementation language. Due to such structural and implementation differences between local and REST APIs, the existing technique and tools for local API documentation cannot be readily
+used to generate REST API documentation.
 

### 2017-11-03 case_study/case_study.tex
footnotes added + removed the dashes
adds changes based on fixes from Nils

@@
 
-\chapter{A Case Study of Web API Evolution}
+\chapter{\texorpdfstring{A Case Study of Web API Evolution \protect\footnote{S.
+Sohan, C.
+Anslow, and F.
+Maurer, ``A case study of Web API evolution,'' in Proc.
+of IEEE World Congress on Services (SERVICES).
+IEEE, 2015, pp.
+245-252.}}{S.
+Sohan, C.
+Anslow, and F.
+Maurer, ``A case study of Web API evolution,'' in Proc.
+of IEEE World Congress on Services (SERVICES).
+IEEE, 2015, pp.
+245–252.}}
 \label{chapter:case_study}
@@
 In this paper we conducted a case study of evolving Web APIs to investigate what changes are made between versions and how the changes are documented and communicated to the API users.
-The findings are a list of recommendations for practitioners and researchers based on API change profiles, versioning, documentation and communication approaches that are observed in practice.
+The findings are a list of recommendations for practitioners and researchers based on API change profiles, versioning, documentation, and communication approaches that are observed in practice.
 This study will help inform developers of evolving Web APIs to make decision about versioning, documentation and communication methods.
@@
 For example, in the case of the aforementioned business search portal, if the Google Maps API changes, it may also require the portal to change.
-This introduces unique challenges to the evolution of Web APIs compared because the evolution is out of the control of the API users.
+This introduces unique challenges to the evolution of Web APIs because the evolution is out of the control of the API users.
 For example, API users of a local API can use a copy of an older version, but for evolving Web APIs, API users are forced to upgrade if older versions are no longer supported.
@@
 Wordpress REST API is developed as a plugin for Wordpress.
-As a result, a new version of the Web API is released with a new version of the plugin and it all API clients are affected when a new version is deployed.
+As a result, a new version of the Web API is released with a new version of the plugin and all API clients are affected when a new version is deployed.
 The source code for OpenStreetMap.org Web API shows the use of configuration files for API Versioning.
@@
 We also investigated how the API documentation is generated.
-We found that, Web API documents are manually generated and updated, unlike the documentation for local APIs which are are typically auto-generated.
+We found that, Web API documents are manually generated and updated, unlike the documentation for local APIs which are typically auto-generated.
 For WordPress and OpenStreetMap, this is apparent from their source code, as we found the use of manually edited wiki and Markdown files to generate the documentation\footnote{\url{http://en.wikipedia.org/wiki/Markdown}}.

### 2017-11-03 cisco_study/cisco_study.tex
adds changes based on fixes from Nils

@@
 Researchers identified the documentation of APIs as both the primary source of information as well as the key obstacle for API usability \cite{Robillard_what_makes}.
-To this regard, researchers have identified the qualities of ``good API documentation'' as follows: complete, correct, includes thorough explanations and code examples, provides consistent presentation and organization \cite{Robillard_what_makes,Myers_study}.
+In this regard, researchers have identified the qualities of ``good API documentation'' as follows: complete, correct, includes thorough explanations and code examples, provides consistent presentation and organization \cite{Robillard_what_makes,Myers_study}.
 In our previous work, we introduced a novel technique and SpyREST, an implementation, based on an HTTP proxy server to automatically intercept example REST API calls and synthesize the data to produce REST API documentation to meet the aforementioned qualities \cite{DBLP:conf/kbse/SohanAM15}.
@@
 The REST API is documented using SpyREST via automated tests.
-To evaluate the outcome of the documentation process using SpyREST, we've incorporated stakeholder feedback as well as collected metrics from the various log files associated with the development and deployment of API docs and the actual API changes.
+To evaluate the outcome of the documentation process using SpyREST, we have incorporated stakeholder feedback as well as collected metrics from the various log files associated with the development and deployment of API docs and the actual API changes.
 25 API stakeholders from Cisco were invited via email to provide free form feedback about the API documentation and its underlying process via emails to the first author of this paper.
@@
 The version numbers v0 and v1 were left unchanged to release backward compatible new features and bug fixes.
-At present, the API serves 80,000 API calls on an average per day (as per the API call logs) and is  used by over a hundred enterprise customers that are globally distributed.
+At present, the API serves 80,000 API calls on average per day (as per the API call logs) and is  used by over a hundred enterprise customers that are globally distributed.
 The customers use the SpyREST generated documentation as the sole information source for learning the API features.
@@
 the $API$ class has capabilities to make the $patch$ HTTP request over an HTTP Proxy server.
-While using the proxy server, the $API$ class also sends the test description ``Moves computer to a group with given $connector_guid$ and $group_guid$'' with a SpyREST specific header to automatically map test description as the description of the API usage example.
+While using the proxy server, the $API$ class also sends the test description ``Moves computer to a group with given \texttt{connector\_guid} and \texttt{group\_guid}'' with a SpyREST specific header to automatically map test description as the description of the API usage example.
 As discussed earlier, the SpyREST analyzers can inspect the HTTP request and response information with the custom headers to produce usable REST API documentation from this automated test code.
@@
 \subsection{Maintaining Custom Content}
-Based on the feedback on SpyREST by Cisco developers, we've learned both the benefits and drawbacks of using a wiki-like editor for custom content with the API documentation.
+Based on the feedback on SpyREST by Cisco developers, we have learned both the benefits and drawbacks of using a wiki-like editor for custom content with the API documentation.
 The wiki within SpyREST allows the API developers to add rich content comprising of proses, images, tables, code fragments, and web links, etc.
 to explain concepts that are required to understand the API.
-One developer feedback (D1) mentioned the following about the wiki: \textbf{``Good customization on the overview of each resource.''} Developer D3 mentioned the following: \textbf{``Being able to edit the documentation in place, similar to a wiki works really well, effectively allowing an organization to crowd-source the job.''} For example, at Cisco, we've added an overview explaining API access information, rate-limit, and common approaches to perform pagination within the API actions as custom content.
+One developer feedback (D1) mentioned the following about the wiki: \textbf{``Good customization on the overview of each resource.''} Developer D3 mentioned the following: \textbf{``Being able to edit the documentation in place, similar to a wiki works really well, effectively allowing an organization to crowd-source the job.''} For example, at Cisco, we have added an overview explaining API access information, rate-limit, and common approaches to perform pagination within the API actions as custom content.
 
@@
 While it provides us with a unique insights, it also introduces a confirmation bias.
-Our results may include a selection bias since the results are based on a single single set of REST APIs.
+Our results may include a selection bias since the results are based on a single set of REST APIs.
 While the aforementioned feedback shows that REST API documentation process is effective for us, we have a lack understanding of the impact of the documentation on the API client developers.

### 2017-11-03 demo_paper/demo_paper.tex
adds changes based on fixes from Nils

@@
 In this paper we present SpyREST which provides the necessary tool support to address the unique needs for RESTful API documentation.
-To this regard, in the next section we have discussed the design goals followed by a HTTP proxy server based prototype implementation to demonstrate an automated RESTful API documentation tool.
+In this regard, in the next section we discuss the design goals followed by a HTTP proxy server based prototype implementation to demonstrate an automated RESTful API documentation tool.
 We then present two use cases to show how API developers can use SpyREST to automatically generate and maintain RESTful API documentation.
@@
 
-The technology choice behind for our implementation of SpyREST components is based on our past experience of using these tools but the concepts behind SpyREST is not dependent on the aforementioned technology choice.
+The technology choice behind our implementation of SpyREST components is based on our past experience of using these tools but the concepts behind SpyREST is not dependent on the aforementioned technology choice.
 As a proof of concept, SpyREST only analyzes JSON based API request and response payloads.

### 2017-11-03 introduction.tex
adds changes based on fixes from Nils

@@
 The overarching goal of my research is to improve the current state of the process of REST API documentation process by automating the repeating manual steps in the aforementioned workflow.
-More specifically, I aimed to solve the following research goals:
+More specifically, I aim to solve the following research goals:
 
@@
   \item \textit{RG1: To identify a set of REST API documentation requirements.} Existing research on API usability has derived a number of API documentation requirements primarily focusing on local APIs.
-Due to the structural and implementation differences between local and REST APIs, I aimed to identity REST API specific documentation requirements in my research.
+Due to the structural and implementation differences between local and REST APIs, I aim to identity REST API specific documentation requirements in my research.
 This has several implications as follows: having access to a scientifically derived list of REST API documentation requirements allows REST API developers to make an informed decision related to planning and execution of their documentation efforts.
@@
 REST API documentation tool developers can also use the requirements to build necessary features.
-  \item \textit{RG2: To develop and evaluate a technique to overcome the commonly observed challenges related to REST API documentation.} I aimed to devise a reusable technique so that REST APIs can be documented following the identified REST API documentation requirements without following manual or bespoke methods.
+  \item \textit{RG2: To develop and evaluate a technique to overcome the commonly observed challenges related to REST API documentation.} I aim to devise a reusable technique so that REST APIs can be documented following the identified REST API documentation requirements without following manual or bespoke methods.
 It has the following implications: REST API developers can utilize the technique to overcome the common challenges with the generation and maintenance of their API documentation, and REST API client developers can access an API documentation.
@@
 \section{Research Objectives}
-Related to RG1 and RG2, I aimed to address the following research questions as my objectives:
+Related to RG1 and RG2, I aim to address the following research questions as my objectives:
 
 \begin{itemize}
-  \item \textit{RO1: How are evolving REST APIs documented?} This question is aimed to analyze the gap between current industry practice around the documentation of evolving REST APIs and the proposed solutions from existing literature.
+  \item \textit{RO1: How are evolving REST APIs documented?} This question is aimed to analyze the gap between current industry practice around the documentation of evolving REST APIs and the propose solutions from existing literature.
 This objective helped me to identify the research opportunity to work on unresolved problems related to REST API documentation.

### 2017-11-03 root.tex
footnotes added + removed the dashes
adds changes based on fixes from Nils

@@
 \usepackage{enumitem}
-
 \title{Automated Example Oriented REST API Documentation}
@@
 
-In this thesis, I present and evaluate of a novel technique to solve REST API documentation requirements.
+In this thesis, I present and evaluate a novel technique to solve REST API documentation requirements.
 I present a set of REST API documentation requirements by studying the existing literature and the current industry practice.
@@
 I present, \textit{SpyREST}, a reusable technique with a prototype implementation to automate the REST API documentation process with executable API usage examples.
-The technique involves the interception of example REST API calls using an HTTP proxy server to auto-generate an accurate and updated REST API documentation.
+The technique involves the interception of example REST API calls from API test code using an HTTP proxy server to auto-generate an accurate and updated REST API documentation.
 I present an industrial evaluation of the proposed technique based on a period of eighteen months of production use.
@@
 
-The findings of this research can be used as a guideline by practitioners to automatically generate REST API documentation with usage examples.
+The findings of this research can be used as a guideline for practitioners to automatically generate REST API documentation with usage examples.
 Researchers can extend the novel concept of using interception to automatically document API usage examples for other forms of APIs.

### 2017-11-03 spy_rest/spy_rest.tex
adds changes based on fixes from Nils

@@
 
-  \textbf{R7 - Collaborative}: SpyREST needs to allow people to collaborate on the API documentation so that questions and answers about APIs can coexist with the API documentation to to overcome the knowledge fragmentation \cite{Chen_who_asked}.
+  \textbf{R7 - Collaborative}: SpyREST needs to allow people to collaborate on the API documentation so that questions and answers about APIs can coexist with the API documentation to overcome the knowledge fragmentation \cite{Chen_who_asked}.
 
@@
 In this section we discuss a case study of using SpyREST to auto generate documentation for 25 API actions randomly sampled from three RESTful API providers.
-These 25 API APIs are documented using 272 lines of code\footnote{\url{https://github.com/smsohan/spyrest_examples}}.
+These 25 APIs are documented using 272 lines of code\footnote{\url{https://github.com/smsohan/spyrest_examples}}.
 These three API providers are: GitHub.com, KISSMetrics.com (Online analytics tool), and LiquidPlanner.com (Online project management tool).

### 2017-11-03 introduction.tex
removes the steps with numbers

@@
 Conceptually, the high level steps involved in the manual process can be identified as shown in Figure \ref{fig:manual}.
-These steps S1-S7 can be described as follows:
+These steps 1-7 can be described as follows:
 
-\begin{itemize}
-  \item S1 - API developer prepares an example API with the required HTTP parameters and request body.
-  \item S2 - API developer uses a REST API client to make the example API call.
-  \item S3 - the API developer records the HTTP data.
-  \item S4 - API developer then edits the recorded data so that only relevant content is selected for documentation.
-  \item S5 - API developers use any custom content to describe the API example.
-  \item S6 - API developers combine the custom content with the edited content from the HTTP traffic into HTML to publish on the web.
-  \item S7 - API developer adds custom overview information to explain API concepts and general rules, and publishes the final documentation so other developers can learn the API.
-\end{itemize}
+\begin{enumerate}
+  \item API developer prepares an example API with the required HTTP parameters and request body.
+  \item API developer uses a REST API client to make the example API call.
+  \item the API developer records the HTTP data.
+  \item API developer then edits the recorded data so that only relevant content is selected for documentation.
+  \item API developers use any custom content to describe the API example.
+  \item API developers combine the custom content with the edited content from the HTTP traffic into HTML to publish on the web.
+  \item API developer adds custom overview information to explain API concepts and general rules, and publishes the final documentation so other developers can learn the API.
+\end{enumerate}
 
-The steps S1-S6 are repeated for every API action that needs to be documented, and S1-S7 need to repeated for every version of the REST API as it evolves.
+The steps 1-6 are repeated for every API action that needs to be documented, and steps 1-7 need to repeated for every version of the REST API as it evolves.
 As depicted here, this manual process can be both time consuming and error-prone.

### 2017-11-03 controlled_study/controlled_study.tex
more spell checks
adds ethics certificate

@@
-%\documentclass[conference]{IEEEtran}
-
-%\ifCLASSINFOpdf
-  %\usepackage{listings}
-  %\usepackage[pdftex]{graphicx}
-  %\usepackage{url}
-  %\usepackage{subcaption}
-  %\usepackage{tikz,pgfplots,pgfplotstable}
-  %\usepackage{mathtools}
-  %\usepgfplotslibrary{statistics}
-  %\usepackage{courier}
-%\else
-%\fi
-
-%\makeatletter
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
-%\def\ps@IEEEtitlepagestyle{%
-  %\def\@oddfoot{\mycopyrightnotice}%
-  %\def\@evenfoot{}%
-%}
-%\def\mycopyrightnotice{%
-  %{\footnotesize 978-1-5386-0443-4/17/\$31.00 ©2017 IEEE\hfill}% <--- Change here
-  %\gdef\mycopyrightnotice{}% just in case
-%}
-%\begin{document}
-%\title{A Study of the Effectiveness of Usage Examples in REST API Documentation}
-\chapter{A Study of the Effectiveness of Usage Examples in REST API Documentation}
+\chapter{\texorpdfstring{A Study of the Effectiveness of Usage Examples in REST API Documentation \protect\footnote{S.
+Sohan, C.
+Anslow, F.
+Maurer, and M.
+Robillard, ``A study of the effectiveness of usage examples in REST API documentation,'' in to appear in Proc.
+of IEEE Symposium on Visual Languages and Human-Centric Computing, 2017.}}{A Study of the Effectiveness of Usage Examples in REST API Documentation}}
 \label{chapter:controlled_study}
-
- %\author{\IEEEauthorblockN{S M Sohan, Frank Maurer\IEEEauthorrefmark{1}}
- %\IEEEauthorblockA{\IEEEauthorrefmark{1}
- %Dept. of Computer Science\\
- %University of Calgary\\
- %Canada\\
- %\{smsohan, frank.maurer\}@ucalgary.ca}
- %\and
- %\IEEEauthorblockN{Craig Anslow}
- %\IEEEauthorblockA{School of Eng. and Computer Science\\
- %Victoria University of Wellington\\
- %New Zealand\\
- %craig@ecs.vuw.ac.nz}
- %\and
- %\IEEEauthorblockN{Martin P. Robillard}
- %\IEEEauthorblockA{
- %School of Computer Science\\
- %McGill University\\
- %Canada\\
- %martin@cs.mcgill.ca}
- %}
-
-%\maketitle
 
@@
 By following the REST API documentation suggestions from this paper, REST API developers can reduce the errors, improve success rate and satisfaction of API client developers.
-
-%\begin{IEEEkeywords}
-%API; REST; Documentation; Usage Examples; Empirical Study; Controlled Study; Productivity;
-
-%\end{IEEEkeywords}
-
-%\IEEEpeerreviewmaketitle
 
@@
 
-To answer the aforementioned research question, the study has the following requirements:
+To answer the aforementioned research question, the study has the following requirements (Ethics Certificate in Appendix~\ref{ethics}):
 
@@
 Participant Selection.} Developers with prior experience on REST APIs need to be recruited as study participants to perform the study within a limited amount of time and in a realistic setup.
-Furthermore, to reduce a learning bias, only participants with no prior experince of using the WordPress REST API V2 are accepted for this study.
+Furthermore, to reduce a learning bias, only participants with no prior experience of using the WordPress REST API V2 are accepted for this study.
 \end{itemize}

### 2017-11-03 root.tex
added extras for future work

@@
 Future research is also needed to investigate techniques for automatically cross-referencing API elements to allow API client developers with an improved navigation experience while reading the REST API documentation.
+In Chapter~\ref{chapter:controlled_study}, I discussed that the study participants faced obstacles with performing API tasks that required the use of multiple related API calls to reach a desired outcome.
+Future research needs to investigate techniques that can be used to provide links to related API elements and usage examples to allow API client developers with an alternative navigation experience.
   \item \textbf{Automated Changelog Generation.} For frequently evolving APIs, future research needs to extend the proposed technique in this thesis to automatically compute human readable changelogs and detect version compatibility to follow a semantic versioning scheme.
 This would help API client developers with an accurate information about the potential impact of an API update on their own application.
+In Chapter~\ref{chapter:case_study}, I discussed that REST APIs evolve frequently, often several times per month, and the change profiles contain different types of changes.
+I also provided a list of change patterns that I observed while studying the change profiles.
+In Chapter~\ref{chapter:cisco_study}, I discussed that manual effort was required to maintain the API changelog for the studied API.
+Future research needs to investigate automated changelog generation to help the evolution of REST APIs.
+The outcome of an automated changelog generation can be applied to automatically infer a semantic version of a REST API based on the inferred compatibility information.
   \item \textbf{Automated API Client Code Generation.} Several tools exist that can convert custom REST API description files to API client code in multiple programming languages.
@@
 The structure of REST APIs are commonly documented using programming language features and formatted comments.
+Using formatted coments to provide API documentation with executable code examples requires manual effort since no automated feedback is available about the accuracy of the comments.
+API developers commonly write and maintain API test code that shows the intended way of using the APIs.
+The API test code provides automated feedback by means of test results.
+However, API test code may not be a substitute for API documentation.
+API test code may contain complex setup, tear-down, and verification logic that are not useful for documentation.
+Also, for closed source applications, organizations may not be able to publish their source code to external API client developers.
+The concept of intereception as described in Chapter~\ref{chapter:spy_rest} can be applied to other forms of API to generate and maintain API documenation by intercepting REST API tests.
+For example, several programming languages support concepts such as Aspect Oriented Programming, where custom code can be written to intercept API method calls, that can be used for API documentation.
 Future research can investigate the feasibility of leveraging the interception technique to source realistic usage examples for local stateless APIs from existing API test code.
+
 \end{itemize}

### 2017-11-08 introduction.tex
final version

@@
 
+Application Programming Interfaces, commonly known as APIs, are used to express a software component in terms of its operations, inputs, outputs, and their types\footnote{\url{https://en.wikipedia.org/wiki/Application_programming_interface}}.
 Robillard defines an API as follows: An API is the interface to implement functionality that developers can access to perform various tasks \cite{Robillard_a_field_study, Robillard_what_makes}.
-Application Programming Interfaces, commonly known as APIs, are used to express a software component in terms of its operations, inputs, outputs, and their types\footnote{\url{https://en.wikipedia.org/wiki/Application_programming_interface}}.
 APIs enable multiple software components to interact with each other.
@@
 Fielding defined Representational State Transfer or REST as an architectural style for developing distributed hypermedia systems \cite{Fielding_rest}.
-He described the design of modern Web as an example of a REST-style architecture.
+Fielding described the design of the modern Web as an example of a REST-style architecture.
 REST-style architecture allows software components to satisfy the data transfer needs of the modern Internet-scale applications by leveraging a generic interface based on URIs and Internet media types that allow intemediate processing such as caching to improve system performance.
 
+The overarching goal of my research is to improve the current state of the REST API documentation process by automating the repeating manual steps.
 REST APIs allow access to application resources over HTTP.
@@
 To this regard, researchers have identified the qualities of ``good API documentation'' as follows: complete, correct, includes thorough explanations and code examples, provides consistent presentation and organization \cite{Robillard_what_makes,Myers_study}.
-Today, there are several tools such as Javadoc\footnote{\url{http://www.oracle.com/technetwork/java/javase/documentation/index-jsp-135444.html}}, UsETeC \cite{zhu2014mining}, Jadeite \cite{Jadeite}, APIMiner \cite{montandon2013documenting}, Roast \cite{Hoffman_api_documentation} that can be used to document local APIs with the aforementioned qualities .
+Today, there are several tools such as Javadoc\footnote{\url{http://www.oracle.com/technetwork/java/javase/documentation/index-jsp-135444.html}}, UsETeC \cite{zhu2014mining}, Jadeite \cite{Jadeite}, APIMiner \cite{montandon2013documenting}, Roast \cite{Hoffman_api_documentation} that can be used to document local APIs with the aforementioned qualities.
 
@@
 Currently, the process of documenting REST APIs is largely manual.
+This is problematic because, it is time a consuming and an error-prone process.
 Conceptually, the high level steps involved in the manual process can be identified as shown in Figure \ref{fig:manual}.
@@
 \section{Research Goals}
-The overarching goal of my research is to improve the current state of the process of REST API documentation process by automating the repeating manual steps in the aforementioned workflow.
+The principal goal of my research is to improve the REST API documentation process by automating the repeating manual steps in the aforementioned workflow.
 More specifically, I aim to solve the following research goals:
@@
 REST API documentation tool developers can also use the requirements to build necessary features.
-  \item \textit{RG2: To develop and evaluate a technique to overcome the commonly observed challenges related to REST API documentation.} I aim to devise a reusable technique so that REST APIs can be documented following the identified REST API documentation requirements without following manual or bespoke methods.
+  \item \textit{RG2: To develop and evaluate a documentation technique to overcome the commonly observed challenges related to REST API documentation.} I aim to devise a reusable technique so that REST APIs can be documented following the identified REST API documentation requirements without following manual or bespoke methods.
 It has the following implications: REST API developers can utilize the technique to overcome the common challenges with the generation and maintenance of their API documentation, and REST API client developers can access an API documentation.
+
 \end{itemize}
 
-\section{Research Objectives}
+\section{Research Questions}
 Related to RG1 and RG2, I aim to address the following research questions as my objectives:
@@
 \begin{itemize}
-  \item \textit{RO1: How are evolving REST APIs documented?} This question is aimed to analyze the gap between current industry practice around the documentation of evolving REST APIs and the propose solutions from existing literature.
+  \item \textit{RQ1: How are evolving REST APIs documented?} This question is aimed to analyze the gap between current industry practice around the documentation of evolving REST APIs and the propose solutions from existing literature.
 This objective helped me to identify the research opportunity to work on unresolved problems related to REST API documentation.
-  \item \textit{RO2: What are the requirements for usable REST API documentation?} This objective provides a necessary background for improving the state of REST API documentation process.
-  \item \textit{RO3: What technique can be used to document REST APIs to meet the requirements?} This is the central research objective.
+  \item \textit{RQ2: What are the requirements for usable REST API documentation?} Answer of this question provides a necessary background for improving the state of REST API documentation process.
+  \item \textit{RQ3: What technique can be used to document REST APIs to meet the requirements?} This is the central research question.
 I researched to develop a new technique that can fill the gap between existing REST API documentation methods and the identified requirements.
-  \item \textit{RO4: Is the proposed technique a practical one for REST API developers?} A qualitative evaluation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
-  \item \textit{RO5: How does a documentation using the proposed technique impact REST API client developers?} This question investigates the impact of the proposed technique from the perspective of the REST API client developer effort and success rate while performing API tasks.
+  \item \textit{RQ4: Is the proposed technique practical for REST API developers?} A qualitative evaluation of the proposed technique in a real-world setup is needed to understand the feasibility and the shortcomings of the proposed technique from the perspective of the REST API developers.
+  \item \textit{RQ5: How does documentation using the proposed technique impact REST API client developers?} This question investigates the impact of the proposed technique from the perspective of the REST API client developer effort and success rate while performing API tasks.
 \end{itemize}
@@
 \section{Cohesion and Unitary Report}
-To address the aforementioned research goals and objectives, my research strategy included five sub-projects, have resulted in publications at peer-reviewed international conferences.
-Table~\ref{table:chapter_summary} shows a high level overview of the flow of arguments between the sub-projects to meet the research goals.
+To address the aforementioned research goals and objectives, my research strategy included five sub-projects, which have resulted in publications at peer-reviewed international conferences.
+Table~\ref{table:chapter_summary} shows a high level overview of the flow of arguments between the sub-projects to meet the research goals and questions.
 
@@
   \hline
-  Chapter & Goal & Objective & Research Project & Primary Outcome \\
+  Chapter & Goal & Question& Research Project & Primary Outcome \\
   \hline
-  Chapter~\ref{chapter:case_study}  & RG1 & RO1 & A case study of Web API evolution \cite{sohan2015case}.
+  Chapter~\ref{chapter:case_study}  & RG1 & RQ1 & A case study of Web API evolution \cite{sohan2015case}.
 & A set of challenging problems related to evolving Web APIs.
@@
   \hline
-  Chapter~\ref{chapter:spy_rest} & RG1 and RG2 & RO2 and RO3 & SpyREST: Automated RESTful API documentation using an HTTP proxy server \cite{sohan2015spyrest}.
+  Chapter~\ref{chapter:spy_rest} & RG1 and RG2 & RQ2 and RQ3 & SpyREST: Automated RESTful API documentation using an HTTP proxy server \cite{sohan2015spyrest}.
 & A list of requirements.
@@
   \hline
-  Chapter~\ref{chapter:demo_paper} & RG2 & RO3 & SpyREST in action: An automated RESTful API documentation tool \cite{sohan2015spyrest_tool}.
+  Chapter~\ref{chapter:demo_paper} & RG2 & RQ3 & SpyREST in action: An automated RESTful API documentation tool \cite{sohan2015spyrest_tool}.
 & A reusable tool that implements the conceptual design and shows how the identified REST API documentation requirements are met.\\
   \hline
-  Chapter~\ref{chapter:cisco_study} & RG2 & RO4 & Automated example oriented REST API documentation at Cisco \cite{sohan_cisco}.
+  Chapter~\ref{chapter:cisco_study} & RG2 & RQ4 & Automated example oriented REST API documentation at Cisco \cite{sohan_cisco}.
 & Empirical evidence that the proposed technique provides a practical documentation solution for REST API developers.\\
   \hline
-  Chapter~\ref{chapter:controlled_study} & RG2 & RO5 & A study of the effectiveness of usage examples in REST API documentation \cite{sohan_vlhcc}.
+  Chapter~\ref{chapter:controlled_study} & RG2 & RQ5 & A study of the effectiveness of usage examples in REST API documentation \cite{sohan_vlhcc}.
 & Empirical evidence that the proposed technique allows REST API client developers to perform API tasks with less effort, higher success and satisfaction ratings.\\
@@
 I discuss my fourth research project, related to RG2, involving an industrial evaluation of the proposed technique and the tool in Chapter~\ref{chapter:cisco_study} \cite{sohan_cisco}.
-The evaluation is carried out based on the data collected from an eighteen month of production use of SpyREST at Cisco, my current employer.
-From the study I found that it is feasible to use the proposed interception technique and automated tests to document evolving REST APIs satisfying the previously presented REST API documentation requirements.
+The evaluation is carried out based on the data collected from an 18 month production use of SpyREST at Cisco, my current employer.
+From the study I found that it is effective to use the proposed interception technique and automated tests to document evolving REST APIs satisfying the previously presented REST API documentation requirements.
 Automated always-updated REST API documentation was found to help establish a quick feedback cycle among the stakeholders.
@@
 
-In Chapter~\ref{chapter:conclusion}, I conclude by discussing how the research objectives are achieved and provide a list of problems for future research.
+In Chapter~\ref{chapter:conclusion}, I conclude by discussing how the research objectives are achieved and provide a list of potential future research problems.
 
@@
 
+\newpage
 \bibliographystyle{IEEEtran}

### 2017-11-08 root.tex
adds a line about client library
final version

@@
 \label{chapter:conclusion}
+In this thesis, I focused on improving the state of REST API documentation process.
+Existing REST API documentation processes rely on largely manual or bespoke approaches that are time consuming and error-prone, specially for evolving REST APIs.
+Through my research, I aimed to bridge the gap between API source code and its documentation by intercepting usage examples from API test code to automatically generate an always updated API documentation for REST APIs.
 
-In this thesis, I presented a set of five primary research objectives related to the topic of REST API documentation.
-I have addressed the first research objective, RO1, of finding the gap between the current industry practices and existing research on REST API documentation by performing a case study in my first research sub-project.
+I presented a set of five primary research questions related to the topic of REST API documentation.
+I have addressed the first research question, RQ1, of finding the gap between the current industry practices and existing research on REST API documentation by performing a case study in my first research sub-project.
 I started by reviewing the existing literature on Web API evolution.
@@
 
-I addressed RO2, finding REST API documentation requirements, in my second research sub-project.
+I addressed RQ2, finding REST API documentation requirements, in my second research sub-project.
 I started by analyzing the existing literature on API usability to summarize a list of general API documentation requirements as derived by other researchers.
@@
 Based on these two analyses, I derived the following list of documentation requirements specific for evolving REST APIs: automated, example based, executable, version aware, customizable, reusable, and collaborative.
-These requirements are used as a background work to address my subsequent research objectives.
+These requirements are used as a background work to address my subsequent research questions.
 
-RO3, the primary objective of finding a new technique to satisfy the documentation requirements of evolving REST APIs, is addressed in my third and fourth sub-projects.
+RQ3, the primary research question of finding a new technique to satisfy the documentation requirements of evolving REST APIs, is addressed in my third and fourth sub-projects.
 I observed that most existing techniques for REST API documentation relied on custom API specification formats as an input.
@@
 
-I addressed RO4, an evaluation of the proposed technique from the perspective of REST API developers, in my fourth sub-project using an industrial case study.
+I addressed RQ4, an evaluation of the proposed technique from the perspective of REST API developers, in my fourth sub-project using an industrial case study.
 I presented the proposed technique and the prototype tool to a team of software engineers at Cisco when we were in the process of selecting a documentation method for a suite of new REST APIs to be built.
@@
 
-My last research objective, RO5, to evaluate the impact of the proposed technique on REST API client developers, was addressed in my fifth sub-project.
+My last research question, RQ5, to evaluate the impact of the proposed technique on REST API client developers, was addressed in my fifth sub-project.
 I performed a controlled study with 26 professional software engineers with prior REST API experience.
@@
 \section{Future Work}
-Based on my findings from the this research, I identified the following list of potential future research problems.
+Based on my findings from the this research, I identified the following list of potential future research directions.
 
@@
   \item \textbf{Automated API Client Code Generation.} Several tools exist that can convert custom REST API description files to API client code in multiple programming languages.
+REST API client libraries help developers to leverage various IDE features such as code completion, and suggested data types for API elements.
 Future research can be carried out to map the SpyREST data model with the custom API description languages to leverage the interception technique to reduce the amount of manual effort required to maintain the custom formatted API descriptions as the API evolves.

### 2017-11-08 appendix.tex
adds the copyright info
adds ethics certificate
overview + questionnaire

@@
 \end{lstlisting}
+
+\chapter{Permissions for Reusing Published Papers in PhD Thesis}
+
+\begin{figure}
+ \centering
+ \includegraphics[width=\textwidth]{copyrights/services.pdf}
+\end{figure}
+\begin{figure}
+ \centering
+ \includegraphics[width=\textwidth]{copyrights/spy_rest.pdf}
+\end{figure}
+\begin{figure}
+ \centering
+ \includegraphics[width=\textwidth]{copyrights/tool.pdf}
+\end{figure}
+\begin{figure}
+ \centering
+ \includegraphics[width=\textwidth]{copyrights/cisco_paper.pdf}
+\end{figure}
+\chapter{Ethics Certificate}
+\label{ethics}
+\begin{figure}
+ \centering
+ \includegraphics[width=\textwidth]{ethics.pdf}
+\end{figure}
+\chapter{API Tasks for Controlled Study}
+\label{chap:appendix_questions}
+\begin{enumerate}
+  \item \textbf{List all Posts.} Use the WordPress REST API to get a list of all the blog posts from the blog at http://wp.spyrest.com"
+  \item \textbf{Find all posts by author Spyrest.} Use the WordPress REST API to find a list of all the blog posts by author Spyrest from the blog at http://wp.spyrest.com"
+  \item \textbf{List all Posts excluding ids 1 and 4} Use the WordPress REST API to find a list of all the blog posts where the post id is not 1 or 4 from the blog at http://wp.spyrest.com"
+  \item \textbf{Find the total number of Posts.} Use the WordPress REST API to find the total number of blog posts from the blog at http://wp.spyrest.com"
+  \item \textbf{Publish a Post} Use the WordPress REST API to publish a new post with title 'a random post', content 'lorem ipsum' and published date of 'January 1, 2016' on the blog at http://wp.spyrest.com"
+  \item \textbf{Update the excerpt of the Post you published to 'My new excerpt'.} Use the WordPress REST API to update the excerpt for the post you created from the blog at http://wp.spyrest.com."
+\end{enumerate}
+
+\chapter{Feedback Questionnaire for Controlled Study}
+
+\begin{enumerate}
+  \item Rate your experience of performing the API tasks using the \textbf{WordPress REST API} documentation (10 being the best possible).
+  \item Share your thoughts about the reasoning behind the above rating.
+\end{enumerate}

### 2017-11-08 introduction.tex
adds box borders
overview + questionnaire

@@
 \begin{figure}[htb]
-  \includegraphics[width=\linewidth]{javadoc.png}
+  \fbox{\includegraphics[width=\linewidth]{javadoc.png}}
   \captionsource{Example Local API Documentation from Formatted Comments Using JavaDoc}{Log4J}
@@
 \begin{figure}[htb]
-  \includegraphics[width=\linewidth]{wordpress_doc.png}
+  \fbox{\includegraphics[width=\linewidth]{wordpress_doc.png}}
   \captionsource{Difference between a REST API Code and its Documentation}{WordPress REST API}
@@
 To address the aforementioned research goals and objectives, my research strategy included five sub-projects, which have resulted in publications at peer-reviewed international conferences.
-Table~\ref{table:chapter_summary} shows a high level overview of the flow of arguments between the sub-projects to meet the research goals and questions.
+Figure~\ref{fig:flow} and Table~\ref{table:chapter_summary} show a high level overview of the flow of arguments between the sub-projects to meet the research goals and questions.
+
+\begin{figure}[htb]
+  \includegraphics[width=\linewidth]{flow.png}
+  \caption{Overview of Flow of Arguments of the Sub-Projects }
+  \label{fig:flow}
+\end{figure}
 
 \begin{table}
-  \caption{Summary and Flow of Arguments of the Sub-projects}
+  \caption{Summary of the Sub-projects}
   \begin{tabular}{|p{1.5cm}|p{1.5cm}|p{1.5cm}|p{4.5cm}|p{6cm}|}
